### 2014-12-10 manuscript/ms.tex
Reorganization

% PRL look and style (easy on the eyes)
\documentclass[aps,pre,twocolumn,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
% Two-column style (for submission/review/editing)
%\documentclass[aps,prl,preprint,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}

\usepackage{palatino}
\usepackage{amsmath}
\usepackage{amssymb}
\pdfoutput=1
\usepackage[pdftex]{graphicx}
%\usepackage[mathbf,mathcal]{euler}
%\usepackage{citesort}
\usepackage{dcolumn}
\usepackage{boxedminipage}
\usepackage{verbatim}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}

% The figures are in a figures/ subdirectory.
\graphicspath{{figures/}}

% \newcommand{\pyitc}{\url{http://www.simtk.org/home/bayesian-itc}} % URL of pyITC project homepage

%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Ensembler: Enabling high-throughput molecular simulations at the superfamily scale}

\author{Daniel L. Parton}
  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
  %\email{daniel.parton@choderalab.org}
\author{Patrick B. Grinaway}
  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
  %\email{patrick.grinaway@choderalab.org}
\author{John D. Chodera}
 \thanks{Corresponding author}
 \email{john.chodera@choderalab.org}
  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}

\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT/pacs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\label{section:methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{section:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DISCUSSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{section:discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
\label{section:acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliographystyle{prsty} 
\bibliography{ms.bib}

\end{document}

### 2015-01-23 manuscript/ms.tex
Completed abstract

@@
 
+The rapidly expanding body of available genomic and protein structural data
+provides a rich resource for the field of biomolecular simulation. However,
+simulations on an omics scale are not yet widely performed, partly because
+software has had trouble keeping pace. For example, it should be possible to
+study proteins across entire (super)families, and to do so in a way which
+exploits the entire variety of available structural biology data. Here, we
+present a new tool for enabling high-throughput simulation in the genomics era.
+Ensembler takes any set of sequences - from a single sequence to an entire
+superfamily of interest - and shepherds them through various stages of:
+comparative modeling to all relevant PDB structures, reconstruction of missing
+loops, addition of missing atoms, culling by close structural similarity,
+assignment of protonation states, solvation, and refinement with molecular
+simulation. The output is an ensemble of structures ready for subsequent
+parallel or distributed molecular simulations. This automates much of the
+time-consuming process of preparing protein models suitable for simulation,
+while also allowing this process to be scaled to the superfamily scale.  A
+particular advantage of this approach can be found in the construction of
+kinetic models of conformational dynamics - such as Markov state models - for
+which a diverse array of starting configurations is expected to aid sampling.
+We demonstrate the power of this approach by constructing initial models for
+all catalytic domains in the human tyrosine kinase family, using all kinase
+catalytic domain structures from any organism as structural templates.
+Ensembler should run on all major operating systems, and has been tested on
+Linux and OS X. The program is free of charge, and is made available under the
+terms of the GNU General Public License (GPL) v2. The latest release can be
+installed via the conda package manager, and the latest source can be
+downloaded from \url{https://github.com/choderalab/ensembler}.
+
+% TODO expand to all human protein kinases?
+
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% METHODS
+% DESIGN AND IMPLEMENTATION
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-\label{section:methods}
+\section{Design and Implementation}
+\label{section:design}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{section:discussion}
+\section{Availability and Future Directions}
+\label{section:availability}
 

### 2015-02-06 manuscript/ms.tex
Wrote introduction

@@
 % The figures are in a figures/ subdirectory.
-\graphicspath{{figures/}}
+% \graphicspath{{figures/}}
 
@@
 \section{Design and Implementation}
+
+% Python
+% Can be used via command-line tool Ensembler, or via the Python API
+
+Ensembler is written in Python, and can be used via a command-line tool ({\tt
+ensembler}), or via the flexible Python API.
+
+% Pipeline architecture, with a series of stages which must be performed in order.
+
+The Ensembler modeling pipeline entails a series of stages which are performed
+in a defined order. A visual overview of the pipeline is shown in Fig.
+\ref{figpipeline}, and a detailed description follows.
+
+\begin{figure*}[h]
+  \centering
+    \includegraphics[width=0.8\textwidth]{pipeline}
+  \caption{Ensembler pipeline}
+  \label{figpipeline}
+\end{figure*}
+
+\subsubsection{Target selection}
+
+The output from this stage is simply a FASTA-formatted sequence file containing
+the selected target sequences and identifiers. The {\tt ensembler} command-line
+tool provides methods for selecting targets from either UniProt (a freely
+accessible resource for protein sequence and functional
+data---\href{http://www.uniprot.org/}{uniprot.org}) or TargetExplorer (a
+database framework for aggregating various types of biological data; work to be
+published). This allows the user to easily select a single protein, many
+proteins, or an entire superfamily. Alternatively, the targets file can be
+generated using any other software, and stored at the appropriate filepath.
+
+The method for selecting targets from UniProt is described here. A query string
+is required as input, using the same syntax as the search function on the
+UniProt website. For example, {\tt'domain:"Protein kinase" AND taxonomy:9606
+AND reviewed:yes'} would select all human protein kinases which have been
+reviewed by a human curator. Ensembler is designed to work with protein
+\emph{domains}, rather than full-length proteins, and the desired protein
+domain(s) can be selected using a regular expression. For example, the string
+{\tt'\^{}Protein kinase(?!; truncated)(?!; inactive)'} would match domains
+annotated as "Protein kinase", "Protein kinase; 1" or "Protein kinase; 2", but
+would exclude the domains "Protein kinase; truncated" and "Protein kinase;
+inactive". The program then extracts sequences and identifiers from the UniProt
+data, and saves these to a FASTA-format text file.
+
+% The selected targets are
+% given identifiers of the form {\tt [UniProt mnemonic]\_D[domain index]}, which
+% consists of the UniProt name for the target protein and a 0-based index for the
+% domain (since a single target protein may contain multiple domains of
+% interest). Example: {\tt EGFR\_HUMAN\_D0}.
+
+\subsubsection{Template selection}
+
+As for target selection, the {\tt ensembler} tool provides methods for
+selecting templates from various resources---UniProt, the Protein Data Bank
+(PDB; \href{http://www.pdb.org}{pdb.org}) or a TargetExplorer database.  From
+the user perspective, selection from UniProt proceeds in a similar fashion as
+described above. The returned data for each UniProt entry includes an
+up-to-date list of PDB structures and their residue spans. PDB files are
+downloaded from the PDB for structures which include the desired domain. Data
+from the SIFTS service
+(\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts})
+(CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings
+between PDB and UniProt entries, is then used to filter out PDB chains with $<$
+70\% resolved residues within the domain span. Template sequences and
+structures are then extracted and written as FASTA-format sequence and
+PDB-format coordinate files respectively.
+
+Selection from the PDB simply requires specifying a list of PDB IDs. These are
+matched to UniProt entries via the SIFTS service, and the same procedure is
+then followed to extract template sequences and structures.
+
+Unresolved template loops can optionally be remodeled with a kinematic closure
+algorithm, which is provided via the loopmodel tool of the Rosetta software
+suite (CITE: Rosetta and/or loopmodel). This tends to provide higher-quality
+models following the subsequent modeling process.
+
+\subsubsection{Modeling}
+
+In this stage, models are generated for each target-template pair, using the
+Modeller automodel function (CITE: Modeller), which implements comparative
+structure modeling by satisfaction of spatial restraints (CITE: Sali Blundell J
+Mol Biol 1993; Fiser Sali Prot Sci 9 2000). Modeller requires the user to first
+provide a target-template sequence alignment. This is implemented in Ensembler
+using the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a
+dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet
+\textit{et al} (CITE: Gaston Gonnet Science 1992).
+
+Non-unique models are then filtered out using structural similarity-based
+clustering. The mdtraj (CITE: mdtraj) Python library is used to calculate RMSD
+with a fast quaternion characteristic polynomial (QCP) implementation, and the
+leader algorithm is then used to populate clusters. A minimum distance cutoff
+(default: 0.6 \AA) is used to retain only a single model per cluster.
+
+\subsubsection{Refinement}
+
+Models are then refined with a steepest descent energy minimization and a short
+molecular dynamics (MD) simulation with implicit solvent. This is implemented
+using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen
+for its flexible Python API, and high performance GPU-acclerated simulation
+code. The Amber99SB-ILDN force field is used (CITE: amber99sbildn refs) with a
+modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC). The
+simulation is run for a default of 100 ps. This refinement process helps to
+prepare models for subsequent production simulation, and also helps to filter
+out poor quality models.
+
+Ensembler also provides optional routines for solvating models with explicit
+solvent and performing a second MD refinement. The solvation step solvates each
+model for a given target with the same number of waters, as this is (currently)
+a requirement for building MSMs from multiple independent MD trajectories. The
+target number of waters is selected by first solvating each model with a
+specified padding distance (default: 10 \AA), then taking a percentile value
+from the distribution (default: 68th percentile). Models are resolvated with
+the target number of waters by first solvating with zero padding, then
+incrementally increasing the box size and resolvating until the target is
+exceeded, then finally deleting sufficient waters to match the target value.
+The explicit solvent MD simulation is also implemented using OpenMM, with the
+Amber99SB-ILDN force field and TIP3P water.
+
+\subsubsection{Packaging}
+
+Finally, Ensembler provides a packaging module, which can be used to compress
+models in preparation for data transfer, or to prepare models with the
+appropriate directory and file structure for subsequent production simulations
+on the distributed computing platform Folding@Home (CITE: F@H).
+
+\subsubsection{Other features}
+
+% quickmodel function performs the entire pipeline in one go, for a single target and a small number of templates.
+The command-line tool also provides a {\tt quickmodel} function, which performs
+the entire Ensembler pipeline for a single target with a small number of
+templates. For larger numbers of models (such as entire protein families), the
+main pipeline functions should be used. The modeling and refinement functions
+use MPI to trivially parallelize computation across each model (or across each
+template, in the case of the loop reconstruction code).
+
+Each pipeline function also outputs a metadata file, which helps to link data
+to the software version used to generate it (both Ensembler and its
+dependencies), and also provides timing and performance information, and other
+data such as hostname.
+
+% ? The API is highly flexible, allowing control over many important parameters. Appropriate choices have been set as default values.
+
+% TODO where do these go?
+% Distribution: conda package or github source
+% Software is open source and designed to be extensible as far as possible.
+For example, it would be very simple to implement further options for providing target and template structures, or for packaging models.
+
 \label{section:design}

### 2015-02-11 manuscript/ms.tex
Edited abstract.

@@
 
-The rapidly expanding body of available genomic and protein structural data
-provides a rich resource for the field of biomolecular simulation. However,
-simulations on an omics scale are not yet widely performed, partly because
-software has had trouble keeping pace. For example, it should be possible to
-study proteins across entire (super)families, and to do so in a way which
-exploits the entire variety of available structural biology data. Here, we
-present a new tool for enabling high-throughput simulation in the genomics era.
-Ensembler takes any set of sequences - from a single sequence to an entire
-superfamily of interest - and shepherds them through various stages of:
-comparative modeling to all relevant PDB structures, reconstruction of missing
-loops, addition of missing atoms, culling by close structural similarity,
-assignment of protonation states, solvation, and refinement with molecular
-simulation. The output is an ensemble of structures ready for subsequent
-parallel or distributed molecular simulations. This automates much of the
-time-consuming process of preparing protein models suitable for simulation,
-while also allowing this process to be scaled to the superfamily scale.  A
-particular advantage of this approach can be found in the construction of
-kinetic models of conformational dynamics - such as Markov state models - for
-which a diverse array of starting configurations is expected to aid sampling.
-We demonstrate the power of this approach by constructing initial models for
-all catalytic domains in the human tyrosine kinase family, using all kinase
-catalytic domain structures from any organism as structural templates.
-Ensembler should run on all major operating systems, and has been tested on
-Linux and OS X. The program is free of charge, and is made available under the
-terms of the GNU General Public License (GPL) v2. The latest release can be
-installed via the conda package manager, and the latest source can be
-downloaded from \url{https://github.com/choderalab/ensembler}.
+The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
+However, simulations on an \emph{omics} scale are not yet widely performed, partly because software infrastructure to enable this has not kept pace. 
+For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and similarities in conformations that homologous proteins can adopt.
+Here, we present a new tool for enabling high-throughput simulation in the genomics era.
+Ensembler takes any set of sequences---from a single sequence to an entire superfamily of interest---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
+This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement with molecular simulation to ensure stable simulation.
+The output of this pipeline is an ensemble of structures ready for subsequent parallel or distributed molecular simulations using clusters, supercomputers, or distributed computing projects like Folding@home.
+Ensembler automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.  
+A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
+We demonstrate the power of this approach by constructing models for all catalytic domains in the human tyrosine kinase family, using all available kinase catalytic domain structures from any organism as structural templates.
+
+Ensembler is free of charge, and is made available under the terms of the GNU General Public License (GPL) v2. 
+It should run on all major operating systems, and has been tested on Linux and OS X.
+The latest release can be installed via the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
+
+\emph{Keywords: molecular dynamics simulation; comparative modeling}
 

### 2015-02-11 manuscript/ms.tex
Minor edit to phrasing in abstract

@@
 
-\usepackage{palatino}
+%\usepackage{palatino}
+
+% Change to a sans serif font.
+\usepackage{sourcesanspro}
+\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
+\usepackage[T1]{fontenc}
+%\usepackage[font=sf,justification=justified]{caption}
+\usepackage[font=sf]{floatrow}
+
+% Rework captions to use sans serif font.
+\makeatletter
+\renewcommand\@make@capt@title[2]{%
+ \@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
+  {\textbf{#1}}\sf\@caption@fignum@sep#2\quad
+}%
+\makeatother
+
+\usepackage{listings} % For code examples
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+
 \usepackage{amsmath}
@@
 However, simulations on an \emph{omics} scale are not yet widely performed, partly because software infrastructure to enable this has not kept pace. 
-For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and similarities in conformations that homologous proteins can adopt.
+For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and conformational similarities across homologous proteins.
 Here, we present a new tool for enabling high-throughput simulation in the genomics era.

### 2015-02-16 manuscript/ms.tex

@@
 
-Ensembler is written in Python, and can be used via a command-line tool ({\tt
-ensembler}), or via the flexible Python API.
+Ensembler is written in Python, and can be used via a command-line tool ({\tt ensembler}), or via the flexible Python API.
 
@@
 
-The Ensembler modeling pipeline entails a series of stages which are performed
-in a defined order. A visual overview of the pipeline is shown in Fig.
-\ref{figpipeline}, and a detailed description follows.
+The Ensembler modeling pipeline entails a series of stages which are performed in a defined order. 
+A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}, and a detailed description follows.
 
@@
 
-The output from this stage is simply a FASTA-formatted sequence file containing
-the selected target sequences and identifiers. The {\tt ensembler} command-line
-tool provides methods for selecting targets from either UniProt (a freely
-accessible resource for protein sequence and functional
-data---\href{http://www.uniprot.org/}{uniprot.org}) or TargetExplorer (a
-database framework for aggregating various types of biological data; work to be
-published). This allows the user to easily select a single protein, many
-proteins, or an entire superfamily. Alternatively, the targets file can be
-generated using any other software, and stored at the appropriate filepath.
+The output from this stage is simply a FASTA-formatted sequence file containing the selected target sequences and identifiers. 
+The {\tt ensembler} command-line tool provides methods for selecting targets from either UniProt (a freely accessible resource for protein sequence and functional data---\href{http://www.uniprot.org/}{uniprot.org}) or TargetExplorer (a database framework for aggregating various types of biological data; work to be published).
+This allows the user to easily select a single protein, many proteins, or an entire superfamily.
+Alternatively, the targets file can be generated using any other software, and stored at the appropriate filepath.
 
-The method for selecting targets from UniProt is described here. A query string
-is required as input, using the same syntax as the search function on the
-UniProt website. For example, {\tt'domain:"Protein kinase" AND taxonomy:9606
-AND reviewed:yes'} would select all human protein kinases which have been
-reviewed by a human curator. Ensembler is designed to work with protein
-\emph{domains}, rather than full-length proteins, and the desired protein
-domain(s) can be selected using a regular expression. For example, the string
-{\tt'\^{}Protein kinase(?!; truncated)(?!; inactive)'} would match domains
-annotated as "Protein kinase", "Protein kinase; 1" or "Protein kinase; 2", but
-would exclude the domains "Protein kinase; truncated" and "Protein kinase;
-inactive". The program then extracts sequences and identifiers from the UniProt
-data, and saves these to a FASTA-format text file.
+The method for selecting targets from UniProt is described here. 
+A query string is required as input, using the same syntax as the search function on the UniProt website. 
+For example, {\tt'domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator. 
+Ensembler is designed to work with protein \emph{domains}, rather than full-length proteins, and the desired protein domain(s) can be selected using a regular expression. 
+For example, the string {\tt'\^{}Protein kinase(?!; truncated)(?!; inactive)'} would match domains annotated as "Protein kinase", "Protein kinase; 1" or "Protein kinase; 2", but would exclude the domains "Protein kinase; truncated" and "Protein kinase; inactive".
+The program then extracts sequences and identifiers from the UniProt data, and saves these to a FASTA-format text file.
 

### 2015-02-16 manuscript/ms.tex
Edits to Design and Implementation section

@@
 
-As for target selection, the {\tt ensembler} tool provides methods for
-selecting templates from various resources---UniProt, the Protein Data Bank
-(PDB; \href{http://www.pdb.org}{pdb.org}) or a TargetExplorer database.  From
-the user perspective, selection from UniProt proceeds in a similar fashion as
-described above. The returned data for each UniProt entry includes an
-up-to-date list of PDB structures and their residue spans. PDB files are
-downloaded from the PDB for structures which include the desired domain. Data
-from the SIFTS service
-(\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts})
-(CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings
-between PDB and UniProt entries, is then used to filter out PDB chains with $<$
-70\% resolved residues within the domain span. Template sequences and
-structures are then extracted and written as FASTA-format sequence and
-PDB-format coordinate files respectively.
+As for target selection, the {\tt ensembler} tool provides methods for selecting templates from various resources---UniProt, the Protein Data Bank (PDB; \href{http://www.pdb.org}{pdb.org}) or a TargetExplorer database.
+From the user perspective, selection from UniProt proceeds in a similar fashion as described above.
+The returned data for each UniProt entry includes an up-to-date list of PDB structures and their residue spans.
+PDB files are downloaded from the PDB for structures which include the desired domain.
+Data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries, is then used to filter out PDB chains with $<$ 70\% resolved residues within the domain span.
+Template sequences and structures are then extracted and written as FASTA-format sequence and PDB-format coordinate files respectively.
 
-Selection from the PDB simply requires specifying a list of PDB IDs. These are
-matched to UniProt entries via the SIFTS service, and the same procedure is
-then followed to extract template sequences and structures.
+Selection from the PDB simply requires specifying a list of PDB IDs.
+These are matched to UniProt entries via the SIFTS service, and the same procedure is then followed to extract template sequences and structures.
 
-Unresolved template loops can optionally be remodeled with a kinematic closure
-algorithm, which is provided via the loopmodel tool of the Rosetta software
-suite (CITE: Rosetta and/or loopmodel). This tends to provide higher-quality
-models following the subsequent modeling process.
+Unresolved template loops can optionally be remodeled with a kinematic closure algorithm, which is provided via the loopmodel tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
+This tends to provide higher-quality models following the subsequent modeling process.
 
@@
 
-In this stage, models are generated for each target-template pair, using the
-Modeller automodel function (CITE: Modeller), which implements comparative
-structure modeling by satisfaction of spatial restraints (CITE: Sali Blundell J
-Mol Biol 1993; Fiser Sali Prot Sci 9 2000). Modeller requires the user to first
-provide a target-template sequence alignment. This is implemented in Ensembler
-using the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a
-dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet
-\textit{et al} (CITE: Gaston Gonnet Science 1992).
+In this stage, models are generated for each target-template pair, using the Modeller automodel function (CITE: Modeller), which implements comparative structure modeling by satisfaction of spatial restraints (CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000).
+Modeller requires the user to first provide a target-template sequence alignment.
+This is implemented in Ensembler using the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al} (CITE: Gaston Gonnet Science 1992).
 
-Non-unique models are then filtered out using structural similarity-based
-clustering. The mdtraj (CITE: mdtraj) Python library is used to calculate RMSD
-with a fast quaternion characteristic polynomial (QCP) implementation, and the
-leader algorithm is then used to populate clusters. A minimum distance cutoff
-(default: 0.6 \AA) is used to retain only a single model per cluster.
+Non-unique models are then filtered out using structural similarity-based clustering.
+The mdtraj (CITE: mdtraj) Python library is used to calculate RMSD with a fast quaternion characteristic polynomial (QCP) implementation, and the leader algorithm is then used to populate clusters.
+A minimum distance cutoff (default: 0.6 \AA) is used to retain only a single model per cluster.
 
@@
 
-Models are then refined with a steepest descent energy minimization and a short
-molecular dynamics (MD) simulation with implicit solvent. This is implemented
-using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen
-for its flexible Python API, and high performance GPU-acclerated simulation
-code. The Amber99SB-ILDN force field is used (CITE: amber99sbildn refs) with a
-modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC). The
-simulation is run for a default of 100 ps. This refinement process helps to
-prepare models for subsequent production simulation, and also helps to filter
-out poor quality models.
+Models are then refined with a steepest descent energy minimization and a short molecular dynamics (MD) simulation with implicit solvent.
+This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
+The Amber99SB-ILDN force field is used (CITE: amber99sbildn refs) with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
+The simulation is run for a default of 100 ps.
+This refinement process helps to prepare models for subsequent production simulation, and also helps to filter out poor quality models.
 
-Ensembler also provides optional routines for solvating models with explicit
-solvent and performing a second MD refinement. The solvation step solvates each
-model for a given target with the same number of waters, as this is (currently)
-a requirement for building MSMs from multiple independent MD trajectories. The
-target number of waters is selected by first solvating each model with a
-specified padding distance (default: 10 \AA), then taking a percentile value
-from the distribution (default: 68th percentile). Models are resolvated with
-the target number of waters by first solvating with zero padding, then
-incrementally increasing the box size and resolvating until the target is
-exceeded, then finally deleting sufficient waters to match the target value.
-The explicit solvent MD simulation is also implemented using OpenMM, with the
-Amber99SB-ILDN force field and TIP3P water.
+Ensembler also provides optional routines for solvating models with explicit solvent and performing a second MD refinement.
+The solvation step solvates each model for a given target with the same number of waters, as this is (currently) a requirement for building MSMs from multiple independent MD trajectories.
+The target number of waters is selected by first solvating each model with a specified padding distance (default: 10 \AA), then taking a percentile value from the distribution (default: 68th percentile).
+Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
+The explicit solvent MD simulation is also implemented using OpenMM, with the Amber99SB-ILDN force field and TIP3P water.
 
@@
 
-Finally, Ensembler provides a packaging module, which can be used to compress
-models in preparation for data transfer, or to prepare models with the
-appropriate directory and file structure for subsequent production simulations
-on the distributed computing platform Folding@Home (CITE: F@H).
+Finally, Ensembler provides a packaging module, which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@Home (CITE: F@H).
 
@@
 % quickmodel function performs the entire pipeline in one go, for a single target and a small number of templates.
-The command-line tool also provides a {\tt quickmodel} function, which performs
-the entire Ensembler pipeline for a single target with a small number of
-templates. For larger numbers of models (such as entire protein families), the
-main pipeline functions should be used. The modeling and refinement functions
-use MPI to trivially parallelize computation across each model (or across each
-template, in the case of the loop reconstruction code).
+The command-line tool also provides a {\tt quickmodel} function, which performs the entire Ensembler pipeline for a single target with a small number of templates.
+For larger numbers of models (such as entire protein families), the main pipeline functions should be used.
+The modeling and refinement functions use MPI to trivially parallelize computation across each model (or across each template, in the case of the loop reconstruction code).
 
-Each pipeline function also outputs a metadata file, which helps to link data
-to the software version used to generate it (both Ensembler and its
-dependencies), and also provides timing and performance information, and other
-data such as hostname.
+Each pipeline function also outputs a metadata file, which helps to link data to the software version used to generate it (both Ensembler and its dependencies), and also provides timing and performance information, and other data such as hostname.
 

### 2015-02-16 manuscript/ms.tex
Added JDC's text for the Introduction

@@
 \label{section:introduction}
+
+Proteins play a diverse variety of roles in biology, and the understanding of their function---and how mutations can cause dysfunction and disease---is the preoccupation of much of modern biology. 
+The diminishing cost of nucleic acid sequencing technologies has produced an enormous wealth of sequence data, yielding a large collection of protein-coding open reading frames that provide basic information about these proteins (at the level of primary amino acid sequences) for numerous organisms [CITE].
+Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have yielded a great number of protein structures, allowing comparative modeling to provide insight into the static structures many of these proteins adopt [CITE review of structural biology efforts or current comparative modeling?].
+
+Static structures, however, provide only a snapshot of the rich dynamical behavior of proteins.
+Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require conformational changes.
+
+Molecular dynamics simulations have proven to be a useful tool for revealing the dynamics of individual proteins, with a number of software packages and forcefields available for biomolecular simulation.
+Advances in computing architectures---especially the availability of GPUs, which provide a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as those that drive the Folding@home distributed computing resource [CITE]) now provide new hardware platforms on which to study the dynamics of these proteins.
+In parallel, techniques for aggregating molecular dynamics simulation data to survey the kinetic landscape of biomolecules, such as Markov state modeling approaches [CITE MSM reviews], are now reaching maturity.
+
+Despite this, a critical gap remains in the ability to bridge genome-scale sequence information and molecular simulations to enable the study of entire families or superfamilies of proteins in a single organism or across organisms.
+Molecular simulations must largely be set up by hand, with little in the way of automation available to provide practitioners a way of studying many members of a family where only a subset may have structural data.
+Complicating matters further, in families of proteins known to be conformationally labile or able to adopt a diversity of structures---such as kinases---existing structural data may only exist for one or two conformations for any individual member of the superfamily for which there is structural data, leading to difficulties in trying to construct consistent kinetic models for comparison among different members of the same superfamily.
+
+Here, we present the first steps toward a resolution of this problem: A fully automated open source framework for building simulation-ready protein models scalable to the superfamily scale.
+
+\ldots
+
+We demonstrate the utility of this tool by constructing models for the entire set of human tyrosine kinase catalytic domains, and demonstrate that the resulting models provide good coverage of the known functional regions of structure space.
+This tool forms the foundation for a new era of superfamily-scale molecular simulations for the Folding@home project.
 

### 2015-02-16 manuscript/ms.tex
DLP comment on Introduction
Edits to Introduction

@@
 Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have yielded a great number of protein structures, allowing comparative modeling to provide insight into the static structures many of these proteins adopt [CITE review of structural biology efforts or current comparative modeling?].
+% TODO Maybe it is not necessary to mention comparative modeling here? If so, this could be changed to:
+% Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have provided insight into the static structures which these proteins can adopt [CITE review of structural biology efforts?].
 
@@
 Molecular simulations must largely be set up by hand, with little in the way of automation available to provide practitioners a way of studying many members of a family where only a subset may have structural data.
-Complicating matters further, in families of proteins known to be conformationally labile or able to adopt a diversity of structures---such as kinases---existing structural data may only exist for one or two conformations for any individual member of the superfamily for which there is structural data, leading to difficulties in trying to construct consistent kinetic models for comparison among different members of the same superfamily.
 
-Here, we present the first steps toward a resolution of this problem: A fully automated open source framework for building simulation-ready protein models scalable to the superfamily scale.
+Complicating matters further, in protein families known to be able to adopt multiple conformations---such as kinases---structural data may only exist for one or two conformations for any individual member of the family for which there is structural data.
+This poses a challenge for biomolecular simulation methods such as Markov state models, which can provide detailed insight, but require global coverage of the conformational landscape to realize their full potential.
 
-\ldots
-
+Here, we present the first steps toward a resolution of this problem: a fully automated open source framework for building simulation-ready protein models scalable to the superfamily scale.
 We demonstrate the utility of this tool by constructing models for the entire set of human tyrosine kinase catalytic domains, and demonstrate that the resulting models provide good coverage of the known functional regions of structure space.

### 2015-02-18 figures/rmsddist/rmsddist.py
Added first version of RMSD distribution figure

@@ -1,0 +1,57 @@
+import os
+import ensembler
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+data_dir = os.path.join('..', '..', 'data')
+targets = open(os.path.join(data_dir, 'TKs.txt')).read().splitlines()
+
+models_target_dir = os.path.join(data_dir, 'models', targets[0])
+rmsds_filepath = os.path.join(models_target_dir, 'rmsds.csv')
+combined_dfs = pd.read_csv(rmsds_filepath, index_col=0)
+
+for target in targets[1:]:
+    models_target_dir = os.path.join(data_dir, 'models', target)
+    rmsds_filepath = os.path.join(models_target_dir, 'rmsds.csv')
+    combined_dfs = pd.concat([combined_dfs, pd.read_csv(rmsds_filepath, index_col=0)]).reset_index(drop=True)
+
+df_has_model = combined_dfs[combined_dfs.has_model]
+
+seqid_mins = np.arange(0., 100., 20.)
+seqid_maxs = np.arange(20., 101., 20.)
+seqid_ranges = zip(seqid_mins, seqid_maxs)
+
+seqid_maxs[-1] = 101.
+
+sns.set(style="white")
+
+# palette = sns.blend_palette([sns.desaturate('royalblue', 0), 'royalblue'], 5)
+palette = sns.color_palette('GnBu_d', n_colors=5)[::-1]
+
+sns.kdeplot(df_has_model.rmsd*10, label='0-100 ($n=$%d)' % len(df_has_model), color='red', bw=02.5)
+
+for i, seqid_range in enumerate(seqid_ranges):
+    df_seqid_bin = df_has_model[df_has_model.seqid >= seqid_range[0]][df_has_model.seqid < seqid_range[1]]
+    print seqid_range, len(df_seqid_bin)
+    label = '%.0f-%.0f ($n=$%d)' % (seqid_range[0], seqid_range[1], len(df_seqid_bin))
+
+    # sns.distplot(df_seqid_bin.rmsd, label=labels[i], hist=False)
+    sns.kdeplot(df_seqid_bin.rmsd*10, label=label, color=palette[i], bw=2.5)
+
+    # if i == 4:
+    #     sns.rugplot(df_seqid_bin.rmsd*10, color=palette[i])
+
[…]

### 2015-02-18 figures/rmsddist/rmsddist.py
Added sequence identity distribution plot

@@ -55,3 +55,14 @@ plt.xlim(0,40)
 plt.xlabel('RMSD ($\AA$)')
 sns.despine(left=True)
 plt.savefig('rmsddist.png')
+
+plt.close()
+
+# Plot sequence identity distribution
+
+sns.set(style="white")
+sns.kdeplot(df_has_model.seqid)
+sns.despine(left=True)
+plt.setp(plt.gca(), yticks=[])
+plt.xlim(0,102)
+plt.savefig('seqid_dist.png')

### 2015-02-19 figures/rmsddist/rmsddist2.py
Added rmsddist figure v2

@@ -1,0 +1,68 @@
+import os
+import ensembler
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+data_dir = os.path.join('..', '..', 'data')
+targets = open(os.path.join(data_dir, 'TKs.txt')).read().splitlines()
+
+models_target_dir = os.path.join(data_dir, 'models', targets[0])
+rmsds_filepath = os.path.join(models_target_dir, 'rmsds.csv')
+combined_dfs = pd.read_csv(rmsds_filepath, index_col=0)
+
+for target in targets[1:]:
+    models_target_dir = os.path.join(data_dir, 'models', target)
+    rmsds_filepath = os.path.join(models_target_dir, 'rmsds.csv')
+    combined_dfs = pd.concat([combined_dfs, pd.read_csv(rmsds_filepath, index_col=0)]).reset_index(drop=True)
+
+df_has_model = combined_dfs[combined_dfs.has_model]
+
+seqid_mins = np.arange(0., 100., 100./3.)
+seqid_maxs = np.arange(100./3., 101., 100./3.)
+seqid_ranges = zip(seqid_mins, seqid_maxs)
+
+seqid_maxs[-1] = 101.
+
+sns.set(style="white")
+
+# palette = sns.blend_palette([sns.desaturate('royalblue', 0), 'royalblue'], 5)
+palette = sns.color_palette('GnBu_d', n_colors=3)[::-1]
+
+# sns.kdeplot(df_has_model.rmsd*10, label='0-100 ($n=$%d)' % len(df_has_model), color='red', bw=02.5)
+
+for i, seqid_range in enumerate(seqid_ranges):
+    df_seqid_bin = df_has_model[df_has_model.seqid >= seqid_range[0]][df_has_model.seqid < seqid_range[1]]
+    print seqid_range, len(df_seqid_bin)
+    label = '%.0f-%.0f ($n=$%d)' % (seqid_range[0], seqid_range[1], len(df_seqid_bin))
+
+    # sns.distplot(df_seqid_bin.rmsd, label=labels[i], hist=False)
+    sns.kdeplot(df_seqid_bin.rmsd*10, label=label, color=palette[i], bw=2.5)
+
+    # if i == 4:
+    #     sns.rugplot(df_seqid_bin.rmsd*10, color=palette[i])
+
[…]

### 2015-02-19 figures/rmsddist/rmsddist.py
Changes to sequence identity distribution figure

@@ -60,9 +60,19 @@ plt.close()
 
 # Plot sequence identity distribution
 
+fig, ax1 = plt.subplots()
+
+palette = sns.color_palette('deep', 2)
+
 sns.set(style="white")
-sns.kdeplot(df_has_model.seqid)
-sns.despine(left=True)
-plt.setp(plt.gca(), yticks=[])
-plt.xlim(0,102)
+sns.kdeplot(df_has_model.seqid, bw=1., ax=ax1, color=palette[0])
+ax1.set_ylabel('probability density', color=palette[0])
+ax1.legend_.remove()
+ax2 = ax1.twinx()
+sns.kdeplot(df_has_model.seqid, bw=1., cumulative=True, ax=ax2, color=palette[1])
+ax2.set_ylabel('cumulative probability density', color=palette[1])
+ax2.legend_.remove()
+# sns.despine(left=True)
+# plt.setp(plt.gca(), yticks=[])
+plt.xlim(0,100)
 plt.savefig('seqid_dist.png')

### 2015-02-19 figures/rmsddist/rmsddist.py
Minor

@@ -68,6 +68,7 @@ sns.set(style="white")
 sns.kdeplot(df_has_model.seqid, bw=1., ax=ax1, color=palette[0])
 ax1.set_ylabel('probability density', color=palette[0])
 ax1.legend_.remove()
+ax1.set_xlabel('sequence identity (%)')
 ax2 = ax1.twinx()
 sns.kdeplot(df_has_model.seqid, bw=1., cumulative=True, ax=ax2, color=palette[1])
 ax2.set_ylabel('cumulative probability density', color=palette[1])

### 2015-02-19 figures/rmsddist/rmsddist.py
Updated sequence identity distribution figure

@@ -69,11 +69,13 @@ sns.kdeplot(df_has_model.seqid, bw=1., ax=ax1, color=palette[0])
 ax1.set_ylabel('probability density', color=palette[0])
 ax1.legend_.remove()
 ax1.set_xlabel('sequence identity (%)')
-ax2 = ax1.twinx()
-sns.kdeplot(df_has_model.seqid, bw=1., cumulative=True, ax=ax2, color=palette[1])
-ax2.set_ylabel('cumulative probability density', color=palette[1])
+ax2 = ax1.twinx().twiny()
+sns.kdeplot(100.-df_has_model.seqid, bw=1., cumulative=True, ax=ax2, color=palette[1])
 ax2.legend_.remove()
 # sns.despine(left=True)
 # plt.setp(plt.gca(), yticks=[])
-plt.xlim(0,100)
+ax1.set_xlim(100,0)
+ax2.set_xlim(0,100)
+ax2.set_ylabel('CDF', color=palette[1])
+plt.setp(ax2, xticks=[])
 plt.savefig('seqid_dist.png')

### 2015-02-19 figures/rmsddist/rmsddist2.py

@@ -38,7 +38,7 @@ for i, seqid_range in enumerate(seqid_ranges):
     label = '%.0f-%.0f ($n=$%d)' % (seqid_range[0], seqid_range[1], len(df_seqid_bin))
 
     # sns.distplot(df_seqid_bin.rmsd, label=labels[i], hist=False)
-    sns.kdeplot(df_seqid_bin.rmsd*10, label=label, color=palette[i], bw=2.5)
+    sns.kdeplot(df_seqid_bin.rmsd*10, label=label, color=palette[i], bw=2.5, shade=True)
 
     # if i == 4:
     #     sns.rugplot(df_seqid_bin.rmsd*10, color=palette[i])

### 2015-02-19 figures/rmsddist/rmsddist2.py
Updated rmsd distribution figure v2

@@ -19,11 +19,18 @@ for target in targets[1:]:
 
 df_has_model = combined_dfs[combined_dfs.has_model]
 
-seqid_mins = np.arange(0., 100., 100./3.)
-seqid_maxs = np.arange(100./3., 101., 100./3.)
-seqid_ranges = zip(seqid_mins, seqid_maxs)
+# seqid_mins = np.arange(0., 100., 100./3.)
+# seqid_maxs = np.arange(100./3., 101., 100./3.)
+# seqid_ranges = zip(seqid_mins, seqid_maxs)
+seqid_ranges = [
+    [0., 35.],
+    [35., 55.],
+    [55., 100.],
+]
 
-seqid_maxs[-1] = 101.
+seqid_ranges_labels = seqid_ranges
+
+seqid_ranges[-1][-1] = 101.
 
 sns.set(style="white")
 
@@ -35,7 +42,7 @@ palette = sns.color_palette('GnBu_d', n_colors=3)[::-1]
 for i, seqid_range in enumerate(seqid_ranges):
     df_seqid_bin = df_has_model[df_has_model.seqid >= seqid_range[0]][df_has_model.seqid < seqid_range[1]]
     print seqid_range, len(df_seqid_bin)
-    label = '%.0f-%.0f ($n=$%d)' % (seqid_range[0], seqid_range[1], len(df_seqid_bin))
+    label = '%.0f-%.0f ($n=$%d)' % (seqid_ranges_labels[i][0], seqid_ranges_labels[i][1], len(df_seqid_bin))
 
     # sns.distplot(df_seqid_bin.rmsd, label=labels[i], hist=False)
     sns.kdeplot(df_seqid_bin.rmsd*10, label=label, color=palette[i], bw=2.5, shade=True)

### 2015-02-19 figures/rmsddist/rmsddist2.py
Minor

@@ -1,5 +1,6 @@
 import os
 import ensembler
+from copy import deepcopy
 import pandas as pd
 import numpy as np
 import seaborn as sns
@@ -28,7 +29,7 @@ seqid_ranges = [
     [55., 100.],
 ]
 
-seqid_ranges_labels = seqid_ranges
+seqid_ranges_labels = deepcopy(seqid_ranges)
 
 seqid_ranges[-1][-1] = 101.
 

### 2015-02-22 figures/attrition/accumulate_attrition_rates.py
Added attrition figure and data

@@ -1,0 +1,49 @@
+import os
+import numpy as np
+import pandas as pd
+
+targetids = open('../../data/TKs.txt').read().splitlines()
+
+seqid_ranges = ['all', '0-35', '35-55', '55-101']
+project_stages = ['templates', 'models', 'unique', 'implicit_refined']
+
+counts = {}
+rates = {}
+
+for seqid_range in seqid_ranges:
+    if seqid_range == 'all':
+        counts_filename = 'counts.csv'
+    else:
+        counts_filename = 'counts-seqid{0}.csv'.format(seqid_range)
+
+    # print 'Working on target', targetids[0]
+    counts_path = os.path.join('../../data/models', targetids[0], counts_filename)
+    df = pd.read_csv(counts_path)
+
+    for targetid in targetids[1:]:
+        # print 'Working on target', targetid
+        counts_path = os.path.join('../../data/models', targetid, counts_filename)
+        new_df = pd.read_csv(counts_path)
+        df += new_df
+
+    counts[seqid_range] = [
+        df.templates.values[0],
+        df.models.values[0],
+        df.unique_models.values[0],
+        df.implicit_refined.values[0],
+    ]
+
+    rates[seqid_range] = [
+        1.0,
+        counts[seqid_range][1] / np.float64(counts[seqid_range][0]),
+        counts[seqid_range][2] / np.float64(counts[seqid_range][1]),
+        counts[seqid_range][3] / np.float64(counts[seqid_range][2]),
+    ]
+
+counts = pd.DataFrame(counts, index=project_stages, columns=seqid_ranges)
+counts.to_csv('counts.csv')
+print counts.to_string()
+
+rates = pd.DataFrame(rates, index=project_stages, columns=seqid_ranges)
+rates.to_csv('rates.csv')
+print rates.to_string(float_format='{:.3f}'.format)

### 2015-02-22 figures/attrition/attrition.py

@@ -1,24 +1,0 @@
-import seaborn as sns
-
-
-counts = {
-    'all': [398970, 382568, 378839, 373513],
-    '0-35': [343393, 333564, 332055, 327250],
-    '35-55': [50643, 44602, 42907, 42431],
-    '55-100': [4934, 4402, 3877, 3832],
-}
-
-rates = {
-    'all': [1.0, 0.95888913953430077, 0.99025271324313591, 0.98594125736790561],
-    '0-35': [1.0, 0.9713768189800025, 0.9954761305176818, 0.98552950565418385],
-    '35-55': [1.0, 0.88071401773196689, 0.96199721985561182, 0.98890623907520914],
-    '55-100': [1.0, 0.89217673287393595, 0.880736029077692, 0.98839308743874132],
-}
-
-seqid_ranges = ['all', '0-35', '35-55', '55-100']
-
-for seqid_range in seqid_ranges:
-    sns.plt.plot(rates[seqid_range], label=seqid_range)
-
-sns.plt.legend()
-sns.plt.savefig('attrition_rates.png')

### 2015-02-22 figures/attrition/mkfig.py

@@ -1,0 +1,39 @@
+import pandas as pd
+import seaborn as sns
+
+counts = pd.read_csv('counts.csv', index_col=0)
+    # 'all': [398970, 382568, 378839, 373513],
+    # '0-35': [343393, 333564, 332055, 327250],
+    # '35-55': [50643, 44602, 42907, 42431],
+    # '55-100': [4934, 4402, 3877, 3832],
+
+print counts
+
+rates = pd.read_csv('rates.csv', index_col=0)
+    # 'all': [1.0, 0.95888913953430077, 0.99025271324313591, 0.98594125736790561],
+    # '0-35': [1.0, 0.9713768189800025, 0.9954761305176818, 0.98552950565418385],
+    # '35-55': [1.0, 0.88071401773196689, 0.96199721985561182, 0.98890623907520914],
+    # '55-100': [1.0, 0.89217673287393595, 0.880736029077692, 0.98839308743874132],
+
+seqid_ranges = ['all', '0-35', '35-55', '55-101']
+project_stages = ['templates', 'models', 'unique', 'implicit_refined']
+
+ax = rates.plot()
+
+# for seqid_range in seqid_ranges:
+#     sns.plt.plot(rates[seqid_range], label=seqid_range)
+
+# sns.plt.legend()
+
+palette = ['k'] + sns.color_palette('GnBu_d', n_colors=3)[::-1]
+
+for i, line in enumerate(ax.lines):
+    sns.plt.setp(line, color=palette[i])
+
+sns.plt.legend(loc='best')
+sns.plt.ylim(0.85,1)
+sns.plt.savefig('attrition_rates.png')
+sns.plt.close()
+
+# counts.plot()
+# sns.plt.savefig('attrition_counts.png')

### 2015-02-24 manuscript/ms.tex
Minor edit to Introduction
Added timing and model_counts tables

@@
 \usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[group-separator={,}]{siunitx}
 
@@
 Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have yielded a great number of protein structures, allowing comparative modeling to provide insight into the static structures many of these proteins adopt [CITE review of structural biology efforts or current comparative modeling?].
-% TODO Maybe it is not necessary to mention comparative modeling here? If so, this could be changed to:
+% TODO Maybe it is not necessary to mention comparative modeling here? The structures themselves provide insight. If so, this could be changed to:
 % Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have provided insight into the static structures which these proteins can adopt [CITE review of structural biology efforts?].
@@
 
+\begin{table}
+\begin{tabular}{l|c|c|c}
+
+    & SRC & ABL & All 90 TKs \\
+    \hline
+    Templates & \num{4433} & \num{4433} & \num{398970} \\
+    Models & \num{4248} & \num{4250} & \num{382568} \\
+    Unique models & \num{4093} & \num{4247} & \num{378839} \\
+    Refined models & \num{4023} & \num{4241} & \num{373513} \\
+
+\end{tabular}
+\label{tab:model_counts}
+\caption{Model counts}
+\end{table}
+
+\begin{table}
+\begin{tabular}{l|c}
+
+    & Average timings per template/model (h:m:s) \\
+    \hline
+    Template reconstruction & 03:00:00 \\
+    Modeling & 00:01:19 \\
+    Refinement (100 ps implicit solvent MD) & 00:01:58 \\
+
+\end{tabular}
+\label{tab:timings}
+\caption{Pipeline timings for 90 TKs}
+\end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-02-25 figures/superposition-src_abl/genmodels.py
Added superposition figure

@@ -1,0 +1,77 @@
+import os
+import shutil
+import tempfile
+import numpy as np
+import pandas as pd
+import mdtraj
+import seaborn as sns
+
+srcdir = '../../data/models/SRC_HUMAN_D0'
+
+df = pd.read_csv(os.path.join(srcdir, 'modelstraj-data.csv'))
+
+traj = mdtraj.load(os.path.join(srcdir, 'modelstraj.xtc'), top=os.path.join(srcdir, 'modelstraj-topol.pdb'))
+
+nmodels = 20
+
+modelsdir = 'models'
+if os.path.exists(modelsdir):
+    shutil.rmtree(modelsdir)
+    os.mkdir(modelsdir)
+else:
+    os.mkdir(modelsdir)
+
+# model_indices = np.arange(nmodels) * (len(traj) / nmodels)
+
+desired_seqid_values = 100. - np.arange(0., 100., 100./nmodels)
+model_indices = [np.abs(df.seqid.values-seqid_value).argmin() for seqid_value in desired_seqid_values]
+model_indices = sorted(list(set(model_indices)))
+nmodels = len(model_indices)
+seqids = [df.seqid.values[m] for m in model_indices]
+
+print 'Number of models selected:', nmodels
+print model_indices
+print seqids
+
+for m,f in enumerate(model_indices):
+    frame = traj[f]
+    frame.save_pdb(os.path.join(modelsdir, '{0}.pdb'.format(m)))
+
+
+# seqids = [df.seqid[f] for f in model_indices]
+transparency_floats = [(1. - seqid*0.01) for seqid in seqids]
+transparency_text = '\n'.join(['set cartoon_transparency, {0}, {1}'.format(transparency_floats[m], m) for m in range(nmodels)])
+
+palette = sns.color_palette('deep', nmodels)
+set_color_text = '\n'.join(['set_color {0}= [{1}, {2}, {3}]'.format(m, palette[m][0], palette[m][1], palette[m][2]) for m in range(nmodels)])
+color_text = '\n'.join(['color {0}, {0}'.format(m) for m in range(nmodels)])
+
+renderscripttext = """\
+bg white
+# set ray_opaque_background, off
+# set ray_trace_mode, 0
+# set ambient, 1
+# set reflect, 0
+set antialias, 1
+set two_sided_lighting, on
+set cartoon_fancy_helices, 1
+
+hide
+show cartoon
+
+{0}
+
+{1}
+
+{2}
+
+color black
+
+zoom all
[…]

### 2015-02-25 figures/superposition-src_abl/genmodels.py
Minor

@@ -30,8 +30,7 @@ nmodels = len(model_indices)
 seqids = [df.seqid.values[m] for m in model_indices]
 
 print 'Number of models selected:', nmodels
-print model_indices
-print seqids
+print '\n'.join(['{:6d} {:6.1f}'.format(x[0], x[1]) for x in zip(model_indices, seqids)])
 
 for m,f in enumerate(model_indices):
     frame = traj[f]

### 2015-02-25 figures/superposition-src_abl/genmodels.py
Superposition fig

@@ -41,9 +41,9 @@ for m,f in enumerate(model_indices):
 transparency_floats = [(1. - seqid*0.01) for seqid in seqids]
 transparency_text = '\n'.join(['set cartoon_transparency, {0}, {1}'.format(transparency_floats[m], m) for m in range(nmodels)])
 
-palette = sns.color_palette('deep', nmodels)
-set_color_text = '\n'.join(['set_color {0}= [{1}, {2}, {3}]'.format(m, palette[m][0], palette[m][1], palette[m][2]) for m in range(nmodels)])
-color_text = '\n'.join(['color {0}, {0}'.format(m) for m in range(nmodels)])
+palette = sns.color_palette('GnBu_d', nmodels)
+set_color_text = '\n'.join(['set_color color{0}, [{1}, {2}, {3}]'.format(m, palette[m][0], palette[m][1], palette[m][2]) for m in range(nmodels)])
+color_text = '\n'.join(['color color{0}, {0}'.format(m) for m in range(nmodels)])
 
 renderscripttext = """\
 bg white
@@ -64,7 +64,7 @@ show cartoon
 
 {2}
 
-color black
+# color black
 
 zoom all
 

### 2015-02-25 figures/superposition-src_abl/genmodels2.py
Superposition fig RdBu

@@ -1,6 +1,7 @@
 import os
 import shutil
 import tempfile
+import random
 import numpy as np
 import pandas as pd
 import mdtraj
@@ -23,11 +24,17 @@ else:
 
 # model_indices = np.arange(nmodels) * (len(traj) / nmodels)
 
-desired_seqid_values = 100. - np.arange(0., 100., 100./nmodels)
-model_indices = [np.abs(df.seqid.values-seqid_value).argmin() for seqid_value in desired_seqid_values]
-model_indices = sorted(list(set(model_indices)))
-nmodels = len(model_indices)
-seqids = [df.seqid.values[m] for m in model_indices]
+seqid_classes = [[0., 35.], [35., 55.], [55., 101.]]
+nmodels_per_class = 3
+model_indices = []
+seqids = []
+for seqid_class in seqid_classes:
+    df_class = df[df.seqid >= seqid[0]][df.seqid < seqid[1]]
+    selected_model_indices = random.sample(df_class.index, nmodels_per_class)
+    selected_models = df_class[selected_model_indices].sort()
+    model_indices += list(selected_models.index)
+    seqids += selected_models.seqid.values
+
 
 print 'Number of models selected:', nmodels
 print '\n'.join(['{:6d} {:6.1f}'.format(x[0], x[1]) for x in zip(model_indices, seqids)])

### 2015-02-25 figures/superposition-src_abl/genmodels.py
Superposition figure, models picked from seqid ranges

@@ -1,3 +1,5 @@
+# Pick models across a wide range of sequence identities
+
 import os
 import shutil
 import tempfile
@@ -41,7 +43,7 @@ for m,f in enumerate(model_indices):
 transparency_floats = [(1. - seqid*0.01) for seqid in seqids]
 transparency_text = '\n'.join(['set cartoon_transparency, {0}, {1}'.format(transparency_floats[m], m) for m in range(nmodels)])
 
-palette = sns.color_palette('RdBu', nmodels)[::-1]
+palette = sns.color_palette('Bluered', nmodels)[::-1]
 set_color_text = '\n'.join(['set_color color{0}, [{1}, {2}, {3}]'.format(m, palette[m][0], palette[m][1], palette[m][2]) for m in range(nmodels)])
 color_text = '\n'.join(['color color{0}, {0}'.format(m) for m in range(nmodels)])
 

### 2015-02-25 figures/superposition-src_abl/genmodels2.py

@@ -1,3 +1,5 @@
+# Randomly pick models from each of three sequence identity classes
+
 import os
 import shutil
 import tempfile
@@ -9,11 +11,11 @@ import seaborn as sns
 
 srcdir = '../../data/models/SRC_HUMAN_D0'
 
+nmodels_per_class = 3
+
 df = pd.read_csv(os.path.join(srcdir, 'modelstraj-data.csv'))
 
 traj = mdtraj.load(os.path.join(srcdir, 'modelstraj.xtc'), top=os.path.join(srcdir, 'modelstraj-topol.pdb'))
-
-nmodels = 20
 
 modelsdir = 'models'
 if os.path.exists(modelsdir):
@@ -22,19 +24,17 @@ if os.path.exists(modelsdir):
 else:
     os.mkdir(modelsdir)
 
-# model_indices = np.arange(nmodels) * (len(traj) / nmodels)
-
-seqid_classes = [[0., 35.], [35., 55.], [55., 101.]]
-nmodels_per_class = 3
+seqid_classes = [[0., 35.], [35., 55.], [55., 101.]][::-1]
 model_indices = []
 seqids = []
 for seqid_class in seqid_classes:
-    df_class = df[df.seqid >= seqid[0]][df.seqid < seqid[1]]
+    df_class = df[df.seqid >= seqid_class[0]][df.seqid < seqid_class[1]]
     selected_model_indices = random.sample(df_class.index, nmodels_per_class)
-    selected_models = df_class[selected_model_indices].sort()
+    selected_models = df_class.loc[selected_model_indices].sort()
     model_indices += list(selected_models.index)
-    seqids += selected_models.seqid.values
+    seqids += list(selected_models.seqid.values)
 
+nmodels = len(model_indices)
 
 print 'Number of models selected:', nmodels
 print '\n'.join(['{:6d} {:6.1f}'.format(x[0], x[1]) for x in zip(model_indices, seqids)])

### 2015-02-25 figures/superposition-src_abl/genmodels3.py
Added clustering version of superposition figure

@@ -1,4 +1,4 @@
-# Randomly pick models from each of three sequence identity classes
+# Performs RMSD-based clustering on all models and picks centroids
 
 import os
 import shutil
@@ -6,16 +6,44 @@ import tempfile
 import random
 import numpy as np
 import pandas as pd
-import mdtraj
+import mdtraj as md
+import msmbuilder.cluster
 import seaborn as sns
 
 srcdir = '../../data/models/SRC_HUMAN_D0'
 
-nmodels_per_class = 3
+nmodels = 8
 
 df = pd.read_csv(os.path.join(srcdir, 'modelstraj-data.csv'))
 
-traj = mdtraj.load(os.path.join(srcdir, 'modelstraj.xtc'), top=os.path.join(srcdir, 'modelstraj-topol.pdb'))
+traj = md.load(os.path.join(srcdir, 'modelstraj.xtc'), top=os.path.join(srcdir, 'modelstraj-topol.pdb'))
+
+atom_indices = traj.topology.select_atom_indices('alpha')
+traj_for_clustering = traj.atom_slice(atom_indices)
+
+X = [traj_for_clustering]
+
+clusterer = msmbuilder.cluster.KMedoids(n_clusters=nmodels, metric='rmsd')
+clusterer.fit(X)
+
+model_indices = clusterer.cluster_ids_
+model_indices.sort()
+seqids = df.loc[model_indices].seqid.values
+
+# seqid_classes = [[0., 35.], [35., 55.], [55., 101.]][::-1]
+# model_indices = []
+# seqids = []
+# for seqid_class in seqid_classes:
+#     df_class = df[df.seqid >= seqid_class[0]][df.seqid < seqid_class[1]]
+#     selected_model_indices = random.sample(df_class.index, nmodels_per_class)
+#     selected_models = df_class.loc[selected_model_indices].sort()
+#     model_indices += list(selected_models.index)
+#     seqids += list(selected_models.seqid.values)
+#
+# nmodels = len(model_indices)
+
+print 'Number of models selected:', nmodels
+print '\n'.join(['{:6d} {:6.1f}'.format(x[0], x[1]) for x in zip(model_indices, seqids)])
 
 modelsdir = 'models'
 if os.path.exists(modelsdir):
@@ -23,21 +51,6 @@ if os.path.exists(modelsdir):
     os.mkdir(modelsdir)
 else:
     os.mkdir(modelsdir)
-
[…]

### 2015-02-26 figures/superposition-src_abl/genmodels2.py
Updated superposition fig

@@ -71,7 +71,7 @@ show cartoon
 
 {2}
 
-color black
+# color black
 
 zoom all
 

### 2015-02-26 figures/superposition-src_abl/genmodels4.py
Updated superposition fig

@@ -1,0 +1,106 @@
+# Three seqid classes (manually chosen). RMSD clustering performed within each class, and centroids picked.
+
+import os
+import shutil
+import tempfile
+import random
+import numpy as np
+import pandas as pd
+import mdtraj as md
+import msmbuilder.cluster
+import seaborn as sns
+
+srcdir = '../../data/models/SRC_HUMAN_D0'
+
+nmodels_per_class = 3
+
+seqid_classes = [[0., 35.], [35., 55.], [55., 101.]][::-1]
+
+traj = md.load(os.path.join(srcdir, 'modelstraj.xtc'), top=os.path.join(srcdir, 'modelstraj-topol.pdb'))
+
+df = pd.read_csv(os.path.join(srcdir, 'modelstraj-data.csv'))
+
+model_indices = []
+seqids = []
+
+for seqid_class in seqid_classes:
+    seqid_class_df = df[df.seqid >= seqid_class[0]][df.seqid < seqid_class[1]]
+    seqid_class_model_indices = list(seqid_class_df.index)
+    seqid_class_traj = traj[seqid_class_model_indices]
+    alpha_atom_indices = seqid_class_traj.topology.select_atom_indices('alpha')
+    seqid_class_traj_alphas = seqid_class_traj.atom_slice(alpha_atom_indices)
+
+    X = [seqid_class_traj_alphas]
+
+    clusterer = msmbuilder.cluster.KMedoids(n_clusters=nmodels_per_class, metric='rmsd')
+    clusterer.fit(X)
+
+    seqid_class_cluster_indices = clusterer.cluster_ids_
+    seqid_class_cluster_indices.sort()
+    seqid_class_selected_model_indices = list(seqid_class_df.iloc[seqid_class_cluster_indices].index)
+    seqid_class_cluster_seqids = list(seqid_class_df.iloc[seqid_class_cluster_indices].seqid.values)
+    model_indices += seqid_class_selected_model_indices
+    seqids += seqid_class_cluster_seqids
+
+# model_indices = []
+# seqids = []
+# for seqid_class in seqid_classes:
+#     df_class = df[df.seqid >= seqid_class[0]][df.seqid < seqid_class[1]]
+#     selected_model_indices = random.sample(df_class.index, nmodels_per_class)
+#     selected_models = df_class.loc[selected_model_indices].sort()
+#     model_indices += list(selected_models.index)
[…]

### 2015-03-04 figures/residue_pair_distances/abl/plot.py
Updated residue_pair_distances

@@ -12,8 +12,8 @@ df = pd.read_csv(os.path.join(srcdir, 'traj-refine_implicit_md-data.csv'))
 
 traj = mdtraj.load(os.path.join(srcdir, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir, 'traj-refine_implicit_md-topol.pdb'))
 
-active_structure = mdtraj.load('2SRC.pdb')
-inactive_structure = mdtraj.load('1Y57.pdb')
+inactive_structure = mdtraj.load('2SRC.pdb')
+active_structure = mdtraj.load('1Y57.pdb')
 inactive_abl_structure = mdtraj.load('2F4J.pdb')
 
 # Src  Abl
@@ -34,13 +34,13 @@ ref_residue_selections = [
     'residue 310 and (name "OE1" or name CD or name "OE2")',
     'residue 409 and (name "NH1" or name CZ or name "NH2")',
 ]
-active_structure_residue_indices = [211, 226, 325]   # 0-based
-inactive_structure_residue_indices = [213, 228, 327]   # 0-based
+inactive_structure_residue_indices = [211, 226, 325]   # 0-based
+active_structure_residue_indices = [213, 228, 327]   # 0-based
 inactive_abl_structure_residue_indices = [44, 59, 159]   # 0-based
 
 traj_residue_pair_indices = np.array([[29, 44], [44, 144]])
-active_structure_residue_pair_indices = np.array([[211, 226], [226, 325]])
-inactive_structure_residue_pair_indices = np.array([[213, 228], [228, 327]])
+inactive_structure_residue_pair_indices = np.array([[211, 226], [226, 325]])
+active_structure_residue_pair_indices = np.array([[213, 228], [228, 327]])
 inactive_abl_structure_residue_pair_indices = np.array([[44, 59], [59, 159]])
 
 
@@ -51,8 +51,8 @@ pairs_dict = {pair: {'indices_traj': traj_residue_pair_indices[x]} for x, pair i
 
 contacts = mdtraj.compute_contacts(traj, contacts=traj_residue_pair_indices)[0].T
 
-active_structure_contacts = mdtraj.compute_contacts(active_structure, contacts=active_structure_residue_pair_indices)[0].T
 inactive_structure_contacts = mdtraj.compute_contacts(inactive_structure, contacts=inactive_structure_residue_pair_indices)[0].T
[…]

### 2015-03-04 figures/residue_pair_distances/src/plot.py

@@ -12,8 +12,8 @@ df = pd.read_csv(os.path.join(srcdir, 'traj-refine_implicit_md-data.csv'))
 
 traj = mdtraj.load(os.path.join(srcdir, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir, 'traj-refine_implicit_md-topol.pdb'))
 
-active_structure = mdtraj.load('2SRC.pdb')
-inactive_structure = mdtraj.load('1Y57.pdb')
+inactive_structure = mdtraj.load('2SRC.pdb')
+active_structure = mdtraj.load('1Y57.pdb')
 
 # K295 = 29
 # E310 = 44
@@ -32,56 +32,22 @@ ref_residue_selections = [
     'residue 310 and (name "OE1" or name CD or name "OE2")',
     'residue 409 and (name "NH1" or name CZ or name "NH2")',
 ]
-active_structure_residue_indices = [211, 226, 325]   # 0-based
-inactive_structure_residue_indices = [213, 228, 327]   # 0-based
+inactive_structure_residue_indices = [211, 226, 325]   # 0-based
+active_structure_residue_indices = [213, 228, 327]   # 0-based
 
 traj_residue_pair_indices = np.array([[28, 43], [43, 142]])
-active_structure_residue_pair_indices = np.array([[211, 226], [226, 325]])
-inactive_structure_residue_pair_indices = np.array([[213, 228], [228, 327]])
+inactive_structure_residue_pair_indices = np.array([[211, 226], [226, 325]])
+active_structure_residue_pair_indices = np.array([[213, 228], [228, 327]])
 
 
-
-# data_dict = {label: {'traj_residue_selection': traj_residue_selections[i], 'ref_residue_selection': ref_residue_selections[i]} for i, label in enumerate(residue_labels)}
-#
-# for label in residue_labels:
-#     residue_dict = data_dict[label]
-#     traj_residue_selection = residue_dict['traj_residue_selection']
-#     ref_residue_selection = residue_dict['ref_residue_selection']
-#     # TODO work out which atom indices you want to keep - prob the last three
-#     residue_dict['indices'] = traj.top.select(traj_residue_selection)
-#
-#     residue_dict['atom_slice_traj'] = traj.atom_slice(residue_dict['indices'])
-#
-#     residue_dict['com'] = mdtraj.compute_center_of_mass(residue_dict['atom_slice_traj'])
-#
[…]

### 2015-03-07 manuscript/ms.tex
Minor
Added superposition figure
Modified pipeline figure
Lots of edits.

--- 2015-02-16 manuscript/ms.tex
@@
 
+%\usepackage{palatino}
 % Change to a sans serif font.
@@
  \@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
-  {\textbf{#1}}\sf\@caption@fignum@sep#2\quad
+  {\sf\textbf{#1}}\sf\@caption@fignum@sep#2\quad
 }%
@@
 \author{Daniel L. Parton}
-  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
   %\email{daniel.parton@choderalab.org}
 \author{Patrick B. Grinaway}
-  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
   %\email{patrick.grinaway@choderalab.org}
@@
  \email{john.chodera@choderalab.org}
-  \affiliation{Computational Biology Center, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
 
@@
 The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
-However, simulations on an \emph{omics} scale are not yet widely performed, partly because software infrastructure to enable this has not kept pace. 
-For example, it should now be possible to study protein dynamics across entire (super)families, exploiting the entire variety of available structural biology data and conformational similarities across homologous proteins.
+While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable this has not kept pace. 
+It should now be possible to study protein dynamics across entire (super)families, exploiting the variety of available structural biology data and conformational similarities across homologous proteins.
 Here, we present a new tool for enabling high-throughput simulation in the genomics era.
-Ensembler takes any set of sequences---from a single sequence to an entire superfamily of interest---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
+{\bf Ensembler} takes any set of sequences---from a single sequence to an entire superfamily---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
 This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement with molecular simulation to ensure stable simulation.
-The output of this pipeline is an ensemble of structures ready for subsequent parallel or distributed molecular simulations using clusters, supercomputers, or distributed computing projects like Folding@home.
-Ensembler automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.  
+The output of this pipeline is an ensemble of structures ready for subsequent molecular simulations using computer clusters, supercomputers, or distributed computing projects like Folding@home.
+{\bf Ensembler} automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.  
 A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
@@
 
-Ensembler is free of charge, and is made available under the terms of the GNU General Public License (GPL) v2. 
+{\bf Ensembler} is free and open source software licensed under the GNU General Public License (GPL) v2. 
 It should run on all major operating systems, and has been tested on Linux and OS X.
@@
 
-Proteins play a diverse variety of roles in biology, and the understanding of their function---and how mutations can cause dysfunction and disease---is the preoccupation of much of modern biology. 
-The diminishing cost of nucleic acid sequencing technologies has produced an enormous wealth of sequence data, yielding a large collection of protein-coding open reading frames that provide basic information about these proteins (at the level of primary amino acid sequences) for numerous organisms [CITE].
+Proteins play a diverse variety of roles in living organisms, and the understanding of their function---and how mutations can cause dysfunction and disease---is the preoccupation of much of modern biology.
+The diminishing cost of nucleic acid sequencing technologies has produced an enormous wealth of genomic data, yielding a large collection of protein-coding open reading frames that provide basic information about these proteins (at the level of primary amino acid sequences) for numerous organisms [CITE].
 Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have yielded a great number of protein structures, allowing comparative modeling to provide insight into the static structures many of these proteins adopt [CITE review of structural biology efforts or current comparative modeling?].
-% TODO Maybe it is not necessary to mention comparative modeling here? If so, this could be changed to:
+% TODO Maybe it is not necessary to mention comparative modeling here? The structures themselves provide insight. If so, this could be changed to:
 % Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have provided insight into the static structures which these proteins can adopt [CITE review of structural biology efforts?].
@@
 Static structures, however, provide only a snapshot of the rich dynamical behavior of proteins.
-Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require conformational changes.
+Many functional properties---such as the ability to bind small molecules or interact with signaling partners---often require conformational changes at many levels, from reorganization of sidechains at binding interfaces to loop motions to large scale folding-unfolding events.
 
-Molecular dynamics simulations have proven to be a useful tool for revealing the dynamics of individual proteins, with a number of software packages and forcefields available for biomolecular simulation.
-Advances in computing architectures---especially the availability of GPUs, which provide a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as those that drive the Folding@home distributed computing resource [CITE]) now provide new hardware platforms on which to study the dynamics of these proteins.
-In parallel, techniques for aggregating molecular dynamics simulation data to survey the kinetic landscape of biomolecules, such as Markov state modeling approaches [CITE MSM reviews], are now reaching maturity.
+Molecular dynamics simulations have proven to be a useful tool for revealing the dynamics of individual proteins, with a number of mature software packages and forcefields available for biomolecular simulation.
+Advances in computing architectures---especially the recent emergence of GPUs as a technology for providing a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as distributed computing platforms like Folding@home [CITE], GPUGrid [CITE], and Compernicus [CITE]) now provide unprecedented hardware platforms on which to study the dynamics of these proteins.
+In parallel, techniques for aggregating molecular dynamics simulation data to survey the conformational and kinetic landscapes of biomolecules, such as Markov state modeling (MSM) approaches [CITE MSM reviews], are now reaching maturity.
 
-Despite this, a critical gap remains in the ability to bridge genome-scale sequence information and molecular simulations to enable the study of entire families or superfamilies of proteins in a single organism or across organisms.
-Molecular simulations must largely be set up by hand, with little in the way of automation available to provide practitioners a way of studying many members of a family where only a subset may have structural data.
+Despite this, a critical gap remains in our ability to bridge genome-scale sequence information and molecular simulations to enable the study of entire families or superfamilies of proteins in a single organism or across organisms.
+Molecular simulations must largely be set up by hand, with little in the way of automation available to provide practitioners a way of studying many members of a family in a manner that exploits their similarity, and especially in cases where only a subset of members may have structural data.
 
 Complicating matters further, in protein families known to be able to adopt multiple conformations---such as kinases---structural data may only exist for one or two conformations for any individual member of the family for which there is structural data.
-This poses a challenge for biomolecular simulation methods such as Markov state models, which can provide detailed insight, but require global coverage of the conformational landscape to realize their full potential.
+This poses a challenge for biomolecular simulation and analysis methods such as MSMs, which can provide detailed insight but require global coverage of the conformational landscape to realize their full potential.
 
-Here, we present the first steps toward a resolution of this problem: a fully automated open source framework for building simulation-ready protein models scalable to the superfamily scale.
-We demonstrate the utility of this tool by constructing models for the entire set of human tyrosine kinase catalytic domains, and demonstrate that the resulting models provide good coverage of the known functional regions of structure space.
-This tool forms the foundation for a new era of superfamily-scale molecular simulations for the Folding@home project.
+Here, we present the first steps toward a resolution of this problem: a fully automated open source framework for building simulation-ready protein models scalable from single sequences to entire superfamilies.
+We demonstrate the utility of this tool by constructing models for the entire set of human tyrosine kinase catalytic domains, and show that the resulting models provide good coverage of known functionally relevant regions of structure space.
+While this tool was originally constructed to form the foundation for a new era of superfamily-scale molecular simulations for the Folding@home project, we anticipate its utility is far broader.
 
@@
 
-Ensembler is written in Python, and can be used via a command-line tool ({\tt ensembler}), or via the flexible Python API.
+{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API.
 
@@
 
-The Ensembler modeling pipeline entails a series of stages which are performed in a defined order. 
-A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}, and a detailed description follows.
+The {\bf Ensembler} modeling pipeline entails a series of stages which are performed in a defined order. 
+A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}.
+The various stages of this pipeline are described in detail below.
 
-\begin{figure*}[h]
+{\color{red}[JDC: We could really help the reader if we preface each section here with a bit of an introduction of what we're trying to accomplish in each stage.
+Otherwise, I worry that each section is a long list of things we do without reference to an overall concept of what the stage is trying to accomplish or why certain decisions were made.]}
+
+\begin{figure*}[tb]
   \centering
-    \includegraphics[width=0.8\textwidth]{pipeline}
-  \caption{Ensembler pipeline}
+    \includegraphics[width=1.0\textwidth]{pipeline2}
+  \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
+  The number of viable models surviving each stage of the pipeline for are shown, either for all tyrosine kinases (\emph{All TKs}) or representative individual kinases (\emph{SRC} and \emph{ABL}).
+  In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to convey resources required per model and for modeling the entire set of tyrosine kinases.
+  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread---parallel execution can reduce wall clock time nearly linearly.
+  }
   \label{figpipeline}
@@
 
-The output from this stage is simply a FASTA-formatted sequence file containing the selected target sequences and identifiers. 
 The {\tt ensembler} command-line tool provides methods for selecting targets from either UniProt (a freely accessible resource for protein sequence and functional data---\href{http://www.uniprot.org/}{uniprot.org}) or TargetExplorer (a database framework for aggregating various types of biological data; work to be published).
-This allows the user to easily select a single protein, many proteins, or an entire superfamily.
-Alternatively, the targets file can be generated using any other software, and stored at the appropriate filepath.
+{\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
+This tool allows the user to easily select a single protein, many proteins, or an entire superfamily. 
+{\color{red}[JDC: Can you give a clearer picture of how this is done?  Does the user just specify different flags for these categories, or would they have to construct different Uniprot searches?]}
+The output from this stage is a FASTA-formatted text file containing the selected target sequences and identifiers {\color{red}[JDC: What kind of identifiers?
+Arbitrary, Uniprot, etc?]}, such that the output of other software that produces FASTA-formatted sequence files can be used as input to the next stage.
 
-The method for selecting targets from UniProt is described here. 
-A query string is required as input, using the same syntax as the search function on the UniProt website. 
+To select targets from UniProt, a query string must be constructed by the user that conforms to the same syntax as the search function on the UniProt website. 
+{\color{red}[JDC: What command-line flag or API call is used to select targets from UniProt?]}
 For example, {\tt'domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator. 
-Ensembler is designed to work with protein \emph{domains}, rather than full-length proteins, and the desired protein domain(s) can be selected using a regular expression. 
+{\bf Ensembler} is designed to work with protein \emph{domains}, rather than full-length proteins, and the desired protein domain(s) can be selected using a regular expression. 
 For example, the string {\tt'\^{}Protein kinase(?!; truncated)(?!; inactive)'} would match domains annotated as "Protein kinase", "Protein kinase; 1" or "Protein kinase; 2", but would exclude the domains "Protein kinase; truncated" and "Protein kinase; inactive".
-The program then extracts sequences and identifiers from the UniProt data, and saves these to a FASTA-format text file.
+The program then performs this search on UniProt, retrieves the data, extracts sequences and identifiers {\color{red}[JDC: UniProt identifiers?]}, and saves these to a FASTA-format text file.
 
@@
 
-As for target selection, the {\tt ensembler} tool provides methods for selecting templates from various resources---UniProt, the Protein Data Bank (PDB; \href{http://www.pdb.org}{pdb.org}) or a TargetExplorer database.
-From the user perspective, selection from UniProt proceeds in a similar fashion as described above.
-The returned data for each UniProt entry includes an up-to-date list of PDB structures and their residue spans.
-PDB files are downloaded from the PDB for structures which include the desired domain.
+As for target selection, the {\tt ensembler} tool provides methods for selecting template structures (from which models will be built) from various sources---UniProt {\color{red}[CITE]}, the Protein Data Bank (PDB; \href{http://www.pdb.org}{pdb.org}) {\color{red}[JDC: Do you use the RCSB or some other PDB site?]}, or a TargetExplorer database {\color{red}[JDC: Again, not sure if it is helpful to leave this in unless it will be immediately useful to the reader]}.
+From the user perspective, selection of templates from UniProt proceeds in a similar fashion as described above; a used-specified UniProt query string is used to retrieve the list of PDB entries and residue spans associated with each UniProt entry mathing the query.
+{\color{red}[JDC: Can you be more precise what is meant by ``residue span'' here?]}
+Structures that include the desired domain {\color{red}[JDC: Can you elaborate on how this is determined?]} are retrieved from the RCSB in PDB format.
 Data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries, is then used to filter out PDB chains with $<$ 70\% resolved residues within the domain span.
-Template sequences and structures are then extracted and written as FASTA-format sequence and PDB-format coordinate files respectively.
+Template sequences and structures are then extracted and written as FASTA-format sequence and PDB-format coordinate files, respectively.
 
-Selection from the PDB simply requires specifying a list of PDB IDs.
+Selection of template structures from the RCSB simply requires specifying a list of PDB IDs.
+{\color{red}[JDC: Who specifies the list of PDB IDs?  How are they specified?  Can this be built automatically, or does the user need to specify these?]}
 These are matched to UniProt entries via the SIFTS service, and the same procedure is then followed to extract template sequences and structures.
 
-Unresolved template loops can optionally be remodeled with a kinematic closure algorithm, which is provided via the loopmodel tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
-This tends to provide higher-quality models following the subsequent modeling process.
+Unresolved template loops can optionally be remodeled with a kinematic closure algorithm {\color{red}[CITE]}, which is provided via the loopmodel tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
+Because shorter loops need to be built by the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models following the subsequent modeling process.
 
@@
 
-In this stage, models are generated for each target-template pair, using the Modeller automodel function (CITE: Modeller), which implements comparative structure modeling by satisfaction of spatial restraints (CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000).
-Modeller requires the user to first provide a target-template sequence alignment.
-This is implemented in Ensembler using the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al} (CITE: Gaston Gonnet Science 1992).
+In this stage, models are generated for each target-template pair, using the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
+While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
 
-Non-unique models are then filtered out using structural similarity-based clustering.
-The mdtraj (CITE: mdtraj) Python library is used to calculate RMSD with a fast quaternion characteristic polynomial (QCP) implementation, and the leader algorithm is then used to populate clusters.
-A minimum distance cutoff (default: 0.6 \AA) is used to retain only a single model per cluster.
+All chains of template structures that contain the template sequence are utilized in the modeling phase, which can sometimes cause models to be nearly identical.
+Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
+The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD with a fast quaternion characteristic polynomial (QCP)~{\color{red}[JDC: Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
+A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
+{\color{red}[JDC: Which atoms are used in the RMSD comparison?  All atoms, heavy atoms, or CA only?]}
 
@@
 
-Models are then refined with a steepest descent energy minimization and a short molecular dynamics (MD) simulation with implicit solvent.
+Models are then refined with a steepest descent energy minimization {\color{red}[JDC: I think OpenMM uses L-BFGS]} and a short molecular dynamics (MD) simulation in implicit solvent.
 This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-The Amber99SB-ILDN force field is used (CITE: amber99sbildn refs) with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
-The simulation is run for a default of 100 ps.
-This refinement process helps to prepare models for subsequent production simulation, and also helps to filter out poor quality models.
+The Amber99SB-ILDN force field is used {\color{red}[CITE: amber99sbildn refs]} with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
+The simulation is run for a default of 100 ps to filter out poor quality models (where atomic overlaps that cannot be resolved by energy minimization would cause the simulation to explode) and help relax models for subsequent production simulation.
+{\color{red}[JDC: What criteria were applied to filter out poor models?  Do we only look for thrown exceptions or NaNs?  Or do we use an energy filtering criteria too?]}
 
-Ensembler also provides optional routines for solvating models with explicit solvent and performing a second MD refinement.
-The solvation step solvates each model for a given target with the same number of waters, as this is (currently) a requirement for building MSMs from multiple independent MD trajectories.
+While protein-only models may be sufficient for structural analysis or implicit solvent simulations, {\bf Ensembler} also provides a stage for solvating models with explicit water and performing a round of explicit-solvent MD refinement/equilibration under isothermal-isobaric (NPT) conditions.
+The solvation step solvates each model for a given target with the same number of waters to facilitate the integration of data from multiple simulations, such as the construction of MSMs.
 The target number of waters is selected by first solvating each model with a specified padding distance (default: 10 \AA), then taking a percentile value from the distribution (default: 68th percentile).
+{\color{red}[JDC: Would be useful to explain why we are doing this.]}
 Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
-The explicit solvent MD simulation is also implemented using OpenMM, with the Amber99SB-ILDN force field and TIP3P water.
+The explicit solvent MD simulation is also implemented using OpenMM, with the Amber99SB-ILDN force field and TIP3P water {\color{red}[JDC: CITE]}.
+{\color{red}[JDC: We should allow other water models in OpenMM too, such as TIP4P-Ew?]}
+
+{\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
+For example, we don't yet handle counterions (e.g.
+structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
+We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
+It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
 
@@
 
-Finally, Ensembler provides a packaging module, which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@Home (CITE: F@H).
+{\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home (CITE: F@H).
 
-\subsubsection{Other features}
+\subsubsection{Provenance}
+
+To aid the user in tracking the provenance of each model, each pipeline function also outputs a metadata file, which helps to link data to the software version used to generate it (both {\bf Ensembler} and its dependencies), and also provides timing and performance information, and other data such as hostname.
+
+\subsubsection{Rapidly modeling a single template}
 
 % quickmodel function performs the entire pipeline in one go, for a single target and a small number of templates.
-The command-line tool also provides a {\tt quickmodel} function, which performs the entire Ensembler pipeline for a single target with a small number of templates.
-For larger numbers of models (such as entire protein families), the main pipeline functions should be used.
-The modeling and refinement functions use MPI to trivially parallelize computation across each model (or across each template, in the case of the loop reconstruction code).
-
-Each pipeline function also outputs a metadata file, which helps to link data to the software version used to generate it (both Ensembler and its dependencies), and also provides timing and performance information, and other data such as hostname.
+For users interested in simply using {\bf Ensembler} to rapidly generate a set of models for a single template sequence, {\bf Ensembler} provides a command-line tool {\tt quickmodel}, which performs the entire pipeline for a single target with a small number of templates.
+For larger numbers of models (such as entire protein families), modeling time is greatly reduced by using the main modeling pipeline parallelized via MPI, which distributes computation across each model (or across each template, in the case of the loop reconstruction code), scaling (in a ``pleasantly parallel'' manner) up to the number of models generated.
 
@@
 
+% \begin{table}
+% \begin{tabular}{l|c|c|c}
+%
+%     & SRC & ABL & All 90 TKs \\
+%     \hline
+%     Templates & \num{4433} & \num{4433} & \num{398970} \\
+%     Models & \num{4248} & \num{4250} & \num{382568} \\
+%     Unique models & \num{4093} & \num{4247} & \num{378839} \\
+%     Refined models & \num{4023} & \num{4241} & \num{373513} \\
+%
+% \end{tabular}
+% \label{tab:model_counts}
+% \caption{Model counts for the kinase domains of human SRC and ABL, and for all of the 90 human TK targets, at various stages of the Ensembler pipeline.}
+% \end{table}
+%
+% \begin{table}
+% \begin{tabular}{l|c}
+%
+%     & Timing ($h$:$m$:$s$) \\
+%     \hline
+%     Template reconstruction & 03:00:00 \\
+%     Modeling & 00:01:19 \\
+%     Refinement (100 ps implicit solvent MD) & 00:01:58 \\
+%
+% \end{tabular}
+% \label{tab:timings}
+% \caption{Average timings per template/model for the compute-intensive stages of the Ensembler pipeline.}
+% \end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+The authors are grateful to Kyle A.~Beauchamp (MSKCC), Robert McGibbon (Stanford), Arien S. Rustenburg (MSKCC) for many excellent software engineering suggestions.
+The authors thank Sonya M.~Hanson (MSKCC), Nicholas M.~Levinson (?), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
+The authors are grateful to Benjamin Webb and Andrej Sali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
+DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-03-10 manuscript/ms.tex
Revised Design and Implementation section

@@
 Molecular dynamics simulations have proven to be a useful tool for revealing the dynamics of individual proteins, with a number of mature software packages and forcefields available for biomolecular simulation.
-Advances in computing architectures---especially the recent emergence of GPUs as a technology for providing a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as distributed computing platforms like Folding@home [CITE], GPUGrid [CITE], and Compernicus [CITE]) now provide unprecedented hardware platforms on which to study the dynamics of these proteins.
+Advances in computing architectures---especially the recent emergence of GPUs as a technology for providing a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as distributed computing platforms like Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE]) now provide unprecedented hardware platforms on which to study the dynamics of these proteins.
 In parallel, techniques for aggregating molecular dynamics simulation data to survey the conformational and kinetic landscapes of biomolecules, such as Markov state modeling (MSM) approaches [CITE MSM reviews], are now reaching maturity.
@@
 
-% Python
-% Can be used via command-line tool Ensembler, or via the Python API
-
 {\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API.
 
-% Pipeline architecture, with a series of stages which must be performed in order.
-
-The {\bf Ensembler} modeling pipeline entails a series of stages which are performed in a defined order. 
+The {\bf Ensembler} modeling pipeline comprises a series of stages which are performed in a defined order. 
 A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}.
@@
 Otherwise, I worry that each section is a long list of things we do without reference to an overall concept of what the stage is trying to accomplish or why certain decisions were made.]}
+{\color{blue}[DLP: Good point. I've added in brief introductions for each section.]}
 
@@
 
-The {\tt ensembler} command-line tool provides methods for selecting targets from either UniProt (a freely accessible resource for protein sequence and functional data---\href{http://www.uniprot.org/}{uniprot.org}) or TargetExplorer (a database framework for aggregating various types of biological data; work to be published).
-{\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
-This tool allows the user to easily select a single protein, many proteins, or an entire superfamily. 
-{\color{red}[JDC: Can you give a clearer picture of how this is done?  Does the user just specify different flags for these categories, or would they have to construct different Uniprot searches?]}
-The output from this stage is a FASTA-formatted text file containing the selected target sequences and identifiers {\color{red}[JDC: What kind of identifiers?
-Arbitrary, Uniprot, etc?]}, such that the output of other software that produces FASTA-formatted sequence files can be used as input to the next stage.
+The first stage entails the selection of a set of target protein sequences.
 
-To select targets from UniProt, a query string must be constructed by the user that conforms to the same syntax as the search function on the UniProt website. 
-{\color{red}[JDC: What command-line flag or API call is used to select targets from UniProt?]}
-For example, {\tt'domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator. 
-{\bf Ensembler} is designed to work with protein \emph{domains}, rather than full-length proteins, and the desired protein domain(s) can be selected using a regular expression. 
-For example, the string {\tt'\^{}Protein kinase(?!; truncated)(?!; inactive)'} would match domains annotated as "Protein kinase", "Protein kinase; 1" or "Protein kinase; 2", but would exclude the domains "Protein kinase; truncated" and "Protein kinase; inactive".
-The program then performs this search on UniProt, retrieves the data, extracts sequences and identifiers {\color{red}[JDC: UniProt identifiers?]}, and saves these to a FASTA-format text file.
+These targets can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
+The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}), using the subcommand {\tt gather\_targets}.
+% {\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
+% {\color{blue}[DLP: Yes, I think TargetExplorer is primarily an internal tool for the moment, so let's omit discussion of that.]}
+The user specifies a query string with the {\tt -{}-query} flag, which conforms to the same syntax as the search function available on the UniProt website.
+% {\color{red}[JDC: What command-line flag or API call is used to select targets from UniProt?]}
+% {\color{red}[DLP: the command-line flag is now described]}
+For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
+In this way, the user may select a single protein, many proteins, or an entire superfamily.
+The program outputs a FASTA file, setting the UniProt mnemonic (e.g. {\tt SRC\_HUMAN}) as the identifier for each target protein.
+% {\color{red}[JDC: Can you give a clearer picture of how this is done?  Does the user just specify different flags for these categories, or would they have to construct different Uniprot searches?]}
+% {\color{blue}[DLP: This is explained more clearly now.]}
 
-% The selected targets are
-% given identifiers of the form {\tt [UniProt mnemonic]\_D[domain index]}, which
-% consists of the UniProt name for the target protein and a 0-based index for the
-% domain (since a single target protein may contain multiple domains of
-% interest). Example: {\tt EGFR\_HUMAN\_D0}.
+In many cases, it will be desirable to build models of an isolated protein domain, rather than the full-length protein.
+The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-domains} flag. 
+For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
+To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
+In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain.
+This is necessary because a single target protein may contain multiple domains of interest.
+Example identifiers: {\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}.
 
@@
 
-As for target selection, the {\tt ensembler} tool provides methods for selecting template structures (from which models will be built) from various sources---UniProt {\color{red}[CITE]}, the Protein Data Bank (PDB; \href{http://www.pdb.org}{pdb.org}) {\color{red}[JDC: Do you use the RCSB or some other PDB site?]}, or a TargetExplorer database {\color{red}[JDC: Again, not sure if it is helpful to leave this in unless it will be immediately useful to the reader]}.
-From the user perspective, selection of templates from UniProt proceeds in a similar fashion as described above; a used-specified UniProt query string is used to retrieve the list of PDB entries and residue spans associated with each UniProt entry mathing the query.
-{\color{red}[JDC: Can you be more precise what is meant by ``residue span'' here?]}
-Structures that include the desired domain {\color{red}[JDC: Can you elaborate on how this is determined?]} are retrieved from the RCSB in PDB format.
-Data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries, is then used to filter out PDB chains with $<$ 70\% resolved residues within the domain span.
-Template sequences and structures are then extracted and written as FASTA-format sequence and PDB-format coordinate files, respectively.
+The second stage entails the selection of templates and storage of associated structures, sequences and identifiers.
 
-Selection of template structures from the RCSB simply requires specifying a list of PDB IDs.
-{\color{red}[JDC: Who specifies the list of PDB IDs?  How are they specified?  Can this be built automatically, or does the user need to specify these?]}
-These are matched to UniProt entries via the SIFTS service, and the same procedure is then followed to extract template sequences and structures.
+This data can be provided manually, by storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
+The structure residues must also match those in the sequence file.
+
+The {\tt ensembler gather\_templates} subcommand also provides methods for selecting template structures from either UniProt or the Protein Data Bank (PDB; \href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
+
+% {\color{red}[JDC: Do you use the RCSB or some other PDB site?]} - DLP: I use the RCSB, but all PDB sites are updated simultaneously so I think it shouldn't matter.
+However, I've updated the link so that it points to the RCSB PDB.
+% {\color{red}[JDC: Again, not sure if it is helpful to leave this in unless it will be immediately useful to the reader]} - DLP: references to TargetExplorer removed
+
+Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g. {\tt -{}-query 2H8H,1Y57}.
+% {\color{red}[JDC: Who specifies the list of PDB IDs?
+How are they specified?
+Can this be built automatically, or does the user need to specify these?]} DLP: these are specified by the user as a comma-separated list, as is now described above.
+Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
+The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries.
+The SIFTS data is used to extract template sequences, retaining only residues which are resolved and match the equivalent residue in the UniProt sequence---non-wildtype residues are thus removed from the template structures.
+Furthermore, PDB chains with less than a given percentage of resolved residues (default: 70\%) are filtered out.
+Sequences are stored in a FASTA file, with identifiers of the form {\tt [UniProt mnemonic]\_D[UniProt domain index]\_[PDB ID]\_[PDB chain ID]}, e.g. {\tt SRC\_HUMAN\_D0\_2H8H\_A}.
+Template structures with residues matching the sequence data are then extracted and stored as PDB-format coordinate files.
+
+Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-domains} flag allows selection of individual domains with a regular expression string.
+The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
+If the {\tt -{}-domains} flag is used, then templates are truncated at the start and end of the domain sequence.
 
 Unresolved template loops can optionally be remodeled with a kinematic closure algorithm {\color{red}[CITE]}, which is provided via the loopmodel tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
-Because shorter loops need to be built by the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models following the subsequent modeling process.
+Because fewer loops need to be built during the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models following the subsequent modeling process.
 
@@
 
-In this stage, models are generated for each target-template pair, using the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
+This stage entails the generation of models via comparative modeling of each target sequence onto each template structure. Non-unique models are filtered out using a RMSD-based clustering scheme.
+
+Modeling is performed with the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
 While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
@@
 Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
-The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD with a fast quaternion characteristic polynomial (QCP)~{\color{red}[JDC: Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
+The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD (for C$\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~{\color{red}[Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
 A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
-{\color{red}[JDC: Which atoms are used in the RMSD comparison?  All atoms, heavy atoms, or CA only?]}
+% {\color{red}[JDC: Which atoms are used in the RMSD comparison?  All atoms, heavy atoms, or CA only?]} - DLP: CA only. This is now specified in the text.
 
@@
 
-Models are then refined with a steepest descent energy minimization {\color{red}[JDC: I think OpenMM uses L-BFGS]} and a short molecular dynamics (MD) simulation in implicit solvent.
+This stage entails the use of molecular dynamics simulations to refine the models built in the previous step.
+This helps to improve model quality and also prepares models for subsequent production simulation, including solvation with explicit water molecules, if desired.
+
+Models are first subjected to energy minimization (using the L-BFGS algorithm {\color{red}[CITE]}), followed by a short molecular dynamics (MD) simulation with an implicit solvent representation.
 This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-The Amber99SB-ILDN force field is used {\color{red}[CITE: amber99sbildn refs]} with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
+By default, the Amber99SB-ILDN force field is used {\color{red}[CITE: amber99sbildn refs]} with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
+The {\bf Ensembler} API allows the use of any of the other force fields implemented in OpenMM.
 The simulation is run for a default of 100 ps to filter out poor quality models (where atomic overlaps that cannot be resolved by energy minimization would cause the simulation to explode) and help relax models for subsequent production simulation.
 {\color{red}[JDC: What criteria were applied to filter out poor models?  Do we only look for thrown exceptions or NaNs?  Or do we use an energy filtering criteria too?]}
+{\color{blue}[DLP: We currently just filter out models which throw exceptions or NaNs.]}
 
@@
 {\color{red}[JDC: Would be useful to explain why we are doing this.]}
+{\color{blue}[DLP: Addressed.]}
+This helps to prevent models with particularly long, extended loops---such as those arising from template structures with unresolved termini---from imposing very large box sizes on the entire set of models.
 Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
-The explicit solvent MD simulation is also implemented using OpenMM, with the Amber99SB-ILDN force field and TIP3P water {\color{red}[JDC: CITE]}.
+The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field and TIP3P water {\color{red}[JDC: CITE]} by default.
+Other force fields or water models such as TIP4P-Ew {\color{red}[CITE]}) can be specified via the {\bf Ensembler} API.
 {\color{red}[JDC: We should allow other water models in OpenMM too, such as TIP4P-Ew?]}
+{\color{blue}[DLP: I forgot to mention this in the text previously - any of the OpenMM force fields can be chosen via the API.
+I've updated the text accordingly.
+Is this functionality sufficient?
+I guess it's ok to leave ff choice as an "advanced" feature which requires use of the API?
+Otherwise I could add a --water\_model flag to the CLI, for example.]}
 
@@
 For users interested in simply using {\bf Ensembler} to rapidly generate a set of models for a single template sequence, {\bf Ensembler} provides a command-line tool {\tt quickmodel}, which performs the entire pipeline for a single target with a small number of templates.
-For larger numbers of models (such as entire protein families), modeling time is greatly reduced by using the main modeling pipeline parallelized via MPI, which distributes computation across each model (or across each template, in the case of the loop reconstruction code), scaling (in a ``pleasantly parallel'' manner) up to the number of models generated.
+For larger numbers of models (such as entire protein families), modeling time is greatly reduced by using the main modeling pipeline, which is parallelized via MPI, distributing computation across each model (or across each template, in the case of the loop reconstruction code), and scaling (in a ``pleasantly parallel'' manner) up to the number of models generated.
 

### 2015-03-12 manuscript/ms.tex
Revised introduction

@@
 
-Proteins play a diverse variety of roles in living organisms, and the understanding of their function---and how mutations can cause dysfunction and disease---is the preoccupation of much of modern biology.
-The diminishing cost of nucleic acid sequencing technologies has produced an enormous wealth of genomic data, yielding a large collection of protein-coding open reading frames that provide basic information about these proteins (at the level of primary amino acid sequences) for numerous organisms [CITE].
-Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have yielded a great number of protein structures, allowing comparative modeling to provide insight into the static structures many of these proteins adopt [CITE review of structural biology efforts or current comparative modeling?].
-% TODO Maybe it is not necessary to mention comparative modeling here? The structures themselves provide insight. If so, this could be changed to:
-% Complementing this, large-scale structural biology efforts such as the Protein Structure Initiative (PSI) and Structural Genomics Consortium (SGC) have provided insight into the static structures which these proteins can adopt [CITE review of structural biology efforts?].
+Recent advances in genomics and structural biology have helped generate an enormous wealth of protein data at the level of amino-acid sequence and three-dimensional structure.
+However, proteins typically exist as an ensemble of thermally accessible conformational states, and static structures provide only a snapshot of their rich dynamical behavior.
+Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require transitions between states, encompassing anything from reorganization of sidechains at binding interfaces to domain motions to large scale folding-unfolding events.
+Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state [CITE Lee Craik Science 2009].
 
-Static structures, however, provide only a snapshot of the rich dynamical behavior of proteins.
-Many functional properties---such as the ability to bind small molecules or interact with signaling partners---often require conformational changes at many levels, from reorganization of sidechains at binding interfaces to loop motions to large scale folding-unfolding events.
+Molecular dynamics (MD) simulations have the capability, in principle, to describe the time evolution of a protein in atomistic detail, and have proven themselves to be a useful tool in the study of protein dynamics.
+A number of mature software packages and forcefields are available, and much recent progress has been driven by advances in computing architecture.
+For example, many MD packages are now able to exploit GPUs, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE] allow scalability on an unprecedented level.
+In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity [CITE MSM reviews].
+MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
+There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates [CITE Modeller and Rosetta and a recent homology modeling review].
 
-Molecular dynamics simulations have proven to be a useful tool for revealing the dynamics of individual proteins, with a number of mature software packages and forcefields available for biomolecular simulation.
-Advances in computing architectures---especially the recent emergence of GPUs as a technology for providing a hundredfold increase in computational power per unit cost for a variety of applications---and the proliferation of scalable computing technologies (such as distributed computing platforms like Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE]) now provide unprecedented hardware platforms on which to study the dynamics of these proteins.
-In parallel, techniques for aggregating molecular dynamics simulation data to survey the conformational and kinetic landscapes of biomolecules, such as Markov state modeling (MSM) approaches [CITE MSM reviews], are now reaching maturity.
+However, it remains difficult for researchers to exploit the full variety of available protein sequence and structural data in simulation studies, largely due to limitations in software architecture.
+For example, the set up of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure, addition of missing residues and atoms, solvation with explicit water and salt buffer, choice of simulation parameters, and system relaxation with energy minimization and one or more short MD simulations.
+For this reason, simulation studies typically consider only one or a few proteins and starting configurations.
 
-Despite this, a critical gap remains in our ability to bridge genome-scale sequence information and molecular simulations to enable the study of entire families or superfamilies of proteins in a single organism or across organisms.
-Molecular simulations must largely be set up by hand, with little in the way of automation available to provide practitioners a way of studying many members of a family in a manner that exploits their similarity, and especially in cases where only a subset of members may have structural data.
+The ability to fully exploit the large base of available protein sequence and structural data in biomolecular simulation studies could open up many interesting avenues for research, enabling the study of entire protein families or superfamilies across multiple organisms.
+The similarity between members of a given protein family could be exploited to generate arrays of conformational models, which could be used as starting configurations to aid sampling in MD simulations.
+This approach would be highly beneficial for many MD methods, such as Markov state modeling, which require global coverage of the conformational landscape to realize their full potential, and would also be particularly useful in cases where structural data is present for only a subset of the members of a protein family.
+It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---, for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
 
-Complicating matters further, in protein families known to be able to adopt multiple conformations---such as kinases---structural data may only exist for one or two conformations for any individual member of the family for which there is structural data.
-This poses a challenge for biomolecular simulation and analysis methods such as MSMs, which can provide detailed insight but require global coverage of the conformational landscape to realize their full potential.
-
-Here, we present the first steps toward a resolution of this problem: a fully automated open source framework for building simulation-ready protein models scalable from single sequences to entire superfamilies.
-We demonstrate the utility of this tool by constructing models for the entire set of human tyrosine kinase catalytic domains, and show that the resulting models provide good coverage of known functionally relevant regions of structure space.
-While this tool was originally constructed to form the foundation for a new era of superfamily-scale molecular simulations for the Folding@home project, we anticipate its utility is far broader.
+Here, we present the first steps toward bridging the gap between biomolecular simulation software and \emph{omics}-scale sequence and structural data: a fully automated open source framework for building simulation-ready protein models scalable from single sequences to entire superfamilies.
+{\bf Ensembler} provides functions for selecting target sequences and homologous template structures, and (by interfacing with a number of external packages) performs pairwise alignments, comparative modeling of target-template pairs, and several stages of model refinement.
+As an example application, we have constructed models for the entire set of human tyrosine kinase catalytic domains, using all available structures of protein kinase domains (from any species) as templates.
+This results in a total of almost 400,000 models, and we demonstrate that these provide wide-ranging coverage of known functionally relevant regions of structure.
+By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
+We anticipate that the tool will prove to be useful in a number of other ways.
+For example, the generated models could represent valuable data sets even without subsequent production simulation, allowing exploration of the conformational diversity present within the available structural data for a given protein family.
+Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
 
@@
 To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
-In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain.
-This is necessary because a single target protein may contain multiple domains of interest.
+In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest.
 Example identifiers: {\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}.

### 2015-03-13 manuscript/ms.tex
Added figures to manuscript

@@
 %\documentclass[aps,prl,preprint,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
+
+\pdfoutput=1
+\usepackage[pdftex]{graphicx}
 
@@
 \usepackage{amssymb}
-\pdfoutput=1
-\usepackage[pdftex]{graphicx}
 %\usepackage[mathbf,mathcal]{euler}
 %\usepackage{citesort}
+\usepackage{subcaption}
 \usepackage{dcolumn}
@@
 % The figures are in a figures/ subdirectory.
-% \graphicspath{{figures/}}
+\graphicspath{{../figures/}}
 
@@
   \centering
-    \includegraphics[width=1.0\textwidth]{pipeline2}
+    \includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
   \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
@@
 I guess it's ok to leave ff choice as an "advanced" feature which requires use of the API?
-Otherwise I could add a --water\_model flag to the CLI, for example.]}
+Otherwise I could add a -{}-water\_model flag to the CLI, for example.]}
 
@@
 
-% \begin{table}
-% \begin{tabular}{l|c|c|c}
-%
-%     & SRC & ABL & All 90 TKs \\
-%     \hline
-%     Templates & \num{4433} & \num{4433} & \num{398970} \\
-%     Models & \num{4248} & \num{4250} & \num{382568} \\
-%     Unique models & \num{4093} & \num{4247} & \num{378839} \\
-%     Refined models & \num{4023} & \num{4241} & \num{373513} \\
-%
-% \end{tabular}
-% \label{tab:model_counts}
-% \caption{Model counts for the kinase domains of human SRC and ABL, and for all of the 90 human TK targets, at various stages of the Ensembler pipeline.}
-% \end{table}
-%
-% \begin{table}
-% \begin{tabular}{l|c}
-%
-%     & Timing ($h$:$m$:$s$) \\
-%     \hline
-%     Template reconstruction & 03:00:00 \\
-%     Modeling & 00:01:19 \\
-%     Refinement (100 ps implicit solvent MD) & 00:01:58 \\
-%
-% \end{tabular}
-% \label{tab:timings}
-% \caption{Average timings per template/model for the compute-intensive stages of the Ensembler pipeline.}
-% \end{table}
+\begin{figure}[tb]
+  \centering
+    \includegraphics[width=1.0\textwidth]{seqid_dist/seqid_dist}
+    \caption{{\bf Sequence identity distribution for human TK models.}
+    Distribution of sequence identities for all 373,513 models generated for the human tyrosine kinases.
+    {\color{blue}[DLP: should I mention the use of KDE smoothing?]}
+    Sequence identities are calculated from pairwise target-template alignments.
+    The cumulative distribution function is shown by the dashed line.
+    }
+  \label{figseqid_dist}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+    \includegraphics[width=1.0\textwidth]{rmsddist/rmsddist2}
+    \caption{{\bf RMSD distribution by sequence identity.}
+    RMSD distributions for all 373,513 human TK models, divided into three sequence identity ranges.
+    % TODO may want to mention KDE smoothing
+    For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
+  }
+  \label{figrmsddist}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+    \includegraphics[width=1.0\textwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+    \caption{{\bf Superposition of clustered models of Src and Abl.}
+    Superposed renderings of nine models each for Src and Abl, giving some indication the diversity of conformations generated by Ensembler.
+    The models for each target were divided into three sequence identity ranges (as in Fig. \ref{figrmsddist}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
+    The models shown are the centroids of each cluster.
+    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
+  }
+  \label{figsuperposition}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+      \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/distances}
+      \caption{Src}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+      \includegraphics[width=1.0\textwidth]{residue_pair_distances/abl/distances}
+      \caption{Abl}
+    \end{subfigure}
+    \caption{{\bf Residue pair distances for Src and Abl.}
+    % ...
+  }
+  \label{figaa_pair_distances_ref_structures}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
+    \caption{{\bf Reference structures for Src.}
+    % ...
+  }
+  \label{figsrc_ref_structures}
+\end{figure}
 

### 2015-03-13 figures/residue_pair_distances/abl/plot.py

@@ -75,4 +75,4 @@ sns.plt.xlim(0,5)
 sns.plt.ylim(0,5)
 sns.plt.axes().set_aspect('equal')
 
-sns.plt.savefig('distances.png')
+sns.plt.savefig('distances.png', bbox_inches='tight')

### 2015-03-13 figures/residue_pair_distances/src/plot.py

@@ -69,4 +69,4 @@ sns.plt.xlim(0,5)
 sns.plt.ylim(0,5)
 sns.plt.axes().set_aspect('equal')
 
-sns.plt.savefig('distances.png')
+sns.plt.savefig('distances.png', bbox_inches='tight')

### 2015-03-14 manuscript/ms.tex
Started writing Results section

@@
   \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
-  The number of viable models surviving each stage of the pipeline for are shown, either for all tyrosine kinases (\emph{All TKs}) or representative individual kinases (\emph{SRC} and \emph{ABL}).
-  In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to convey resources required per model and for modeling the entire set of tyrosine kinases.
+  The number of viable models surviving each stage of the pipeline are shown, either for all tyrosine kinases (\emph{All TKs}) or representative individual kinases (\emph{SRC} and \emph{ABL}).
+  In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to exemplify the resources required per model and for modeling the entire set of tyrosine kinases.
   Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread---parallel execution can reduce wall clock time nearly linearly.
@@
 The {\tt ensembler gather\_templates} subcommand also provides methods for selecting template structures from either UniProt or the Protein Data Bank (PDB; \href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
+Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.
+for crystal unit cells with multiple asymmetric units) would thus give rise to multiple template structures.
 
@@
 Sequences are stored in a FASTA file, with identifiers of the form {\tt [UniProt mnemonic]\_D[UniProt domain index]\_[PDB ID]\_[PDB chain ID]}, e.g. {\tt SRC\_HUMAN\_D0\_2H8H\_A}.
-Template structures with residues matching the sequence data are then extracted and stored as PDB-format coordinate files.
+Matching residues then extracted from the original coordinate files and stored as PDB-format coordinate files.
 
@@
 
-Unresolved template loops can optionally be remodeled with a kinematic closure algorithm {\color{red}[CITE]}, which is provided via the loopmodel tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
-Because fewer loops need to be built during the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models following the subsequent modeling process.
+Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm {\color{red}[CITE]} provided via the {\tt loopmodel} tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
+Because fewer loops need to be built during the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
+Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
+Furthermore, the Rosetta {\tt loopmodel} program will not model missing residues at the termini of a structure---such residues spans are modeled in the subsequent stage.
 
@@
 
-% ? The API is highly flexible, allowing control over many important parameters. Appropriate choices have been set as default values.
+% TODO? The API is highly flexible, allowing control over many important parameters. Appropriate choices have been set as default values.
+
+% TODO? other tools for analyzing project etc.
 
@@
 
+As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TKs) domains listed in UniProt.
+TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer.
+For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl cancer involvement]}.
+Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
+Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach could benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.
+by generating large numbers of starting configurations to be used in highly parallel MD simulation.
+
+We selected all available structures of protein kinase domains (of any species) as templates, for a total of 4433 (\num{398970} target-template pairs).
+The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
+
+The templates were first subjected to loop remodeling, using the {\tt loopmodel} subcommand.
+Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled.
+
+Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
+In addition, the solvation stage was performed for two representative individual kinases (\emph{Src} and \emph{Abl}).
+The number of models which survived each stage are shown in Fig. \ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
+Fig. \ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage.
+
+The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig. \ref{figrmsddist}.
+The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in lower RMSD models.
+The sequence identity stratifications were selected based on the sequence identity distribution (Fig.
+\ref{figseqiddist}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
+
+% superpositions etc.
+
 \begin{figure}[tb]
@@
     }
-  \label{figseqid_dist}
+  \label{figseqiddist}
 \end{figure}
@@
     \end{subfigure}
-    \caption{{\bf Residue pair distances for Src and Abl.}
-    % ...
+    \caption{{\bf E310-R409 and K295-E310 distances for models of Src and Abl, colored by sequence identity.}
   }
@@
     \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
-    \caption{{\bf Reference structures for Src.}
-    % ...
+    % \begin{subfigure}[b]{0.45\textwidth}
+    %   \includegraphics[width=1.0\textwidth, clip=true, trim=30em 0em 40em 0em]{residue_pair_distances/src/inactive}
+    %   \caption{Inactive}
+    % \end{subfigure}
+    % \begin{subfigure}[b]{0.45\textwidth}
+    %   \includegraphics[width=1.0\textwidth, clip=true, trim=30em 0em 27em 10em]{residue_pair_distances/src/active}
+    %   \caption{Active}
+    % \end{subfigure}
+    \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
+    In the inactive state, E310 forms a salt bridge with R409.
+    During activation, the C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
+    This positions K295 in the appropriate position for catalysis.
   }
@@
 \label{section:availability}
+
+The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python.
+Instructions can be found at \url{https://github.com/choderalab/ensembler}.
+This will install all dependencies except for Modeller and (optionally) Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
+The latest source can be downloaded from the above GitHub repository, which also contains instructions for building and installing the code.
+
+% TODO future directions
+% TODO what's on the Ensembler 2 roadmap?
+% TODO MCCE?
 

### 2015-03-15 manuscript/ms.tex
Revisions to abstract.

@@
 
+% Pretty-printing of shell commands
+\newcommand{\shellcmd}[1]{\\\ \texttt{\scriptsize\# #1}\\}
+
 % The figures are in a figures/ subdirectory.
@@
 The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
-While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable this has not kept pace. 
-It should now be possible to study protein dynamics across entire (super)families, exploiting the variety of available structural biology data and conformational similarities across homologous proteins.
+While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable simulations at this scale have not kept pace.
+It should now be possible to study protein dynamics across entire (super)families, exploiting both available structural biology data and conformational similarities across homologous proteins.
 Here, we present a new tool for enabling high-throughput simulation in the genomics era.
 {\bf Ensembler} takes any set of sequences---from a single sequence to an entire superfamily---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
-This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement with molecular simulation to ensure stable simulation.
+This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement and filtering with molecular simulation to ensure stable simulation.
 The output of this pipeline is an ensemble of structures ready for subsequent molecular simulations using computer clusters, supercomputers, or distributed computing projects like Folding@home.
 {\bf Ensembler} automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.
+{\color{red}[JDC: Prior sentence is redundant?]}
 A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
@@
 
-\emph{Keywords: molecular dynamics simulation; comparative modeling}
+\emph{Keywords: molecular dynamics simulation; comparative modeling; distributed simulation}
 
@@
 
-The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python.
-Instructions can be found at \url{https://github.com/choderalab/ensembler}.
+The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}.
+\shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
+Up to date instructions can be found at \url{https://github.com/choderalab/ensembler}.
 This will install all dependencies except for Modeller and (optionally) Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.

### 2015-03-15 manuscript/ms.tex
Fixed some issues with figures and caption justification. Removed subcaption because it is incompatible with revtex. Multiple figure issues remain, like font size issues.

@@
 %\usepackage{citesort}
-\usepackage{subcaption}
+\usepackage[caption=false]{subfig}
 \usepackage{dcolumn}
@@
 \usepackage[group-separator={,}]{siunitx}
+
+% Justification
+\captionsetup{singlelinecheck=off}
 
@@
 Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require transitions between states, encompassing anything from reorganization of sidechains at binding interfaces to domain motions to large scale folding-unfolding events.
-Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state [CITE Lee Craik Science 2009].
+Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state~\cite{craik:2009:science:trapping-moving-targets}.
 
@@
 \begin{figure*}[tb]
-  \centering
     \includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
+
   \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
@@
 
-The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig. \ref{figrmsddist}.
+The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig. \ref{figure:rmsd-distribution-by-sequence-identity}.
 The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in lower RMSD models.
 The sequence identity stratifications were selected based on the sequence identity distribution (Fig.
-\ref{figseqiddist}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
+\ref{figure:sequence-identity-distribution}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Sequence identity distribution
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[tb]
-  \centering
-    \includegraphics[width=1.0\textwidth]{seqid_dist/seqid_dist}
+    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist}
+
     \caption{{\bf Sequence identity distribution for human TK models.}
@@
     }
-  \label{figseqiddist}
+  \label{figure:sequence-identity-distribution}
 \end{figure}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: RMSD distribution by sequence identity
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[tb]
-  \centering
-    \includegraphics[width=1.0\textwidth]{rmsddist/rmsddist2}
+    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2}
+
     \caption{{\bf RMSD distribution by sequence identity.}
@@
   }
-  \label{figrmsddist}
+  \label{figure:rmsd-distribution-by-sequence-identity}
 \end{figure}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Superposition of Src and Abl sequence identity classes
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[tb]
-  \centering
-    \includegraphics[width=1.0\textwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+    \includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+
     \caption{{\bf Superposition of clustered models of Src and Abl.}
     Superposed renderings of nine models each for Src and Abl, giving some indication the diversity of conformations generated by Ensembler.
-    The models for each target were divided into three sequence identity ranges (as in Fig. \ref{figrmsddist}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
+    The models for each target were divided into three sequence identity ranges (as in Fig.
+\ref{figure:rmsd-distribution-by-sequence-identity}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
     The models shown are the centroids of each cluster.
@@
   }
-  \label{figsuperposition}
+  \label{figure:superposition}
 \end{figure}
 
-\begin{figure}
-  \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/distances}
-      \caption{Src}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=1.0\textwidth]{residue_pair_distances/abl/distances}
-      \caption{Abl}
-    \end{subfigure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Residue pair distance plots
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}
+
+\subfloat[Src]{\label{figure:src-distances}%
+   \includegraphics[width=0.45\textwidth]{residue_pair_distances/src/distances}
+}
+\subfloat[Abl]{\label{figure:abl-distances}%
+   \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
+}
+
     \caption{{\bf E310-R409 and K295-E310 distances for models of Src and Abl, colored by sequence identity.}
   }
-  \label{figaa_pair_distances_ref_structures}
-\end{figure}
+  \label{figure:pair-distances-ref-structures}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Pairwise distances with reference structures for Src
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{figure}[tb]
-  \centering
     \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
-    % \begin{subfigure}[b]{0.45\textwidth}
-    %   \includegraphics[width=1.0\textwidth, clip=true, trim=30em 0em 40em 0em]{residue_pair_distances/src/inactive}
-    %   \caption{Inactive}
-    % \end{subfigure}
-    % \begin{subfigure}[b]{0.45\textwidth}
-    %   \includegraphics[width=1.0\textwidth, clip=true, trim=30em 0em 27em 10em]{residue_pair_distances/src/active}
-    %   \caption{Active}
-    % \end{subfigure}
+
     \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
@@
   }
-  \label{figsrc_ref_structures}
+  \label{figure:src-ref-structrues}
 \end{figure}

### 2015-03-15 manuscript/ms.tex

@@
 The similarity between members of a given protein family could be exploited to generate arrays of conformational models, which could be used as starting configurations to aid sampling in MD simulations.
-This approach would be highly beneficial for many MD methods, such as Markov state modeling, which require global coverage of the conformational landscape to realize their full potential, and would also be particularly useful in cases where structural data is present for only a subset of the members of a protein family.
-It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---, for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
+This approach would be highly beneficial for many MD methods, such as MSM construction, which require global coverage of the conformational landscape to realize their full potential, and would also be particularly useful in cases where structural data is present for only a subset of the members of a protein family.
+It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
 
-Here, we present the first steps toward bridging the gap between biomolecular simulation software and \emph{omics}-scale sequence and structural data: a fully automated open source framework for building simulation-ready protein models scalable from single sequences to entire superfamilies.
+Here, we present the first steps toward bridging the gap between biomolecular simulation software and \emph{omics}-scale sequence and structural data: a fully automated open source framework for building simulation-ready protein models in multiple conformational substates scalable from single sequences to entire superfamilies.
 {\bf Ensembler} provides functions for selecting target sequences and homologous template structures, and (by interfacing with a number of external packages) performs pairwise alignments, comparative modeling of target-template pairs, and several stages of model refinement.
 As an example application, we have constructed models for the entire set of human tyrosine kinase catalytic domains, using all available structures of protein kinase domains (from any species) as templates.
-This results in a total of almost 400,000 models, and we demonstrate that these provide wide-ranging coverage of known functionally relevant regions of structure.
+This results in a total of almost 400,000 models, and we demonstrate that these provide wide-ranging coverage of known functionally relevant conformations.
 By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
@@
 {\color{blue}[DLP: Good point. I've added in brief introductions for each section.]}
+{\color{red}[JDC: Can you do a bit more here?
+I feel that the reader may need more orientation.
+You've essentially just added a sentence or two at the beginning of each stage that doesn't really enlighten the user as to your terminology (\emph{tempalte} and \emph{target}), your motivation for \emph{why} things are done each stage, or \emph{how} the user is to select among the various options available.]}
 
@@
 The first stage entails the selection of a set of target protein sequences.
+{\color{red}[JDC: Maybe explain what is meant by ``target protein sequences''?  These are the sequences the user is interested in modeling.]}
 
 These targets can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
-The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}), using the subcommand {\tt gather\_targets}.
+The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Isn't there a real citation for UniProt?]}, using the subcommand {\tt gather\_targets}.
 % {\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
@@
 Example identifiers: {\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}.
+{\color{red}[JDC: Does it make sense to set some of these coded examples off on their own lines?]}
 
@@
 The second stage entails the selection of templates and storage of associated structures, sequences and identifiers.
+{\color{red}[JDC: Again, can you provide more information about \emph{why} this is being done? What the motivation is, and how the user might expect to select these?]}
 
@@
 
-As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TKs) domains listed in UniProt.
+{\color{red}[JDC: It would be useful to have some subheadings in this section to give it some internal organization.]}
+
+As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TK) domains listed in UniProt.
+{\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
 TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer.
+{\color{red}[JDC: CITE]}
 For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl cancer involvement]}.
 Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
-Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach could benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.
-by generating large numbers of starting configurations to be used in highly parallel MD simulation.
+{\color{red}[JDC: Lee and Craik do not discuss kinases, I don't believe; you'll have to find an accurate reference on kinase conformations.]}
+Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach stands to benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.~by generating large numbers of starting configurations to be used in highly parallel MD simulation.
 
@@
 Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled.
+{\color{red}[JDC: Can you give some statistics on the distribution of loop lengths modeled?  Why did loop modeling fail in the cases it did?  Anything else you can say here beyond this one sentence?]}
 
@@
 In addition, the solvation stage was performed for two representative individual kinases (\emph{Src} and \emph{Abl}).
-The number of models which survived each stage are shown in Fig. \ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
-Fig. \ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage.
+The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
+Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage.
 
-The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig. \ref{figure:rmsd-distribution-by-sequence-identity}.
+The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
 The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in lower RMSD models.
-The sequence identity stratifications were selected based on the sequence identity distribution (Fig.
-\ref{figure:sequence-identity-distribution}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
+The sequence identity stratifications were selected based on the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
 
@@
     {\color{blue}[DLP: should I mention the use of KDE smoothing?]}
+    {\color{red}[JDC: Yes.]}
     Sequence identities are calculated from pairwise target-template alignments.
     The cumulative distribution function is shown by the dashed line.    
+    {\color{red}[JDC: Font size too small.]}
     }
@@
     For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
+    {\color{red}[JDC: Font size too small.]}
   }
@@
     \caption{{\bf Superposition of clustered models of Src and Abl.}
-    Superposed renderings of nine models each for Src and Abl, giving some indication the diversity of conformations generated by Ensembler.
+    Superposed renderings of nine models each for Src and Abl, 
+    {\color{red}[JDC: Src and Abl, or Src and Abl1? The description should match the captions above.]}
+    giving some indication the diversity of conformations generated by Ensembler.
     The models for each target were divided into three sequence identity ranges (as in Fig.
@@
     Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
+    {\color{red}[JDC: Font size too small.]}
   }
@@
     \caption{{\bf E310-R409 and K295-E310 distances for models of Src and Abl, colored by sequence identity.}
+  {\color{red}[JDC: Font size too small if single-column, so I made this double-column. Not sure if that's what you want.]}
+  {\color{red}[JDC: Fill in rest of caption.]}
   }

### 2015-03-17 manuscript/ms.tex
Updates to Results section.

@@
 The output of this pipeline is an ensemble of structures ready for subsequent molecular simulations using computer clusters, supercomputers, or distributed computing projects like Folding@home.
-{\bf Ensembler} automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.
+{\bf Ensembler} thus automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.
 {\color{red}[JDC: Prior sentence is redundant?]}
@@
 While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
+Models are output as PDB-format coordinate files.
+To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files.
 
@@
 
-The templates were first subjected to loop remodeling, using the {\tt loopmodel} subcommand.
-Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled.
+Unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
+The number of missing residues in each template ranged from 0 to 102, with a median of 11 and a standard deviation of 13.
+Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled, with most remodeling failures attributable to spatial constraints imposed by the original template structure.
+There was some correlation between remodeling failures and the number of missing residues; templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
+The distributions are plotted in Fig. S1.
+% TODO Work out if there is an easy way to reference SI figures
 {\color{red}[JDC: Can you give some statistics on the distribution of loop lengths modeled?  Why did loop modeling fail in the cases it did?  Anything else you can say here beyond this one sentence?]}
+{\color{blue}[DLP: Addressed in the text, and a SI figure.]}
 
 Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
-In addition, the solvation stage was performed for two representative individual kinases (\emph{Src} and \emph{Abl}).
+To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for just two representative individual kinases (\emph{Src} and \emph{Abl}).
 The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
-Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage.
+The number of refined models for each target ranged from 4005 to 4248, with a median of 4160 and standard deviation of 60.
+Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
+
+Each model generated about 513 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 1.7 GB per TK target or 149 GB for all 90 TKs.
+The data generated per model breaks down as 436 kB for the output from the modeling stage---with the largest contribution arising from the Modeller restraint files---and 77 kB for the implicit solvent MD refinement stage.
 
 The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
-The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in lower RMSD models.
-The sequence identity stratifications were selected based on the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
+The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in models with lower RMSDs.
+The sequence identity stratifications were selected based on the sequence identity distribution plotted in Fig.~\ref{figure:sequence-identity-distribution}, which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
 
-% superpositions etc.
+To provide a more complete evaluation of the models generated, we have analyzed two example TKs (\emph{Src} and \emph{Abl}) in detail.
+Due to their importance in cancer, as outlined above, these kinases have been the subject of numerous studies, encompassing many different methodologies.
+In terms of structural data, a large number of crystal structures have been solved (with or without ligands such as nucleotide substrate or inhibitor drugs), showing the kinases in a number of different conformations.
+These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src {\color{blue}[CITE:Shukla Pande Nat Commun 2014]}.
+
+Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl}.
+Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
+Each model is colored and given a transparency based on the sequence identity between the target and template sequence.
+The figure gives an idea of the variance present in the generated models.
+High sequence identity models (in opaque blue) tend to be quite structurally similar, with some variation in loops or changes in domain orientation.
+The Abl1 renderings indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
+While such models are not necessarily incorrect or undersirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
+Lower sequence identity models (in transparent white or red) indicate much greater variation in all parts of the structure.
+We believe the mix of high and low sequence identity models to be particularly useful for methods such as MSM building, which require thorough sampling of the conformational landscape.
+The high sequence identity models could be considered to be the most likely to accurately represent true metastable states.
+Conversely, the lower sequence identity models could be expected to help push a simulation into regions of conformation space which might take intractably long to reach if starting a single metastable conformation.
+
+To evaluate the models of \emph{Src} and \emph{Abl} in the context of the published literature, we have focused on two residue pair distances thought to be important for the activation process in protein kinase domains.
+Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
+A number of articles and review articles suggest that the activation process entails (amongst other processes) the transfer of a salt-bridge from E310 and R409 to a salt-bridge between E310 and K295, thus positioning K295 in the required position for catalysis.
+% TODO CITE
+These residues are also known to be conserved in protein kinase domains, % TODO CITE
+and thus likely also play a role in the activation process of Abl1. % TODO CITE
+% TODO NOTE: distances for Abl1 structures all look like ?inactive or ?active
+As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
+The models appear to have strong coverage of regions in which either of the salt-bridges is formed, as well as a wide range of regions inbetween.
+We thus expect that such a set of models, if used as starting configurations for a parallel MD simulation, could greastly aid in sampling of the activation process.
 
@@
     Distribution of sequence identities for all 373,513 models generated for the human tyrosine kinases.
-    {\color{blue}[DLP: should I mention the use of KDE smoothing?]}
-    {\color{red}[JDC: Yes.]}
     Sequence identities are calculated from pairwise target-template alignments.
     The cumulative distribution function is shown by the dashed line.
+    The plotted distributions have been smoothed using kernel density estimation.
     {\color{red}[JDC: Font size too small.]}
@@
     RMSD distributions for all 373,513 human TK models, divided into three sequence identity ranges.
-    % TODO may want to mention KDE smoothing
     For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
+    The plotted distributions have been smoothed using kernel density estimation.
     {\color{red}[JDC: Font size too small.]}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Reference structures for Src
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[tb]
+    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
+
+    \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
+    In the inactive state, E310 forms a salt bridge with R409.
+    During activation, the C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
+    This positions K295 in the appropriate position for catalysis.
+  }
+  \label{figure:src-ref-structures}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % FIGURE: Residue pair distance plots
@@
 
-    \caption{{\bf E310-R409 and K295-E310 distances for models of Src and Abl, colored by sequence identity.}
+    \caption{{\bf Src and Abl1 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
+    Two Src structures (PDB entries 1Y57 [CITE] and 2SRC [CITE]) are projected onto the plots for reference, representing active and inactive states respectively.
+    These structures and the residue pairs analyzed here are depicted in Fig.~\ref{figure:src-ref-structures}.
+    Distances are measured between the center of masses of the three terminal sidechain heavy atoms of each residue.
+    The atom names for these atoms, according to the PDB coordinate files for both reference structures, are---Lys: NZ, CD, CE (ethylamine); Glu: OE1, CD, OE2 (carboxylate); Arg: NH1, CZ, NH2 (part of guanidine).
     {\color{red}[JDC: Font size too small if single-column, so I made this double-column. Not sure if that's what you want.]}
@@
     }
-  \label{figure:pair-distances-ref-structures}
+  \label{figure:pair-distances}
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Pairwise distances with reference structures for Src
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
-
-    \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
-    In the inactive state, E310 forms a salt bridge with R409.
-    During activation, the C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
-    This positions K295 in the appropriate position for catalysis.
-  }
-  \label{figure:src-ref-structrues}
-\end{figure}
 

### 2015-03-17 supporting-info/si.tex
Updates to Results section.

--- 2015-03-15 manuscript/ms.tex
@@
 %% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Ensembler: Enabling high-throughput molecular simulations at the superfamily scale}
+\title{Supporting Information for "Ensembler: Enabling high-throughput molecular simulations at the superfamily scale"}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABSTRACT/pacs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-
-The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
-While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable simulations at this scale have not kept pace.
-It should now be possible to study protein dynamics across entire (super)families, exploiting both available structural biology data and conformational similarities across homologous proteins.
-Here, we present a new tool for enabling high-throughput simulation in the genomics era.
-{\bf Ensembler} takes any set of sequences---from a single sequence to an entire superfamily---and shepherds them through various stages of modeling and refinement to produce simulation-ready structures.
-This includes comparative modeling to all relevant PDB structures (which may span multiple conformational states of interest), reconstruction of missing loops, addition of missing atoms, culling of nearly identical structures, assignment of appropriate protonation states, solvation in explicit solvent, and refinement and filtering with molecular simulation to ensure stable simulation.
-The output of this pipeline is an ensemble of structures ready for subsequent molecular simulations using computer clusters, supercomputers, or distributed computing projects like Folding@home.
-{\bf Ensembler} automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.
-{\color{red}[JDC: Prior sentence is redundant?]}
-A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
-We demonstrate the power of this approach by constructing models for all catalytic domains in the human tyrosine kinase family, using all available kinase catalytic domain structures from any organism as structural templates.
-
-{\bf Ensembler} is free and open source software licensed under the GNU General Public License (GPL) v2. 
-It should run on all major operating systems, and has been tested on Linux and OS X.
-The latest release can be installed via the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
-
-\emph{Keywords: molecular dynamics simulation; comparative modeling; distributed simulation}
-
-% TODO expand to all human protein kinases?
-
-\end{abstract}
-
 \maketitle
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% INTRODUCTION
+% MAIN
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{section:introduction}
-
-Recent advances in genomics and structural biology have helped generate an enormous wealth of protein data at the level of amino-acid sequence and three-dimensional structure.
-However, proteins typically exist as an ensemble of thermally accessible conformational states, and static structures provide only a snapshot of their rich dynamical behavior.
-Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require transitions between states, encompassing anything from reorganization of sidechains at binding interfaces to domain motions to large scale folding-unfolding events.
-Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state~\cite{craik:2009:science:trapping-moving-targets}.
-
-Molecular dynamics (MD) simulations have the capability, in principle, to describe the time evolution of a protein in atomistic detail, and have proven themselves to be a useful tool in the study of protein dynamics.
-A number of mature software packages and forcefields are available, and much recent progress has been driven by advances in computing architecture.
-For example, many MD packages are now able to exploit GPUs, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE] allow scalability on an unprecedented level.
-In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity [CITE MSM reviews].
-MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
-There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates [CITE Modeller and Rosetta and a recent homology modeling review].
-
-However, it remains difficult for researchers to exploit the full variety of available protein sequence and structural data in simulation studies, largely due to limitations in software architecture.
-For example, the set up of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure, addition of missing residues and atoms, solvation with explicit water and salt buffer, choice of simulation parameters, and system relaxation with energy minimization and one or more short MD simulations.
-For this reason, simulation studies typically consider only one or a few proteins and starting configurations.
-
-The ability to fully exploit the large base of available protein sequence and structural data in biomolecular simulation studies could open up many interesting avenues for research, enabling the study of entire protein families or superfamilies across multiple organisms.
-The similarity between members of a given protein family could be exploited to generate arrays of conformational models, which could be used as starting configurations to aid sampling in MD simulations.
-This approach would be highly beneficial for many MD methods, such as Markov state modeling, which require global coverage of the conformational landscape to realize their full potential, and would also be particularly useful in cases where structural data is present for only a subset of the members of a protein family.
-It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---, for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
-
-Here, we present the first steps toward bridging the gap between biomolecular simulation software and \emph{omics}-scale sequence and structural data: a fully automated open source framework for building simulation-ready protein models scalable from single sequences to entire superfamilies.
-{\bf Ensembler} provides functions for selecting target sequences and homologous template structures, and (by interfacing with a number of external packages) performs pairwise alignments, comparative modeling of target-template pairs, and several stages of model refinement.
-As an example application, we have constructed models for the entire set of human tyrosine kinase catalytic domains, using all available structures of protein kinase domains (from any species) as templates.
-This results in a total of almost 400,000 models, and we demonstrate that these provide wide-ranging coverage of known functionally relevant regions of structure.
-By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
-We anticipate that the tool will prove to be useful in a number of other ways.
-For example, the generated models could represent valuable data sets even without subsequent production simulation, allowing exploration of the conformational diversity present within the available structural data for a given protein family.
-Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DESIGN AND IMPLEMENTATION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design and Implementation}
-
-{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API.
-
-The {\bf Ensembler} modeling pipeline comprises a series of stages which are performed in a defined order. 
-A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}.
-The various stages of this pipeline are described in detail below.
-
-{\color{red}[JDC: We could really help the reader if we preface each section here with a bit of an introduction of what we're trying to accomplish in each stage.
-Otherwise, I worry that each section is a long list of things we do without reference to an overall concept of what the stage is trying to accomplish or why certain decisions were made.]}
-{\color{blue}[DLP: Good point. I've added in brief introductions for each section.]}
 
 \begin{figure*}[tb]
-    \includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
+    \includegraphics[width=0.7\textwidth]{loopmodel_analysis/nmissing_resis_distributions}
 
-  \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
-  The number of viable models surviving each stage of the pipeline are shown, either for all tyrosine kinases (\emph{All TKs}) or representative individual kinases (\emph{SRC} and \emph{ABL}).
-  In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to exemplify the resources required per model and for modeling the entire set of tyrosine kinases.
-  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread---parallel execution can reduce wall clock time nearly linearly.
+  \caption{{\bf Distributions of the number of missing residues for templates for which remodeling (with the {\tt loopmodel} command) was either successful or unsuccessful.}
+  The plotted distributions are smoothed using kernel density estimation, and the raw data points are shown as a rug plot.
   }
@@
 \end{figure*}
-
-\subsubsection{Target selection}
-
-The first stage entails the selection of a set of target protein sequences.
-
-These targets can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
-The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}), using the subcommand {\tt gather\_targets}.
-% {\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
-% {\color{blue}[DLP: Yes, I think TargetExplorer is primarily an internal tool for the moment, so let's omit discussion of that.]}
-The user specifies a query string with the {\tt -{}-query} flag, which conforms to the same syntax as the search function available on the UniProt website.
-% {\color{red}[JDC: What command-line flag or API call is used to select targets from UniProt?]}
-% {\color{red}[DLP: the command-line flag is now described]}
-For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
-In this way, the user may select a single protein, many proteins, or an entire superfamily.
-The program outputs a FASTA file, setting the UniProt mnemonic (e.g. {\tt SRC\_HUMAN}) as the identifier for each target protein.
-% {\color{red}[JDC: Can you give a clearer picture of how this is done?  Does the user just specify different flags for these categories, or would they have to construct different Uniprot searches?]}
-% {\color{blue}[DLP: This is explained more clearly now.]}
-
-In many cases, it will be desirable to build models of an isolated protein domain, rather than the full-length protein.
-The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-domains} flag. 
-For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
-To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
-In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest.
-Example identifiers: {\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}.
-
-\subsubsection{Template selection}
-
-The second stage entails the selection of templates and storage of associated structures, sequences and identifiers.
-
-This data can be provided manually, by storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
-The structure residues must also match those in the sequence file.
-
-The {\tt ensembler gather\_templates} subcommand also provides methods for selecting template structures from either UniProt or the Protein Data Bank (PDB; \href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
-Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.
-for crystal unit cells with multiple asymmetric units) would thus give rise to multiple template structures.
-
-% {\color{red}[JDC: Do you use the RCSB or some other PDB site?]} - DLP: I use the RCSB, but all PDB sites are updated simultaneously so I think it shouldn't matter.
-However, I've updated the link so that it points to the RCSB PDB.
-% {\color{red}[JDC: Again, not sure if it is helpful to leave this in unless it will be immediately useful to the reader]} - DLP: references to TargetExplorer removed
-
-Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g. {\tt -{}-query 2H8H,1Y57}.
-% {\color{red}[JDC: Who specifies the list of PDB IDs?
-How are they specified?
-Can this be built automatically, or does the user need to specify these?]} DLP: these are specified by the user as a comma-separated list, as is now described above.
-Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
-The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries.
-The SIFTS data is used to extract template sequences, retaining only residues which are resolved and match the equivalent residue in the UniProt sequence---non-wildtype residues are thus removed from the template structures.
-Furthermore, PDB chains with less than a given percentage of resolved residues (default: 70\%) are filtered out.
-Sequences are stored in a FASTA file, with identifiers of the form {\tt [UniProt mnemonic]\_D[UniProt domain index]\_[PDB ID]\_[PDB chain ID]}, e.g. {\tt SRC\_HUMAN\_D0\_2H8H\_A}.
-Matching residues then extracted from the original coordinate files and stored as PDB-format coordinate files.
-
-Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-domains} flag allows selection of individual domains with a regular expression string.
-The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
-If the {\tt -{}-domains} flag is used, then templates are truncated at the start and end of the domain sequence.
-
-Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm {\color{red}[CITE]} provided via the {\tt loopmodel} tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
-Because fewer loops need to be built during the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
-Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
-Furthermore, the Rosetta {\tt loopmodel} program will not model missing residues at the termini of a structure---such residues spans are modeled in the subsequent stage.
-
-\subsubsection{Modeling}
-
-This stage entails the generation of models via comparative modeling of each target sequence onto each template structure. Non-unique models are filtered out using a RMSD-based clustering scheme.
-
-Modeling is performed with the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
-While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
-
-All chains of template structures that contain the template sequence are utilized in the modeling phase, which can sometimes cause models to be nearly identical.
-Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
-The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD (for C$\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~{\color{red}[Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
-A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
-% {\color{red}[JDC: Which atoms are used in the RMSD comparison?  All atoms, heavy atoms, or CA only?]} - DLP: CA only. This is now specified in the text.
-
-\subsubsection{Refinement}
-
-This stage entails the use of molecular dynamics simulations to refine the models built in the previous step.
-This helps to improve model quality and also prepares models for subsequent production simulation, including solvation with explicit water molecules, if desired.
-
-Models are first subjected to energy minimization (using the L-BFGS algorithm {\color{red}[CITE]}), followed by a short molecular dynamics (MD) simulation with an implicit solvent representation.
-This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-By default, the Amber99SB-ILDN force field is used {\color{red}[CITE: amber99sbildn refs]} with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
-The {\bf Ensembler} API allows the use of any of the other force fields implemented in OpenMM.
-The simulation is run for a default of 100 ps to filter out poor quality models (where atomic overlaps that cannot be resolved by energy minimization would cause the simulation to explode) and help relax models for subsequent production simulation.
-{\color{red}[JDC: What criteria were applied to filter out poor models?  Do we only look for thrown exceptions or NaNs?  Or do we use an energy filtering criteria too?]}
-{\color{blue}[DLP: We currently just filter out models which throw exceptions or NaNs.]}
-
-While protein-only models may be sufficient for structural analysis or implicit solvent simulations, {\bf Ensembler} also provides a stage for solvating models with explicit water and performing a round of explicit-solvent MD refinement/equilibration under isothermal-isobaric (NPT) conditions.
-The solvation step solvates each model for a given target with the same number of waters to facilitate the integration of data from multiple simulations, such as the construction of MSMs.
-The target number of waters is selected by first solvating each model with a specified padding distance (default: 10 \AA), then taking a percentile value from the distribution (default: 68th percentile).
-{\color{red}[JDC: Would be useful to explain why we are doing this.]}
-{\color{blue}[DLP: Addressed.]}
-This helps to prevent models with particularly long, extended loops---such as those arising from template structures with unresolved termini---from imposing very large box sizes on the entire set of models.
-Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
-The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field and TIP3P water {\color{red}[JDC: CITE]} by default.
-Other force fields or water models such as TIP4P-Ew {\color{red}[CITE]}) can be specified via the {\bf Ensembler} API.
-{\color{red}[JDC: We should allow other water models in OpenMM too, such as TIP4P-Ew?]}
-{\color{blue}[DLP: I forgot to mention this in the text previously - any of the OpenMM force fields can be chosen via the API.
-I've updated the text accordingly.
-Is this functionality sufficient?
-I guess it's ok to leave ff choice as an "advanced" feature which requires use of the API?
-Otherwise I could add a -{}-water\_model flag to the CLI, for example.]}
-
-{\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
-For example, we don't yet handle counterions (e.g.
-structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
-We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
-It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
-
-\subsubsection{Packaging}
-
-{\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home (CITE: F@H).
-
-\subsubsection{Provenance}
-
-To aid the user in tracking the provenance of each model, each pipeline function also outputs a metadata file, which helps to link data to the software version used to generate it (both {\bf Ensembler} and its dependencies), and also provides timing and performance information, and other data such as hostname.
-
-\subsubsection{Rapidly modeling a single template}
-
-% quickmodel function performs the entire pipeline in one go, for a single target and a small number of templates.
-For users interested in simply using {\bf Ensembler} to rapidly generate a set of models for a single template sequence, {\bf Ensembler} provides a command-line tool {\tt quickmodel}, which performs the entire pipeline for a single target with a small number of templates.
-For larger numbers of models (such as entire protein families), modeling time is greatly reduced by using the main modeling pipeline, which is parallelized via MPI, distributing computation across each model (or across each template, in the case of the loop reconstruction code), and scaling (in a ``pleasantly parallel'' manner) up to the number of models generated.
-
-% TODO? The API is highly flexible, allowing control over many important parameters. Appropriate choices have been set as default values.
-
-% TODO? other tools for analyzing project etc.
-
-% TODO where do these go?
-% Distribution: conda package or github source
-% Software is open source and designed to be extensible as far as possible.
-For example, it would be very simple to implement further options for providing target and template structures, or for packaging models.
-
-\label{section:design}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% RESULTS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{section:results}
-
-As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TKs) domains listed in UniProt.
-TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer.
-For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl cancer involvement]}.
-Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
-Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach could benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.
-by generating large numbers of starting configurations to be used in highly parallel MD simulation.
-
-We selected all available structures of protein kinase domains (of any species) as templates, for a total of 4433 (\num{398970} target-template pairs).
-The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
-
-The templates were first subjected to loop remodeling, using the {\tt loopmodel} subcommand.
-Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled.
-
-Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
-In addition, the solvation stage was performed for two representative individual kinases (\emph{Src} and \emph{Abl}).
-The number of models which survived each stage are shown in Fig. \ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
-Fig. \ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage.
-
-The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig. \ref{figure:rmsd-distribution-by-sequence-identity}.
-The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in lower RMSD models.
-The sequence identity stratifications were selected based on the sequence identity distribution (Fig.
-\ref{figure:sequence-identity-distribution}), which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
-
-% superpositions etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Sequence identity distribution
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist}
-
-    \caption{{\bf Sequence identity distribution for human TK models.}
-    Distribution of sequence identities for all 373,513 models generated for the human tyrosine kinases.
-    {\color{blue}[DLP: should I mention the use of KDE smoothing?]}
-    Sequence identities are calculated from pairwise target-template alignments.
-    The cumulative distribution function is shown by the dashed line.    
-    }
-  \label{figure:sequence-identity-distribution}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: RMSD distribution by sequence identity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2}
-
-    \caption{{\bf RMSD distribution by sequence identity.}
-    RMSD distributions for all 373,513 human TK models, divided into three sequence identity ranges.
-    % TODO may want to mention KDE smoothing
-    For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
-  }
-  \label{figure:rmsd-distribution-by-sequence-identity}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Superposition of Src and Abl sequence identity classes
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
-
-    \caption{{\bf Superposition of clustered models of Src and Abl.}
-    Superposed renderings of nine models each for Src and Abl, giving some indication the diversity of conformations generated by Ensembler.
-    The models for each target were divided into three sequence identity ranges (as in Fig.
-\ref{figure:rmsd-distribution-by-sequence-identity}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
-    The models shown are the centroids of each cluster.
-    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
-  }
-  \label{figure:superposition}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Residue pair distance plots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-
-\subfloat[Src]{\label{figure:src-distances}%
-   \includegraphics[width=0.45\textwidth]{residue_pair_distances/src/distances}
-}
-\subfloat[Abl]{\label{figure:abl-distances}%
-   \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
-}
-
-    \caption{{\bf E310-R409 and K295-E310 distances for models of Src and Abl, colored by sequence identity.}
-  }
-  \label{figure:pair-distances-ref-structures}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Pairwise distances with reference structures for Src
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
-
-    \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
-    In the inactive state, E310 forms a salt bridge with R409.
-    During activation, the C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
-    This positions K295 in the appropriate position for catalysis.
-  }
-  \label{figure:src-ref-structrues}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Availability and Future Directions}
-\label{section:availability}
-
-The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}.
-\shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
-Up to date instructions can be found at \url{https://github.com/choderalab/ensembler}.
-This will install all dependencies except for Modeller and (optionally) Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
-The latest source can be downloaded from the above GitHub repository, which also contains instructions for building and installing the code.
-
-% TODO future directions
-% TODO what's on the Ensembler 2 roadmap?
-% TODO MCCE?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ACKNOWLEDGMENTS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-\label{section:acknowledgments}
-
-The authors are grateful to Kyle A.~Beauchamp (MSKCC), Robert McGibbon (Stanford), Arien S. Rustenburg (MSKCC) for many excellent software engineering suggestions.
-The authors thank Sonya M.~Hanson (MSKCC), Nicholas M.~Levinson (?), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
-The authors are grateful to Benjamin Webb and Andrej Sali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
-DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
 

### 2015-03-17 figures/loopmodel_analysis/plot_dists.py

@@ -1,0 +1,34 @@
+import pandas as pd
+import seaborn as sns
+
+sns.set_context('paper', font_scale=1.7)
+
+df = pd.read_csv('loopmodel_data.csv')
+
+successful = df[df.no_missing_residues == False][df.successful == True]
+unsuccessful = df[df.no_missing_residues == False][df.successful == False]
+
+sns.distplot(successful.nmissing_resis, hist=False, kde=True, rug=True, label='successful',
+    rug_kws={'height': 0.002},
+)
+sns.distplot(unsuccessful.nmissing_resis, hist=False, kde=True, rug=True, label='unsuccessful', color='red',
+    # kde_kws={'bw': 2.5},
+    rug_kws={'height': 0.001},
+)
+
+# sns.violinplot([successful.nmissing_resis, unsuccessful.nmissing_resis], inner='points',
+#     inner_kws={'markersize': 5, 'marker': 'o'},
+#     alpha=0.5,
+# )
+# sns.violinplot([successful.nmissing_resis, unsuccessful.nmissing_resis], inner='stick',
+#     inner_kws={'width': 5},
+#     alpha=0.5,
+# )
+
+ax = sns.plt.gca()
+sns.plt.setp(ax, yticks=[])
+sns.plt.xlim(0,120)
+sns.plt.tight_layout()
+sns.plt.xlabel('Number of missing residues')
+
+sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)

### 2015-03-17 figures/seqid_dist/seqid_dist.py

@@ -3,8 +3,11 @@ import ensembler
 import pandas as pd
 import numpy as np
 import seaborn as sns
-import matplotlib.pyplot as plt
 from matplotlib import lines
+
+# ====================
+# Data
+# ====================
 
 data_dir = os.path.join('..', '..', 'data')
 targets = open(os.path.join(data_dir, 'TKs.txt')).read().splitlines()
@@ -20,10 +23,16 @@ for target in targets[1:]:
 
 df_has_model = combined_dfs[combined_dfs.has_model]
 
-# Plot sequence identity distribution
+# ====================
+# Plotting
+# ====================
+
+# sns.set_context('paper', font_scale=1.7)
 
 sns.set(style="white")
-fig, ax1 = plt.subplots()
+fig = sns.plt.figure()
+ax1 = sns.plt.axes()
+# fig, ax1 = sns.plt.subplots()
 
 palette = sns.color_palette('deep', 2)
 
@@ -32,7 +41,7 @@ sns.kdeplot(df_has_model.seqid, bw=1., ax=ax1, color='black', shade=True)
 ax1.legend_.remove()
 ax1.set_xlabel('sequence identity (%)')
 ax1.set_xlim(100,0)
-plt.setp(ax1, yticks=[])
+sns.plt.setp(ax1, yticks=[])
 
 ax2 = ax1.twinx().twiny()
 sns.kdeplot(100.-df_has_model.seqid[::100], bw=1., cumulative=True, ax=ax2, color='black', linestyle='--')
@@ -40,7 +49,7 @@ ax2.set_ylabel('CDF')
 ax2.legend_.remove()
 
 ax2.set_xlim(0,100)
-plt.setp(ax2, xticks=[])
+sns.plt.setp(ax2, xticks=[])
 
 x1  = [-3.0, -3.0]
 y1  = [0.68, 0.78]
@@ -55,5 +64,5 @@ ax2.add_line(line2)
 ax2.text(-3., 0.5, 'probability density', rotation='vertical', ha='center', va='center')
 ax2.text(107.5, 0.5, 'CDF', rotation='vertical', ha='center', va='center')
 
-# plt.tight_layout()
-plt.savefig('seqid_dist.png')
+# sns.plt.tight_layout()
+sns.plt.savefig('seqid_dist.png')

### 2015-03-18 manuscript/ms.tex
Added residue numbering schemes

@@
 
-\subsubsection{Target selection}
+\subsubsection*{Target selection}
 
-The first stage entails the selection of a set of target protein sequences.
+The first stage entails the selection of a set of target protein sequences, i.e. the sequences the user is interested in modeling.
 {\color{red}[JDC: Maybe explain what is meant by ``target protein sequences''?  These are the sequences the user is interested in modeling.]}
+{\color{blue}[DLP: Addressed.]}
 
@@
 The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Isn't there a real citation for UniProt?]}, using the subcommand {\tt gather\_targets}.
-% {\color{red}[JDC: Does it make sense to omit discussion of TargetExplorer since it's not clear how useful this will be to users?]}
-% {\color{blue}[DLP: Yes, I think TargetExplorer is primarily an internal tool for the moment, so let's omit discussion of that.]}
 The user specifies a query string with the {\tt -{}-query} flag, which conforms to the same syntax as the search function available on the UniProt website.
-% {\color{red}[JDC: What command-line flag or API call is used to select targets from UniProt?]}
-% {\color{red}[DLP: the command-line flag is now described]}
 For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
@@
 The program outputs a FASTA file, setting the UniProt mnemonic (e.g. {\tt SRC\_HUMAN}) as the identifier for each target protein.
-% {\color{red}[JDC: Can you give a clearer picture of how this is done?  Does the user just specify different flags for these categories, or would they have to construct different Uniprot searches?]}
-% {\color{blue}[DLP: This is explained more clearly now.]}
 
@@
 
-\subsubsection{Template selection}
+\subsubsection*{Template selection}
 
-The second stage entails the selection of templates and storage of associated structures, sequences and identifiers.
+{\bf Ensembler} uses comparative modeling to build models, and as such requires a set of structures to be used as templates.
+The second stage thus entails the selection of templates and storage of associated sequences, structures and identifiers.
+These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates from the Protein Data Bank (PDB) or from UniProt.
+A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing some degree of homology between targets and templates.
 {\color{red}[JDC: Again, can you provide more information about \emph{why} this is being done? What the motivation is, and how the user might expect to select these?]}
+{\color{blue}[DLP: Addressed]}
 
-This data can be provided manually, by storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
+Manual selection of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
 The structure residues must also match those in the sequence file.
 
-The {\tt ensembler gather\_templates} subcommand also provides methods for selecting template structures from either UniProt or the Protein Data Bank (PDB; \href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
+The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
 Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.
@@
 
-% {\color{red}[JDC: Do you use the RCSB or some other PDB site?]} - DLP: I use the RCSB, but all PDB sites are updated simultaneously so I think it shouldn't matter.
-However, I've updated the link so that it points to the RCSB PDB.
-% {\color{red}[JDC: Again, not sure if it is helpful to leave this in unless it will be immediately useful to the reader]} - DLP: references to TargetExplorer removed
-
 Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g. {\tt -{}-query 2H8H,1Y57}.
-% {\color{red}[JDC: Who specifies the list of PDB IDs?
-How are they specified?
-Can this be built automatically, or does the user need to specify these?]} DLP: these are specified by the user as a comma-separated list, as is now described above.
 Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
@@
 
-\subsubsection{Modeling}
+\subsubsection*{Modeling}
 
-This stage entails the generation of models via comparative modeling of each target sequence onto each template structure. Non-unique models are filtered out using a RMSD-based clustering scheme.
+This stage entails the generation of models via comparative modeling of each target sequence onto each template structure.
+Non-unique models are subsequently filtered out using a RMSD-based clustering scheme.
 
 Modeling is performed with the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
-While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module (CITE: BioPython)---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
+While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module [CITE: BioPython]---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
 Models are output as PDB-format coordinate files.
-To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files.
+A list of all model identifiers sorted by sequence identity is also written to a text file.
+To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
 
@@
 A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
-% {\color{red}[JDC: Which atoms are used in the RMSD comparison?  All atoms, heavy atoms, or CA only?]} - DLP: CA only. This is now specified in the text.
 
-\subsubsection{Refinement}
+\subsubsection*{Refinement}
 
@@
 
-{\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
-For example, we don't yet handle counterions (e.g.
-structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
-We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
-It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
-
-\subsubsection{Packaging}
+\subsubsection*{Packaging}
 
@@
 
-\subsubsection{Provenance}
+\subsubsection*{Provenance}
 
@@
 
-\subsubsection{Rapidly modeling a single template}
+\subsubsection*{Rapidly modeling a single template}
 
-% quickmodel function performs the entire pipeline in one go, for a single target and a small number of templates.
 For users interested in simply using {\bf Ensembler} to rapidly generate a set of models for a single template sequence, {\bf Ensembler} provides a command-line tool {\tt quickmodel}, which performs the entire pipeline for a single target with a small number of templates.
@@
 
+\subsubsection*{Modeling of all human tyrosine kinase catalytic domains}
+
 As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TK) domains listed in UniProt.
@@
 {\color{red}[JDC: CITE]}
-For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl cancer involvement]}.
+For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl1 cancer involvement]}.
 Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
@@
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
+
+\subsubsection*{Ensembler modeling statistics}
 
@@
 Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
-To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for just two representative individual kinases (\emph{Src} and \emph{Abl}).
+To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
 The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
@@
 
+\subsubsection*{Evaluation of model quality}
+
 The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
@@
 
-To provide a more complete evaluation of the models generated, we have analyzed two example TKs (\emph{Src} and \emph{Abl}) in detail.
+To provide a more complete evaluation of the models generated, we have analyzed two example TKs (\emph{Src} and \emph{Abl1}) in detail.
 Due to their importance in cancer, as outlined above, these kinases have been the subject of numerous studies, encompassing many different methodologies.
@@
 
-Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl}.
+Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
 Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
@@
 
-To evaluate the models of \emph{Src} and \emph{Abl} in the context of the published literature, we have focused on two residue pair distances thought to be important for the activation process in protein kinase domains.
+To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
+We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S1.
 Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
-A number of articles and review articles suggest that the activation process entails (amongst other processes) the transfer of a salt-bridge from E310 and R409 to a salt-bridge between E310 and K295, thus positioning K295 in the required position for catalysis.
-% TODO CITE
-These residues are also known to be conserved in protein kinase domains, % TODO CITE
-and thus likely also play a role in the activation process of Abl1. % TODO CITE
-% TODO NOTE: distances for Abl1 structures all look like ?inactive or ?active
-As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
-The models appear to have strong coverage of regions in which either of the salt-bridges is formed, as well as a wide range of regions inbetween.
-We thus expect that such a set of models, if used as starting configurations for a parallel MD simulation, could greastly aid in sampling of the activation process.
+One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
+These three residues are also well conserved [CITE Kannan Neuwald JMB 2005] and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
+As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
+The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
+We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of the activation process.
 
@@
     \caption{{\bf Sequence identity distribution for human TK models.}
-    Distribution of sequence identities for all 373,513 models generated for the human tyrosine kinases.
+    Distribution of sequence identities for all \num{373513} models generated for the human tyrosine kinases.
     Sequence identities are calculated from pairwise target-template alignments.
@@
     \caption{{\bf RMSD distribution by sequence identity.}
-    RMSD distributions for all 373,513 human TK models, divided into three sequence identity ranges.
+    RMSD distributions for all \num{373513} human TK models, divided into three sequence identity ranges.
     For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Superposition of Src and Abl sequence identity classes
+% FIGURE: Superposition of Src and Abl1 sequence identity classes
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-    \caption{{\bf Superposition of clustered models of Src and Abl.}
-    Superposed renderings of nine models each for Src and Abl, 
+    \caption{{\bf Superposition of clustered models of Src and Abl1.}
+    Superposed renderings of nine models each for Src and Abl1, 
     {\color{red}[JDC: Src and Abl, or Src and Abl1? The description should match the captions above.]}
+    {\color{blue}[DLP: Addressed. Using Abl1, as this is the HGNC recommended symbol.]}
     giving some indication the diversity of conformations generated by Ensembler.
@@
     In the inactive state, E310 forms a salt bridge with R409.
-    During activation, the C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
+    During activation, the $\alpha$C-helix (green) moves and rotates, orienting E310 towards the ATP-binding site and allowing it to instead form a salt bridge with K295.
     This positions K295 in the appropriate position for catalysis.
@@
 }
-\subfloat[Abl]{\label{figure:abl-distances}%
+\subfloat[Abl1]{\label{figure:abl-distances}%
    \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
@@
 Up to date instructions can be found at \url{https://github.com/choderalab/ensembler}.
-This will install all dependencies except for Modeller and (optionally) Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
+This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
 The latest source can be downloaded from the above GitHub repository, which also contains instructions for building and installing the code.
+
+% Limitations:
+% counterions
+% PTMs
+% long loops
+
+{\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
+For example, we don't yet handle counterions (e.g.
+structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
+We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
+It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
 
@@
 % TODO MCCE?
+
+% TODO CITE: Shan Shaw Proton-dependent switch Abl1 PNAS 2009
 

### 2015-03-18 manuscript/ms.tex
Moved residue numbering schemes into an appendix in the main text

@@
 \usepackage[pdftex]{graphicx}
+
+\usepackage{alltt}
 
@@
 To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
-We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S1.
+We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
 Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% APPENDICES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\onecolumngrid
+\newpage
+\appendix
+
+\renewcommand{\thesection}{\arabic{section}}
+\section{Sequences and residue numbering schemes for Src and Abl1}
+\label{si:residue-numbering-schemes}
+
+Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
+
+\subsubsection*{Human Abl1 sequence}
+
+\begin{alltt}
+1     MLEICLKLVG CKSKKGLSSS SSCYLEEALQ RPVASDFEPQ GLSEAARWNS KENLLAGPSE    60
+61    NDPNLFVALY DFVASGDNTL SITKGEKLRV LGYNHNGEWC EAQTKNGQGW VPSNYITPVN   120
+121   SLEKHSWYHG PVSRNAAEYL LSSGINGSFL VRESESSPGQ RSISLRYEGR VYHYRINTAS   180
+181   DGKLYVSSES RFNTLAELVH HHSTVADGLI TTLHYPAPKR NKPTVYGVSP NYDKWEMERT   240
+241   D{\color{red}ITMKHKLGG GQYGEVYEGV WKKYSLTVAV {\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TLKEDTMEV EEFLK{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AAVM KEIKHPNLVQ}   300
+301   {\color{red}LLGVCTREPP FYIITEFMTY GNLLDYLREC NRQEVNAVVL LYMATQISSA MEYLEKKNFI}   360
+361   {\color{red}HRDLAARNCL VGENHLVKVA DFGLS{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LMTG DTYTAHAGAK FPIKWTAPES LAYNKFSIKS}   420
+421   {\color{red}DVWAFGVLLW EIATYGMSPY PGIDLSQVYE LLEKDYRMER PEGCPEKVYE LMRACWQWNP}   480
+481   {\color{red}SDRPSFAEIH QAF}ETMFQES SISDEVEKEL GKQGVRGAVS TLLQAPELPT KTRTSRRAAE   540
+541   HRDTTDVPEM PHSKGQGESD PLDHEPAVSP LLPRKERGPP EGGLNEDERL LPKDKKTNLF   600
+601   SALIKKKKKT APTPPKRSSS FREMDGQPER RGAGEEEGRD ISNGALAFTP LDTADPAKSP   660
+661   KPSNGAGVPN GALRESGGSG FRSPHLWKKS STLTSSRLAT GEEEGGGSSS KRFLRSCSAS   720
+721   CVPHGAKDTE WRSVTLPRDL QSTGRQFDSS TFGGHKSEKP ALPRKRAGEN RSDQVTRGTV   780
+781   TPPPRLVKKN EEAADEVFKD IMESSPGSSP PNLTPKPLRR QVTVAPASGL PHKEEAGKGS   840
+841   ALGTPAAAEP VTPTSKAGSG APGGTSKGPA EESRVRRHKH SSESPGRDKG KLSRLKPAPP   900
+901   PPPAASAGKA GGKPSQSPSQ EAAGEAVLGA KTKATSLVDA VNSDAAKPSQ PGEGLKKPVL   960
+961   PATPKPQSAK PSGTPISPAP VPSTLPSASS ALAGDQPSST AFIPLISTRV SLRKTRQPPE  1020
+1021  RIASGAITKG VVLDSTEALC LAISRNSEQM ASHSAVLEAG KNLYTFCVSY VDSIQQMRNK  1080
+1081  FAFREAINKL ENNLRELQIC PATAGSGPAA TQDFSKLLSS VKEISDIVQR             1130
+\end{alltt}
+
+\subsubsection*{Sequences for human and chicken Src, aligned using Clustal Omega}
+
+\begin{alltt}
+SRC_HUMAN   1     MGSNKSKPKD ASQRRRSLEP AENVHGAGGG AFPASQTPSK PASADGHRGP SAAFAPAAAE    60
+SRC_CHICK   1     MGSSKSKPKD PSQRRRSLEP PDSTH---HG GFPASQTPNK TAAPDTHRTP SRSFGTVATE    57
+                  ***.******  *********  :..*    * .*******.*  *: * ** * * :*. .*:*  
+SRC_HUMAN   61    PKLFGGFNSS DTVTSPQRAG PLAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   120
+SRC_CHICK   58    PKLFGGFNTS DTVTSPQRAG ALAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   117
+                  ********:* **********  ********* ********** ********** **********  
+SRC_HUMAN   121   WWLAHSLSTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNAEN PRGTFLVRES   180
+SRC_CHICK   118   WWLAHSLTTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNPEN PRGTFLVRES   177
+                  *******:** ********** ********** ********** ******* ** **********  
+SRC_HUMAN   181   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFNSLQQLV AYYSKHADGL   240
+SRC_CHICK   178   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFSSLQQLV AYYSKHADGL   237
+                  ********** ********** ********** ********** ***.****** **********  
+SRC_HUMAN   241   CHRLTTVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   300
+SRC_CHICK   238   CHRLTNVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   297
+                  *****.**** ********** ********** ********** ********** **********  
+SRC_HUMAN   301   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGETGKY}   360
+SRC_CHICK   298   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGEMGKY}   357
+                  ********** ********** ********** ********** ********** ****** ***  
+SRC_HUMAN   361   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   420
+SRC_CHICK   358   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   417
+                  ********** ********** ********** ********** ********** **********  
+SRC_HUMAN   421   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   480
+SRC_CHICK   418   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   477
+                  ********** ********** ********** ********** ********** **********  
+SRC_HUMAN   481   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKEPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       536
+SRC_CHICK   478   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKDPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       533
+                  ********** ********** ****:***** ********** ********** ******
+\end{alltt}
+
 \end{document}

### 2015-03-18 manuscript/ms.tex
Sorted out figure font sizes

@@
 There was some correlation between remodeling failures and the number of missing residues; templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
-The distributions are plotted in Fig. S1.
+The distributions are plotted in Fig. S\ref{si:loopmodel-nmissing-residues}.
 % TODO Work out if there is an easy way to reference SI figures
@@
     The plotted distributions have been smoothed using kernel density estimation.
-    {\color{red}[JDC: Font size too small.]}
+    % {\color{red}[JDC: Font size too small.]}
     }
@@
     The plotted distributions have been smoothed using kernel density estimation.
-    {\color{red}[JDC: Font size too small.]}
+    % {\color{red}[JDC: Font size too small.]}
   }
@@
     Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
-    {\color{red}[JDC: Font size too small.]}
+    % {\color{red}[JDC: Font size too small.]}
   }
@@
 
-\subfloat[Src]{\label{figure:src-distances}%
-   \includegraphics[width=0.45\textwidth]{residue_pair_distances/src/distances}
-}
-\subfloat[Abl1]{\label{figure:abl-distances}%
-   \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
-}
+% \subfloat[Src]{\label{figure:src-distances}%
+%    \includegraphics[width=0.45\textwidth]{residue_pair_distances/src/distances}
+% }
+% \subfloat[Abl1]{\label{figure:abl-distances}%
+%    \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
+% }
+
+    \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances}
 
@@
     The atom names for these atoms, according to the PDB coordinate files for both reference structures, are---Lys: NZ, CD, CE (ethylamine); Glu: OE1, CD, OE2 (carboxylate); Arg: NH1, CZ, NH2 (part of guanidine).
-    {\color{red}[JDC: Font size too small if single-column, so I made this double-column. Not sure if that's what you want.]}
-    {\color{red}[JDC: Fill in rest of caption.]}
+    % {\color{red}[JDC: Font size too small if single-column, so I made this double-column. Not sure if that's what you want.]}
+    % {\color{red}[JDC: Fill in rest of caption.]}
     }
@@
 \appendix
+\setcounter{figure}{0}
 
@@
 
+\section{Figures}
+
+\begin{figure*}[h]
+    \includegraphics[width=0.5\textwidth]{loopmodel_analysis/nmissing_resis_distributions}
+    \caption{{\bf Distributions for the number of missing residues for templates for which remodeling (with the {\tt loopmodel} command) was either successful or unsuccessful.}
+    The plotted distributions are smoothed using kernel density estimation, and the raw data points are shown as a rug plot.
+    }
+    \label{si:loopmodel-nmissing-residues}
+\end{figure*}
+
 \end{document}

### 2015-03-18 figures/loopmodel_analysis/plot_dists.py

@@ -1,12 +1,17 @@
 import pandas as pd
 import seaborn as sns
 
-sns.set_context('paper', font_scale=1.7)
-
 df = pd.read_csv('loopmodel_data.csv')
 
 successful = df[df.no_missing_residues == False][df.successful == True]
 unsuccessful = df[df.no_missing_residues == False][df.successful == False]
+
+# ====
+# plot
+# ====
+
+sns.set_context('paper', font_scale=1)
+sns.plt.figure(figsize=(3.5, 2.625))
 
 sns.distplot(successful.nmissing_resis, hist=False, kde=True, rug=True, label='successful',
     rug_kws={'height': 0.002},
@@ -28,7 +33,8 @@ sns.distplot(unsuccessful.nmissing_resis, hist=False, kde=True, rug=True, label=
 ax = sns.plt.gca()
 sns.plt.setp(ax, yticks=[])
 sns.plt.xlim(0,120)
-sns.plt.tight_layout()
 sns.plt.xlabel('Number of missing residues')
+
+sns.plt.tight_layout()
 
 sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)

### 2015-03-18 figures/residue_pair_distances/src_abl_combined/plot.py

@@ -12,8 +12,8 @@ df = pd.read_csv(os.path.join(srcdir, 'traj-refine_implicit_md-data.csv'))
 
 traj = mdtraj.load(os.path.join(srcdir, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir, 'traj-refine_implicit_md-topol.pdb'))
 
-inactive_structure = mdtraj.load('2SRC.pdb')
-active_structure = mdtraj.load('1Y57.pdb')
+inactive_structure = mdtraj.load('../src/2SRC.pdb')
+active_structure = mdtraj.load('../src/1Y57.pdb')
 
 # K295 = 29
 # E310 = 44
@@ -54,28 +54,47 @@ active_structure_contacts = mdtraj.compute_contacts(active_structure, contacts=a
 # plot
 # ====
 
-sns.set_context('paper', font_scale=1.)
+sns.set_context('paper', font_scale=0.8)
 sns.set_style('whitegrid')
-fig = sns.plt.figure(figsize=(3.5,2.625))
+fig = sns.plt.figure(figsize=(7.25,2.625))
+ax1 = sns.plt.subplot(1,2,1)
 
 seqids = df.seqid
 
 # ax_models = sns.plt.scatter(pairs_dict[pairs[0]]['distances'][::-1], pairs_dict[pairs[1]]['distances'][::-1], c=seqids[::-1], cmap=sns.plt.cm.coolwarm_r, marker='o', alpha=0.7, vmin=0, vmax=100)
 ax_models = sns.plt.scatter(contacts[0][::-1], contacts[1][::-1], c=seqids[::-1], cmap=sns.plt.cm.coolwarm_r, marker='o', alpha=0.7, vmin=0, vmax=100, s=15.)
-cb = sns.plt.colorbar(ax_models, label='sequence identity (%)')
-# cb.solids.set_edgecolor("face")   # makes sure colorbar is smooth
 
 sns.plt.scatter(active_structure_contacts[0], active_structure_contacts[1], facecolor='g', marker='*', s=80., linewidth=0.5, label='1Y57 (SRC, active)')
 sns.plt.scatter(inactive_structure_contacts[0], inactive_structure_contacts[1], facecolor='r', marker='*', s=80., linewidth=0.5, label='2SRC (SRC, inactive)')
 
 sns.plt.xlabel('-'.join(pairs[0]) + ' (nm)')
[…]

### 2015-03-18 figures/rmsddist/rmsddist2.py

@@ -33,7 +33,13 @@ seqid_ranges_labels = deepcopy(seqid_ranges)
 
 seqid_ranges[-1][-1] = 101.
 
-sns.set(style="white")
+# ========
+# Plotting
+# ========
+
+sns.set_context('paper', font_scale=1.)
+sns.set_style('white')
+fig = sns.plt.figure(figsize=(3.5,2.625))
 
 # palette = sns.blend_palette([sns.desaturate('royalblue', 0), 'royalblue'], 5)
 palette = sns.color_palette('GnBu_d', n_colors=3)[::-1]
@@ -58,11 +64,15 @@ for i, seqid_range in enumerate(seqid_ranges):
 ax = plt.gca()
 plt.setp(ax, yticks=[])
 
-plt.legend(title='Sequence identity (%)')
+legend = plt.legend(title='Sequence identity (%)', fontsize='xx-small')
+plt.setp(legend.get_title(),fontsize='x-small')
 plt.xlim(0,40)
 plt.xlabel('RMSD ($\AA$)')
 sns.despine(left=True)
-plt.savefig('rmsddist2.png')
+
+plt.tight_layout()
+
+plt.savefig('rmsddist2.png', dpi=300)
 
 plt.close()
 

### 2015-03-18 figures/seqid_dist/seqid_dist.py

@@ -27,12 +27,11 @@ df_has_model = combined_dfs[combined_dfs.has_model]
 # Plotting
 # ====================
 
-# sns.set_context('paper', font_scale=1.7)
+sns.set_context('paper', font_scale=1.)
+sns.set_style('white')
+fig = sns.plt.figure(figsize=(3.5,2.625))
 
-sns.set(style="white")
-fig = sns.plt.figure()
 ax1 = sns.plt.axes()
-# fig, ax1 = sns.plt.subplots()
 
 palette = sns.color_palette('deep', 2)
 
@@ -51,18 +50,18 @@ ax2.legend_.remove()
 ax2.set_xlim(0,100)
 sns.plt.setp(ax2, xticks=[])
 
-x1  = [-3.0, -3.0]
-y1  = [0.68, 0.78]
+ax2.text(-3.5, 0.45, 'probability density', rotation='vertical', ha='center', va='center', fontsize='x-small')
+ax2.text(112.0, 0.45, 'CDF', rotation='vertical', ha='center', va='center', fontsize='x-small')
+x1  = [-3.5, -3.5]
+y1  = [0.78, 0.88]
 line1 = lines.Line2D(x1, y1, color='k')
 line1.set_clip_on(False)
-x2  = [107.0, 107.0]
-y2  = [0.57, 0.67]
+x2  = [111.2, 111.2]
+y2  = [0.54, 0.69]
 line2 = lines.Line2D(x2, y2, color='k', linestyle='--')
 line2.set_clip_on(False)
 ax2.add_line(line1)
 ax2.add_line(line2)
-ax2.text(-3., 0.5, 'probability density', rotation='vertical', ha='center', va='center')
-ax2.text(107.5, 0.5, 'CDF', rotation='vertical', ha='center', va='center')
 
-# sns.plt.tight_layout()
-sns.plt.savefig('seqid_dist.png')
+sns.plt.tight_layout()
+sns.plt.savefig('seqid_dist.png', dpi=300)

### 2015-03-18 figures/superposition-src_abl/src/drawcolorbar.py

@@ -1,0 +1,11 @@
+import seaborn as sns
+
+# draw colorbar
+cmap = sns.plt.cm.coolwarm_r
+norm = sns.mpl.colors.Normalize(vmin=0, vmax=100)
+fig = sns.plt.figure(figsize=(1.5,4))
+ax1 = fig.add_axes([0.05, 0.05, 0.3, 0.9])
+cb = sns.mpl.colorbar.ColorbarBase(ax1, cmap=cmap, norm=norm)
+cb.ax.tick_params(labelsize=20)
+cb.set_label('sequence identity (%)', fontsize=27.)
+sns.plt.savefig('colorbar.png', dpi=300)

### 2015-03-18 figures/superposition-src_abl/src/genmodels4.py

@@ -9,21 +9,9 @@ import numpy as np
 import pandas as pd
 import mdtraj as md
 import msmbuilder.cluster
-import seaborn as sns
 import matplotlib as mpl
 
 srcdir = '../../../data/models/SRC_HUMAN_D0'
-
-# draw colorbar
-cmap = sns.plt.cm.coolwarm_r
-norm = sns.mpl.colors.Normalize(vmin=0, vmax=100)
-fig = sns.plt.figure(figsize=(2,8))
-ax1 = fig.add_axes([0.05, 0.05, 0.3, 0.9])
-cb = sns.mpl.colorbar.ColorbarBase(ax1, cmap=cmap, norm=norm)
-cb.ax.tick_params(labelsize=25.)
-cb.set_label('sequence identity (%)', fontsize=35.)
-sns.plt.savefig('colorbar.png')
-
 
 nmodels_per_class = 3
 

### 2015-03-18 manuscript/ms.tex
Updated some figures to pdf format

@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist}
+    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
 
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2}
+    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2.pdf}
 
@@
 
-% \subfloat[Src]{\label{figure:src-distances}%
-%    \includegraphics[width=0.45\textwidth]{residue_pair_distances/src/distances}
-% }
-% \subfloat[Abl1]{\label{figure:abl-distances}%
-%    \includegraphics[width=0.45\textwidth]{residue_pair_distances/abl/distances}
-% }
-
-    \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances}
+    \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances.pdf}
 
@@
 \begin{figure*}[h]
-    \includegraphics[width=0.5\textwidth]{loopmodel_analysis/nmissing_resis_distributions}
+    \includegraphics[width=0.5\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
     \caption{{\bf Distributions for the number of missing residues for templates for which remodeling (with the {\tt loopmodel} command) was either successful or unsuccessful.}

### 2015-03-18 figures/loopmodel_analysis/plot_dists.py

@@ -11,6 +11,7 @@ unsuccessful = df[df.no_missing_residues == False][df.successful == False]
 # ====
 
 sns.set_context('paper', font_scale=1)
+sns.set_style('whitegrid')
 sns.plt.figure(figsize=(3.5, 2.625))
 
 sns.distplot(successful.nmissing_resis, hist=False, kde=True, rug=True, label='successful',
@@ -37,4 +38,5 @@ sns.plt.xlabel('Number of missing residues')
 
 sns.plt.tight_layout()
 
-sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)
+sns.plt.savefig('nmissing_resis_distributions.pdf')
+# sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)

### 2015-03-18 figures/residue_pair_distances/src_abl_combined/plot.py

@@ -96,5 +96,6 @@ cbar.set_alpha=1
 fig.subplots_adjust(hspace=0)
 # sns.plt.tight_layout()
 
-sns.plt.savefig('distances.png', bbox_inches='tight', dpi=300)
+sns.plt.savefig('distances.pdf', bbox_inches='tight')
+# sns.plt.savefig('distances.png', bbox_inches='tight', dpi=300)
 # sns.plt.savefig('distances.png', dpi=300)

### 2015-03-18 figures/rmsddist/rmsddist2.py

@@ -72,7 +72,8 @@ sns.despine(left=True)
 
 plt.tight_layout()
 
-plt.savefig('rmsddist2.png', dpi=300)
+plt.savefig('rmsddist2.pdf')
+# plt.savefig('rmsddist2.png', dpi=300)
 
 plt.close()
 

### 2015-03-18 figures/seqid_dist/seqid_dist.py

@@ -64,4 +64,5 @@ ax2.add_line(line1)
 ax2.add_line(line2)
 
 sns.plt.tight_layout()
-sns.plt.savefig('seqid_dist.png', dpi=300)
+sns.plt.savefig('seqid_dist.pdf')
+# sns.plt.savefig('seqid_dist.png', dpi=300)

### 2015-03-18 manuscript/ms.tex
Now using Oxford comma

@@
 {\bf Ensembler} uses comparative modeling to build models, and as such requires a set of structures to be used as templates.
-The second stage thus entails the selection of templates and storage of associated sequences, structures and identifiers.
+The second stage thus entails the selection of templates and storage of associated sequences, structures, and identifiers.
 These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates from the Protein Data Bank (PDB) or from UniProt.
@@
 {\color{red}[JDC: CITE]}
-For example, mutations of Src are associated with colon, breast and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl1 cancer involvement]}.
+For example, mutations of Src are associated with colon, breast, and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl1 cancer involvement]}.
 Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
@@
 One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
-These three residues are also well conserved [CITE Kannan Neuwald JMB 2005] and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
+These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
 As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).

### 2015-03-18 manuscript/ms.tex
Added Availability and Future Directions section

@@
 The second stage thus entails the selection of templates and storage of associated sequences, structures, and identifiers.
-These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates from the Protein Data Bank (PDB) or from UniProt.
+These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates based on a search of the Protein Data Bank (PDB) or UniProt.
 A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing some degree of homology between targets and templates.
@@
 The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
+Only structures solved by X-ray crystallography or NMR are selected, thus excluding computer-generated models available from the PDB.
 If the {\tt -{}-domains} flag is used, then templates are truncated at the start and end of the domain sequence.
@@
 The distributions are plotted in Fig. S\ref{si:loopmodel-nmissing-residues}.
-% TODO Work out if there is an easy way to reference SI figures
 {\color{red}[JDC: Can you give some statistics on the distribution of loop lengths modeled?  Why did loop modeling fail in the cases it did?  Anything else you can say here beyond this one sentence?]}
@@
 
+\subsubsection*{Availability}
+
 The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}.
@@
 
-% Limitations:
-% counterions
-% PTMs
-% long loops
+\subsubsection*{Future Directions}
 
@@
 It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
+
+Comparative protein modeling and MD simulation set-up can be approached in a number of different ways, with varying degrees of complexity, and there are a number of obvious additions and improvements which we plan to implement in future versions of {\bf Ensembler}.
+
+Some amino acids can exist in different protonation states, depending on pH and on their local environment.
+These protonation states can have important effects on biological processes.
+For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism[CITE: Abl1 DFG flip evidence]---is controlled by protonation of the aspartate [CITE: Shan Shaw Proton-dependent switch Abl1 PNAS 2009].
+Currently, protonation states are assigned simply based on pH (a user-controllable parameter).
+At neutral pH, histidines have two protonation states which are approximately equally likely, and in this situation the selection is therefore made based on which state results in a better hydrogen bond.
+It would be highly desirable to instead use a method which assigns amino acid protonation states based on a rigorous assessment of the local environment.
+We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.
+
+Many proteins require the presence of various types of non-protein atoms and molecules for proper function, such as metal ions (e.g.~Mg$^{+2}$), cofactors (e.g.~ATP) or post-translational modifications (e.g.~phosphorylation, methylation, glycosylation, etc.), and we thus plan for {\bf Ensembler} to eventually have the capability to include such entities in the generated models.
+Binding sites for metal ions are frequently found in proteins, often playing a role in catalysis.
+For example, protein kinase domains contain two binding sites for divalent metal cations, and display significantly increased activity in the presence of Mg$^{2+}$~[CITE: Adams Taylor Protein Sci 1993], the divalent cation with highest concentration in mammalian cells.
+Metal ions are often not resolved in experimental structures of proteins, but by taking into account the full range of available structural data, it should be possible in many cases to include metal ions based on the structures of homologous proteins.
+We are careful to point out, however, that metal ion parameters in classical MD force fields have significant limitations, particularly in their interactions with proteins~[CITE: Sousa Ramos chapter 11 of Kinetics and Dynamics: From Nano- to Bio-Scale, Springer, 2010].
+Cofactors and post-translational modifications are also often not fully resolved in experimental structures, and endogenous cofactors are frequently substituted with other molecules to facilitate experimental structural analysis.
+Again, {\bf Ensembler} could exploit structural data from a set of homologous proteins to model in these molecules, although there will be likely be a number of challenges to overcome in the design and implementation of such functionality.
+
+Another limitation with the present version of {\bf Ensembler} involves the treatment of members of a protein family with especially long residue insertions or deletions.
+For example, the set of all human protein kinase domains listed in UniProt have a median length of 265 residues and a standard deviation of 45, yet the minimum and maximum lengths are 102 and 801 respectively.
+The latter value corresponds to the protein kinase domain of serine/threonine-kinase \emph{greatwall}, which includes a long insertion between the two main lobes of the catalytic domain.
+In principle, such insertions could be excluded from the generated models, though a number of questions would arise as to how best to approach this.
+
+\subsubsection*{Conclusion}
+
+We believe {\bf Ensembler} to be an important first step toward enabling computational modeling and simulation of proteins on the scale of entire protein families, and suggest that it could likely prove useful for tasks beyond its original aim of providing diverse starting configurations for MD simulations.
+The code is open source and has been developed with extensibility in mind, in order to facilitate its customization for a wide range of potential uses by the wider scientific community.
+
+% Limitations:
+% protonation states
+% structural ions, e.g. Mg2+
+% prosthetic groups, e.g. heme
+% cofactors, e.g. ATP
+% PTMs - phosphorylation, methylation, glycosylation
+% long loops
 

### 2015-03-19 manuscript/ms.tex
Some edits to Abstract and Introduction.

@@
 
-% \newcommand{\pyitc}{\url{http://www.simtk.org/home/bayesian-itc}} % URL of pyITC project homepage
-
 %% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
-While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable simulations at this scale have not kept pace.
+While computational infrastructure has grown rapidly, simulations on an \emph{omics} scale are not yet widespread, primarily because software infrastructure to enable simulations at this scale has not kept pace.
 It should now be possible to study protein dynamics across entire (super)families, exploiting both available structural biology data and conformational similarities across homologous proteins.
@@
 {\bf Ensembler} thus automates much of the time-consuming process of preparing protein models suitable for simulation, while allowing scalability up to entire superfamilies.
-{\color{red}[JDC: Prior sentence is redundant?]}
-A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
+A particular advantage of this approach can be found in the construction of kinetic models of conformational dynamics---such as Markov state models (MSMs)---which benefit from a diverse array of initial configurations that span the accessible conformational states to aid sampling.
 We demonstrate the power of this approach by constructing models for all catalytic domains in the human tyrosine kinase family, using all available kinase catalytic domain structures from any organism as structural templates.
@@
 Many functional properties---such as the ability to bind small molecules or interact with signaling partners---require transitions between states, encompassing anything from reorganization of sidechains at binding interfaces to domain motions to large scale folding-unfolding events.
-Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state~\cite{craik:2009:science:trapping-moving-targets}.
+Drug discovery could also benefit from a more extensive consideration of protein dynamics, whereby small molecules might be selected based on their predicted ability to bind and trap a protein target in an inactive state~\cite{craik:science:2009:trapping-moving-targets}.
 
 Molecular dynamics (MD) simulations have the capability, in principle, to describe the time evolution of a protein in atomistic detail, and have proven themselves to be a useful tool in the study of protein dynamics.
-A number of mature software packages and forcefields are available, and much recent progress has been driven by advances in computing architecture.
-For example, many MD packages are now able to exploit GPUs, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE] allow scalability on an unprecedented level.
-In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity [CITE MSM reviews].
+A number of mature software packages and forcefields are now available, and much recent progress has been driven by advances in computing architecture.
+For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE] allow scalability on an unprecedented level.
+In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2010:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
 MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
@@
 The authors are grateful to Kyle A.~Beauchamp (MSKCC), Robert McGibbon (Stanford), Arien S. Rustenburg (MSKCC) for many excellent software engineering suggestions.
-The authors thank Sonya M.~Hanson (MSKCC), Nicholas M.~Levinson (?), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
-The authors are grateful to Benjamin Webb and Andrej Sali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
+The authors thank Sonya M.~Hanson (MSKCC), Nicholas M.~Levinson (University of Minnesota), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
+The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.

### 2015-03-19 manuscript/ms.tex
Finished edits to Introduction.

@@
 However, it remains difficult for researchers to exploit the full variety of available protein sequence and structural data in simulation studies, largely due to limitations in software architecture.
-For example, the set up of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure, addition of missing residues and atoms, solvation with explicit water and salt buffer, choice of simulation parameters, and system relaxation with energy minimization and one or more short MD simulations.
-For this reason, simulation studies typically consider only one or a few proteins and starting configurations.
+For example, the set up of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure(s), addition of missing residues and atoms, solvation with explicit water and counterions (and potentially buffer components and cosolvents), choice of simulation parameters (or parameterization schemes for components where parameters do not yet exist),  system relaxation with energy minimization, and one or more short preparatory MD simulations to equilibrate the system and relax the simulation cell.
+Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations.
+Worse still, studies (or collections of studies) that \emph{do} consider multiple proteins often suffer from the lack of consistent best practices in this preparation process, making comparisons between related proteins unnecessarily difficult.
 
-The ability to fully exploit the large base of available protein sequence and structural data in biomolecular simulation studies could open up many interesting avenues for research, enabling the study of entire protein families or superfamilies across multiple organisms.
+The ability to fully exploit the large quantity of available protein sequence and structural data in biomolecular simulation studies could open up many interesting avenues for research, enabling the study of entire protein families or superfamilies within a single organism or across multiple organisms.
 The similarity between members of a given protein family could be exploited to generate arrays of conformational models, which could be used as starting configurations to aid sampling in MD simulations.
@@
 By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
+
 We anticipate that the tool will prove to be useful in a number of other ways.
@@
 Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
+{\color{red}[JDC: Can we also add the URL of where to get the code and TK models here?]}
 
@@
 
-{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API.
+{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
 
@@
 The various stages of this pipeline are described in detail below.
-
-{\color{red}[JDC: We could really help the reader if we preface each section here with a bit of an introduction of what we're trying to accomplish in each stage.
-Otherwise, I worry that each section is a long list of things we do without reference to an overall concept of what the stage is trying to accomplish or why certain decisions were made.]}
-{\color{blue}[DLP: Good point. I've added in brief introductions for each section.]}
-{\color{red}[JDC: Can you do a bit more here?
-I feel that the reader may need more orientation.
-You've essentially just added a sentence or two at the beginning of each stage that doesn't really enlighten the user as to your terminology (\emph{tempalte} and \emph{target}), your motivation for \emph{why} things are done each stage, or \emph{how} the user is to select among the various options available.]}
 
@@
 
-\subsubsection*{Target selection}
+\subsubsection*{Target selection and retrieval}
 
-The first stage entails the selection of a set of target protein sequences, i.e. the sequences the user is interested in modeling.
-{\color{red}[JDC: Maybe explain what is meant by ``target protein sequences''?  These are the sequences the user is interested in modeling.]}
-{\color{blue}[DLP: Addressed.]}
+The first stage entails the selection of a set of \emph{target} protein sequences---the sequences the user is interested in generating simulation-ready structural models for.
+This may be a single sequence---such as a full-length protein or a construct representing a single domain---or a collection of sequences, such as a particular domain from an entire family of proteins.
 
-These targets can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
-The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Isn't there a real citation for UniProt?]}, using the subcommand {\tt gather\_targets}.
-The user specifies a query string with the {\tt -{}-query} flag, which conforms to the same syntax as the search function available on the UniProt website.
+Target sequences can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
+The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Isn't there a real citation for UniProt?]}---via a UniProt search query.
+To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} us used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
 For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
@@
 
-\subsubsection*{Template selection}
+\subsubsection*{Template selection and retrieval}
 
@@
 
-\subsubsection*{Modeling}
+\subsubsection*{Modeling of targets}
 
@@
 
-\subsubsection*{Refinement}
+\subsubsection*{Refinement of models}
 
@@
 {\color{blue}[DLP: We currently just filter out models which throw exceptions or NaNs.]}
+
+\subsubsection*{Solvation and NPT equilibration}
 

### 2015-03-19 manuscript/ms.tex
More edits to Design and Implementation.

@@
   In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to exemplify the resources required per model and for modeling the entire set of tyrosine kinases.
-  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread---parallel execution can reduce wall clock time nearly linearly.
+  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread and \emph{GPU-h} on a single GPU---parallel execution can reduce wall clock time nearly linearly.
   }
@@
 This may be a single sequence---such as a full-length protein or a construct representing a single domain---or a collection of sequences, such as a particular domain from an entire family of proteins.
+The output of this stage is a FASTA-formatted text file containing the desired target sequences with corresponding arbitrary identifiers.
 
-Target sequences can be defined manually, simply by providing a FASTA-formatted text file containing the desired target sequences with arbitrary identifiers.
-The {\tt ensembler} command-line tool also allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Isn't there a real citation for UniProt?]}---via a UniProt search query.
+The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Cite \href{http://www.uniprot.org/help/publications}{one of these UniProt citations?}]}---via a UniProt search query.
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} us used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
 For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
-In this way, the user may select a single protein, many proteins, or an entire superfamily.
+In this way, the user may select a single protein, many proteins, or an entire superfamily from UniProt.
 The program outputs a FASTA file, setting the UniProt mnemonic (e.g. {\tt SRC\_HUMAN}) as the identifier for each target protein.
@@
 To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
-In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest.
-Example identifiers: {\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}.
-{\color{red}[JDC: Does it make sense to set some of these coded examples off on their own lines?]}
+In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~{\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}).
+
+Target sequences can also be defined manually (or from another program) by providing a FASTA-formatted text file containing the desired target sequences with corresponding arbitrary identifiers.
 
@@
 A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing some degree of homology between targets and templates.
-{\color{red}[JDC: Again, can you provide more information about \emph{why} this is being done? What the motivation is, and how the user might expect to select these?]}
-{\color{blue}[DLP: Addressed]}
 
-Manual selection of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
-The structure residues must also match those in the sequence file.
-
-The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\href{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
+The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\url{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
 Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.
@@
 If the {\tt -{}-domains} flag is used, then templates are truncated at the start and end of the domain sequence.
+
+Templates can also be defined manually.
+Manual selection of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
+The structure residues must also match those in the sequence file.
+
+\subsubsection*{Template refinement}
 
@@
 Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
-The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD (for C$\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~{\color{red}[Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
+The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~{\color{red}[Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
 A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.

### 2015-03-19 manuscript/ms.tex

@@
 
-Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm {\color{red}[CITE]} provided via the {\tt loopmodel} tool of the Rosetta software suite (CITE: Rosetta and/or loopmodel).
-Because fewer loops need to be built during the subsequent model-building stage, prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
+Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm {\color{red}[CITE]} provided via the {\tt loopmodel} tool of the Rosetta software suite {\color{red}[CITE: Rosetta and/or loopmodel]}.
+Because fewer loops need to be built during the subsequent model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
+{\color{red}[JDC: Can you show the distribution of missing loop lengths for the TKs?]}
+
 Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
@@
 {\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home (CITE: F@H).
+{\color{red}[JDC: Is there a way we can make this more generally useful to others?
+For example, is there a different system they might want to use, such as \href{http://copernicus.readthedocs.org/en/latest/tutorials/md-tutorial.html#creating-a-workflow-for-molecular-simulations}{Copernicus}?]}
 
@@
 
-{\color{red}[JDC: It would be useful to have some subheadings in this section to give it some internal organization.]}
-
 \subsubsection*{Modeling of all human tyrosine kinase catalytic domains}
@@
 The number of missing residues in each template ranged from 0 to 102, with a median of 11 and a standard deviation of 13.
+{\color{red}[JDC: Any chance you can generate a plot of the distribution of loop lengths?  I'm guessing this is pretty non-normal since the standard deviation is larger than the median!]}
 Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled, with most remodeling failures attributable to spatial constraints imposed by the original template structure.
@@
 Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
+{\color{red}[JDC: What percentage is this?]}
 To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
@@
 Each model generated about 513 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 1.7 GB per TK target or 149 GB for all 90 TKs.
-The data generated per model breaks down as 436 kB for the output from the modeling stage---with the largest contribution arising from the Modeller restraint files---and 77 kB for the implicit solvent MD refinement stage.
+The data generated per model breaks down as 436 kB for the output from the modeling stage---with the largest contribution arising from the stored Modeller restraint files---and 77 kB for the implicit solvent MD refinement stage.
+{\color{red}[JDC: Maybe we want to add a flag to make retaining the Modeller restraint files optional?
+I had originally just saved these so we could do subsequent OpenMM-based model refinement if desired, but we don't actually do that yet.]}
 
-\subsubsection*{Evaluation of model quality}
+\subsubsection*{Evaluation of model quality and utility}
 
@@
 Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
+{\color{red}[JDC: Explain how $k$-medoids clustering was done either here or in figure caption.]}
 Each model is colored and given a transparency based on the sequence identity between the target and template sequence.
@@
 The Abl1 renderings indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
-While such models are not necessarily incorrect or undersirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
+While such models are not necessarily incorrect or undesirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
 Lower sequence identity models (in transparent white or red) indicate much greater variation in all parts of the structure.
@@
 
-To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
+To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
 We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
@@
 
-The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}.
+The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub,\\
+\url{http://github.com/choderalab/ensembler} \\
+The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}:
 \shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
-Up to date instructions can be found at \url{https://github.com/choderalab/ensembler}.
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
-The latest source can be downloaded from the above GitHub repository, which also contains instructions for building and installing the code.
+The latest source can be downloaded from the GitHub repository, which also contains up-to-date instructions for building and installing the code.
+{\color{red}[JDC: Where does documentation reside?  And example inputs for generating the results in this paper?]}
 
@@
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
+{\color{red}[Add PBG support statement.]}
 

### 2015-03-21 manuscript/ms.tex
Various updates

@@
 A number of mature software packages and forcefields are now available, and much recent progress has been driven by advances in computing architecture.
-For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home [CITE], GPUGrid [CITE], and Copernicus [CITE] allow scalability on an unprecedented level.
+For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, GPUGrid~\cite{buch:2010:gpugrid}, and Copernicus~\cite{pronk:2011:copernicus} allow scalability on an unprecedented level.
 In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2010:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
 MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
-There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates [CITE Modeller and Rosetta and a recent homology modeling review].
+There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates~\cite{moult:proteins:2014:casp10,baker:science:2001:structure-prediction-review}.
 
@@
 In this way, the user may select a single protein, many proteins, or an entire superfamily from UniProt.
-The program outputs a FASTA file, setting the UniProt mnemonic (e.g. {\tt SRC\_HUMAN}) as the identifier for each target protein.
+The program outputs a FASTA file, setting the UniProt mnemonic (e.g.~{\tt SRC\_HUMAN}) as the identifier for each target protein.
 
@@
 The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\url{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
-Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.
-for crystal unit cells with multiple asymmetric units) would thus give rise to multiple template structures.
+Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.~for crystal unit cells with multiple asymmetric units) would thus give rise to multiple template structures.
 
-Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g. {\tt -{}-query 2H8H,1Y57}.
+Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g.~{\tt -{}-query 2H8H,1Y57}.
 Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
@@
 
-Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm {\color{red}[CITE]} provided via the {\tt loopmodel} tool of the Rosetta software suite {\color{red}[CITE: Rosetta and/or loopmodel]}.
+Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
 Because fewer loops need to be built during the subsequent model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
@@
 
-\subsubsection*{Modeling of targets}
+\subsubsection*{Modeling}
 
@@
 
-Modeling is performed with the Modeller automodel function {\color{red}[CITE: Modeller]}, which implements comparative structure modeling by satisfaction of spatial restraints {\color{red}[CITE: Sali Blundell J Mol Biol 1993; Fiser Sali Prot Sci 9 2000]}.
+Modeling is performed with the automodel function of the Modeller software package, which implements comparative structure modeling by satisfaction of spatial restraints~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
 While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module [CITE: BioPython]---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
@@
 To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
+The restraints used by Modeller could potentially be used in alternative refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
+This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 KB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
 
@@
 {\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home (CITE: F@H).
+The module could be easily extended to add methods for preparing models for use with other software, such as the Copernicus platform for running automated, distributed MD simulations.
 {\color{red}[JDC: Is there a way we can make this more generally useful to others?
 For example, is there a different system they might want to use, such as \href{http://copernicus.readthedocs.org/en/latest/tutorials/md-tutorial.html#creating-a-workflow-for-molecular-simulations}{Copernicus}?]}
+{\color{blue}[DLP: addressed]}
 
@@
 
-Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} surviving models.
-{\color{red}[JDC: What percentage is this?]}
+Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} (94\%) surviving models.
 To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
@@
 
-Each model generated about 513 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 1.7 GB per TK target or 149 GB for all 90 TKs.
-The data generated per model breaks down as 436 kB for the output from the modeling stage---with the largest contribution arising from the stored Modeller restraint files---and 77 kB for the implicit solvent MD refinement stage.
+Each model generated about 116 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 0.5 GB per TK target or 41 GB for all 90 TKs.
+The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints to file) and 77 kB for the implicit solvent MD refinement stage.
 {\color{red}[JDC: Maybe we want to add a flag to make retaining the Modeller restraint files optional?
 I had originally just saved these so we could do subsequent OpenMM-based model refinement if desired, but we don't actually do that yet.]}
+{\color{blue}[DLP: I've added this flag and discussed it in the Design and Implementation section, and also updated the file storage details here.]}
 
@@
 {\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
-For example, we don't yet handle counterions (e.g.
-structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
+For example, we don't yet handle counterions (e.g.~structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
 We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
@@
     \includegraphics[width=0.5\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
-    \caption{{\bf Distributions for the number of missing residues for templates for which remodeling (with the {\tt loopmodel} command) was either successful or unsuccessful.}
-    The plotted distributions are smoothed using kernel density estimation, and the raw data points are shown as a rug plot.
+    \caption{{\bf Distributions for the number of missing residues in the TK templates.}
+    The upper plot shows the distribution of the total number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling (with the {\tt loopmodel} subcommand) failed (red).
+The raw data points for all templates are shown as a rug plot.
+    The lower plot shows the distribution of the number of residues per template gap, normalized and smoothed using kernel density estimation. The raw data points are shown as a rug plot.
     }

### 2015-03-21 figures/loopmodel_analysis/plot_dists.py

@@ -1,26 +1,45 @@
+import numpy as np
 import pandas as pd
 import seaborn as sns
+from ast import literal_eval
 
-df = pd.read_csv('loopmodel_data.csv')
+df = pd.read_csv('loopmodel_data.csv', converters={'missing_resi_spans': literal_eval})
 
 successful = df[df.no_missing_residues == False][df.successful == True]
 unsuccessful = df[df.no_missing_residues == False][df.successful == False]
+
+missing_resis_spans = np.array([x for sublist in df.missing_resi_spans for x in sublist])
 
 # ====
 # plot
 # ====
 
+palette = sns.color_palette()
+
 sns.set_context('paper', font_scale=1)
 sns.set_style('whitegrid')
-sns.plt.figure(figsize=(3.5, 2.625))
+figsize=(3.5, 4)
+# sns.plt.figure(figsize=figsize)
 
-sns.distplot(successful.nmissing_resis, hist=False, kde=True, rug=True, label='successful',
-    rug_kws={'height': 0.002},
-)
-sns.distplot(unsuccessful.nmissing_resis, hist=False, kde=True, rug=True, label='unsuccessful', color='red',
-    # kde_kws={'bw': 2.5},
-    rug_kws={'height': 0.001},
-)
+fig, (ax1, ax2) = sns.plt.subplots(2, 1, sharex=True, figsize=figsize)
+
+bins = np.linspace(0, df.nmissing_resis.max(), 10)
+# sns.distplot(df.nmissing_resis, kde=False, bins=bins, label='total')
+# sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins, label='successful')
+# same with log y-axis:
+sns.distplot(df.nmissing_resis, kde=False, bins=bins, label='all templates', hist_kws={'log': True}, rug=True, rug_kws={'height': 0.7, 'lw': 0.5}, ax=ax1)
+sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins, label='failed remodeling', hist_kws={'log': True}, color=palette[2], ax=ax1)
+# sns.plt.ylim(0.8,10000)
+
+ax1.legend()
+
+# sns.distplot(successful.nmissing_resis, hist=False, kde=True, rug=True, label='successful',
+#     rug_kws={'height': 0.002},
+# )
[…]

### 2015-03-21 manuscript/ms.tex
Some edits to figures and references.

@@
 This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-By default, the Amber99SB-ILDN force field is used {\color{red}[CITE: amber99sbildn refs]} with a modified generalized Born solvent model (GBSA-OBC) (CITE: GBSA-OBC).
+By default, the Amber99SB-ILDN force field~\cite{amber99sb-ildn} is used with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
 The {\bf Ensembler} API allows the use of any of the other force fields implemented in OpenMM.
@@
 It would be highly desirable to instead use a method which assigns amino acid protonation states based on a rigorous assessment of the local environment.
-We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.
+We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{alexov-gunner:biophys-j:1997:mcce2,georgescu-alexov-gunner:biophys-j:2002:mcce2,song-mao-gunner:j-comput-chem:2009:mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.
+{\color{red}[JDC: I think we may want to consider doing that at this stage.  Let's discuss.]}
 
@@
 
-%\bibliographystyle{prsty} 
+\bibliographystyle{prsty} 
 \bibliography{ms.bib}
@@
 Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
+{\color{red}[JDC: Very cool!  Did you write a script to generate this, or did you have to do this by hand?]}
 
@@
     The lower plot shows the distribution of the number of residues per template gap, normalized and smoothed using kernel density estimation. The raw data points are shown as a rug plot.
+    {\color{red}[JDC: Some ideas for cleaning this up: Either the tick marks are being misrendered in that they are not taller if there are multiple data points with the same number or the data is really funky, since I would expect there to be a few examples in some bins.
+Also, is there a big drawback to making the top histogram bin size unity, since the values are integral?
+I don't think transparency is needed for the histogram bars either.
+I would also make the x-axes for the top and bottom plots different, since the data ranges are different.
+Finally, I'd see if a histogram with unit bin size might be more appropriate for the bottom plot as well---the KDE just doesn't feel right for this kind of data, since we are trying to report exact statistics from a specific example rather than estimate a general density for problems of this sort.
+Finally, I like ``remodeled loop length'' or ``missing loop length'' much better than ``Number of missing residues per template gap'', which seems unnecessarily verbose.]}
 }

### 2015-03-22 manuscript/ms.tex
Added some more references.

@@
 Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
-The mdtraj {\color{red}[CITE: mdtraj]} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~{\color{red}[Cite Theobald QCP papers]} implementation, and the leader algorithm is then used to populate clusters.
+The mdtraj~\cite{mdtraj} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation, and the leader algorithm is then used to populate clusters.
 A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
@@
 
-Models are first subjected to energy minimization (using the L-BFGS algorithm {\color{red}[CITE]}), followed by a short molecular dynamics (MD) simulation with an implicit solvent representation.
-This is implemented using the OpenMM molecular simulation toolkit (link and CITE: OpenMM), chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
+Models are first subjected to energy minimization (using the L-BFGS algorithm~\cite{lbfgs}, followed by a short molecular dynamics (MD) simulation with an implicit solvent representation.
+This is implemented using the OpenMM molecular simulation toolkit~\cite{eastman:jctc:2012:openmm}, chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
 By default, the Amber99SB-ILDN force field~\cite{amber99sb-ildn} is used with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
@@
 Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
-The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field and TIP3P water {\color{red}[JDC: CITE]} by default.
-Other force fields or water models such as TIP4P-Ew {\color{red}[CITE]}) can be specified via the {\bf Ensembler} API.
+The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field~\cite{amber99sb-ildn} and TIP3P water~\cite{tip3p} by default.
+Other force fields or water models such as TIP4P-Ew~\cite{tip4p-ew} can be specified via the {\bf Ensembler} API.
 {\color{red}[JDC: We should allow other water models in OpenMM too, such as TIP4P-Ew?]}

### 2015-03-22 manuscript/ms.tex
Moved loop length figure into main body text.

@@
 Modeling is performed with the automodel function of the Modeller software package, which implements comparative structure modeling by satisfaction of spatial restraints~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
-While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module [CITE: BioPython]---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~{\color{red}[CITE: Gaston Gonnet Science 1992]}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
+While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module [CITE: BioPython]---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
+{\color{red}[JDC: What evidence can we present or cite to support this claim?]}
 Models are output as PDB-format coordinate files.
@@
 
-Unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
-The number of missing residues in each template ranged from 0 to 102, with a median of 11 and a standard deviation of 13.
-{\color{red}[JDC: Any chance you can generate a plot of the distribution of loop lengths?  I'm guessing this is pretty non-normal since the standard deviation is larger than the median!]}
-Out of 3666 templates with one or more missing residues, 3134 were successfully remodeled, with most remodeling failures attributable to spatial constraints imposed by the original template structure.
-There was some correlation between remodeling failures and the number of missing residues; templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
-The distributions are plotted in Fig. S\ref{si:loopmodel-nmissing-residues}.
-{\color{red}[JDC: Can you give some statistics on the distribution of loop lengths modeled?  Why did loop modeling fail in the cases it did?  Anything else you can say here beyond this one sentence?]}
-{\color{blue}[DLP: Addressed in the text, and a SI figure.]}
+\begin{figure}[tbp]
+    \includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
+    \caption{{\bf Distributions for the number of missing residues in the TK templates.}
+    The upper plot shows the distribution of the total number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling (with the {\tt loopmodel} subcommand) failed (red).
+The raw data points for all templates are shown as a rug plot.
+    The lower plot shows the distribution of the number of residues per template gap, normalized and smoothed using kernel density estimation. The raw data points are shown as a rug plot.
+    {\color{red}[JDC: Some ideas for cleaning this up: Either the tick marks are being misrendered in that they are not taller if there are multiple data points with the same number or the data is really funky, since I would expect there to be a few examples in some bins.
+Also, is there a big drawback to making the top histogram bin size unity, since the values are integral?
+I don't think transparency is needed for the histogram bars either.
+We can also ditch the semilogy axis.
+I would also make the x-axes for the top and bottom plots different, since the data ranges are different.
+I'd see if a histogram with unit bin size might be more appropriate for the bottom plot as well---the KDE just doesn't feel right for this kind of data, since we are trying to report exact statistics from a specific example rather than estimate a general density for problems of this sort.
+Finally, I like ``remodeled loop length'' or ``missing loop length'' much better than ``Number of missing residues per template gap'', which seems unnecessarily verbose.]}
+}
+    \label{figure:loopmodel-nmissing-residues}
+\end{figure}
+
+Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
+To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
+Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination with out error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
+There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
 
@@
 The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints to file) and 77 kB for the implicit solvent MD refinement stage.
-{\color{red}[JDC: Maybe we want to add a flag to make retaining the Modeller restraint files optional?
-I had originally just saved these so we could do subsequent OpenMM-based model refinement if desired, but we don't actually do that yet.]}
-{\color{blue}[DLP: I've added this flag and discussed it in the Design and Implementation section, and also updated the file storage details here.]}
+{\color{red}[JDC: Can you rework this paragraph to specify total space requirements \emph{without} saving MODELLER restraint files (since this is now the default), plus how much each restraint file adds if they are retained as well?
+I took a stab at it, but couldn't figure out how the numbers worked out.]}
 
@@
 
-\begin{figure*}[h]
-    \includegraphics[width=0.5\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
-    \caption{{\bf Distributions for the number of missing residues in the TK templates.}
-    The upper plot shows the distribution of the total number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling (with the {\tt loopmodel} subcommand) failed (red).
-The raw data points for all templates are shown as a rug plot.
-    The lower plot shows the distribution of the number of residues per template gap, normalized and smoothed using kernel density estimation. The raw data points are shown as a rug plot.
-    {\color{red}[JDC: Some ideas for cleaning this up: Either the tick marks are being misrendered in that they are not taller if there are multiple data points with the same number or the data is really funky, since I would expect there to be a few examples in some bins.
-Also, is there a big drawback to making the top histogram bin size unity, since the values are integral?
-I don't think transparency is needed for the histogram bars either.
-I would also make the x-axes for the top and bottom plots different, since the data ranges are different.
-Finally, I'd see if a histogram with unit bin size might be more appropriate for the bottom plot as well---the KDE just doesn't feel right for this kind of data, since we are trying to report exact statistics from a specific example rather than estimate a general density for problems of this sort.
-Finally, I like ``remodeled loop length'' or ``missing loop length'' much better than ``Number of missing residues per template gap'', which seems unnecessarily verbose.]}
-}
-    \label{si:loopmodel-nmissing-residues}
-\end{figure*}
-
 \end{document}

### 2015-03-22 manuscript/ms.tex
Reworking some figure captions.

@@
 As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TK) domains listed in UniProt.
+{\color{red}[JDC: What query was used to yield this set?]}
 {\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
@@
 Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
+{\color{red}[JDC: Can you add statistics (median, stddev, max) for loop lengths too?]}
 To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
@@
 
-The distribution of RMSDs of the final models (relative to the highest sequence identity model for a given target) is shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
-The distributions are stratified based on the sequence identity between target and template, indicating that higher sequence identity templates result in models with lower RMSDs.
-The sequence identity stratifications were selected based on the sequence identity distribution plotted in Fig.~\ref{figure:sequence-identity-distribution}, which suggests an intuitive division into three categories, with \num{307753} models in the 0-35\% sequence identity range, \num{69922} models in the 35-55\% range, and \num{4893} models in the 55-100\% range.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: RMSD distribution by sequence identity
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[tbp]
+    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2.pdf}
+
+    \caption{{\bf Distribution of RMSDs to all tyrosine kinase catalytic domain models relative to model derived from highest sequence identity template.}    
+    Distributions are averaged over all 90 tyrosine kinase catalytic domains.
+    {\color{red}[JDC: Is this accurate?]}
+    To better illustrate how conformational similarity depends on sequence identity, these were separated into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
+    The plotted distributions have been smoothed using kernel density estimation.
+    {\color{red}[JDC: Are these computed for the templates or the resulting models?]}
+    {\color{red}[JDC: Can we also show the overall distribution without stratification (e.g.~in grey in a separate panel)?]}
+  }
+  \label{figure:rmsd-distribution-by-sequence-identity}
+\end{figure}
+
+To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
+To better understand the influence of sequence identity on the conformational similarities of resulting models, the sequence identities were stratified based on the sequence identity distribution plotted in Fig.~\ref{figure:sequence-identity-distribution}, which suggests an intuitive division into three categories, with \num{307753} models in the 0--35\% sequence identity range, \num{69922} models in the 35--55\% range, and \num{4893} models in the 55--100\% range.
+It is clear that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average.
 
@@
 In terms of structural data, a large number of crystal structures have been solved (with or without ligands such as nucleotide substrate or inhibitor drugs), showing the kinases in a number of different conformations.
-These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src {\color{blue}[CITE:Shukla Pande Nat Commun 2014]}.
-
+These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}.
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
@@
 High sequence identity models (in opaque blue) tend to be quite structurally similar, with some variation in loops or changes in domain orientation.
+
 The Abl1 renderings indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
@@
 One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
-These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family [CITE Foda Shan Seeliger Src Nat Commun 2015; Shukla Pande Nat Commun 2014; Ozkirimli Post Prot Sci 2008].
+These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm} [CITE Foda Shan Seeliger Src Nat Commun 2015; Ozkirimli Post Prot Sci 2008].
 As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
@@
 
-    \caption{{\bf Sequence identity distribution for human TK models.}
-    Distribution of sequence identities for all \num{373513} models generated for the human tyrosine kinases.
-    Sequence identities are calculated from pairwise target-template alignments.
-    The cumulative distribution function is shown by the dashed line.
-    The plotted distributions have been smoothed using kernel density estimation.
-    % {\color{red}[JDC: Font size too small.]}
+    \caption{{\bf Template-target sequence identity distribution for human tyrosine kinase catalytic domains.}
+    Sequence identities are calculated from all pairwise target-template alignments, where targets are human kinase catalytic domain sequences and templates are all kinase catalytic domains from any organism with structures in the PDB, as described in the text.
+    A kernel density estimate of the target-template sequence identity probability density function is shown as a solid line with shaded region, while the corresponding cumulative distribution function is shown as a dashed line.
     }
   \label{figure:sequence-identity-distribution}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: RMSD distribution by sequence identity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2.pdf}
-
-    \caption{{\bf RMSD distribution by sequence identity.}
-    RMSD distributions for all \num{373513} human TK models, divided into three sequence identity ranges.
-    For a given target, model RMSDs are calculated relative to the highest sequence identity model for that target.
-    The plotted distributions have been smoothed using kernel density estimation.
-    % {\color{red}[JDC: Font size too small.]}
-  }
-  \label{figure:rmsd-distribution-by-sequence-identity}
 \end{figure}
@@
 \url{http://github.com/choderalab/ensembler} \\
-The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python~\cite{conda}:
-\shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
+The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python [\url{http://conda.pydata.org}]:
+%\shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
+\shellcmd{conda config --add channels https://conda.binstar.org/omnia}
+\shellcmd{conda ensembler}
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.

### 2015-03-22 manuscript/ms.tex
More edits to Design and Implementation.

@@
 Templates can also be defined manually.
-Manual selection of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
+Manual specification of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
 The structure residues must also match those in the sequence file.
@@
 
-Unresolved template residues can optionally be remodeled with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
-Because fewer loops need to be built during the subsequent model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
-{\color{red}[JDC: Can you show the distribution of missing loop lengths for the TKs?]}
+Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
+Because fewer loops need to be built during the subsequent target model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
+{\color{red}[JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]}
 
@@
 
-This stage entails the generation of models via comparative modeling of each target sequence onto each template structure.
-Non-unique models are subsequently filtered out using a RMSD-based clustering scheme.
+In the modeling stage, structural models of the target sequence are generated from the template structures, with the goal of modeling the target in a variety of conformations that could be significantly populated under equilibrium conditions.
 
-Modeling is performed with the automodel function of the Modeller software package, which implements comparative structure modeling by satisfaction of spatial restraints~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
-While Modeller can generate alignments automatically, we utilize the BioPython {\tt pairwise2} module [CITE: BioPython]---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
-{\color{red}[JDC: What evidence can we present or cite to support this claim?]}
+Modeling is performed using the automodel function of the Modeller software package~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller} to rapidly generate a single model of the target sequence from each template structure.
+Modeller uses simulated annealing cycles along with a minimal forcefield and spatial restraints---generally Gaussian interatomic probability densities extracted from the template structure with database-derived statistics determining the distribution width---to rapidly generate candidate structures of the target sequence from the provided template sequence~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
+
+While Modeller's automodel function can generate its own alignments automatically, we utilize the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
+{\color{red}[JDC: Can we give some evidence to support this claim, or soften our language on this?]}
+
 Models are output as PDB-format coordinate files.
-A list of all model identifiers sorted by sequence identity is also written to a text file.
+A list of all model identifiers sorted by sequence identity is also written to a text file, so that users may select models from the desired range of sequence identities.
+{\color{red}[JDC: I wonder if we should add an option ``exclude models with sequence identities poorer than X'', since this could be something people might want to do.]}
 To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
-The restraints used by Modeller could potentially be used in alternative refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
+The restraints used by Modeller could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
 This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 KB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
 
-All chains of template structures that contain the template sequence are utilized in the modeling phase, which can sometimes cause models to be nearly identical.
-Since the goal is to provide good coverage of conformation space, {\bf Ensembler} filters out nearly identical models using structural similarity-based clustering.
-The mdtraj~\cite{mdtraj} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation, and the leader algorithm is then used to populate clusters.
+\subsubsection*{Filtering of nearly identical models}
+
+Because {\bf Ensembler} treats individual chains from source PDB structures as individual templates, a number of models may be generated with very similar structures if these individual chains are nearly identical in conformation.
+For this reason, and also to allow users to select for high diversity if they so choose, {\bf Ensembler} provides a way to filter out models that are very similar in RMSD.
+A fast clustering scheme is used to identify models differing by a user-specified minimum RMSD.
+The mdtraj~\cite{mdtraj} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation, and the leader algorithm {\color{red}[JDC: Citation for leader algorithm?]} is then used to populate clusters.
 A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
@@
 Otherwise I could add a -{}-water\_model flag to the CLI, for example.]}
+{\color{red}[JDC: Would be good to let at least water model and forcefield name be selectable via the CLI!]}
 

### 2015-03-22 manuscript/ms.tex
Revised Fig 1 caption.

@@
 
-  \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline.}
-  The number of viable models surviving each stage of the pipeline are shown, either for all tyrosine kinases (\emph{All TKs}) or representative individual kinases (\emph{SRC} and \emph{ABL}).
-  In addition, the typical timing on a cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to exemplify the resources required per model and for modeling the entire set of tyrosine kinases.
-  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single hyperthread and \emph{GPU-h} on a single GPU---parallel execution can reduce wall clock time nearly linearly.
+  \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline and illustrative statistics for modeling all human tyrosine kinase catalytic domains.}
+  On the left, the various stages of the {\bf Ensembler} pipeline are shown.
+  On the right, the number of viable models surviving each stage of the pipeline is shown for modeling all 90 tyrosine kinases (\emph{All TKs}) and representative individual tyrosine kinases (\emph{SRC} and \emph{ABL}).
+  Typical timings on a computer cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to illustrate resource requirements per model for modeling the entire set of tyrosine kinases.
+  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution via MPI reduces wall clock time nearly linearly.
   }
@@
 Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
-The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) (CITE: Velankar Nucleic Acids Res 2013), which provides residue-level mappings between PDB and UniProt entries.
+The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) {\color{red}(CITE: Velankar Nucleic Acids Res 2013)}, which provides residue-level mappings between PDB and UniProt entries.
 The SIFTS data is used to extract template sequences, retaining only residues which are resolved and match the equivalent residue in the UniProt sequence---non-wildtype residues are thus removed from the template structures.

### 2015-03-22 manuscript/ms.tex
For some reason, document was structures as section -> subsubsection. Fixed this so it is section -> subsection -> subsubsection.

@@
 
-\subsubsection*{Target selection and retrieval}
+\subsection*{Target selection and retrieval}
 
@@
 
-\subsubsection*{Template selection and retrieval}
+\subsection*{Template selection and retrieval}
 
@@
 
-\subsubsection*{Template refinement}
+\subsection*{Template refinement}
 
@@
 
-\subsubsection*{Modeling}
+\subsection*{Modeling}
 
@@
 
-\subsubsection*{Filtering of nearly identical models}
+\subsection*{Filtering of nearly identical models}
 
@@
 
-\subsubsection*{Refinement of models}
+\subsection*{Refinement of models}
 
@@
 
-\subsubsection*{Solvation and NPT equilibration}
+\subsection*{Solvation and NPT equilibration}
 
@@
 
-\subsubsection*{Packaging}
+\subsection*{Packaging}
 
-{\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home (CITE: F@H).
+{\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home {\color{red}[CITE: F@H]}.
 The module could be easily extended to add methods for preparing models for use with other software, such as the Copernicus platform for running automated, distributed MD simulations.
@@
 {\color{blue}[DLP: addressed]}
+{\color{red}[JDC: How was this addressed?]}
 
-\subsubsection*{Provenance}
+\subsection*{Other features}
+
+\subsubsection*{Tracking provenance information}
 
@@
 
-\subsubsection*{Modeling of all human tyrosine kinase catalytic domains}
+\subsection*{Modeling of all human tyrosine kinase catalytic domains}
 
@@
 
-\subsubsection*{Ensembler modeling statistics}
+\subsection*{Ensembler modeling statistics}
 
@@
 
-Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} (94\%) surviving models.
-To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
+Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} (94\%) surviving models across all TKs.
+To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data {\color{red}[JDC: State how big the solvated model PDBs are?]}, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
+
 The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
-The number of refined models for each target ranged from 4005 to 4248, with a median of 4160 and standard deviation of 60.
+The number of refined models for each target ranged from \num{4005} to \num{4248}, with a median of \num{4160} and standard deviation of \num{60}.
 Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
@@
 
-\subsubsection*{Evaluation of model quality and utility}
+\subsection*{Evaluation of model quality and utility}
 
@@
 
+\subsubsection*{All tyrosine kinases}
+
 To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
@@
 It is clear that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average.
+
+\subsubsection*{Src and Abl1}
 
@@
 
-\subsubsection*{Availability}
+\subsection*{Availability}
 
@@
 
-\subsubsection*{Future Directions}
+\subsection*{Future Directions}
 
@@
 
-\subsubsection*{Conclusion}
+\subsection*{Conclusion}
 

### 2015-03-22 manuscript/ms.tex
More edits.

@@
 As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TK) domains listed in UniProt.
-{\color{red}[JDC: What query was used to yield this set?]}
 {\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
@@
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
+{\color{red}[JDC: Shouldn't we state which options we used and what Uniprot searches we used for templates and targets?
+How would someone reproduce what we did here?
+Can we communicate this beyond just saying ``here are the scripts''?]}
 
@@
 
+{\color{red}[JDC: This section looks pretty anemic.
+What other kinds of analyses can we do for all the TKs?
+There is so much data here!
+There must be something neat we can do to examine it, right?]}
+
 \subsubsection*{Src and Abl1}
@@
 To provide a more complete evaluation of the models generated, we have analyzed two example TKs (\emph{Src} and \emph{Abl1}) in detail.
-Due to their importance in cancer, as outlined above, these kinases have been the subject of numerous studies, encompassing many different methodologies.
+Due to their importance in cancer, these kinases have been the subject of numerous studies, encompassing many different methodologies.
 In terms of structural data, a large number of crystal structures have been solved (with or without ligands such as nucleotide substrate or inhibitor drugs), showing the kinases in a number of different conformations.
 These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}.
+
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
@@
 
-The Abl1 renderings indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
+The Abl1 renderings in Fig.~\ref{figure:superposition} indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
 While such models are not necessarily incorrect or undesirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
@@
 
-To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domains.
+To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
 We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
+
 Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
@@
 The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
-We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of the activation process.
+We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
 

### 2015-03-24 manuscript/ms.tex
Minor edits

@@
 A number of mature software packages and forcefields are now available, and much recent progress has been driven by advances in computing architecture.
-For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, GPUGrid~\cite{buch:2010:gpugrid}, and Copernicus~\cite{pronk:2011:copernicus} allow scalability on an unprecedented level.
+For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, Copernicus~\cite{pronk:2011:copernicus}, and GPUGrid~\cite{buch:2010:gpugrid}, allow scalability on an unprecedented level.
 In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2010:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
@@
 
-While Modeller's automodel function can generate its own alignments automatically, we utilize the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, which we have empirically found to produce better quality alignments for purposes of high-throughput model building.
-{\color{red}[JDC: Can we give some evidence to support this claim, or soften our language on this?]}
+While Modeller's automodel function can generate its own alignments automatically, it was more convenient from a software architecture standpoint to use a standalone function.
+As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
 
@@
 By default, the Amber99SB-ILDN force field~\cite{amber99sb-ildn} is used with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
-The {\bf Ensembler} API allows the use of any of the other force fields implemented in OpenMM.
+Any of the other force fields or implicit water models implemented in OpenMM can be specified using the {\tt -{}-ff} and {\tt -{}-water\_model} flags respectively.
 The simulation is run for a default of 100 ps to filter out poor quality models (where atomic overlaps that cannot be resolved by energy minimization would cause the simulation to explode) and help relax models for subsequent production simulation.
@@
 The target number of waters is selected by first solvating each model with a specified padding distance (default: 10 \AA), then taking a percentile value from the distribution (default: 68th percentile).
-{\color{red}[JDC: Would be useful to explain why we are doing this.]}
-{\color{blue}[DLP: Addressed.]}
 This helps to prevent models with particularly long, extended loops---such as those arising from template structures with unresolved termini---from imposing very large box sizes on the entire set of models.
+The TIP3P water model~\cite{tip3p} is used by default, but any of the other explicit water models available in OpenMM, such as TIP4P-Ew~\cite{tip4p-ew}, can be specified using the {\tt -{}-water\_model} flag.
 Models are resolvated with the target number of waters by first solvating with zero padding, then incrementally increasing the box size and resolvating until the target is exceeded, then finally deleting sufficient waters to match the target value.
 The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field~\cite{amber99sb-ildn} and TIP3P water~\cite{tip3p} by default.
-Other force fields or water models such as TIP4P-Ew~\cite{tip4p-ew} can be specified via the {\bf Ensembler} API.
-{\color{red}[JDC: We should allow other water models in OpenMM too, such as TIP4P-Ew?]}
-{\color{blue}[DLP: I forgot to mention this in the text previously - any of the OpenMM force fields can be chosen via the API.
-I've updated the text accordingly.
-Is this functionality sufficient?
-I guess it's ok to leave ff choice as an "advanced" feature which requires use of the API?
-Otherwise I could add a -{}-water\_model flag to the CLI, for example.]}
-{\color{red}[JDC: Would be good to let at least water model and forcefield name be selectable via the CLI!]}
+The force field and water model can again be specified using the {\tt -{}-ff} and {\tt -{}-water\_model} flags respectively.
 
@@
 
-{\bf Ensembler} provides a packaging module which can be used to compress models in preparation for data transfer, or to prepare models with the appropriate directory and file structure for subsequent production simulations on the distributed computing platform Folding@home {\color{red}[CITE: F@H]}.
-The module could be easily extended to add methods for preparing models for use with other software, such as the Copernicus platform for running automated, distributed MD simulations.
-{\color{red}[JDC: Is there a way we can make this more generally useful to others?
-For example, is there a different system they might want to use, such as \href{http://copernicus.readthedocs.org/en/latest/tutorials/md-tutorial.html#creating-a-workflow-for-molecular-simulations}{Copernicus}?]}
-{\color{blue}[DLP: addressed]}
-{\color{red}[JDC: How was this addressed?]}
+{\bf Ensembler} provides a packaging module which can be used to prepare models for other uses.
+The {\tt package\_models} subcommand currently provides functions (specified \emph{via} the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home {\color{red}[CITE: F@H]}.
+The module could easily be extended to add methods for preparing models for other purposes.
+For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
 
@@
 For example, mutations of Src are associated with colon, breast, and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl1 cancer involvement]}.
-Protein kinase domains are thought to have multiple accessible metastable conformation states, with a single active conformation, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations [CITE: Lee and Craik Science 2009].
-{\color{red}[JDC: Lee and Craik do not discuss kinases, I don't believe; you'll have to find an accurate reference on kinase conformations.]}
+Protein kinase domains are thought to have multiple accessible metastable conformation states, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations~\cite{liu:ncb:2006:kinase-inhibitors-inactive-conformations}.
 Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach stands to benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.~by generating large numbers of starting configurations to be used in highly parallel MD simulation.
@@
 
+\subsubsection*{All tyrosine kinases}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Sequence identity distribution
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[tb]
+    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
+
+    \caption{{\bf Template-target sequence identity distribution for human tyrosine kinase catalytic domains.}
+    Sequence identities are calculated from all pairwise target-template alignments, where targets are human kinase catalytic domain sequences and templates are all kinase catalytic domains from any organism with structures in the PDB, as described in the text.
+    A kernel density estimate of the target-template sequence identity probability density function is shown as a solid line with shaded region, while the corresponding cumulative distribution function is shown as a dashed line.
+    }
+  \label{figure:sequence-identity-distribution}
+\end{figure}
+
+{\color{red}[JDC: DIscuss Fig.~\ref{figure:sequence-identity-distribution} first.]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection*{All tyrosine kinases}
-
 To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
@@
 These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}.
-
-Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
-Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
-{\color{red}[JDC: Explain how $k$-medoids clustering was done either here or in figure caption.]}
-Each model is colored and given a transparency based on the sequence identity between the target and template sequence.
-The figure gives an idea of the variance present in the generated models.
-High sequence identity models (in opaque blue) tend to be quite structurally similar, with some variation in loops or changes in domain orientation.
-
-The Abl1 renderings in Fig.~\ref{figure:superposition} indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
-While such models are not necessarily incorrect or undesirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
-Lower sequence identity models (in transparent white or red) indicate much greater variation in all parts of the structure.
-We believe the mix of high and low sequence identity models to be particularly useful for methods such as MSM building, which require thorough sampling of the conformational landscape.
-The high sequence identity models could be considered to be the most likely to accurately represent true metastable states.
-Conversely, the lower sequence identity models could be expected to help push a simulation into regions of conformation space which might take intractably long to reach if starting a single metastable conformation.
-
-To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
-
-Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
-One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
-These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm} [CITE Foda Shan Seeliger Src Nat Commun 2015; Ozkirimli Post Prot Sci 2008].
-As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
-The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
-We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Sequence identity distribution
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
-
-    \caption{{\bf Template-target sequence identity distribution for human tyrosine kinase catalytic domains.}
-    Sequence identities are calculated from all pairwise target-template alignments, where targets are human kinase catalytic domain sequences and templates are all kinase catalytic domains from any organism with structures in the PDB, as described in the text.
-    A kernel density estimate of the target-template sequence identity probability density function is shown as a solid line with shaded region, while the corresponding cumulative distribution function is shown as a dashed line.
-    }
-  \label{figure:sequence-identity-distribution}
-\end{figure}
 
@@
 
+Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
+Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
+{\color{red}[JDC: Explain how $k$-medoids clustering was done either here or in figure caption.]}
+Each model is colored and given a transparency based on the sequence identity between the target and template sequence.
+The figure gives an idea of the variance present in the generated models.
+High sequence identity models (in opaque blue) tend to be quite structurally similar, with some variation in loops or changes in domain orientation.
+
+The Abl1 renderings in Fig.~\ref{figure:superposition} indicate one high sequence identity model with a long unstructured region at one of the termini, which was unresolved in the original template structure.
+While such models are not necessarily incorrect or undesirable, it is important to be aware of the effects they may have on production simulations performed under periodic boundary conditions, as long unstructured termini can be prone to interact with a protein's periodic image.
+Lower sequence identity models (in transparent white or red) indicate much greater variation in all parts of the structure.
+We believe the mix of high and low sequence identity models to be particularly useful for methods such as MSM building, which require thorough sampling of the conformational landscape.
+The high sequence identity models could be considered to be the most likely to accurately represent true metastable states.
+Conversely, the lower sequence identity models could be expected to help push a simulation into regions of conformation space which might take intractably long to reach if starting a single metastable conformation.
+
+To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
+We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure}
+
+Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
+One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
+These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm} [CITE Foda Shan Seeliger Src Nat Commun 2015; Ozkirimli Post Prot Sci 2008].
+As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
+The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
+We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
 

### 2015-03-25 manuscript/ms.tex
Minor

@@
 
-While Modeller's automodel function can generate its own alignments automatically, it was more convenient from a software architecture standpoint to use a standalone function.
+While Modeller's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
 As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.

### 2015-03-25 figures/loopmodel_analysis/plot_dists.py
Minor

@@ -59,5 +59,5 @@ ax2.set_xlabel('Number of missing residues per template gap')
 
 sns.plt.tight_layout()
 
-sns.plt.savefig('nmissing_resis_distributions.pdf')
-# sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)
+# sns.plt.savefig('nmissing_resis_distributions.pdf')
+sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)

### 2015-03-25 figures/loopmodel_analysis/plot_dists.py
Updated nmissing_resis_distributions figure

@@ -21,14 +21,16 @@ sns.set_style('whitegrid')
 figsize=(3.5, 4)
 # sns.plt.figure(figsize=figsize)
 
-fig, (ax1, ax2) = sns.plt.subplots(2, 1, sharex=True, figsize=figsize)
+# fig, (ax1, ax2) = sns.plt.subplots(2, 1, sharex=True, figsize=figsize)
+fig, (ax1, ax2) = sns.plt.subplots(2, 1, figsize=figsize)
 
-bins = np.linspace(0, df.nmissing_resis.max(), 10)
-# sns.distplot(df.nmissing_resis, kde=False, bins=bins, label='total')
-# sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins, label='successful')
+# bins = np.linspace(0, df.nmissing_resis.max(), 10)
+bins1 = np.arange(0, df.nmissing_resis.max()+1, 1)
+sns.distplot(df.nmissing_resis, kde=False, bins=bins1, label='all templates', ax=ax1, hist_kws={'alpha': 1.})
+sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins1, label='failed remodeling', ax=ax1, color=palette[2], hist_kws={'alpha': 1.})
 # same with log y-axis:
-sns.distplot(df.nmissing_resis, kde=False, bins=bins, label='all templates', hist_kws={'log': True}, rug=True, rug_kws={'height': 0.7, 'lw': 0.5}, ax=ax1)
-sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins, label='failed remodeling', hist_kws={'log': True}, color=palette[2], ax=ax1)
+# sns.distplot(df.nmissing_resis, kde=False, bins=bins1, label='all templates', hist_kws={'log': True}, rug=True, rug_kws={'height': 0.7, 'lw': 0.5}, ax=ax1)
+# sns.distplot(unsuccessful.nmissing_resis, kde=False, bins=bins1, label='failed remodeling', hist_kws={'log': True}, color=palette[2], ax=ax1)
 # sns.plt.ylim(0.8,10000)
 
 ax1.legend()
@@ -50,12 +52,13 @@ ax1.legend()
 #     alpha=0.5,
 # )
 
-sns.distplot(missing_resis_spans, kde=True, hist=False, rug=True, color='k', kde_kws={'shade': True}, rug_kws={'lw': 0.5}, ax=ax2)
-ax2.set_yticks([])
[…]

### 2015-03-25 figures/loopmodel_analysis/plot_dists.py
Updated loopmodel analysis figure

@@ -54,13 +54,19 @@ ax1.legend()
 
 # sns.distplot(missing_resis_spans, kde=True, hist=False, rug=True, color='k', kde_kws={'shade': True}, rug_kws={'lw': 0.5}, ax=ax2)
 bins2 = np.arange(0, missing_resis_spans.max()+1, 1)
-sns.distplot(missing_resis_spans, kde=False, hist=True, bins=bins2, color='gray', ax=ax2, hist_kws={'alpha': 1.})
+sns.distplot(missing_resis_spans-2, kde=False, hist=True, bins=bins2, color='gray', ax=ax2, hist_kws={'alpha': 1.})
 # ax2.set_yticks([])
 
 ax1.set_xlabel('Remodeled residues per template')
 ax2.set_xlabel('Remodeled loop length')
 
+ax1.set_xlim(0, 60)
+ax2.set_xlim(0, 30)
+
+ax1.grid()
+ax2.grid()
+
 sns.plt.tight_layout()
 
-# sns.plt.savefig('nmissing_resis_distributions.pdf')
+sns.plt.savefig('nmissing_resis_distributions.pdf')
 sns.plt.savefig('nmissing_resis_distributions.png', dpi=300)

### 2015-03-25 figures/loopmodel_analysis/plot_dists.py
Updated loopmodel analysis figure

@@ -58,7 +58,9 @@ sns.distplot(missing_resis_spans-2, kde=False, hist=True, bins=bins2, color='gra
 # ax2.set_yticks([])
 
 ax1.set_xlabel('Remodeled residues per template')
+ax1.set_ylabel('Number of templates')
 ax2.set_xlabel('Remodeled loop length')
+ax2.set_ylabel('Number of loops')
 
 ax1.set_xlim(0, 60)
 ax2.set_xlim(0, 30)

### 2015-03-26 manuscript/ms.tex
Added energy analysis figure

@@
     \caption{{\bf Distributions for the number of missing residues in the TK templates.}
-    The upper plot shows the distribution of the total number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling (with the {\tt loopmodel} subcommand) failed (red).
-The raw data points for all templates are shown as a rug plot.
-    The lower plot shows the distribution of the number of residues per template gap, normalized and smoothed using kernel density estimation. The raw data points are shown as a rug plot.
-    {\color{red}[JDC: Some ideas for cleaning this up: Either the tick marks are being misrendered in that they are not taller if there are multiple data points with the same number or the data is really funky, since I would expect there to be a few examples in some bins.
-Also, is there a big drawback to making the top histogram bin size unity, since the values are integral?
-I don't think transparency is needed for the histogram bars either.
-We can also ditch the semilogy axis.
-I would also make the x-axes for the top and bottom plots different, since the data ranges are different.
-I'd see if a histogram with unit bin size might be more appropriate for the bottom plot as well---the KDE just doesn't feel right for this kind of data, since we are trying to report exact statistics from a specific example rather than estimate a general density for problems of this sort.
-Finally, I like ``remodeled loop length'' or ``missing loop length'' much better than ``Number of missing residues per template gap'', which seems unnecessarily verbose.]}
+    The upper histograms show the number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling with the {\tt loopmodel} subcommand failed (red).
+    The lower histogram shows the number of residues in each missing loop, for all templates.
 }

### 2015-03-26 figures/energies/plot.py

@@ -1,0 +1,28 @@
+import numpy as np
+import pandas as pd
+import seaborn as sns
+
+df = pd.read_csv('final_energies-implicit.csv', index_col=0)
+successful = df[df.successful == True]
+
+sns.set_context('paper', font_scale=1.)
+sns.set_style('white')
+figsize=(3.5, 5)
+fig, (ax1, ax2) = sns.plt.subplots(2, 1, figsize=figsize)
+
+sns.distplot(successful.final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, color='k', ax=ax1, label='all')
+sns.distplot(successful[successful.seqid<35].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='0-35', ax=ax2)
+sns.distplot(successful[successful.seqid>=35][successful.seqid<55].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='35-55', ax=ax2)
+sns.distplot(successful[successful.seqid>=55].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='55-100', ax=ax2)
+ax1.legend()
+ax2.legend()
+
+ax1.set_yticks([])
+ax2.set_yticks([])
+
+ax1.set_xlabel('Final potential energy (kT)')
+ax2.set_xlabel('Final potential energy (kT)')
+
+sns.plt.tight_layout()
+
+sns.plt.savefig('energies.png', dpi=300)

### 2015-03-27 manuscript/ms.tex
Fixed residue-pair-distances figure

@@
 As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
+The alignments are carried out with the {\tt align} subcommand, prior to the modeling step which is carried out with the {\tt build\_models} subcommand.
+The {\tt align} subcommand also writes a list of the sequence identities for each template to a text file, and this can be used to select models from a desired range of sequence identities.
+The {\tt build\_models} subcommand and all subsequent pipeline functions have a {\tt -{}-template\_seqid\_cutoff} flag which can be used to select only models with sequence identities greater than the given value.
 
 Models are output as PDB-format coordinate files.
-A list of all model identifiers sorted by sequence identity is also written to a text file, so that users may select models from the desired range of sequence identities.
-{\color{red}[JDC: I wonder if we should add an option ``exclude models with sequence identities poorer than X'', since this could be something people might want to do.]}
 To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
@@
     Superposed renderings of nine models each for Src and Abl1, 
-    {\color{red}[JDC: Src and Abl, or Src and Abl1? The description should match the captions above.]}
-    {\color{blue}[DLP: Addressed. Using Abl1, as this is the HGNC recommended symbol.]}
     giving some indication the diversity of conformations generated by Ensembler.
@@
     Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
-    % {\color{red}[JDC: Font size too small.]}
   }
@@
 
-    \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances.pdf}
+    % \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances.pdf}
+    \subfloat[Src]{\label{figure:src-distances}%
+       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_src.png}
+    }
+    \subfloat[Abl]{\label{figure:abl-distances}%
+       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_abl.png}
+    }
 
@@
     The atom names for these atoms, according to the PDB coordinate files for both reference structures, are---Lys: NZ, CD, CE (ethylamine); Glu: OE1, CD, OE2 (carboxylate); Arg: NH1, CZ, NH2 (part of guanidine).
-    % {\color{red}[JDC: Font size too small if single-column, so I made this double-column. Not sure if that's what you want.]}
-    % {\color{red}[JDC: Fill in rest of caption.]}
     }

### 2015-03-27 figures/energies/plot.py

@@ -4,6 +4,7 @@ import seaborn as sns
 
 df = pd.read_csv('final_energies-implicit.csv', index_col=0)
 successful = df[df.successful == True]
+unsuccessful_with_energies = df[df.successful == False][df.has_energies == True]
 
 sns.set_context('paper', font_scale=1.)
 sns.set_style('white')

### 2015-03-27 figures/residue_pair_distances/src_abl_combined/plot.py

@@ -1,101 +1,0 @@
-import os
-import shutil
-import tempfile
-import numpy as np
-import pandas as pd
-import mdtraj
-import seaborn as sns
-
-srcdir = '../../../data/models/SRC_HUMAN_D0'
-
-df = pd.read_csv(os.path.join(srcdir, 'traj-refine_implicit_md-data.csv'))
-
-traj = mdtraj.load(os.path.join(srcdir, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir, 'traj-refine_implicit_md-topol.pdb'))
-
-inactive_structure = mdtraj.load('../src/2SRC.pdb')
-active_structure = mdtraj.load('../src/1Y57.pdb')
-
-# K295 = 29
-# E310 = 44
-# R409 = 143
-residue_labels = ['K295', 'E310', 'R409']
-# traj_resids = [29, 44, 143]
-# ref_structure_residue_ids = [295, 310, 409]
-traj_residue_selections = [
-    'residue 29 and (name NZ or name CD or name CE)',
-    'residue 44 and (name "OE1" or name CD or name "OE2")',
-    'residue 143 and (name "NH1" or name CZ or name "NH2")',
-]
-traj_residue_indices = [28, 43, 142]   # 0-based
-ref_residue_selections = [
-    'residue 295 and (name NZ or name CD or name CE)',
-    'residue 310 and (name "OE1" or name CD or name "OE2")',
-    'residue 409 and (name "NH1" or name CZ or name "NH2")',
-]
-inactive_structure_residue_indices = [211, 226, 325]   # 0-based
-active_structure_residue_indices = [213, 228, 327]   # 0-based
-
-traj_residue_pair_indices = np.array([[28, 43], [43, 142]])
-inactive_structure_residue_pair_indices = np.array([[211, 226], [226, 325]])
-active_structure_residue_pair_indices = np.array([[213, 228], [228, 327]])
-
-
-pairs = [('K295', 'E310'), ('E310', 'R409')]
-
-pairs_dict = {pair: {'indices_traj': traj_residue_pair_indices[x]} for x, pair in enumerate(pairs)}
-
-contacts = mdtraj.compute_contacts(traj, contacts=traj_residue_pair_indices)[0].T
-
[…]

### 2015-03-27 figures/residue_pair_distances/src_abl_combined/plot_separate.py

@@ -1,0 +1,111 @@
+import os
+import shutil
+import tempfile
+import numpy as np
+import pandas as pd
+import mdtraj
+import seaborn as sns
+
+srcdir_src = '../../../data/models/SRC_HUMAN_D0'
+srcdir_abl = '../../../data/models/ABL1_HUMAN_D0'
+
+df_src = pd.read_csv(os.path.join(srcdir_src, 'traj-refine_implicit_md-data.csv'))
+df_abl = pd.read_csv(os.path.join(srcdir_abl, 'traj-refine_implicit_md-data.csv'))
+
+traj_src = mdtraj.load(os.path.join(srcdir_src, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir_src, 'topol-renumbered-implicit.pdb'))
+traj_abl = mdtraj.load(os.path.join(srcdir_abl, 'traj-refine_implicit_md.xtc'), top=os.path.join(srcdir_abl, 'topol-renumbered-implicit.pdb'))
+
+inactive_structure = mdtraj.load('../src/2SRC.pdb')
+active_structure = mdtraj.load('../src/1Y57.pdb')
+
+# hSrc cSrc
+# K298 K295 = 29
+# E313 E310 = 44
+# R412 R409 = 143
+
+# Abl
+# K271 = 30
+# E286 = 45
+# R386 = 145
+
+resSeqs_hsrc = [298, 313, 412]
+resSeqs_csrc = [295, 310, 409]
+resSeqs_abl = [271, 286, 386]
+residue_indices_src = [r for r, resi in enumerate(traj_src.topology.residues) if resi.resSeq in resSeqs_hsrc]
+residue_indices_abl = [r for r, resi in enumerate(traj_abl.topology.residues) if resi.resSeq in resSeqs_abl]
+residue_indices_src_inactive = [r for r, resi in enumerate(inactive_structure.topology.residues) if resi.resSeq in resSeqs_csrc]
+residue_indices_src_active = [r for r, resi in enumerate(active_structure.topology.residues) if resi.resSeq in resSeqs_csrc]
+
+traj_residue_pair_indices_src = np.array([residue_indices_src[0:2], residue_indices_src[1:3]])
+traj_residue_pair_indices_abl = np.array([residue_indices_abl[0:2], residue_indices_abl[1:3]])
[…]

### 2015-03-27 manuscript/ms.tex
Minor changes to ms.tex

@@
 
-Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
-{\color{red}[JDC: Can you add statistics (median, stddev, max) for loop lengths too?]}
+Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
 To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
@@
 Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} (94\%) surviving models across all TKs.
-To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data {\color{red}[JDC: State how big the solvated model PDBs are?]}, the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
+To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data (with solvated PDB coordinate files taking up about 0.9 MB each), the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
 
@@
 Each model generated about 116 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 0.5 GB per TK target or 41 GB for all 90 TKs.
-The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints to file) and 77 kB for the implicit solvent MD refinement stage.
-{\color{red}[JDC: Can you rework this paragraph to specify total space requirements \emph{without} saving MODELLER restraint files (since this is now the default), plus how much each restraint file adds if they are retained as well?
-I took a stab at it, but couldn't figure out how the numbers worked out.]}
+The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints files, which are about 397 kB per model) and 77 kB for the implicit solvent MD refinement stage.
 
@@
 
-    \caption{{\bf Distribution of RMSDs to all tyrosine kinase catalytic domain models relative to model derived from highest sequence identity template.}    
-    Distributions are averaged over all 90 tyrosine kinase catalytic domains.
-    {\color{red}[JDC: Is this accurate?]}
+    \caption{{\bf Distribution of RMSDs to all TK catalytic domain models relative to the model derived from the highest sequence identity template.}    
+    Distributions are built from data from all 90 TK targets.
     To better illustrate how conformational similarity depends on sequence identity, these were separated into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
@@
     {\color{red}[JDC: Are these computed for the templates or the resulting models?]}
+    {\color{blue}[DLP: Models.]}
     {\color{red}[JDC: Can we also show the overall distribution without stratification (e.g.~in grey in a separate panel)?]}
@@
 {\color{red}[JDC: Very cool!  Did you write a script to generate this, or did you have to do this by hand?]}
+{\color{blue}[DLP: The alignments come from UniProt. I did the latex formatting by hand (vi).]}
 

### 2015-03-27 manuscript/ms.tex
Added energies plot in ms. Various other changes

@@
 {\color{red}[JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]}
+{\color{blue}[DLP: Sikander mentioned to me that someone has developed an algorithm called pokefind (or knotfind, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
+DOI: 10.1093/bioinformatics/btp198
+It sounds like these algorithms have actually been implemented in Rosetta, so this could explain why Rosetta seems to do better at avoiding making these knotted structures.
+Would be useful to check this out further first, and then decide whether or not to discuss the knotted structures in the manuscript.]}
 
@@
 {\color{blue}[DLP: We currently just filter out models which throw exceptions or NaNs.]}
+{\color{blue}[DLP: Also note that the distribution of final energies for implicit refinement of TK models has been added in the Results section.]}
 
@@
 There must be something neat we can do to examine it, right?]}
+
+We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
+These ranged from -14180 kT to -3590 kT, with a median of -9533 kT and a standard deviation of 1058 kT.
+The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
+Of the \num{25457} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
+
+{\color{blue}[DLP: for further analysis, a good option might be to try to make a more rigorous assessment of model quality via comparison to reference crystal structures, based on features such as RMSD, phi/psi angles, H-bonds etc.
+We could also try using the Rosetta heuristic scoring function for this purpose.]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Final energies following implicit refinement
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[tbp]
+    \includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
+
+    \caption{{\bf Distribution of final energies from implicit solvent MD refinement of TK catalytic domain models.}
+    To illustrate how the energies are affected by sequence identity, the models are separated into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
+    The plotted distributions have been smoothed using kernel density estimation.
+  }
+  \label{figure:energies-implicit}
+\end{figure}
 

### 2015-03-27 figures/energies/plot.py

@@ -6,24 +6,32 @@ df = pd.read_csv('final_energies-implicit.csv', index_col=0)
 successful = df[df.successful == True]
 unsuccessful_with_energies = df[df.successful == False][df.has_energies == True]
 
-sns.set_context('paper', font_scale=1.)
+sns.set_context('paper', font_scale=0.9)
 sns.set_style('white')
-figsize=(3.5, 5)
-fig, (ax1, ax2) = sns.plt.subplots(2, 1, figsize=figsize)
+# figsize=(3.5, 5)
+figsize=(3.5, 2.5)
+# fig, (ax1, ax2) = sns.plt.subplots(2, 1, figsize=figsize)
+fig, (ax2) = sns.plt.subplots(1, 1, figsize=figsize)
 
-sns.distplot(successful.final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, color='k', ax=ax1, label='all')
+# sns.distplot(successful.final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, color='k', ax=ax1, label='all')
+print successful.final_energy.describe()
+
 sns.distplot(successful[successful.seqid<35].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='0-35', ax=ax2)
 sns.distplot(successful[successful.seqid>=35][successful.seqid<55].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='35-55', ax=ax2)
 sns.distplot(successful[successful.seqid>=55].final_energy, hist=False, kde=True, kde_kws={'shade': True, 'bw': 300}, label='55-100', ax=ax2)
-ax1.legend()
+# ax1.legend()
 ax2.legend()
 
-ax1.set_yticks([])
+# ax1.set_yticks([])
 ax2.set_yticks([])
 
-ax1.set_xlabel('Final potential energy (kT)')
+# ax1.set_xlim(-14000, -6000)
+ax2.set_xlim(-14000, -6000)
+
+# ax1.set_xlabel('Final potential energy (kT)')
 ax2.set_xlabel('Final potential energy (kT)')
 
 sns.plt.tight_layout()
 
 sns.plt.savefig('energies.png', dpi=300)
+sns.plt.savefig('energies.pdf', dpi=300)

### 2015-03-31 manuscript/ms.tex
Filled in various citations

@@
 Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
-The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts}) {\color{red}(CITE: Velankar Nucleic Acids Res 2013)}, which provides residue-level mappings between PDB and UniProt entries.
+The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts})~\cite{velankar:2013:sifts}, which provides residue-level mappings between PDB and UniProt entries.
 The SIFTS data is used to extract template sequences, retaining only residues which are resolved and match the equivalent residue in the UniProt sequence---non-wildtype residues are thus removed from the template structures.
@@
 While Modeller's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
-As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module {\color{red}[CITE: BioPython]}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
+As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
 The alignments are carried out with the {\tt align} subcommand, prior to the modeling step which is carried out with the {\tt build\_models} subcommand.
@@
 {\bf Ensembler} provides a packaging module which can be used to prepare models for other uses.
-The {\tt package\_models} subcommand currently provides functions (specified \emph{via} the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home {\color{red}[CITE: F@H]}.
+The {\tt package\_models} subcommand currently provides functions (specified \emph{via} the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
 The module could easily be extended to add methods for preparing models for other purposes.
@@
 To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)[CITE: 2SRC, 1Y57] and human Abl1 isoform A[CITE: 2F4J, 2HYY, 2G1T] respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
+We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
 
@@
 
-Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC) {\color{red}[CITE: 2SRC]} and active (PDB code: 1Y57) {\color{red}[CITE: 1Y57]} states.
+Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC)~\cite{xu:1999:2src} and active (PDB code: 1Y57)~\cite{cowan-jacob:2005:1y57} states.
 One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
-These three residues are also well conserved [CITE Kannan Neuwald JMB 2005], and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm} [CITE Foda Shan Seeliger Src Nat Commun 2015; Ozkirimli Post Prot Sci 2008].
+These three residues are also well conserved~\cite{kannan:jmb:2005:kinase-regulation}, and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm, foda:nat-comm:2015:src, ozkirimli:prot-sci:2008:src}.
 As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
@@
     \caption{{\bf Src and Abl1 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
-    Two Src structures (PDB entries 1Y57 [CITE] and 2SRC [CITE]) are projected onto the plots for reference, representing active and inactive states respectively.
+    Two Src structures (PDB entries 1Y57~\cite{cowan-jacob:2005:1y57} and 2SRC~\cite{xu:1999:2src}) are projected onto the plots for reference, representing active and inactive states respectively.
     These structures and the residue pairs analyzed here are depicted in Fig.~\ref{figure:src-ref-structures}.
@@
 These protonation states can have important effects on biological processes.
-For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism[CITE: Abl1 DFG flip evidence]---is controlled by protonation of the aspartate [CITE: Shan Shaw Proton-dependent switch Abl1 PNAS 2009].
+For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism [CITE: Abl1 DFG flip evidence]---is controlled by protonation of the aspartate~\cite{shan:pnas:2009:abl}.
 Currently, protonation states are assigned simply based on pH (a user-controllable parameter).
@@
 Binding sites for metal ions are frequently found in proteins, often playing a role in catalysis.
-For example, protein kinase domains contain two binding sites for divalent metal cations, and display significantly increased activity in the presence of Mg$^{2+}$~[CITE: Adams Taylor Protein Sci 1993], the divalent cation with highest concentration in mammalian cells.
+For example, protein kinase domains contain two binding sites for divalent metal cations, and display significantly increased activity in the presence of Mg$^{2+}$~\cite{adams:prot-sci:1993:kinase-metal-ions}, the divalent cation with highest concentration in mammalian cells.
 Metal ions are often not resolved in experimental structures of proteins, but by taking into account the full range of available structural data, it should be possible in many cases to include metal ions based on the structures of homologous proteins.
-We are careful to point out, however, that metal ion parameters in classical MD force fields have significant limitations, particularly in their interactions with proteins~[CITE: Sousa Ramos chapter 11 of Kinetics and Dynamics: From Nano- to Bio-Scale, Springer, 2010].
+We are careful to point out, however, that metal ion parameters in classical MD force fields have significant limitations, particularly in their interactions with proteins~\cite{sousa:book:2010:metal-ion-parameters-difficulties}.
 Cofactors and post-translational modifications are also often not fully resolved in experimental structures, and endogenous cofactors are frequently substituted with other molecules to facilitate experimental structural analysis.

### 2015-04-02 manuscript/ms.tex
Various updates

@@
 The mdtraj~\cite{mdtraj} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation, and the leader algorithm {\color{red}[JDC: Citation for leader algorithm?]} is then used to populate clusters.
-A minimum distance cutoff (which defaults to 0.6 \AA) is used to retain only a single model per cluster.
+A minimum distance cutoff (which defaults to 0.6~\AA) is used to retain only a single model per cluster.
 
@@
 
-This stage entails the use of molecular dynamics simulations to refine the models built in the previous step.
-This helps to improve model quality and also prepares models for subsequent production simulation, including solvation with explicit water molecules, if desired.
+While the utility of comparative modeling methods has been greatly enhanced by the recent explosion in the availability of protein structural data, the structures generated are generally considered "low-resolution" in comparison to those derived using experimental techniques such as X-ray crystallography.
+RMS errors of $\sim$3~\AA for C$_\alpha$ atoms relative to a native crystal structure are typical~\cite{raval:prot:2012:long-timescale-md-refinement,maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}.
+A number of refinement methods have been developed to help steer homology models toward more "native-like" conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}, of which MD simulations are an important example.
+While long-timescale unrestrained MD simulations (on the order of 100 $\mu$s) have been found to be ineffective for recapitulating native-like conformations, possibly due to forcefield issues~\cite{raval:prot:2012:long-timescale-md-refinement}, even relatively short simulations can be useful for relaxing structural elements such as sidechain orientation~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction}.
+
+{\bf Ensembler} thus includes a refinement module, which uses short molecular dynamics simulations to refine the models built in the previous step.
+As well improving model quality, this also prepares models for subsequent production MD simulation, including solvation with explicit water molecules, if desired.
 
@@
 This is implemented using the OpenMM molecular simulation toolkit~\cite{eastman:jctc:2012:openmm}, chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-By default, the Amber99SB-ILDN force field~\cite{amber99sb-ildn} is used with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
+The simulation is run for a default of 100 ps, which in our example applications has been sufficient to filter out poor models (i.e.~those with atomic overlaps unresolved by energy minimization, which result in an unstable simulation), as well as helping to relax model conformations.
+As discussed in the Results section, our example application of the {\bf Ensembler} pipeline to the human tyrosine kinase family indicated that of the models which failed implicit solvent MD refinement, the vast majority failed within the first 1~ps of simulation.
+
+The simulation protocol and default parameter values have been chosen to represent current "best practices" for the refinement simulations carried out here.
+As such, the simulation is performed using Langevin dynamics, with a default force field choice of Amber99SB-ILDN~\cite{amber99sb-ildn}, along with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
 Any of the other force fields or implicit water models implemented in OpenMM can be specified using the {\tt -{}-ff} and {\tt -{}-water\_model} flags respectively.
-The simulation is run for a default of 100 ps to filter out poor quality models (where atomic overlaps that cannot be resolved by energy minimization would cause the simulation to explode) and help relax models for subsequent production simulation.
-{\color{red}[JDC: What criteria were applied to filter out poor models?  Do we only look for thrown exceptions or NaNs?  Or do we use an energy filtering criteria too?]}
-{\color{blue}[DLP: We currently just filter out models which throw exceptions or NaNs.]}
-{\color{blue}[DLP: Also note that the distribution of final energies for implicit refinement of TK models has been added in the Results section.]}
+The simulation length can also be controlled \emph{via} the {\tt -{}-simlength} flag, and many other important simulation parameters can be controlled from either the API or CLI (\emph{via} the {\tt -{}-api\_params} flag).
+The default values are set as follows---timestep: 2~ps; temperature: 300~K; Langevin collision rate: 20~ps$^{-1}$; pH (used by OpenMM for protonation state assignment): 7.
+We also draw attention to a recent paper which indicates that lower Langevin collision rates may result in faster phase space exploration~\cite{basconi:jctc:2013:collision-rates}.
 
@@
 While protein-only models may be sufficient for structural analysis or implicit solvent simulations, {\bf Ensembler} also provides a stage for solvating models with explicit water and performing a round of explicit-solvent MD refinement/equilibration under isothermal-isobaric (NPT) conditions.
-The solvation step solvates each model for a given target with the same number of waters to facilitate the integration of data from multiple simulations, such as the construction of MSMs.
+The solvation step solvates each model for a given target with the same number of waters to facilitate the integration of data from multiple simulations, which is important for methods such as the construction of MSMs.
 The target number of waters is selected by first solvating each model with a specified padding distance (default: 10 \AA), then taking a percentile value from the distribution (default: 68th percentile).
@@
 The explicit solvent MD simulation is also implemented using OpenMM, using the Amber99SB-ILDN force field~\cite{amber99sb-ildn} and TIP3P water~\cite{tip3p} by default.
-The force field and water model can again be specified using the {\tt -{}-ff} and {\tt -{}-water\_model} flags respectively.
+The force field, water model, and simulation length can again be specified using the {\tt -{}-ff}, {\tt -{}-water\_model}, and {\tt -{}-simlength} flags respectively.
+Further simulation parameters can be controlled \emph{via} the API or \emph{via} the CLI {\tt -{}-api\_params} flag.
+Pressure control is performed with a Monte Carlo barostat as implemented in OpenMM, with a default pressure of 1~atm and a period of 50 timesteps.
+The remaining simulation parameters have default values set to the same as for the implicit solvent MD refinement.
 
@@
 \begin{figure}[tbp]
-    \includegraphics[width=1.0\columnwidth]{rmsddist/rmsddist2.pdf}
+    \centering
+    {\label{figure:rmsd-distribution-joint}%
+       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
+    }
+
+    {\label{figure:rmsd-distribution-by-sequence-identity}%
+       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
+    }
 
@@
     Distributions are built from data from all 90 TK targets.
-    To better illustrate how conformational similarity depends on sequence identity, these were separated into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
+    To better illustrate how conformational similarity depends on sequence identity, the lower plot illustrates the distributions as stratified into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
     The plotted distributions have been smoothed using kernel density estimation.
-    {\color{red}[JDC: Are these computed for the templates or the resulting models?]}
-    {\color{blue}[DLP: Models.]}
-    {\color{red}[JDC: Can we also show the overall distribution without stratification (e.g.~in grey in a separate panel)?]}
   }
-  \label{figure:rmsd-distribution-by-sequence-identity}
+  \label{figure:rmsd-distributions}
 \end{figure}
 
-To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distribution-by-sequence-identity}.
+To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distributions}.
 To better understand the influence of sequence identity on the conformational similarities of resulting models, the sequence identities were stratified based on the sequence identity distribution plotted in Fig.~\ref{figure:sequence-identity-distribution}, which suggests an intuitive division into three categories, with \num{307753} models in the 0--35\% sequence identity range, \num{69922} models in the 35--55\% range, and \num{4893} models in the 55--100\% range.
@@
     The models for each target were divided into three sequence identity ranges (as in Fig.
-\ref{figure:rmsd-distribution-by-sequence-identity}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
+\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
     The models shown are the centroids of each cluster.

### 2015-04-02 figures/rmsddist/rmsddist2.py

@@ -73,7 +73,7 @@ sns.despine(left=True)
 plt.tight_layout()
 
 plt.savefig('rmsddist2.pdf')
-# plt.savefig('rmsddist2.png', dpi=300)
+plt.savefig('rmsddist2.png', dpi=300)
 
 plt.close()
 
@@ -85,3 +85,26 @@ plt.close()
 # plt.setp(plt.gca(), yticks=[])
 # plt.xlim(0,102)
 # plt.savefig('seqid_dist.png')
+
+# Plot distribution without stratification
+
+sns.set_context('paper', font_scale=1.)
+sns.set_style('white')
+fig = sns.plt.figure(figsize=(3.5,2.625))
+
+sns.kdeplot(df_has_model.rmsd*10, label=label, bw=2.5, shade=True, color='k')
+
+ax = plt.gca()
+plt.setp(ax, yticks=[])
+
+plt.xlim(0,40)
+plt.xlabel('RMSD ($\AA$)')
+sns.despine(left=True)
+
+legend = plt.legend()
+legend.set_visible(False)
+
+plt.tight_layout()
+
+plt.savefig('rmsddist2-joint.png', dpi=300)
+plt.savefig('rmsddist2-joint.pdf')

### 2015-04-03 manuscript/ms.tex
Various updates

@@
 For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, Copernicus~\cite{pronk:2011:copernicus}, and GPUGrid~\cite{buch:2010:gpugrid}, allow scalability on an unprecedented level.
-In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2010:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
+In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2011:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
 MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
@@
 
-The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org}) {\color{red}[JDC: Cite \href{http://www.uniprot.org/help/publications}{one of these UniProt citations?}]}---via a UniProt search query.
+The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---via a UniProt search query.
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} us used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
@@
 For this reason, and also to allow users to select for high diversity if they so choose, {\bf Ensembler} provides a way to filter out models that are very similar in RMSD.
-A fast clustering scheme is used to identify models differing by a user-specified minimum RMSD.
-The mdtraj~\cite{mdtraj} Python library is used to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation, and the leader algorithm {\color{red}[JDC: Citation for leader algorithm?]} is then used to populate clusters.
+The {\tt cluster} subcommand can thus be used to identify models which differ from other models in terms of RMSD distance by a user-specified cutoff.
+Clustering is performed using the regular spatial clustering algorithm~\cite{noe:jcp:2011:msm-review}, as implemented in the MSMBuilder Python library~\cite{msmbuilder}, which uses mdtraj~\cite{mdtraj} to calculate RMSD (for C$_\alpha$ atoms only) with a fast quaternion characteristic polynomial (QCP)~\cite{theobald:acta-cryst-a:2005:qcp,theobald:j-comput-chem:2010:qcp,theobald:j-comput-chem:2011:qcp} implementation.
 A minimum distance cutoff (which defaults to 0.6~\AA) is used to retain only a single model per cluster.

### 2015-04-03 manuscript/ms.tex
Documented commands used to select TK targets and templates

@@
 % Pretty-printing of shell commands
-\newcommand{\shellcmd}[1]{\\\ \texttt{\scriptsize\# #1}\\}
+\newcommand{\shellcmd}[1]{\\\ \texttt{\scriptsize #1}}
 
@@
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} us used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
-For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND taxonomy:9606 AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
+For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND organism:"homo sapiens" AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
 In this way, the user may select a single protein, many proteins, or an entire superfamily from UniProt.
@@
 In many cases, it will be desirable to build models of an isolated protein domain, rather than the full-length protein.
-The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-domains} flag. 
+The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-uniprot\_domain\_regex} flag. 
 For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
 To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
-In this case, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~{\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}).
+If the {\tt -{}-uniprot\_domain\_regex} flag is used, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~{\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}).
 
@@
 
-Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-domains} flag allows selection of individual domains with a regular expression string.
+Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-uniprot\_domain\_regex} flag allows selection of individual domains with a regular expression string.
 The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
 Only structures solved by X-ray crystallography or NMR are selected, thus excluding computer-generated models available from the PDB.
-If the {\tt -{}-domains} flag is used, then templates are truncated at the start and end of the domain sequence.
+If the {\tt -{}-uniprot\_domain\_regex} flag is used, then templates are truncated at the start and end of the domain sequence.
 
 Templates can also be defined manually.
-Manual specification of templates simply requires storing the sequences and identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
+Manual specification of templates simply requires storing the sequences and arbitrary identifiers in a FASTA file, and the structures as PDB-format coordinate files with filenames matching the identifiers in the sequence file.
 The structure residues must also match those in the sequence file.
@@
 
-As a first application of {\bf Ensembler}, we have built models for all 90 human tyrosine kinase (TK) domains listed in UniProt.
+As a first application of {\bf Ensembler}, we have built models for the human tyrosine kinase (TK) family.
 {\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
@@
 
-We selected all available structures of protein kinase domains (of any species) as templates, for a total of 4433 (\num{398970} target-template pairs).
+We selected all 90 TK domains annotated in UniProt as targets, using the following command:
+\shellcmd{gather\_targets -{}-query `family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes' -{}-uniprot\_domain\_regex `\^{}Protein kinase(?!; truncated)(?!; inactive)'}\\
+The {\tt `reviewed:yes'} expression in the UniProt query ensured that only UniProt entries which have been reviewed by a human expert were included in the results.
+The {\tt -{}-uniprot\_domain\_regex} regular expression resulted in the selection of domains annotated "Protein kinase", "Protein kinase 1", and "Protein kinase 2", while excluding domains "Protein kinase; truncated", "Protein kinase; inactive", "Alpha-type protein kinase", and many types of non-kinase domain.
+We selected all available structures of protein kinase domains (of any species) as templates, using the following command:
+\shellcmd{gather\_templates -{}-gather\_from uniprot -{}-query `domain:"Protein kinase" AND reviewed:yes' -{}-uniprot\_domain\_regex `\^{}Protein kinase(?!; truncated)(?!; inactive)'}\\
+This returned 4433 templates, giving a total of \num{398970} target-template pairs.
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
-{\color{red}[JDC: Shouldn't we state which options we used and what Uniprot searches we used for templates and targets?
-How would someone reproduce what we did here?
-Can we communicate this beyond just saying ``here are the scripts''?]}
 
@@
 \shellcmd{conda config --add channels https://conda.binstar.org/omnia}
-\shellcmd{conda ensembler}
+\shellcmd{conda ensembler}\\
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.

### 2015-04-03 manuscript/ms.tex
Various updates

@@
 \author{Patrick B. Grinaway}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+\author{Sonya M. Hanson}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+\author{Kyle A. Beauchamp}
   \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
@@
 {\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
-TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer.
-{\color{red}[JDC: CITE]}
-For example, mutations of Src are associated with colon, breast, and prostate cancer {\color{red}[CITE: Src cancer involvement]}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia {\color{red}[CITE: Abl1 cancer involvement]}.
+TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer~\cite{krause:nejm:2005:tks-cancer}.
+For example, mutations of Src are associated with colon, breast, prostate, lung, and pancreatic cancers~\cite{kim:2009:src-cancer-review}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia~\cite{greuber:2013:abl-cancer-review}.
 Protein kinase domains are thought to have multiple accessible metastable conformation states, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations~\cite{liu:ncb:2006:kinase-inhibitors-inactive-conformations}.
-Kinases are thus a particularly interesting subject for study with MSM methods [CITE: recent kinase MSM papers], and this approach stands to benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.~by generating large numbers of starting configurations to be used in highly parallel MD simulation.
+Kinases are thus a particularly interesting subject for study with MSM methods~\cite{shukla-pande:nature-commun:2014:src-activation-msm}, and this approach stands to benefit greatly from the ability to exploit the full body of available genomic and structural data within the kinase family, e.g.~by generating large numbers of starting configurations to be used in highly parallel MD simulation.
 
@@
 
-{\color{red}[JDC: DIscuss Fig.~\ref{figure:sequence-identity-distribution} first.]}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To evaluate the diversity of conformations captured for each target sequence, we first computed the RMSD distributions for all models for each target (relative to the model derived from the highest-identity template) are shown in Fig.~\ref{figure:rmsd-distributions}.
-To better understand the influence of sequence identity on the conformational similarities of resulting models, the sequence identities were stratified based on the sequence identity distribution plotted in Fig.~\ref{figure:sequence-identity-distribution}, which suggests an intuitive division into three categories, with \num{307753} models in the 0--35\% sequence identity range, \num{69922} models in the 35--55\% range, and \num{4893} models in the 55--100\% range.
-It is clear that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average.
+To evaluate the variety of template sequence similarities relative to each target sequence, we calculated sequence identity distributions, as shown in Fig.~\ref{figure:sequence-identity-distribution}.
+This suggests an intuitive division into three categories, with \num{307753} models in the 0--35\% sequence identity range, \num{69922} models in the 35--55\% range, and \num{4893} models in the 55--100\% range.
+We then computed the RMSD distributions for the models created for each target (relative to the model derived from the template with highest sequence identity) Fig.~\ref{figure:rmsd-distributions}, to assess the diversity of conformations captured by the modeling pipeline.
+Furthermore, to understand the influence of sequence identity on the conformational similarities of the resulting models, the RMSD distributions were stratified based on the three sequence identity categories described above.
+This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average.
+
+We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
+These ranged from -14180 kT to -3590 kT, with a median of -9533 kT and a standard deviation of 1058 kT.
+The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
+Of the \num{25457} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
 
@@
 There must be something neat we can do to examine it, right?]}
-
-We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
-These ranged from -14180 kT to -3590 kT, with a median of -9533 kT and a standard deviation of 1058 kT.
-The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
-Of the \num{25457} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
 
@@
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
-{\color{red}[Add PBG support statement.]}
+{\color{red}[Add PBG, SMH, KAB support statements.]}
 
@@
 Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
-{\color{red}[JDC: Very cool!  Did you write a script to generate this, or did you have to do this by hand?]}
-{\color{blue}[DLP: The alignments come from UniProt. I did the latex formatting by hand (vi).]}
 

### 2015-04-03 manuscript/ms.tex
Various updates

@@
 
-{\color{red}[JDC: This section looks pretty anemic.
-What other kinds of analyses can we do for all the TKs?
-There is so much data here!
-There must be something neat we can do to examine it, right?]}
-
-{\color{blue}[DLP: for further analysis, a good option might be to try to make a more rigorous assessment of model quality via comparison to reference crystal structures, based on features such as RMSD, phi/psi angles, H-bonds etc.
-We could also try using the Rosetta heuristic scoring function for this purpose.]}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-{\color{red}[JDC: In the Discussion, let's be sure to talk about the limitations and what could be improved or added in the future.
-For example, we don't yet handle counterions (e.g.~structural Zn$^{2+}$), prosthetic groups (e.g.~heme), or cofactors (e.g.~ATP) yet.
-We don't handle post-translational modifications either (such as phosphorylation, methylation, glycosylation, etc.).
-It's a good idea to suggest that this is an important first step toward enabling superfamily- and genomics-scale modeling, but there's a lot of work yet to be done.]}
-
 Comparative protein modeling and MD simulation set-up can be approached in a number of different ways, with varying degrees of complexity, and there are a number of obvious additions and improvements which we plan to implement in future versions of {\bf Ensembler}.
@@
 We thus plan to implement an interface and command-line function for assigning protonation states with MCCE2~\cite{alexov-gunner:biophys-j:1997:mcce2,georgescu-alexov-gunner:biophys-j:2002:mcce2,song-mao-gunner:j-comput-chem:2009:mcce2}, which uses electrostatics calculations combined with Monte Carlo sampling of side chain conformers to calculate pKa values.
-{\color{red}[JDC: I think we may want to consider doing that at this stage.  Let's discuss.]}
 

### 2015-04-03 manuscript/ms.tex
Various updates

@@
 Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
-{\color{red}[JDC: Can we also add the URL of where to get the code and TK models here?]}
+{\color{blue}[TODO: Add URL of where to get the code and TK models here]}
 
@@
 {\color{red}[JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]}
-{\color{blue}[DLP: Sikander mentioned to me that someone has developed an algorithm called pokefind (or knotfind, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
+{\color{blue}[DLP: Might be worth investigating an algorithm called \emph{pokefind} (or \emph{knotfind}, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
 DOI: 10.1093/bioinformatics/btp198
@@
 As a first application of {\bf Ensembler}, we have built models for the human tyrosine kinase (TK) family.
-{\color{red}[JDC: Is there a complete list of these somewhere?  Maybe reference supplementary data?]}
+{\color{blue}[TODO: Add list of TK UniProt identifiers and gene names, probably in SI.]}
 TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer~\cite{krause:nejm:2005:tks-cancer}.
@@
 The latest source can be downloaded from the GitHub repository, which also contains up-to-date instructions for building and installing the code.
-{\color{red}[JDC: Where does documentation reside?  And example inputs for generating the results in this paper?]}
+{\color{blue}[TODO: Add link to documentation. What about example inputs for generating the results in this paper? Put this in the TK model data set?]}
 

### 2015-04-09 manuscript/ms.tex
Various updates

@@
 {\bf Ensembler} is free and open source software licensed under the GNU General Public License (GPL) v2. 
-It should run on all major operating systems, and has been tested on Linux and OS X.
+It is compatible with Linux and OS X.
 The latest release can be installed via the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
@@
 {\bf Ensembler} provides functions for selecting target sequences and homologous template structures, and (by interfacing with a number of external packages) performs pairwise alignments, comparative modeling of target-template pairs, and several stages of model refinement.
-As an example application, we have constructed models for the entire set of human tyrosine kinase catalytic domains, using all available structures of protein kinase domains (from any species) as templates.
+As an example application, we have constructed models for the entire set of human tyrosine kinase (TK) catalytic domains, using all available structures of protein kinase domains (from any species) as templates.
 This results in a total of almost 400,000 models, and we demonstrate that these provide wide-ranging coverage of known functionally relevant conformations.
 By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
+For example, models with high target-template sequence identity are expected to be the most likely to represent native metastable states, while lower sequence identity models would aid in sampling of more distant regions of accessible phase space.
+It is also important to note that some models (especially low sequence identity models) may not represent natively accessible conformations.
+However, MSM methods benefit from the ability to remove outlier MD trajectories which start from non-natively accessible conformations, and which would thus be unconnected with the phase space sampled in other trajectories.
+These methods essentially identify the largest subset of Markov nodes which constitute an ergodic network~\cite{msmbuilder,caflisch:jpcb:2011:msm-subgraph}.
 
-We anticipate that the tool will prove to be useful in a number of other ways.
+We anticipate that {\bf Ensembler} will prove to be useful in a number of other ways.
 For example, the generated models could represent valuable data sets even without subsequent production simulation, allowing exploration of the conformational diversity present within the available structural data for a given protein family.
@@
 
-  \caption{{\bf Diagrammatic representation of the various stages of the Ensembler pipeline and illustrative statistics for modeling all human tyrosine kinase catalytic domains.}
+  \caption{{\bf Diagrammatic representation of the stages of the Ensembler pipeline and illustrative statistics for modeling all human tyrosine kinase catalytic domains.}
   On the left, the various stages of the {\bf Ensembler} pipeline are shown.
@@
 
-The first stage entails the selection of a set of \emph{target} protein sequences---the sequences the user is interested in generating simulation-ready structural models for.
+The first stage entails the selection of a set of \emph{target} protein sequences---the sequences for which the user is interested in generating simulation-ready structural models.
 This may be a single sequence---such as a full-length protein or a construct representing a single domain---or a collection of sequences, such as a particular domain from an entire family of proteins.
@@
 The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---via a UniProt search query.
-To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} us used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
+To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} is used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
 For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND organism:"homo sapiens" AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
@@
 Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
-Furthermore, the Rosetta {\tt loopmodel} program will not model missing residues at the termini of a structure---such residues spans are modeled in the subsequent stage.
+Furthermore, the Rosetta {\tt loopmodel} program will not model missing residues at the termini of a structure---such residue spans are modeled in the subsequent stage.
 
@@
 While Modeller's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
-As such, we implemented pairwise alignment functionality using the the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
+As such, we implemented pairwise alignment functionality using the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
 The alignments are carried out with the {\tt align} subcommand, prior to the modeling step which is carried out with the {\tt build\_models} subcommand.
@@
 The {\tt build\_models} subcommand and all subsequent pipeline functions have a {\tt -{}-template\_seqid\_cutoff} flag which can be used to select only models with sequence identities greater than the given value.
+We also note that alternative approaches could be used for the alignment stage.
+For example, multiple sequence alignment algorithms~\cite{thompson:2011:msa-review}, allow alignments to be guided using sequence data from across the entire protein family of interest, while (multiple) structural alignment algorithms such as Modeller's salign routine~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}, PROMALS3D~\cite{pei:2008:promals3d}, and Expresso and 3DCoffee~\cite{expresso,3dcoffee}, can additionally exploit structural data.
+{\bf Ensembler's} modular architecture facilitates the implementation of alternative alignment approaches, and we plan to implement some of these in future versions, to allow exploration of the influence of different alignment methods on model quality.
 
@@
 The restraints used by Modeller could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
-This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 KB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
+This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 kB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
 
@@
 
-While the utility of comparative modeling methods has been greatly enhanced by the recent explosion in the availability of protein structural data, the structures generated are generally considered "low-resolution" in comparison to those derived using experimental techniques such as X-ray crystallography.
-RMS errors of $\sim$3~\AA for C$_\alpha$ atoms relative to a native crystal structure are typical~\cite{raval:prot:2012:long-timescale-md-refinement,maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}.
-A number of refinement methods have been developed to help steer homology models toward more "native-like" conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}, of which MD simulations are an important example.
+A number of refinement methods have been developed to help guide comparative modeling techniques toward more "native-like" and physically consistent conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}, of which MD simulations are an important example.
 While long-timescale unrestrained MD simulations (on the order of 100 $\mu$s) have been found to be ineffective for recapitulating native-like conformations, possibly due to forcefield issues~\cite{raval:prot:2012:long-timescale-md-refinement}, even relatively short simulations can be useful for relaxing structural elements such as sidechain orientation~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction}.
@@
 {\bf Ensembler} thus includes a refinement module, which uses short molecular dynamics simulations to refine the models built in the previous step.
-As well improving model quality, this also prepares models for subsequent production MD simulation, including solvation with explicit water molecules, if desired.
+As well as improving model quality, this also prepares models for subsequent production MD simulation, including solvation with explicit water molecules, if desired.
 
@@
 
-As a first application of {\bf Ensembler}, we have built models for the human tyrosine kinase (TK) family.
+As a first application of {\bf Ensembler}, we have built models for the human TK family.
 {\color{blue}[TODO: Add list of TK UniProt identifiers and gene names, probably in SI.]}
 TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer~\cite{krause:nejm:2005:tks-cancer}.
-For example, mutations of Src are associated with colon, breast, prostate, lung, and pancreatic cancers~\cite{kim:2009:src-cancer-review}, while a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia~\cite{greuber:2013:abl-cancer-review}.
+For example, a translocation between the TK Abl1 and the pseudokinase Bcr is closely associated with chronic myelogenous leukemia~\cite{greuber:2013:abl-cancer-review}, while mutations of Src are associated with colon, breast, prostate, lung, and pancreatic cancers~\cite{kim:2009:src-cancer-review}.
 Protein kinase domains are thought to have multiple accessible metastable conformation states, and much effort is directed at developing kinase inhibitor drugs which bind to and stabilize inactive conformations~\cite{liu:ncb:2006:kinase-inhibitors-inactive-conformations}.
@@
 
-Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
+Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, mean 14, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, mean 7, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
 To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
@@
 The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
-The number of refined models for each target ranged from \num{4005} to \num{4248}, with a median of \num{4160} and standard deviation of \num{60}.
+The number of refined models for each target ranged from \num{4005} to \num{4248}, with a median of \num{4160}, mean of \num{4150}, and standard deviation of \num{60}.
 Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
 
-Each model generated about 116 KB of file data (up to and including the implicit solvent MD refinement stage), totalling 0.5 GB per TK target or 41 GB for all 90 TKs.
+Each model generated about 116 kB of file data (up to and including the implicit solvent MD refinement stage), totalling 0.5 GB per TK target or 41 GB for all 90 TKs.
 The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints files, which are about 397 kB per model) and 77 kB for the implicit solvent MD refinement stage.
@@
 We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
-These ranged from -14180 kT to -3590 kT, with a median of -9533 kT and a standard deviation of 1058 kT.
+These ranged from -14180 kT to -3590 kT, with a median of -9533 kT, mean of -9564 kT, and a standard deviation of 1058 kT.
 The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
@@
 As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
-The models show strong coverage of regions in which either of the electrostatic interactions is formed, as well as a wide range of regions inbetween.
+The models show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions inbetween (mainly models with low sequence identity).
 We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
@@
 \shellcmd{conda config --add channels https://conda.binstar.org/omnia}
-\shellcmd{conda ensembler}\\
+\shellcmd{conda install ensembler}\\
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
@@
 Another limitation with the present version of {\bf Ensembler} involves the treatment of members of a protein family with especially long residue insertions or deletions.
-For example, the set of all human protein kinase domains listed in UniProt have a median length of 265 residues and a standard deviation of 45, yet the minimum and maximum lengths are 102 and 801 respectively.
+For example, the set of all human protein kinase domains listed in UniProt have a median length of 265 residues (mean 277) and a standard deviation of 45, yet the minimum and maximum lengths are 102 and 801 respectively.
 The latter value corresponds to the protein kinase domain of serine/threonine-kinase \emph{greatwall}, which includes a long insertion between the two main lobes of the catalytic domain.
@@
 
-The authors are grateful to Kyle A.~Beauchamp (MSKCC), Robert McGibbon (Stanford), Arien S. Rustenburg (MSKCC) for many excellent software engineering suggestions.
-The authors thank Sonya M.~Hanson (MSKCC), Nicholas M.~Levinson (University of Minnesota), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
+The authors are grateful to Robert McGibbon (Stanford) and Arien S. Rustenburg (MSKCC) for many excellent software engineering suggestions.
+The authors thank Nicholas M.~Levinson (University of Minnesota), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
 The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
-{\color{red}[Add PBG, SMH, KAB support statements.]}
+{\color{red}[Add PBG, ?SMH, ?KAB support statements.]}
 

### 2015-04-10 manuscript/ms.tex
Various updates

@@
 \usepackage{alltt}
+\usepackage{fancyvrb}
 
@@
 \usepackage[font=sf]{floatrow}
+
+% To allow code "Boxes" as well as figures
+\usepackage{newfloat}
+\DeclareFloatingEnvironment[name={Box}]{codebox}
 
@@
 By using these models as starting configurations for highly parallel MD simulations, we expect their structural diversity to greatly aid in sampling of conformational space.
-For example, models with high target-template sequence identity are expected to be the most likely to represent native metastable states, while lower sequence identity models would aid in sampling of more distant regions of accessible phase space.
+We further suggest that models with high target-template sequence identity are the most likely to represent native metastable states, while lower sequence identity models would aid in sampling of more distant regions of accessible phase space.
 It is also important to note that some models (especially low sequence identity models) may not represent natively accessible conformations.
@@
 {\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
+The specific command-line flags and API details discussed in this paper correspond to the version [TODO] release (TODO: link).
+Up-to-date documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
 
 The {\bf Ensembler} modeling pipeline comprises a series of stages which are performed in a defined order. 
-A visual overview of the pipeline is shown in Fig.~\ref{figpipeline}.
+A visual overview of the pipeline is shown in Fig.~\ref{figure:pipeline}.
 The various stages of this pipeline are described in detail below.
@@
   On the left, the various stages of the {\bf Ensembler} pipeline are shown.
+  The red labels indicate the corresponding text description provided for each stage in the Design and Implementation section.
   On the right, the number of viable models surviving each stage of the pipeline is shown for modeling all 90 tyrosine kinases (\emph{All TKs}) and representative individual tyrosine kinases (\emph{SRC} and \emph{ABL}).
@@
   }
-  \label{figpipeline}
+  \label{figure:pipeline}
 \end{figure*}
 
-\subsection*{Target selection and retrieval}
+\subsection{Target selection and retrieval}
 
@@
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} is used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
-For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while {\tt -{}-query `domain:"Protein kinase" AND organism:"homo sapiens" AND reviewed:yes'} would select all human protein kinases which have been reviewed by a human curator.
+For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while the query shown in Box~\ref{box:tk-gather-targets-templates} would select all human tyrosine protein kinases which have been reviewed by a human curator.
 In this way, the user may select a single protein, many proteins, or an entire superfamily from UniProt.
@@
 For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
-To select only domains of the first three types, the following regular expression could be used: {\tt`\^{}Protein kinase(?!; truncated)(?!; inactive)'}.
+The regular expression shown in Box~\ref{box:tk-gather-targets-templates} selects only domains of the first three types.
 If the {\tt -{}-uniprot\_domain\_regex} flag is used, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~{\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}).
@@
 
-\subsection*{Template selection and retrieval}
+\subsection{Template selection and retrieval}
 
@@
 
-Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-uniprot\_domain\_regex} flag allows selection of individual domains with a regular expression string.
+Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-uniprot\_domain\_regex} flag allows selection of individual domains with a regular expression string (Box~\ref{box:tk-gather-targets-templates}).
 The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
@@
 
-\subsection*{Template refinement}
+\subsection{Template refinement}
 
 Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
-Because fewer loops need to be built during the subsequent target model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
-{\color{red}[JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]}
-{\color{blue}[DLP: Might be worth investigating an algorithm called \emph{pokefind} (or \emph{knotfind}, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
-DOI: 10.1093/bioinformatics/btp198
-It sounds like these algorithms have actually been implemented in Rosetta, so this could explain why Rosetta seems to do better at avoiding making these knotted structures.
-Would be useful to check this out further first, and then decide whether or not to discuss the knotted structures in the manuscript.]}
-
+We expect that in certain cases, pre-building template loops with Rosetta {\tt loopmodel} prior to the main modeling stage (with Modeller) may result in improved model quality.
 Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
@@
 
-\subsection*{Modeling}
+% Because fewer loops need to be built during the subsequent target model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
+% [JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]
+% [DLP: Might be worth investigating an algorithm called \emph{pokefind} (or \emph{knotfind}, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
+% DOI: 10.1093/bioinformatics/btp198
+% It sounds like these algorithms have actually been implemented in Rosetta, so this could explain why Rosetta seems to do better at avoiding making these knotted structures.
+% Would be useful to check this out further first, and then decide whether or not to discuss the knotted structures in the manuscript.]
+
+\subsection{Modeling}
 
@@
 We also note that alternative approaches could be used for the alignment stage.
-For example, multiple sequence alignment algorithms~\cite{thompson:2011:msa-review}, allow alignments to be guided using sequence data from across the entire protein family of interest, while (multiple) structural alignment algorithms such as Modeller's salign routine~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}, PROMALS3D~\cite{pei:2008:promals3d}, and Expresso and 3DCoffee~\cite{expresso,3dcoffee}, can additionally exploit structural data.
+For example, multiple sequence alignment algorithms~\cite{thompson:2011:msa-review}, allow alignments to be guided using sequence data from across the entire protein family of interest, while (multiple) structural alignment algorithms such as Modeller's {\tt salign} routine~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}, PROMALS3D~\cite{pei:2008:promals3d}, and Expresso and 3DCoffee~\cite{expresso,3dcoffee}, can additionally exploit structural data.
 {\bf Ensembler's} modular architecture facilitates the implementation of alternative alignment approaches, and we plan to implement some of these in future versions, to allow exploration of the influence of different alignment methods on model quality.
@@
 
-\subsection*{Refinement of models}
+\subsection{Refinement of models}
 
@@
 
-\subsection*{Solvation and NPT equilibration}
+\subsection{Solvation and NPT equilibration}
 
@@
 
-% TODO where do these go?
-% Distribution: conda package or github source
-% Software is open source and designed to be extensible as far as possible.
-For example, it would be very simple to implement further options for providing target and template structures, or for packaging models.
-
 \label{section:design}
@@
 As a first application of {\bf Ensembler}, we have built models for the human TK family.
-{\color{blue}[TODO: Add list of TK UniProt identifiers and gene names, probably in SI.]}
+{\color{blue}[TODO: Add list of TK UniProt identifiers and gene names. There will be a file with this info within the TK data set.]}
 TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer~\cite{krause:nejm:2005:tks-cancer}.
@@
 
-We selected all 90 TK domains annotated in UniProt as targets, using the following command:
-\shellcmd{gather\_targets -{}-query `family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes' -{}-uniprot\_domain\_regex `\^{}Protein kinase(?!; truncated)(?!; inactive)'}\\
-The {\tt `reviewed:yes'} expression in the UniProt query ensured that only UniProt entries which have been reviewed by a human expert were included in the results.
-The {\tt -{}-uniprot\_domain\_regex} regular expression resulted in the selection of domains annotated "Protein kinase", "Protein kinase 1", and "Protein kinase 2", while excluding domains "Protein kinase; truncated", "Protein kinase; inactive", "Alpha-type protein kinase", and many types of non-kinase domain.
-We selected all available structures of protein kinase domains (of any species) as templates, using the following command:
-\shellcmd{gather\_templates -{}-gather\_from uniprot -{}-query `domain:"Protein kinase" AND reviewed:yes' -{}-uniprot\_domain\_regex `\^{}Protein kinase(?!; truncated)(?!; inactive)'}\\
-This returned 4433 templates, giving a total of \num{398970} target-template pairs.
+We selected all human TK domains annotated in UniProt as targets, and all available structures of protein kinase domains (of any species) as templates, using the commands shown in Box~\ref{box:tk-gather-targets-templates}.
+This returned 90 target sequences and 4433 template structures, giving a total of \num{398970} target-template pairs.
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
+
+\begin{codebox*}[tbp]
+\scriptsize
+\begin{Verbatim}[frame=single]
+ensembler gather_targets --query `family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes'
+                         --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
+ensembler gather_templates --gather_from uniprot --query `domain:"Protein kinase" AND reviewed:yes'
+                           --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
+\end{Verbatim}
+\caption{{\bf Ensembler command-line functions used to select targets and templates.}
+The commands retrieve target and template data by querying UniProt.
+The query string provided to the {\tt gather\_targets} command selects all human tyrosine protein kinases which have been reviewed by a curator, while the query string provided to the {\tt gather\_templates} command selects all reviewed protein kinases of any species.
+The {\tt-{}-uniprot\_domain\_regex} flag is used to select a subset of the domains belonging to the returned UniProt protein entries, by matching the domain annotations against a given regular expression.
+In this example, domains of type "Protein kinase", "Protein kinase 1", and "Protein kinase 2" were selected, while excluding many other domain types such as "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
+Target selection simply entails the selection of sequences corresponding to each matching UniProt domain.
+Template selection entails the selection of the sequences and structures of any PDB entries corresponding to the matching UniProt domains.
+}
+\label{box:tk-gather-targets-templates}
+\end{codebox*}
 
@@
 
-The number of models which survived each stage are shown in Fig.~\ref{figpipeline}, indicating that the greatest attrition occurred during the modeling stage.
+The number of models which survived each stage are shown in Fig.~\ref{figure:pipeline}, indicating that the greatest attrition occurred during the modeling stage.
 The number of refined models for each target ranged from \num{4005} to \num{4248}, with a median of \num{4160}, mean of \num{4150}, and standard deviation of \num{60}.
-Fig.~\ref{figpipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
+Fig.~\ref{figure:pipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
 
@@
 We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
-These ranged from -14180 kT to -3590 kT, with a median of -9533 kT, mean of -9564 kT, and a standard deviation of 1058 kT.
+These ranged from -14180 kT to -3590 kT, with a median of -9533 kT, mean of -9564 kT, and a standard deviation of 1058 kT (with a simulation temperature of 300~K).
 The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
@@
     The plotted distributions have been smoothed using kernel density estimation.
+    Refinement simulations were carried out at the default temperature of 300~K.
   }
@@
     giving some indication the diversity of conformations generated by Ensembler.
-    The models for each target were divided into three sequence identity ranges (as in Fig.
-\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed to select three clusters from each.
+    The models for each target were divided into three sequence identity ranges (as in Fig.~\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed (using the msmbuilder clustering package~\cite{msmbuilder}) to select three clusters from each.
     The models shown are the centroids of each cluster.
@@
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
-Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to $k$-medoids clustering to pick three representative models from each sequence identity range.
-{\color{red}[JDC: Explain how $k$-medoids clustering was done either here or in figure caption.]}
+Models were first stratified into three ranges, based on the structure of the sequence identity distribution (Fig.~\ref{figure:sequence-identity-distribution}), then subjected to RMSD-based $k$-medoids clustering (using the msmbuilder clustering package~\cite{msmbuilder}) to pick three representative models from each sequence identity range.
 Each model is colored and given a transparency based on the sequence identity between the target and template sequence.
@@
 
-The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub,\\
-\url{http://github.com/choderalab/ensembler} \\
-The latest release of {\bf Ensembler} can be installed via the {\tt conda} package manager for Python [\url{http://conda.pydata.org}]:
-%\shellcmd{conda install -c https://conda.binstar.org/omnia ensembler}
-\shellcmd{conda config --add channels https://conda.binstar.org/omnia}
-\shellcmd{conda install ensembler}\\
+The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub (\href{http://github.com/choderalab/ensembler}{github.com/choderalab/ensembler}).
+The latest release can be installed via the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
 The latest source can be downloaded from the GitHub repository, which also contains up-to-date instructions for building and installing the code.
-{\color{blue}[TODO: Add link to documentation. What about example inputs for generating the results in this paper? Put this in the TK model data set?]}
+Documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
+{\color{blue}[TODO: Make the commands used to generate the TK data available. Perhaps in the TK model data set.]}
+
+\begin{codebox}[tbp]
+\scriptsize
+\begin{Verbatim}[frame=single]
+conda config -add channels https://conda.binstar.org/omnia
+conda install ensembler
+\end{Verbatim}
+\caption{{\bf Ensembler installation using {\tt conda}.}
+}
+\label{box:installation}
+\end{codebox}
 
@@
 
-% Limitations:
-% protonation states
-% structural ions, e.g. Mg2+
-% prosthetic groups, e.g. heme
-% cofactors, e.g. ATP
-% PTMs - phosphorylation, methylation, glycosylation
-% long loops
-
-% TODO future directions
 % TODO what's on the Ensembler 2 roadmap?
-% TODO MCCE?
-
-% TODO CITE: Shan Shaw Proton-dependent switch Abl1 PNAS 2009
 
@@
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
-{\color{red}[Add PBG, ?SMH, ?KAB support statements.]}
+{\color{blue}[Add PBG, ?SMH, ?KAB support statements.]}
 
@@
 
-\section{Figures}
+% \section{Figures}
 

### 2015-04-13 manuscript/ms.tex
Minor

@@
 Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
-{\color{blue}[TODO: Add URL of where to get the code and TK models here]}
 
@@
 As a first application of {\bf Ensembler}, we have built models for the human TK family.
-{\color{blue}[TODO: Add list of TK UniProt identifiers and gene names. There will be a file with this info within the TK data set.]}
 TKs (and protein kinases in general) play important roles in many cellular processes and are involved in a number of types of cancer~\cite{krause:nejm:2005:tks-cancer}.
@@
 Documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
-{\color{blue}[TODO: Make the commands used to generate the TK data available. Perhaps in the TK model data set.]}
 
@@
 DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
-{\color{blue}[Add PBG, ?SMH, ?KAB support statements.]}
 

### 2015-04-13 manuscript/ms.tex
Updated support statements

@@
 {\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
-The specific command-line flags and API details discussed in this paper correspond to the version [TODO] release (TODO: link).
 Up-to-date documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
@@
 The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
-DLP and this work was supported in part by the generous support of a Louis V.~Gerstner Young Investigator Award.
+JDC, KAB, and DLP acknowledge partial support from NIH grant P30 CA008748.
+JDC and DLP also acknowledge the generous support of a Louis V.~Gerstner Young Investigator Award.
+KAB was also supported in part by Starr Foundation grant I8-A8-058.
+PBG acknowledges partial funding support from the Weill Cornell Graduate School of Medical Sciences.
 

### 2015-04-15 manuscript/ms.tex
Fixed DFG-flip citation

@@
 These protonation states can have important effects on biological processes.
-For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism [CITE: Abl1 DFG flip evidence]---is controlled by protonation of the aspartate~\cite{shan:pnas:2009:abl}.
+For example, long timescale MD simulations have suggested that the conformation of the DFG motif of the TK Abl1---believed to be an important regulatory mechanism~\cite{nagar:cell:2003:abl-1opl}---is controlled by protonation of the aspartate~\cite{shan:pnas:2009:abl}.
 Currently, protonation states are assigned simply based on pH (a user-controllable parameter).

### 2015-06-05 manuscript/ms.tex
Updated funding acknowledgements
Completed manuscript dataset

@@
   The red labels indicate the corresponding text description provided for each stage in the Design and Implementation section.
-  On the right, the number of viable models surviving each stage of the pipeline is shown for modeling all 90 tyrosine kinases (\emph{All TKs}) and representative individual tyrosine kinases (\emph{SRC} and \emph{ABL}).
+  On the right, the number of viable models surviving each stage of the pipeline is shown for the 93 target TK domains and for two representative individual TK domains (\emph{SRC} and \emph{ABL}).
   Typical timings on a computer cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to illustrate resource requirements per model for modeling the entire set of tyrosine kinases.
@@
 We selected all human TK domains annotated in UniProt as targets, and all available structures of protein kinase domains (of any species) as templates, using the commands shown in Box~\ref{box:tk-gather-targets-templates}.
-This returned 90 target sequences and 4433 template structures, giving a total of \num{398970} target-template pairs.
+This returned 93 target sequences and 4433 template structures, giving a total of \num{412269} target-template pairs.
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
@@
 To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
-Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination with out error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
+Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
 There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
 
-Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{373513} (94\%) surviving models across all TKs.
+Following loop remodeling, the {\bf Ensembler} pipeline was performed up to and including the implicit solvent MD refinement stage, which completed with \num{389067} (94\%) surviving models across all TKs.
 To obtain statistics for the solvation stage without generating a sizeable amount of coordinate data (with solvated PDB coordinate files taking up about 0.9 MB each), the {\tt solvate} subcommand was performed for two representative individual kinases (\emph{Src} and \emph{Abl1}).
@@
 The number of models which survived each stage are shown in Fig.~\ref{figure:pipeline}, indicating that the greatest attrition occurred during the modeling stage.
-The number of refined models for each target ranged from \num{4005} to \num{4248}, with a median of \num{4160}, mean of \num{4150}, and standard deviation of \num{60}.
+The number of refined models for each target ranged from \num{4046} to \num{4289}, with a median of \num{4185}, mean of \num{4184}, and standard deviation of \num{57}.
 Fig.~\ref{figure:pipeline} also indicates the typical timing achieved on a cluster for each stage, showing that the {\tt build\_models} and {\tt refine\_implicit\_md} stages are by far the most compute-intensive.
 
-Each model generated about 116 kB of file data (up to and including the implicit solvent MD refinement stage), totalling 0.5 GB per TK target or 41 GB for all 90 TKs.
-The data generated per model breaks down as 39 kB for the output from the modeling stage (without saving Modeller restraints files, which are about 397 kB per model) and 77 kB for the implicit solvent MD refinement stage.
+The files generated for each model (up to and including the implicit solvent MD refinement stage) totaled $\sim$116~kB in size, totalling 0.5~GB per TK target or 42~GB for all~93 targets.
+The data generated per model breaks down as 39~kB for the output from the modeling stage (without saving Modeller restraints files, which are about 397~kB per model) and 77~kB for the implicit solvent MD refinement stage.
 
@@
     \caption{{\bf Distribution of RMSDs to all TK catalytic domain models relative to the model derived from the highest sequence identity template.}    
-    Distributions are built from data from all 90 TK targets.
+    Distributions are built from data from all 93 TK domain targets.
     To better illustrate how conformational similarity depends on sequence identity, the lower plot illustrates the distributions as stratified into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
@@
 To evaluate the variety of template sequence similarities relative to each target sequence, we calculated sequence identity distributions, as shown in Fig.~\ref{figure:sequence-identity-distribution}.
-This suggests an intuitive division into three categories, with \num{307753} models in the 0--35\% sequence identity range, \num{69922} models in the 35--55\% range, and \num{4893} models in the 55--100\% range.
+This suggests an intuitive division into three categories, with \num{355712} models in the 0--35\% sequence identity range, \num{51330} models in the 35--55\% range, and \num{5227} models in the 55--100\% range.
 We then computed the RMSD distributions for the models created for each target (relative to the model derived from the template with highest sequence identity) Fig.~\ref{figure:rmsd-distributions}, to assess the diversity of conformations captured by the modeling pipeline.
@@
 We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
-These ranged from -14180 kT to -3590 kT, with a median of -9533 kT, mean of -9564 kT, and a standard deviation of 1058 kT (with a simulation temperature of 300~K).
+These ranged from -14180~kT to -3160~kT, with a median of -9501~kT, mean of -9418~kT, and a standard deviation of 1198~kT (with a simulation temperature of 300~K).
 The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
-Of the \num{25457} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
+Of the \num{4973} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
 
@@
 The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
+All authors acknowledge support from the Sloan Kettering Institute.
 JDC, KAB, and DLP acknowledge partial support from NIH grant P30 CA008748.

### 2015-06-05 data/scripts/gen-dataset-for-publication.py

@@ -1,0 +1,59 @@
+import os
+import shutil
+import pandas as pd
+
+
+def mk_models_data_file(traj_csv_filepath, rmsd_csv_filepath, output_csv_filepath):
+    traj_df = pd.read_csv(traj_csv_filepath)
+    rmsd_df = pd.read_csv(rmsd_csv_filepath)
+    merged_df = pd.merge(traj_df, rmsd_df, on='templateid')
+    merged_df.set_index('templateid', inplace=True)
+    merged_df.rename(columns={'seqid_x': 'seqid'}, inplace=True)
+    merged_df.to_csv(output_csv_filepath, columns=['seqid', 'rmsd'], index_label='templateid')
+    return merged_df
+
+
+if __name__ == '__main__':
+    targetids = open('TKs.txt').read().splitlines()
+
+    shutil.copy('targets_general_data/targets-data.txt', '../dataset-for-publication/')
+    shutil.copy('targets_general_data/targets-data.csv', '../dataset-for-publication/')
+    shutil.copy('TKs.txt', '../dataset-for-publication/targets-IDs.txt')
+    shutil.copy('templates_general_data/templates-data.txt', '../dataset-for-publication/')
+    shutil.copy('templates_general_data/templates-data.csv', '../dataset-for-publication/')
+
+    if not os.path.exists('../dataset-for-publication/models'):
+        os.mkdir('../dataset-for-publication/models')
+
+    for targetid in targetids:
+        src_dir = os.path.join('models', targetid)
+        dest_dir = os.path.join('../dataset-for-publication', 'models', targetid)
+        if not os.path.exists(dest_dir):
+            os.mkdir(dest_dir)
+
+        src_filenames = [
+            'topol-renumbered-implicit.pdb',
+            'traj-refine_implicit_md.xtc',
+        ]
+        dest_filenames = [
+            'topology.pdb',
+            'models.xtc',
+        ]
+
+        for src_filename, dest_filename in zip(src_filenames, dest_filenames):
+            shutil.copy(
+                os.path.join(
+                    src_dir,
+                    src_filename
+                ),
+                os.path.join(
+                    dest_dir,
+                    dest_filename
+                )
+            )
+
+        merged_df = mk_models_data_file(
+            os.path.join(src_dir, 'traj-refine_implicit_md-data.csv'),
+            os.path.join(src_dir, 'rmsds.csv'),
+            os.path.join(dest_dir, 'models-data.csv'),
+        )

### 2015-06-05 data/targets_general_data/mkdata.py

@@ -1,5 +1,5 @@
 from ensembler.initproject import GatherTargetsFromUniProt
-from ensembler.UniProt import get_uniprot_xml
+from ensembler.uniprot import get_uniprot_xml
 import pandas as pd
 
 # ============
@@ -60,8 +60,10 @@ df['UniProt recommended name'] = uniprot_recommended_names
 df['UniProt domain description'] = gtfup.domain_descriptions
 df['Organism'] = ['human'] * len(df)
 
-df.to_csv('targets_data.csv')
+df.sort(columns=['Ensembler target id'], inplace=True)
+
+df.to_csv('targets-data.csv')
 
 table_cols = ['Ensembler target id', 'UniProt entry name', 'Residue span (UniProt coords; 1-based)', 'HGNC HUGO symbol', 'UniProt recommended name', 'UniProt domain description', 'Organism']
-with open('targets_data.txt', 'w') as tablefile:
+with open('targets-data.txt', 'w') as tablefile:
     tablefile.write(df.to_string(columns=table_cols))

### 2015-06-05 data/templates_general_data/gen-templates-datafile.py

@@ -1,0 +1,143 @@
+import os
+import Bio.SeqIO
+import Bio.SubsMat.MatrixInfo
+import Bio.pairwise2
+import pandas as pd
+import re
+from ensembler.core import sequnwrap, xpath_match_regex_case_sensitive
+from ensembler.uniprot import get_uniprot_xml
+from ensembler.initproject import extract_template_pdbchains_from_uniprot_xml
+
+templates_full_seq_fasta_filepath = 'templates_general_data/templates-full-seq.fa'
+templates_full_seq = list(Bio.SeqIO.parse(templates_full_seq_fasta_filepath, 'fasta'))
+
+templateids = []
+seqs = []
+uniprot_mnemonics = []
+domain_ids = []
+pdbids = []
+chainids = []
+
+for template in templates_full_seq:
+    templateid = template.id
+    seq = template.seq
+
+    match = re.match('(.*_.*)_D([0-9]+)_(....)_(.)', templateid)
+    uniprot_mnemonic, domain_id, pdbid, chainid = match.groups()
+    domain_id = int(domain_id)
+
+    templateids.append(templateid)
+    seqs.append(str(seq))
+    uniprot_mnemonics.append(uniprot_mnemonic)
+    domain_ids.append(domain_id)
+    pdbids.append(pdbid)
+    chainids.append(chainid)
+
+df = pd.DataFrame({
+    'Ensembler template id': templateids,
+    'Template sequence': seqs,
+    'UniProt entry name': uniprot_mnemonics,
+    'domain id': domain_ids,
+    'PDB ID': pdbids,
+    'PDB chain ID': chainids,
+})
+
+unique_uniprot_mnemonics = list(set(uniprot_mnemonics))
+uniprot_seqs = {}
+hgnc_symbols = {}
+uniprot_recommended_names = {}
+gene_names = {}
+organisms = {}
+domain_descriptions_by_uniprot_entry = {}
+resi_spans_by_pdbchain = {}
+
+for uniprot_mnemonic in unique_uniprot_mnemonics:
+    uniprot_search = 'mnemonic:{0}'.format(uniprot_mnemonic)
+    uniprot_xml = get_uniprot_xml(uniprot_search)
+
+    entry = uniprot_xml.find('entry')
+    uniprot_seqs[uniprot_mnemonic] = sequnwrap(entry.find('sequence').text).replace('X', '')
+
+    organism_id = int(entry.find('organism/dbReference[@type="NCBI Taxonomy"]').get('id'))
[…]

### 2015-06-05 dataset-for-publication/commands.sh
Update to dataset

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+ensembler init
+ensembler gather_targets --query 'family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes' --uniprot_domain '^Protein kinase(?!; truncated)(?!; inactive)'
+ensembler gather_templates --gather_from uniprot --query 'domain:"Protein kinase" AND reviewed:yes' --uniprot_domain_regex '^Protein kinase(?!; truncated)(?!; inactive)'
+ensembler loopmodel
+ensembler align
+ensembler build_models
+ensembler cluster
+ensembler refine_implicit

### 2015-06-05 manuscript/ms.tex
Updated rmsds

@@
 A number of mature software packages and forcefields are now available, and much recent progress has been driven by advances in computing architecture.
-For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, Copernicus~\cite{pronk:2011:copernicus}, and GPUGrid~\cite{buch:2010:gpugrid}, allow scalability on an unprecedented level.
+For example, many MD packages are now able to exploit GPUs~\cite{eastman:jctc:2012:openmm,walker:jctc:2013:amber-gpu}, which provide greatly improved simulation efficiency per unit cost relative to CPUs, while distributed computing platforms such as Folding@home~\cite{shirts-pande:science:2000:fah}, Copernicus~\cite{pronk:2011:copernicus,pronk:2015:copernicus}, and GPUGrid~\cite{buch:2010:gpugrid}, allow scalability on an unprecedented level.
 In parallel, methods for building human-understandable models of protein dynamics from noisy simulation data, such as Markov state modeling (MSM) approaches, are now reaching maturity~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2011:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
@@
 The module could easily be extended to add methods for preparing models for other purposes.
-For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
+For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus,pronk:2015:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
 

### 2015-06-05 figures/rmsddist/rmsddist2.py

@@ -66,7 +66,7 @@ plt.setp(ax, yticks=[])
 
 legend = plt.legend(title='Sequence identity (%)', fontsize='xx-small')
 plt.setp(legend.get_title(),fontsize='x-small')
-plt.xlim(0,40)
+plt.xlim(0,30)
 plt.xlabel('RMSD ($\AA$)')
 sns.despine(left=True)
 
@@ -97,7 +97,7 @@ sns.kdeplot(df_has_model.rmsd*10, label=label, bw=2.5, shade=True, color='k')
 ax = plt.gca()
 plt.setp(ax, yticks=[])
 
-plt.xlim(0,40)
+plt.xlim(0,30)
 plt.xlabel('RMSD ($\AA$)')
 sns.despine(left=True)
 

### 2015-06-18 dataset-for-publication/commands.sh
Updated dataset README and commands.sh

@@ -1,5 +1,8 @@
 #!/bin/bash
 
+conda create -c https://conda.binstar.org/omnia -n ensembler1.0 python=2.7 ensembler=1.0 --yes
+source activate ensembler1.0
+
 ensembler init
 ensembler gather_targets --query 'family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes' --uniprot_domain '^Protein kinase(?!; truncated)(?!; inactive)'
 ensembler gather_templates --gather_from uniprot --query 'domain:"Protein kinase" AND reviewed:yes' --uniprot_domain_regex '^Protein kinase(?!; truncated)(?!; inactive)'

### 2015-06-19 manuscript/ms.tex
Updated to ms

@@
 It is compatible with Linux and OS X.
-The latest release can be installed via the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
+The latest release can be installed \emph{via} the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
 
@@
 
-{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
+{\bf Ensembler} is written in Python, and can be used \emph{via} a command-line tool ({\tt ensembler}) or \emph{via} a flexible Python API to allow integration of its components into other applications.
+All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/tree/v1.0}{version 1.0 release of Ensembler}.
 Up-to-date documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
@@
   Typical timings on a computer cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to illustrate resource requirements per model for modeling the entire set of tyrosine kinases.
-  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution via MPI reduces wall clock time nearly linearly.
+  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution \emph{via} MPI reduces wall clock time nearly linearly.
   }
@@
 
-The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---via a UniProt search query.
+The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---\emph{via} a UniProt search query.
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} is used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
@@
 Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g.~{\tt -{}-query 2H8H,1Y57}.
-Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
+Specific PDB chain IDs can optionally also be selected \emph{via} the {\tt -{}-chainids} flag.
 The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts})~\cite{velankar:2013:sifts}, which provides residue-level mappings between PDB and UniProt entries.
@@
 
-Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
+Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided \emph{via} the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
 We expect that in certain cases, pre-building template loops with Rosetta {\tt loopmodel} prior to the main modeling stage (with Modeller) may result in improved model quality.
@@
 The templates were derived from 3028 individual PDB entries and encompassed 23 different species, with 3634 template structures from human kinase constructs.
+
+The resultant models are available as part of a supplementary dataset which can be accessed \emph{via} the digital object identifier (DOI) \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}.
 
@@
 The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub (\href{http://github.com/choderalab/ensembler}{github.com/choderalab/ensembler}).
-The latest release can be installed via the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
+The latest release can be installed \emph{via} the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
 This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
@@
 Documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
+
+A supplementary dataset can also be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
+This contains the TK models described in the \ref{section:results} section, general information on the targets and templates, plus a script and instructions for regenerating the same dataset.
 

### 2015-06-21 data/scripts/archive-dataset-for-publication.py
Added script for archiving dataset

@@ -1,0 +1,55 @@
+import os
+import subprocess
+
+cwd = os.getcwd()
+
+os.chdir('..')
+
+dataset_dir = 'ensembler_dataset'
+
+toplevel_filenames = [
+    'templates-data.txt',
+    'templates-data.csv',
+    'targets-data.txt',
+    'targets-data.csv',
+    'targets-IDs.txt',
+    'commands.sh',
+]
+toplevel_filepaths = [os.path.join(dataset_dir, fname) for fname in toplevel_filenames]
+
+subprocess.check_output([
+    'tar',
+    'cvf',
+    'ensembler_dataset-datafiles.tar',
+] + toplevel_filepaths)
+
+models_dir = os.path.join(dataset_dir, 'models')
+targetids = os.listdir(models_dir)
+targetids = sorted(targetids)
+
+targets_to_tar = [[]]
+tarsize = 0
+for targetid in targetids:
+    targetdir = os.path.join(models_dir, targetid)
+    targetdir_size = sum([
+        os.path.getsize(os.path.join(targetdir, f)) for f in os.listdir(targetdir)
+    ])
+    tarsize += targetdir_size
+    if tarsize > 999999999:
+        targets_to_tar.append([])
+        tarsize = 0
+    targets_to_tar[-1].append(targetid)
+
+for targets in targets_to_tar:
+    first_letters = targets[0].split('_')[0]
+    last_letters = targets[-1].split('_')[0]
+    tarfile = 'ensembler_dataset-models-{0}-{1}.tar'.format(first_letters, last_letters)
+    targetdirs = [os.path.join(models_dir, targetid) for targetid in targets]
+
+    subprocess.check_output([
+        'tar',
+        'cvf',
+        tarfile,
+    ] + targetdirs)
+
+os.chdir(cwd)

### 2015-06-22 manuscript/ms.tex
Minor updates to ms

@@
 \emph{Keywords: molecular dynamics simulation; comparative modeling; distributed simulation}
-
-% TODO expand to all human protein kinases?
 
@@
 
-% TODO? The API is highly flexible, allowing control over many important parameters. Appropriate choices have been set as default values.
-
-% TODO? other tools for analyzing project etc.
-
 \label{section:design}
@@
 
-The resultant models are available as part of a supplementary dataset which can be accessed \emph{via} the digital object identifier (DOI) \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}.
+The resultant models are available as part of a supplementary dataset which can be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
 
@@
 
-% TODO what's on the Ensembler 2 roadmap?
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-06-23 manuscript/ms.tex
Typo fix

@@
 This is implemented using the OpenMM molecular simulation toolkit~\cite{eastman:jctc:2012:openmm}, chosen for its flexible Python API, and high performance GPU-acclerated simulation code.
-The simulation is run for a default of 100 ps, which in our example applications has been sufficient to filter out poor models (i.e.~those with atomic overlaps unresolved by energy minimization, which result in an unstable simulation), as well as helping to relax model conformations.
+The simulation is run for a default of 100~ps, which in our example applications has been sufficient to filter out poor models (i.e.~those with atomic overlaps unresolved by energy minimization, which result in an unstable simulation), as well as helping to relax model conformations.
 As discussed in the Results section, our example application of the {\bf Ensembler} pipeline to the human tyrosine kinase family indicated that of the models which failed implicit solvent MD refinement, the vast majority failed within the first 1~ps of simulation.
@@
 The simulation length can also be controlled \emph{via} the {\tt -{}-simlength} flag, and many other important simulation parameters can be controlled from either the API or CLI (\emph{via} the {\tt -{}-api\_params} flag).
-The default values are set as follows---timestep: 2~ps; temperature: 300~K; Langevin collision rate: 20~ps$^{-1}$; pH (used by OpenMM for protonation state assignment): 7.
+The default values are set as follows---timestep: 2~fs; temperature: 300~K; Langevin collision rate: 20~ps$^{-1}$; pH (used by OpenMM for protonation state assignment): 7.
 We also draw attention to a recent paper which indicates that lower Langevin collision rates may result in faster phase space exploration~\cite{basconi:jctc:2013:collision-rates}.
@@
 The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
-Of the \num{4973} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1 ps of simulation.
+Of the \num{4973} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1~ps of simulation.
 

### 2015-06-25 manuscript/ms.tex
Minor edits

@@
 It is compatible with Linux and OS X.
-The latest release can be installed \emph{via} the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
+The latest release can be installed via the {\tt conda} package manager, and the latest source can be downloaded from \url{https://github.com/choderalab/ensembler}.
 
@@
 
-{\bf Ensembler} is written in Python, and can be used \emph{via} a command-line tool ({\tt ensembler}) or \emph{via} a flexible Python API to allow integration of its components into other applications.
+{\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
 All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/tree/v1.0}{version 1.0 release of Ensembler}.
@@
   Typical timings on a computer cluster (containing Intel Xeon E5-2665 2.4GHz hyperthreaded processors and NVIDIA GTX-680 or GTX-Titan GPUs) is reported to illustrate resource requirements per model for modeling the entire set of tyrosine kinases.
-  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution \emph{via} MPI reduces wall clock time nearly linearly.
+  Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution via MPI reduces wall clock time nearly linearly.
   }
@@
 
-The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---\emph{via} a UniProt search query.
+The {\tt ensembler} command-line tool allows targets to be selected from UniProt---a freely accessible resource for protein sequence and functional data (\href{http://www.uniprot.org/}{uniprot.org})~\cite{uniprot:2015}---via a UniProt search query.
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} is used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
@@
 Selection of templates from the PDB simply requires passing a list of PDB IDs as a comma-separated string, e.g.~{\tt -{}-query 2H8H,1Y57}.
-Specific PDB chain IDs can optionally also be selected \emph{via} the {\tt -{}-chainids} flag.
+Specific PDB chain IDs can optionally also be selected via the {\tt -{}-chainids} flag.
 The program retrieves structures from the PDB server, as well as associated data from the SIFTS service (\href{http://www.ebi.ac.uk/pdbe/docs/sifts/}{www.ebi.ac.uk/pdbe/docs/sifts})~\cite{velankar:2013:sifts}, which provides residue-level mappings between PDB and UniProt entries.
@@
 
-Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided \emph{via} the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
-We expect that in certain cases, pre-building template loops with Rosetta {\tt loopmodel} prior to the main modeling stage (with Modeller) may result in improved model quality.
+Unresolved template residues can optionally be modeled into template structures with the {\tt loopmodel} subcommand, which employs a kinematic closure algorithm provided via the {\tt loopmodel} tool of the Rosetta software suite~\cite{qian:nature:2007:modeller,wang:jmb:2007:modeller}.
+We expect that in certain cases, pre-building template loops with Rosetta {\tt loopmodel} prior to the main modeling stage (with MODELLER) may result in improved model quality.
 Loop remodeling may fail for a small proportion of templates due to spatial constraints imposed by the original structure; the subsequent modeling step thus automatically uses the remodeled version of a template if available, but otherwise falls back to using the non-remodeled version.
@@
 
-Modeling is performed using the automodel function of the Modeller software package~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller} to rapidly generate a single model of the target sequence from each template structure.
-Modeller uses simulated annealing cycles along with a minimal forcefield and spatial restraints---generally Gaussian interatomic probability densities extracted from the template structure with database-derived statistics determining the distribution width---to rapidly generate candidate structures of the target sequence from the provided template sequence~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
+Modeling is performed using the automodel function of the MODELLER software package~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller} to rapidly generate a single model of the target sequence from each template structure.
+MODELLER uses simulated annealing cycles along with a minimal forcefield and spatial restraints---generally Gaussian interatomic probability densities extracted from the template structure with database-derived statistics determining the distribution width---to rapidly generate candidate structures of the target sequence from the provided template sequence~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}.
 
-While Modeller's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
+While MODELLER's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
 As such, we implemented pairwise alignment functionality using the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
@@
 We also note that alternative approaches could be used for the alignment stage.
-For example, multiple sequence alignment algorithms~\cite{thompson:2011:msa-review}, allow alignments to be guided using sequence data from across the entire protein family of interest, while (multiple) structural alignment algorithms such as Modeller's {\tt salign} routine~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}, PROMALS3D~\cite{pei:2008:promals3d}, and Expresso and 3DCoffee~\cite{expresso,3dcoffee}, can additionally exploit structural data.
+For example, multiple sequence alignment algorithms~\cite{thompson:2011:msa-review}, allow alignments to be guided using sequence data from across the entire protein family of interest, while (multiple) structural alignment algorithms such as MODELLER's {\tt salign} routine~\cite{fiser:prot-sci:2000:modeller,sali:jmb:1993:modeller}, PROMALS3D~\cite{pei:2008:promals3d}, and Expresso and 3DCoffee~\cite{expresso,3dcoffee}, can additionally exploit structural data.
 {\bf Ensembler's} modular architecture facilitates the implementation of alternative alignment approaches, and we plan to implement some of these in future versions, to allow exploration of the influence of different alignment methods on model quality.
@@
 To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
-The restraints used by Modeller could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
+The restraints used by MODELLER could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
 This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 kB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
@@
 Any of the other force fields or implicit water models implemented in OpenMM can be specified using the {\tt -{}-ff} and {\tt -{}-water\_model} flags respectively.
-The simulation length can also be controlled \emph{via} the {\tt -{}-simlength} flag, and many other important simulation parameters can be controlled from either the API or CLI (\emph{via} the {\tt -{}-api\_params} flag).
+The simulation length can also be controlled via the {\tt -{}-simlength} flag, and many other important simulation parameters can be controlled from either the API or CLI (via the {\tt -{}-api\_params} flag).
 The default values are set as follows---timestep: 2~fs; temperature: 300~K; Langevin collision rate: 20~ps$^{-1}$; pH (used by OpenMM for protonation state assignment): 7.
@@
 The force field, water model, and simulation length can again be specified using the {\tt -{}-ff}, {\tt -{}-water\_model}, and {\tt -{}-simlength} flags respectively.
-Further simulation parameters can be controlled \emph{via} the API or \emph{via} the CLI {\tt -{}-api\_params} flag.
+Further simulation parameters can be controlled via the API or via the CLI {\tt -{}-api\_params} flag.
 Pressure control is performed with a Monte Carlo barostat as implemented in OpenMM, with a default pressure of 1~atm and a period of 50 timesteps.
@@
 {\bf Ensembler} provides a packaging module which can be used to prepare models for other uses.
-The {\tt package\_models} subcommand currently provides functions (specified \emph{via} the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
+The {\tt package\_models} subcommand currently provides functions (specified via the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
 The module could easily be extended to add methods for preparing models for other purposes.
@@
 Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, mean 14, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, mean 7, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
-To reduce the reliance on the Modeller rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
+To reduce the reliance on the MODELLER rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
 Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
@@
 The files generated for each model (up to and including the implicit solvent MD refinement stage) totaled $\sim$116~kB in size, totalling 0.5~GB per TK target or 42~GB for all~93 targets.
-The data generated per model breaks down as 39~kB for the output from the modeling stage (without saving Modeller restraints files, which are about 397~kB per model) and 77~kB for the implicit solvent MD refinement stage.
+The data generated per model breaks down as 39~kB for the output from the modeling stage (without saving MODELLER restraints files, which are about 397~kB per model) and 77~kB for the implicit solvent MD refinement stage.
 
@@
 As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
-The models show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions inbetween (mainly models with low sequence identity).
+The models show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions in-between (mainly models with low sequence identity).
 We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
@@
 The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub (\href{http://github.com/choderalab/ensembler}{github.com/choderalab/ensembler}).
-The latest release can be installed \emph{via} the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
-This will install all dependencies except for Modeller and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
+The latest release can be installed via the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
+This will install all dependencies except for MODELLER and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
 The latest source can be downloaded from the GitHub repository, which also contains up-to-date instructions for building and installing the code.
@@
 Cofactors and post-translational modifications are also often not fully resolved in experimental structures, and endogenous cofactors are frequently substituted with other molecules to facilitate experimental structural analysis.
-Again, {\bf Ensembler} could exploit structural data from a set of homologous proteins to model in these molecules, although there will be likely be a number of challenges to overcome in the design and implementation of such functionality.
+Again, {\bf Ensembler} could exploit structural data from a set of homologous proteins to model in these molecules, although there will likely be a number of challenges to overcome in the design and implementation of such functionality.
 

### 2015-06-25 figures/superposition-src_abl/src/drawcolorbar-CoolWarmRAlpha.py
Updated Fig. 6 colorbar

@@ -1,0 +1,52 @@
+import numpy as np
+# import seaborn as sns
+import matplotlib.pyplot as plt
+from matplotlib.colors import LinearSegmentedColormap
+
+# data = np.random.uniform(0,100, [1000,1000])
+data = np.array([np.linspace(100,0,101)] * 13).T
+
+# fig = plt.figure(figsize=(1.5,4))
+# ax1 = fig.add_axes([0.05, 0.05, 0.3, 0.9])
+
+plt.imshow(data)
+
+coolwarm_r = plt.cm.coolwarm_r
+
+coolwarm_r_alpha_dict = coolwarm_r._segmentdata.copy()
+
+coolwarm_r_alpha_dict['alpha'] = [
+    (0.0, 0.0, 0.0),
+    (1.0, 1.0, 1.0)
+]
+
+coolwarm_r_alpha = LinearSegmentedColormap('coolwarm_r_alpha', coolwarm_r_alpha_dict)
+plt.register_cmap(cmap=coolwarm_r_alpha)
+
+
+# cmap = sns.plt.cm.coolwarm_r
+# norm = sns.mpl.colors.Normalize(vmin=0, vmax=100)
+# fig = sns.plt.figure(figsize=(1.5,4))
+# ax1 = fig.add_axes([0.05, 0.05, 0.3, 0.9])
+# cb = sns.mpl.colorbar.ColorbarBase(ax1, cmap=cmap, norm=norm)
+# cb.ax.tick_params(labelsize=20)
+# cb.set_label('sequence identity (%)', fontsize=27.)
+# cbar = plt.colorbar()
+plt.set_cmap('coolwarm_r_alpha')
+# cbar.solids.set_edgecolor('face')
+
+ax1 = plt.gca()
+ax1.set_xticks([])
+# yticklocs, yticklabels = plt.yticks()
+yticklocs = np.array([0.,   20.,   40.,   60.,   80.,  100.])
+yticklabels = ['100', '80', '60', '40', '20', '0']
+plt.yticks(yticklocs, yticklabels)
+ax1.yaxis.tick_right()
+plt.setp(ax1.get_yticklabels(), fontsize=28.)
+
+plt.ylabel('sequence identity (%)', fontsize=28.)
+ax1.yaxis.set_label_position('right')
+
+# plt.colorbar()
+
+plt.savefig('colorbar-CoolWarmRAlpha.png', dpi=300)

### 2015-07-09 manuscript/ms.tex
Updated Fig. 6 colorbar
Added allostery MODELLER citation

@@
 \graphicspath{{../figures/}}
+
+% Do not hyphenate the following words
+\hyphenation{MODELLER}
 
@@
 There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates~\cite{moult:proteins:2014:casp10,baker:science:2001:structure-prediction-review}.
+One such piece of software, MODELLER, has also been used recently to study protein allostery by generating and refining configurational models, sampled by interpolating between two user-defined metastable structures~\cite{sali:2012:allostery-modeling}.
 
@@
     This positions K295 in the appropriate position for catalysis.
+    Note that ANP (phosphoaminophosphonic acid-adenylate ester; an analog of ATP) is only physically present in the 2SRC structure.
+    To aid visualization of the active site in 1Y57, it has been included in the rendering by structurally aligning the surrounding homologous protein residues.
   }

### 2015-07-13 manuscript/ms.tex
Minor

@@
 {\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
-All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/tree/v1.0}{version 1.0 release of Ensembler}.
+All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/tree/v1.0.2}{version 1.0.2 release of Ensembler}.
 Up-to-date documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
@@
 
-The resultant models are available as part of a supplementary dataset which can be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
+The resultant models are available as part of a supplementary dataset which can be downloaded from the Dryad Digital Repository (DOI: \href{http://datadryad.org/review?doi=doi:10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
 

### 2015-07-14 manuscript/ms.tex
Added cover letter

@@
 To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Supporting Information S\ref{si:residue-numbering-schemes}.
+We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Appendix \ref{si:residue-numbering-schemes}.
 

### 2015-08-23 manuscript/ensembler.tex
Partway through manuscript revisions and referee response letter.

--- 2015-07-14 manuscript/ms.tex
@@
 MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
-There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates~\cite{moult:proteins:2014:casp10,baker:science:2001:structure-prediction-review}.
-One such piece of software, MODELLER, has also been used recently to study protein allostery by generating and refining configurational models, sampled by interpolating between two user-defined metastable structures~\cite{sali:2012:allostery-modeling}.
+\color{red}
+There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates~\cite{moult:proteins:2014:casp10,baker:science:2001:structure-prediction-review,baker:structure:2013:rosettacm,yang:nature-methods:2015:i-tasser-suite}.
+\color{black}
 
-However, it remains difficult for researchers to exploit the full variety of available protein sequence and structural data in simulation studies, largely due to limitations in software architecture.
-For example, the set up of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure(s), addition of missing residues and atoms, solvation with explicit water and counterions (and potentially buffer components and cosolvents), choice of simulation parameters (or parameterization schemes for components where parameters do not yet exist),  system relaxation with energy minimization, and one or more short preparatory MD simulations to equilibrate the system and relax the simulation cell.
+\color{red}
+However, it remains difficult for researchers to exploit the full variety of available protein sequence data (in simulating groups of related proteins) and structural data (exploiting multiple structures for each protein and its homologs/orthologs) in simulation studies in molecular simulations, largely due to limitations in software architecture.
+For example, the preparation of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure(s), addition of missing residues and atoms, solvation with explicit water and counterions (and potentially buffer components and cosolvents), choice of simulation parameters (or parameterization schemes for components where parameters do not yet exist),  system relaxation with energy minimization, and one or more short preparatory MD simulations to equilibrate the system and relax the simulation cell.
 Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations.
 Worse still, studies (or collections of studies) that \emph{do} consider multiple proteins often suffer from the lack of consistent best practices in this preparation process, making comparisons between related proteins unnecessarily difficult.
+While notable exceptions exist---such as the Dynameomics effort of Daggett and coworkers, which has simulated 100 proteins so far using a single initial configuration for each~\cite{daggett:structure:2010:dynameomics}---it is nevertheless clear that the lack of scalability of existing pipelines is a hinderance to making facile use of all available structural data.
 
 The ability to fully exploit the large quantity of available protein sequence and structural data in biomolecular simulation studies could open up many interesting avenues for research, enabling the study of entire protein families or superfamilies within a single organism or across multiple organisms.
-The similarity between members of a given protein family could be exploited to generate arrays of conformational models, which could be used as starting configurations to aid sampling in MD simulations.
+The similarity between members of a given protein family could be exploited to generate arrays of conformational models for related sequences, which could be used as starting configurations to aid sampling in MD simulations.
+The conformations captured in structures of related members has been shown to provide useful information about the conformations accessible to all members of the family~\cite{friedland:plos-comput-biol:2009:conformational-dynamics,sali:2012:allostery-modeling}, though energetic differences between individuals will modify the populations and dynamics of individual conformational states.
 This approach would be highly beneficial for many MD methods, such as MSM construction, which require global coverage of the conformational landscape to realize their full potential, and would also be particularly useful in cases where structural data is present for only a subset of the members of a protein family.
 It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
+\color{black}
 
@@
 
+\color{red}
 We anticipate that {\bf Ensembler} will prove to be useful in a number of other ways.
 For example, the generated models could represent valuable data sets even without subsequent production simulation, allowing exploration of the conformational diversity present within the available structural data for a given protein family.
-Furthermore, the automation of simulation set up provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters.
+Furthermore, automation of simulation preparation provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters, approach to the treatment of protonation states, treatment of cofactors and structural ions, and pre-simulation refinement and equilibration procedures.
+While the current version of {\bf Ensembler} only codifies some of these choices as default parameters, its modular nature allows additional stages to be easily added in the future.
+\color{black}
 
@@
 These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates based on a search of the Protein Data Bank (PDB) or UniProt.
-A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing some degree of homology between targets and templates.
+\color{red}
+A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing homology and some degree of sequence identity between targets and templates.
+\color{black}
 
 The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\url{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
-Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.~for crystal unit cells with multiple asymmetric units) would thus give rise to multiple template structures.
+\color{red}
+Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.~for crystals with non-crystallographic symmetry giving rise to independent conformations of the protein within the asymmetric unit) would thus give rise to multiple template structures.
+\color{black}
 
@@
 
-\subsection{Modeling}
+\subsection{Alignment and comparative modeling}
 
@@
 While MODELLER's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
-As such, we implemented pairwise alignment functionality using the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}.
+\color{red}
+As such, we implemented pairwise alignment functionality using the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, though other choices of scoring matrices available within the module can be selected.
+\color{black}
 The alignments are carried out with the {\tt align} subcommand, prior to the modeling step which is carried out with the {\tt build\_models} subcommand.
@@
 
+\color{orange}
+[Clarify how nonstandard amino acids are treated at this stage.]
+\color{black}
+
 \subsection*{Filtering of nearly identical models}
@@
 
+\subsection*{{\color{red} Filtering by MolProbity score}}
+
+\color{orange}
+[Insert section about optional filtering by model quality scores, such as MolProbity or DOPE.]
+\color{black}
+
 \subsection{Refinement of models}
@@
 While long-timescale unrestrained MD simulations (on the order of 100 $\mu$s) have been found to be ineffective for recapitulating native-like conformations, possibly due to forcefield issues~\cite{raval:prot:2012:long-timescale-md-refinement}, even relatively short simulations can be useful for relaxing structural elements such as sidechain orientation~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction}.
+\color{red}
+We stress that the limited refinement by molecular simulation here is primarily intended as initial relaxation and filtering stages, where implausible models might cause simulations to immediately fail, crash, or generate implausibly high energies or unstable dynamics.
+Exploration of conformational dynamics to derive MSMs, for example, will inevitably require orders of magnitude more simulation effort---very likely tens of microseconds to milliseconds of aggregate dynamics~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
+\color{black}
 
@@
 We also draw attention to a recent paper which indicates that lower Langevin collision rates may result in faster phase space exploration~\cite{basconi:jctc:2013:collision-rates}.
+
+\color{red}
+Currently, {\bf Ensembler} only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions in a consistent way.
+\color{orange}
+[Explain how user-specified overrides can be used to specify specific protonation states or alternative/non-natural residues/mutations at this stage.]
+\color{black}
 
@@
 
-To provide a more complete evaluation of the models generated, we have analyzed two example TKs (\emph{Src} and \emph{Abl1}) in detail.
-Due to their importance in cancer, these kinases have been the subject of numerous studies, encompassing many different methodologies.
-In terms of structural data, a large number of crystal structures have been solved (with or without ligands such as nucleotide substrate or inhibitor drugs), showing the kinases in a number of different conformations.
-These two kinases are thus also interesting targets for MSM studies, with one recent study focusing on modeling the states which constitute the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}.
+\color{red}
+To provide a more detailed evaluation of the variety and utility of generated models, we have analyzed two specific TKs (\emph{Src} and \emph{Abl1}) in depth.
+Due to their importance in cancer, these kinases have been the subject of numerous detailed structural and simulation studies.
+In terms of structural data, a large number of crystal structures have been solved (with or without ligands such as nucleotide substrate mimetics or small-molecule inhibitors), revealing a variety of conformations accessible to these kinases.
+A recent large-scale MSM study has also studied the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}, while a separate study employed biased sampling techniques to dissect the role of conformational changes in selectivity and affinity of imatinib recognition of Abl~\cite{roux:pnas:2013:abl-imatinib}.
+\color{black}
 
@@
 
+{\bf \color{red} Visualizing model structural diversity.}
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
@@
 
-To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (which is commonly used in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Appendix \ref{si:residue-numbering-schemes}.
+\color{red}
+{\bf Comparison with known biochemically relevant conformations.}
+To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important order parameters for the regulation of protein kinase domain activity.
+We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Appendix \ref{si:residue-numbering-schemes}.
+\color{black}
 
@@
 
-Comparative protein modeling and MD simulation set-up can be approached in a number of different ways, with varying degrees of complexity, and there are a number of obvious additions and improvements which we plan to implement in future versions of {\bf Ensembler}.
+\color{red}
+We recognize that the current version of {\bf Ensembler} has a number of limitations that limits its domain of applicability: Support for nonnatural amino acids is currently rudimentary and confined to those already appearing in the forcefield; cofactors cannot currently be automatically modeled in; ligands, cofactors, and nonnatural amino acids cannot yet be automatically parameterized; protonation state assignment is limited to selection of the most populated state based on the intrinsic p$K_a$ or user-specified overrides; the modeling of missing loops is rudimentary, relying on the subsequent dynamics for relaxation; there is not yet support for modeling of distinct domains from different templates, or the use of multiple templates to model a single domain.
+Nevertheless, there are a great number of use cases for this first version of an automated tool for simulation preparation at the superfamily scale.
+To expand this domain of applicability, there are a number of obvious additions and improvements which we plan to implement in future versions of {\bf Ensembler}.
 
+{\bf Comparative modeling.}
+Comparative protein modeling can be approached in a number of different ways, with varying degrees of complexity.
+The comparative modeling stage of {\bf Ensembler} currently uses MODELLER, but a number of excellent alternatives---such as RosettaCM~\cite{baker:structure:2013:rosettacm} and the I-TASSER Suite~\cite{yang:nature-methods:2015:i-tasser-suite}---can be added as user-selectable alternative choices.
+Additional options could be added to allow more expensive loop-modeling approaches to be employed to handle long insertions.
+
+{\bf Protonation states.}
 Some amino acids can exist in different protonation states, depending on pH and on their local environment.
@@
 
+{\bf Cofactors, structural ions, and ligands.}
 Many proteins require the presence of various types of non-protein atoms and molecules for proper function, such as metal ions (e.g.~Mg$^{+2}$), cofactors (e.g.~ATP) or post-translational modifications (e.g.~phosphorylation, methylation, glycosylation, etc.), and we thus plan for {\bf Ensembler} to eventually have the capability to include such entities in the generated models.
@@
 
+{\bf Long insertions and deletions.}
 Another limitation with the present version of {\bf Ensembler} involves the treatment of members of a protein family with especially long residue insertions or deletions.
@@
 In principle, such insertions could be excluded from the generated models, though a number of questions would arise as to how best to approach this.
+\color{black}
 
@@
 The authors thank Nicholas M.~Levinson (University of Minnesota), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
-The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2.
+The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2, {\color{red} as well as the anonymous referees for constructive feedback on this manuscript.}
 All authors acknowledge support from the Sloan Kettering Institute.
@@
 \bibliographystyle{prsty} 
-\bibliography{ms.bib}
+\bibliography{ensembler.bib}
 

### 2015-08-23 manuscript/ensembler.tex
More edits.

@@
 
-{\bf Ensembler} provides a packaging module which can be used to prepare models for other uses.
+\color{red}
+{\bf Ensembler} provides a packaging module which can be used to prepare models for subsequent downstream use, such as the use of distributed or cluster computing resources for the generation of MSMs~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2011:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
 The {\tt package\_models} subcommand currently provides functions (specified via the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
+For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus,pronk:2015:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
+
 The module could easily be extended to add methods for preparing models for other purposes.
-For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus,pronk:2015:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
+For example, models can be exported into pseudotrajectories for the purpose of performing structural analyses across model ensembles using tools like MDTraj~\cite{mdtraj}.
+\color{orange}
+[Describe how packaging for distribution of models as trajectories works?]
+
+\color{red}
+We stress that, despite evidence suggesting that there is a correspondence between solution-state dynamics and structural diversity of related template proteins~\cite{friedland:plos-comput-biol:2009:conformational-dynamics}, all models---especially those derived from low sequence identity templates---are not necessarily representative of conformations thermally accessible to the template proteins of interest.
+Care must be exercised in the use and analysis of these models.
+\color{black}
 
@@
 
+\color{orange}
+[JDC: I think we need a plot of the number of structures available for each kinase.  This could be sorted from most structures to fewest, shown as a bar chart.]
+\color{black}
 We selected all human TK domains annotated in UniProt as targets, and all available structures of protein kinase domains (of any species) as templates, using the commands shown in Box~\ref{box:tk-gather-targets-templates}.
@@
 Furthermore, to understand the influence of sequence identity on the conformational similarities of the resulting models, the RMSD distributions were stratified based on the three sequence identity categories described above.
-This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average.
+\color{red}
+This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average, recapitulating the observation made years ago by Chothia and Lesk~\cite{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}.
+\color{black}
 
@@
 In principle, such insertions could be excluded from the generated models, though a number of questions would arise as to how best to approach this.
+
+{\bf Markov state model (MSM) construction and model utility.}
+We are actively utilizing {\bf Ensembler}-generated models to seed the construction of Markov state models (MSMs)~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
+While the observation that high sequence identity templates are likely to reflect accessible solution-phase conformations suggests that a number of these models occupy thermally accessible regions of configuration space~\cite{friedland:plos-comput-biol:2009:conformational-dynamics}, many models---especially those derived from very low sequence identity templates---are likely to be highly unrepresentative of conformations populated at equilibrium by the target protein.
+It is likely that even with hundreds of microseconds to milliseconds of aggregated dynamics, many of these poor quality models will remain trapped in inaccessible and irrelevant regions of configuration space.
+Standard approaches to MSM construction now employ an \emph{ergodic trimming} step~\cite{caflisch:jpcb:2011:msm-subgraph,msmbuilder} to prune away disconnected minor regions of configuration space, and this step is expected to be essential in the successful construction of MSMs using {\bf Ensembler}-derived models.
+
 \color{black}

### 2015-08-23 manuscript/ensembler.tex
Finished updates pass.

@@
 For example, the preparation of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure(s), addition of missing residues and atoms, solvation with explicit water and counterions (and potentially buffer components and cosolvents), choice of simulation parameters (or parameterization schemes for components where parameters do not yet exist),  system relaxation with energy minimization, and one or more short preparatory MD simulations to equilibrate the system and relax the simulation cell.
-Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations.
+Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations,  though notable exceptions exist, ssuch as the Dynameomics effort of Daggett and coworkers in which over 100 proteins have been simulated so far using a single initial configuration for each~\cite{daggett:structure:2010:dynameomics}
 Worse still, studies (or collections of studies) that \emph{do} consider multiple proteins often suffer from the lack of consistent best practices in this preparation process, making comparisons between related proteins unnecessarily difficult.
-While notable exceptions exist---such as the Dynameomics effort of Daggett and coworkers, which has simulated 100 proteins so far using a single initial configuration for each~\cite{daggett:structure:2010:dynameomics}---it is nevertheless clear that the lack of scalability of existing pipelines is a hinderance to making facile use of all available structural data.
 
@@
 
-\subsection{Refinement of models}
+\color{red}
+\subsection{Refinement of models and filtering of poor models by simulation}
 
-A number of refinement methods have been developed to help guide comparative modeling techniques toward more "native-like" and physically consistent conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}, of which MD simulations are an important example.
-While long-timescale unrestrained MD simulations (on the order of 100 $\mu$s) have been found to be ineffective for recapitulating native-like conformations, possibly due to forcefield issues~\cite{raval:prot:2012:long-timescale-md-refinement}, even relatively short simulations can be useful for relaxing structural elements such as sidechain orientation~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction}.
-\color{red}
+A number of refinement methods have been developed to help guide comparative modeling techniques toward more "native-like" and physically consistent conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}.
+Both short~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction} and long~\cite{raval:prot:2012:long-timescale-md-refinement} molecular dynamics simulations have been employed for this purpose.
+Here, we utilize short molecular dynamic simulations for two purposes: both to slightly relax the initial comparative models and to eliminate those comparative models that result in highly implausible conformations.
+This is especially critical here due to the inclusion of even very low sequence identity template structures.
 We stress that the limited refinement by molecular simulation here is primarily intended as initial relaxation and filtering stages, where implausible models might cause simulations to immediately fail, crash, or generate implausibly high energies or unstable dynamics.
@@
     \caption{{\bf Distributions for the number of missing residues in the TK templates.}
-    The upper histograms show the number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling with the {\tt loopmodel} subcommand failed (red).
-    The lower histogram shows the number of residues in each missing loop, for all templates.
+    \color{red}
+    \emph{Upper:} The number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling with the {\tt loopmodel} subcommand failed (red).
+    Templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
+    \emph{Lower:} The number of residues in each missing loop, for all templates.    
 }
@@
 Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
+\color{orange}
 There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
+[JDC: Are you sure there is a correlation here?
+I think the difference in median was just due to the fact that the red (failed remodeling) templates could \emph{not} have 0 remodeled residues, or else they would not have been subjected to remodeling.
+Or do you mean that you already excluded the 0 remodeled residue lengths from the blue (all templates) before computing the median, in which case you should mention that here?]
+\color{black}
 
@@
 
+{\bf Template remodeling.}
+The lack of crystallographically-resolved regions of template structures presents a challenge to deriving structures from these templates by comparative modeling, especially in kinases, where loops are frequently unresolved.
+Improvements over the Rosetta-based strategy described here are likely possible, especially given the number of modeling failures observed in the template refinement stage (Fig.~\ref{figure:loopmodel-nmissing-residues}).
+An alternative approach could be to re-refine complete-chain template structures to the experimentally-derived electron density or scattering data deposited in the RCSB using methods capable of exploiting the scattering data and crystallographic symmetry~\cite{schnieders:biophys-j:2010:x-ray-refinement}.
+Even if definitive placement of these unresolved regions is impossible, plausible locations constrained by weak scattering data and strong steric exclusion of crystallographic neighbors may provide a great deal of useful information, especially when combined with forcefield priors~\cite{schnieders:jctc:2011:space-group-pme}.
+
 {\bf Comparative modeling.}
@@
 Cofactors and post-translational modifications are also often not fully resolved in experimental structures, and endogenous cofactors are frequently substituted with other molecules to facilitate experimental structural analysis.
-Again, {\bf Ensembler} could exploit structural data from a set of homologous proteins to model in these molecules, although there will likely be a number of challenges to overcome in the design and implementation of such functionality.
+Future extensions to {\bf Ensembler} could transfer cofactor and ion coordinates from homologous proteins in which these components are resolved.
+
+{\bf Post-translationally modified amino acids and other molecules without forcefield parameters.}
+A major challenge in the preparation of simulations of proteins of interest is the wide variety of post-translational modifications possible that are often functionally or structurally relevant.
+Often, forcefields lack parameters for these residues, or for other cofactors or ligands that might be vital to probing the relevant structural dynamics of these systems.
+While tools such as Antechamber~\cite{gaff1,gaff2} can rapidly generate small molecule parameters in an automated manner, the parameterization of polymeric residues or covalently attached cofactors is much more challenging.
+In addition, small molecule forcefields are generally tied to specific corresponding protein and nucleic acid forcefields, meaning that different procedures may be needed to generate consistent parameters.
 

### 2015-09-25 manuscript-double-spaced/ms.tex
Added double spaced manuscript (original submission version)

--- 2015-07-14 manuscript/ms.tex
@@
 % PRL look and style (easy on the eyes)
-\documentclass[aps,pre,twocolumn,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
+%\documentclass[aps,pre,twocolumn,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
 % Two-column style (for submission/review/editing)
-%\documentclass[aps,prl,preprint,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
+\documentclass[aps,prl,preprint,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
 
@@
 
-  \caption{{\bf Diagrammatic representation of the stages of the Ensembler pipeline and illustrative statistics for modeling all human tyrosine kinase catalytic domains.}
+  \caption{\footnotesize {\bf Diagrammatic representation of the stages of the Ensembler pipeline and illustrative statistics for modeling all human tyrosine kinase catalytic domains.}
   On the left, the various stages of the {\bf Ensembler} pipeline are shown.
@@
 To retrieve target sequences from UniProt, the subcommand {\tt gather\_targets} is used with the {\tt -{}-query} flag followed by a UniProt query string conforming to the same syntax as the search function available on the UniProt website.
-For example, {\tt -{}-query `mnemonic:SRC\_HUMAN'} would select the full-length human Src sequence, while the query shown in Box~\ref{box:tk-gather-targets-templates} would select all human tyrosine protein kinases which have been reviewed by a human curator.
+For example, \path{--query} \path{`mnemonic:SRC_HUMAN'} would select the full-length human Src sequence, while the query shown in Box~\ref{box:tk-gather-targets-templates} would select all human tyrosine protein kinases which have been reviewed by a human curator.
 In this way, the user may select a single protein, many proteins, or an entire superfamily from UniProt.
@@
 In many cases, it will be desirable to build models of an isolated protein domain, rather than the full-length protein.
-The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-uniprot\_domain\_regex} flag. 
+The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the \path{--uniprot_domain_regex} flag. 
 For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
 The regular expression shown in Box~\ref{box:tk-gather-targets-templates} selects only domains of the first three types.
-If the {\tt -{}-uniprot\_domain\_regex} flag is used, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~{\tt JAK1\_HUMAN\_D0}, {\tt JAK1\_HUMAN\_D1}).
+If the {\tt -{}-uniprot\_domain\_regex} flag is used, target identifiers are set with the form {\tt [UniProt mnemonic]\_D[domain index]}, where the latter part represents a 0-based index for the domain---necessary because a single target protein may contain multiple domains of interest (e.g.~\path{JAK1_HUMAN_D0}, \path{JAK1_HUMAN_D1}).
 
@@
 
-Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional {\tt -{}-uniprot\_domain\_regex} flag allows selection of individual domains with a regular expression string (Box~\ref{box:tk-gather-targets-templates}).
+Selection of templates from UniProt proceeds in a similar fashion as for target selection; the {\tt -{}-query} flag is used to select full-length proteins from UniProt, while the optional \path{--uniprot_domain_regex} flag allows selection of individual domains with a regular expression string (Box~\ref{box:tk-gather-targets-templates}).
 The returned UniProt data for each protein includes a list of associated PDB chains and their residue spans, and this information is used to select template structures, using the same method as for template selection from the PDB.
@@
 To minimize file storage requirements, {\bf Ensembler} uses the Python {\tt gzip} library to apply compression to all sizeable text files from the modeling stage onwards.
-The restraints used by MODELLER could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag ({\tt -{}-write\_modeller\_restraints\_file}) for optionally saving these restraints to file.
+The restraints used by MODELLER could potentially be used in alternative additional refinement schemes, and {\bf Ensembler} thus provides a flag (\path{--write_modeller_restraints_file}) for optionally saving these restraints to file.
 This option is turned off by default, as the restraint files are relatively large (e.g.~$\sim$400 kB per model for protein kinase domain targets), and are not expected to be used by the majority of users.
@@
 {\bf Ensembler} provides a packaging module which can be used to prepare models for other uses.
-The {\tt package\_models} subcommand currently provides functions (specified via the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
+The {\tt package\_models} subcommand currently provides functions (specified via the \path{--package_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
 The module could easily be extended to add methods for preparing models for other purposes.
@@
 \end{Verbatim}
-\caption{{\bf Ensembler command-line functions used to select targets and templates.}
+\caption{\footnotesize {\bf Ensembler command-line functions used to select targets and templates.}
 The commands retrieve target and template data by querying UniProt.
@@
 \begin{figure}[tbp]
-    \includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
-    \caption{{\bf Distributions for the number of missing residues in the TK templates.}
+    \includegraphics[width=0.7\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
+    \caption{\footnotesize {\bf Distributions for the number of missing residues in the TK templates.}
     The upper histograms show the number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling with the {\tt loopmodel} subcommand failed (red).
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
+    \includegraphics[width=0.7\columnwidth]{seqid_dist/seqid_dist.pdf}
 
-    \caption{{\bf Template-target sequence identity distribution for human tyrosine kinase catalytic domains.}
+    \caption{\footnotesize {\bf Template-target sequence identity distribution for human tyrosine kinase catalytic domains.}
     Sequence identities are calculated from all pairwise target-template alignments, where targets are human kinase catalytic domain sequences and templates are all kinase catalytic domains from any organism with structures in the PDB, as described in the text.
@@
     {\label{figure:rmsd-distribution-joint}%
-       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
+       \includegraphics[height=0.43\textwidth]{rmsddist/rmsddist2-joint.pdf}
     }
@@
     {\label{figure:rmsd-distribution-by-sequence-identity}%
-       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
+       \includegraphics[height=0.43\textwidth]{rmsddist/rmsddist2.pdf}
     }
 
-    \caption{{\bf Distribution of RMSDs to all TK catalytic domain models relative to the model derived from the highest sequence identity template.}    
+    \caption{\footnotesize {\bf Distribution of RMSDs to all TK catalytic domain models relative to the model derived from the highest sequence identity template.}    
     Distributions are built from data from all 93 TK domain targets.
@@
 \begin{figure}[tbp]
-    \includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
+    \includegraphics[width=0.6\columnwidth]{energies/energies.pdf}
 
-    \caption{{\bf Distribution of final energies from implicit solvent MD refinement of TK catalytic domain models.}
+    \caption{\footnotesize {\bf Distribution of final energies from implicit solvent MD refinement of TK catalytic domain models.}
     To illustrate how the energies are affected by sequence identity, the models are separated into three sequence identity classes: high identity (55--100\%), moderate identity (35--55\%), and remote identity (0--35\%).
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+    \includegraphics[width=0.8\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
 
-    \caption{{\bf Superposition of clustered models of Src and Abl1.}
+    \caption{\footnotesize {\bf Superposition of clustered models of Src and Abl1.}
     Superposed renderings of nine models each for Src and Abl1, 
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
+    \includegraphics[width=0.8\textwidth]{residue_pair_distances/src/src_ref_structures}
 
-    \caption{{\bf Two structures of Src, indicating certain residues involved in activation.}
+    \caption{\footnotesize {\bf Two structures of Src, indicating certain residues involved in activation.}
     In the inactive state, E310 forms a salt bridge with R409.
@@
     \subfloat[Src]{\label{figure:src-distances}%
-       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_src.png}
+       \includegraphics[height=0.44\textwidth]{residue_pair_distances/src_abl_combined/distances_src.png}
     }
     \subfloat[Abl]{\label{figure:abl-distances}%
-       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_abl.png}
+       \includegraphics[height=0.44\textwidth]{residue_pair_distances/src_abl_combined/distances_abl.png}
     }
 
-    \caption{{\bf Src and Abl1 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
+    \caption{\footnotesize {\bf Src and Abl1 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
     Two Src structures (PDB entries 1Y57~\cite{cowan-jacob:2005:1y57} and 2SRC~\cite{xu:1999:2src}) are projected onto the plots for reference, representing active and inactive states respectively.
@@
 \end{Verbatim}
-\caption{{\bf Ensembler installation using {\tt conda}.}
+\caption{{\bf \footnotesize Ensembler installation using {\tt conda}.}
 }
@@
 \begin{alltt}
+\footnotesize
 1     MLEICLKLVG CKSKKGLSSS SSCYLEEALQ RPVASDFEPQ GLSEAARWNS KENLLAGPSE    60
@@
 
+\break
+
 \subsubsection*{Sequences for human and chicken Src, aligned using Clustal Omega}
@@
 \begin{alltt}
+\footnotesize
 SRC_HUMAN   1     MGSNKSKPKD ASQRRRSLEP AENVHGAGGG AFPASQTPSK PASADGHRGP SAAFAPAAAE    60

### 2015-09-25 manuscript/ensembler.tex
Minor edits to ms and response letter to fix typos

@@
 For example, the preparation of a biomolecular simulation is typically performed manually, encompassing a series of fairly standard (yet time-consuming) steps such as the choice of protein sequence construct and starting structure(s), addition of missing residues and atoms, solvation with explicit water and counterions (and potentially buffer components and cosolvents), choice of simulation parameters (or parameterization schemes for components where parameters do not yet exist),  system relaxation with energy minimization, and one or more short preparatory MD simulations to equilibrate the system and relax the simulation cell.
-Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations,  though notable exceptions exist, ssuch as the Dynameomics effort of Daggett and coworkers in which over 100 proteins have been simulated so far using a single initial configuration for each~\cite{daggett:structure:2010:dynameomics}
+Due to the laborious and manual nature of this process, simulation studies typically consider only one or a few proteins and starting configurations,  though notable exceptions exist, such as the Dynameomics effort of Daggett and coworkers in which over 100 proteins have been simulated so far using a single initial configuration for each~\cite{daggett:structure:2010:dynameomics}.
 Worse still, studies (or collections of studies) that \emph{do} consider multiple proteins often suffer from the lack of consistent best practices in this preparation process, making comparisons between related proteins unnecessarily difficult.
@@
 \color{red}
-We recognize that the current version of {\bf Ensembler} has a number of limitations that limits its domain of applicability: Support for nonnatural amino acids is currently rudimentary and confined to those already appearing in the forcefield; cofactors cannot currently be automatically modeled in; ligands, cofactors, and nonnatural amino acids cannot yet be automatically parameterized; protonation state assignment is limited to selection of the most populated state based on the intrinsic p$K_a$ or user-specified overrides; the modeling of missing loops is rudimentary, relying on the subsequent dynamics for relaxation; there is not yet support for modeling of distinct domains from different templates, or the use of multiple templates to model a single domain.
+We recognize that the current version of {\bf Ensembler} has a number of limitations that bound its domain of applicability: support for nonnatural amino acids is currently rudimentary and confined to those already appearing in the forcefield; cofactors cannot currently be automatically modeled in; ligands, cofactors, and nonnatural amino acids cannot yet be automatically parameterized; protonation state assignment is limited to selection of the most populated state based on the intrinsic p$K_a$ or user-specified overrides; the modeling of missing loops is rudimentary, relying on the subsequent dynamics for relaxation; there is not yet support for modeling of distinct domains from different templates, or the use of multiple templates to model a single domain.
 Nevertheless, there are a great number of use cases for this first version of an automated tool for simulation preparation at the superfamily scale.

### 2015-10-13 manuscript/ensembler.tex
Manuscript revisions by DLP

@@
 
-\color{orange}
-[Clarify how nonstandard amino acids are treated at this stage.]
+\color{blue}
+At this time, the alignment and modeling functions cannot be used to model non-standard amino acids, though we plan to be able to provide this functionality in future versions.
+Note that the Ensembler functions for target and template selection only include standard amino acids which match the UniProt canonical isoform sequence, and thus any set of targets and templates selected this way should be compatible with the Ensembler alignment and modeling functions.
 \color{black}
@@
 A minimum distance cutoff (which defaults to 0.6~\AA) is used to retain only a single model per cluster.
-
-\subsection*{{\color{red} Filtering by MolProbity score}}
-
-\color{orange}
-[Insert section about optional filtering by model quality scores, such as MolProbity or DOPE.]
-\color{black}
 
@@
 
+\color{blue}
+For some studies, it may be useful to specify the protonation states of individual amino acids, rather than rely only on automatic protonation state assignment by OpenMM.
+The user can do this by listing the residue numbers and their protonation states in a configuration file ({\tt manual\_overrides.yaml}).
+The necessary formatting for the configuration file is specified in the software documentation, and a template file is written when initializing an {\bf Ensembler} project.
+Protonation states are specified by naming the appropriate residue variant type in the force field, e.g. {\tt `ASH'} for an aspartic acid residue, as opposed to the aspartate base {\tt `ASP'}.
+Any residues which do not have specific protonation states listed in the configuration file will have protonation states assigned automatically by OpenMM.
 \color{red}
-Currently, {\bf Ensembler} only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions in a consistent way.
-\color{orange}
-[Explain how user-specified overrides can be used to specify specific protonation states or alternative/non-natural residues/mutations at this stage.]
+Note that {\bf Ensembler} currently only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions in a consistent way.
 \color{black}
@@
 
+\subsection*{{\color{blue} Model validation with MolProbity score}}
+
+\color{blue}
+{\bf Ensembler} provides a function for validating model quality using MolProbity~\cite{molprobity:2007,molprobity:2010}---a widely used tool for validation of protein models, which provides a numerical score derived from features such as steric clashes between atoms, bond geometry, Ramachandran angles, sidechain rotamer outliers, backbone deviations, and the presence of cis-peptides.
+This function is accessed via the {\tt validate} subcommand, which for a given target will output a text file containing a list of model IDs sorted by validation score.
+The optional {\tt -{}-modeling\_stage} flag specifies which of the three main Ensembler modeling stages to validate---the initial comparative modeling stage, the implicit MD refinement stage, or the explicit MD refinement stage.
+If this flag is not used, Ensembler defaults to selecting the latest stage for which models have been generated.
+The output text file can be used to filter models based on validation score, for example by using the {\tt package\_models} subcommand.
+Protein model validation is a challenging problem and an active area of research for many groups, including the developers of {\bf Ensembler}.
+We plan to implement further validation methods in future versions of Ensembler.
+
+\color{black}
+
 \subsection*{Packaging}
@@
 The {\tt package\_models} subcommand currently provides functions (specified via the {\tt -{}-package\_for} flag) for compressing models in preparation for data transfer, or for organizing them with the appropriate directory and file structure for production simulation on the distributed computing platform Folding@home~\cite{shirts-pande:science:2000:fah}.
+The module could easily be extended to add methods for preparing models for other purposes.
 For example, production simulations could alternatively be run using Copernicus~\cite{pronk:2011:copernicus,pronk:2015:copernicus}---a framework for performing parallel adaptive MD simulations--- or GPUGrid~\cite{buch:2010:gpugrid}---a distributing computing platform which relies on computational power voluntarily donated by the owners of nondedicated GPU-equipped computers.
 
-The module could easily be extended to add methods for preparing models for other purposes.
-For example, models can be exported into pseudotrajectories for the purpose of performing structural analyses across model ensembles using tools like MDTraj~\cite{mdtraj}.
-\color{orange}
-[Describe how packaging for distribution of models as trajectories works?]
+\color{blue}
+An important use of the packaging stage is to filter models based on model quality.
+At the current time, the available filtering options are based on either target-template sequence identity or MolProbity validation score.
+The {\tt package\_models} subcommand includes optional flags for specifying a sequence identity cutoff (so that only models with a target-template sequence identity above the specified percentage are chosen), a MolProbity validation score cutoff (to choose only models with lower validation scores, which indicate better model quality), or a MolProbity validation score percentile (to choose only models with validation scores lower than the value at the given percentile).
+
+Models can also be exported into trajectory files for the purpose of performing structural analyses across model ensembles using tools like MDTraj~\cite{mdtraj}.
+This is done using the {\tt mktraj} subcommand, which writes model coordinates for a given target to a Gromacs~\cite{berendsen:1995:gromacs,lindahl:2001:gromacs} XTC format trajectory (chosen for its wide usage and data compression).
+Each frame in the trajectory represents a single model, and  models are sorted in descending order of target-template sequence identity.
+Also output for each target are a PDB coordinate file (for use as a topology input file) and a CSV file containing model IDs (in the same order as the frames in the trajectory file) and other data such as target-template sequence identity.
+Using the {\tt -{}-modeling\_stage} flag, models can be selected from any of three Ensembler modeling stages - after the initial comparative modeling stage, after implicit MD refinement, or after explicit MD refinement.
+If this flag is not used, Ensembler defaults to selecting the latest stage for which models have been generated.
 
@@
 
-\color{orange}
-[JDC: I think we need a plot of the number of structures available for each kinase.  This could be sorted from most structures to fewest, shown as a bar chart.]
-\color{black}
 We selected all human TK domains annotated in UniProt as targets, and all available structures of protein kinase domains (of any species) as templates, using the commands shown in Box~\ref{box:tk-gather-targets-templates}.
@@
 The resultant models are available as part of a supplementary dataset which can be downloaded from the Dryad Digital Repository (DOI: \href{http://datadryad.org/review?doi=doi:10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
+
+\color{blue}
+Figure~\ref{figure:nstructures-per-tk-target} shows the number of PDB structures available for each of the 93 target TK domains.
+This illustrates how a wide range of structural data is available for certain TK domains, while for others there is none.
+{\bf Ensembler} thus helps to overcome this unequal distribution of structural information when building protein models for simulation by exploiting homologous structural data from a wider range of protein kinase domains and species.
+\color{black}
+
+\begin{figure*}[tbp]
+    \includegraphics[width=\textwidth]{nstructures_per_tk_target/nstructures_per_tk_target.pdf}
+    \caption{{\bf Number of PDB structures available for each TK target.}
+    \color{blue}
+    Data is shown for each of the 93 TK kinase domains.
+The labels indicate the UniProt name for the target protein plus an index for the kinase domain (three of the selected proteins have two kinase domains).
+Each PDB chain is counted individually, and only chains which contain the target domain are counted.
+}
+    \label{figure:nstructures-per-tk-target}
+\end{figure*}
 
@@
 Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
-\color{orange}
-There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful.
-[JDC: Are you sure there is a correlation here?
-I think the difference in median was just due to the fact that the red (failed remodeling) templates could \emph{not} have 0 remodeled residues, or else they would not have been subjected to remodeling.
-Or do you mean that you already excluded the 0 remodeled residue lengths from the blue (all templates) before computing the median, in which case you should mention that here?]
+\color{blue}
+There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful (when excluding templates with no missing residues).
 \color{black}
@@
 This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average, recapitulating the observation made years ago by Chothia and Lesk~\cite{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}.
+
+\color{blue}
+We also used the {\tt ensembler validate} subcommand to subject the refined models to analysis with MolProbity~\cite{molprobity:2007,molprobity:2010}.
+The MolProbity scores varied from 0.92 to 4.80, with a median of 3.84, a mean of 3.22, and a standard deviation of 1.07.
+Lower numbers represent better quality models.
+When stratified by the same sequence identity ranges as above, the mean scores were as follows: 2.96 (55--100\% sequence identity), 3.13 (35--55\% sequence identity), 3.24 (0--35\% sequence identity).
+This indicates that models with lower target-template sequence identities tend to be lower quality according to MolProbity analysis, as would be expected.
+
 \color{black}
-
 We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
 These ranged from -14180~kT to -3160~kT, with a median of -9501~kT, mean of -9418~kT, and a standard deviation of 1198~kT (with a simulation temperature of 300~K).
-The distributions---stratified using the same sequence identity ranges as above---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
+The distributions---stratified using the same sequence identity ranges again---are plotted in Fig.~\ref{figure:energies-implicit}, indicating that higher sequence identity templates tend to result in slightly lower energy models.
 Of the \num{4973} models which failed to complete the implicit refinement MD stage, all except 9 failed within the first 1~ps of simulation.
@@
 These three residues are also well conserved~\cite{kannan:jmb:2005:kinase-regulation}, and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm, foda:nat-comm:2015:src, ozkirimli:prot-sci:2008:src}.
-As such, we have projected the {\bf Ensembler} models for \emph{Src} and \emph{Abl1} onto a space consisting of the distances between these two residue pairs (Fig.~\ref{figure:pair-distances}).
-The models show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions in-between (mainly models with low sequence identity).
+As such, we have plotted the distance between these two residue pairs for the {\bf Ensembler} models for \emph{Src} and \emph{Abl1}, as well as \emph{Flt4} (Fig.~\ref{figure:pair-distances}).
+The models all show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions in-between (mainly models with low sequence identity).
 We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
+\color{blue}
+The \emph{Flt4} models are of particular note, as there are no available crystal structures of the kinase domain of this TK protein (which is involved in tumor angiogenesis and lymphangiogenesis~\cite{Su:2006:flt4cancer}), yet the models generated here include structural motifs which are conserved and of known importance to other proteins of the same family.
+\color{black}
 
@@
     \subfloat[Src]{\label{figure:src-distances}%
-       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_src.png}
+       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_src.png}
     }
     \subfloat[Abl]{\label{figure:abl-distances}%
-       \includegraphics[height=0.42\textwidth]{residue_pair_distances/src_abl_combined/distances_abl.png}
+       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_abl.png}
+    }
+    \subfloat[Flt4]{\label{figure:flt4-distances}%
+       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_flt4.png}
     }
 
-    \caption{{\bf Src and Abl1 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
+    \caption{{\bf Src, Abl1, and Flt4 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
+    \color{blue}
     Two Src structures (PDB entries 1Y57~\cite{cowan-jacob:2005:1y57} and 2SRC~\cite{xu:1999:2src}) are projected onto the plots for reference, representing active and inactive states respectively.
@@
 Often, forcefields lack parameters for these residues, or for other cofactors or ligands that might be vital to probing the relevant structural dynamics of these systems.
-While tools such as Antechamber~\cite{gaff1,gaff2} can rapidly generate small molecule parameters in an automated manner, the parameterization of polymeric residues or covalently attached cofactors is much more challenging.
+While tools such as Antechamber~\cite{gaff,gaff2} can rapidly generate small molecule parameters in an automated manner, the parameterization of polymeric residues or covalently attached cofactors is much more challenging.
 In addition, small molecule forcefields are generally tied to specific corresponding protein and nucleic acid forcefields, meaning that different procedures may be needed to generate consistent parameters.

### 2015-10-13 figures/loopmodel_analysis/plot_dists.py

@@ -8,6 +8,10 @@ df = pd.read_csv('loopmodel_data.csv', converters={'missing_resi_spans': literal
 successful = df[df.no_missing_residues == False][df.successful == True]
 unsuccessful = df[df.no_missing_residues == False][df.successful == False]
 
+print df.describe()
+print successful.describe()
+print unsuccessful.describe()
+
 missing_resis_spans = np.array([x for sublist in df.missing_resi_spans for x in sublist])
 
 # ====

### 2015-10-13 figures/molprobity_dist/plot.py

@@ -14,7 +14,7 @@ import matplotlib.pyplot as plt
 data_dir = os.path.join('..', '..', 'data')
 targets = open(os.path.join(data_dir, 'TKs.txt')).read().splitlines()
 
-# Get data from rmsds.csv and validation file, and do left join on templateid
+# Get data from rmsds.csv and validation file, and do inner join on templateid
 
 for target in targets:
     models_target_dir = os.path.join(data_dir, 'models', target)
@@ -50,6 +50,11 @@ seqid_ranges_labels = deepcopy(seqid_ranges)
 
 seqid_ranges[-1][-1] = 101.
 
+print combined_dfs.molprobity_score.describe()
+
+for seqid_range in seqid_ranges:
+    print combined_dfs[combined_dfs.seqid >= seqid_range[0]][combined_dfs.seqid < seqid_range[1]].molprobity_score.describe()
+
 # ====================
 # Plotting
 # ====================
@@ -62,7 +67,6 @@ palette = sns.color_palette('GnBu_d', n_colors=3)[::-1]
 
 for i, seqid_range in enumerate(seqid_ranges):
     df_seqid_bin = combined_dfs[combined_dfs.seqid >= seqid_range[0]][combined_dfs.seqid < seqid_range[1]]
-    print seqid_range, len(df_seqid_bin)
     label = '%.0f-%.0f ($n=$%d)' % (seqid_ranges_labels[i][0], seqid_ranges_labels[i][1], len(df_seqid_bin))
 
     sns.kdeplot(df_seqid_bin.molprobity_score, label=label, color=palette[i], shade=True)

### 2015-10-13 manuscript/ensembler.tex

@@
 
-\color{blue}
+\color{red}
 At this time, the alignment and modeling functions cannot be used to model non-standard amino acids, though we plan to be able to provide this functionality in future versions.
@@
 
-\color{blue}
+\color{red}
 For some studies, it may be useful to specify the protonation states of individual amino acids, rather than rely only on automatic protonation state assignment by OpenMM.
@@
 \color{red}
-Note that {\bf Ensembler} currently only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions in a consistent way.
+Note that {\bf Ensembler} currently only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions not provided by the forcefield.
 \color{black}
@@
 
-\color{blue}
-{\bf Ensembler} provides a function for validating model quality using MolProbity~\cite{molprobity:2007,molprobity:2010}---a widely used tool for validation of protein models, which provides a numerical score derived from features such as steric clashes between atoms, bond geometry, Ramachandran angles, sidechain rotamer outliers, backbone deviations, and the presence of cis-peptides.
+\color{red}
+{\bf Ensembler} provides a function for validating model quality and filtering models using MolProbity~\cite{molprobity:2007,molprobity:2010}---a widely used tool for validation of protein models, which provides a numerical score derived from features such as steric clashes between atoms, bond geometry, Ramachandran angles, sidechain rotamer outliers, backbone deviations, and the presence of cis-peptides.
 This function is accessed via the {\tt validate} subcommand, which for a given target will output a text file containing a list of model IDs sorted by validation score.
@@
 
-\color{blue}
+\color{red}
 An important use of the packaging stage is to filter models based on model quality.
@@
 
-\color{blue}
+\color{red}
 Figure~\ref{figure:nstructures-per-tk-target} shows the number of PDB structures available for each of the 93 target TK domains.
-This illustrates how a wide range of structural data is available for certain TK domains, while for others there is none.
+While a number of experimental structures are available for some TK domains, many TKs have few or no structures. 
 {\bf Ensembler} thus helps to overcome this unequal distribution of structural information when building protein models for simulation by exploiting homologous structural data from a wider range of protein kinase domains and species.
@@
     \caption{{\bf Number of PDB structures available for each TK target.}
-    \color{blue}
+    \color{red}
     Data is shown for each of the 93 TK kinase domains.
@@
 Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
-\color{blue}
+\color{red}
 There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful (when excluding templates with no missing residues).
@@
 
-\color{blue}
+\color{red}
 We also used the {\tt ensembler validate} subcommand to subject the refined models to analysis with MolProbity~\cite{molprobity:2007,molprobity:2010}.
@@
 We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
-\color{blue}
-The \emph{Flt4} models are of particular note, as there are no available crystal structures of the kinase domain of this TK protein (which is involved in tumor angiogenesis and lymphangiogenesis~\cite{Su:2006:flt4cancer}), yet the models generated here include structural motifs which are conserved and of known importance to other proteins of the same family.
+\color{red}
+The \emph{Flt4} models (Fig.~\ref{figure:src-ref-structures}c) are of particular note, as there are no available crystal structures of the kinase domain of this TK protein (which is involved in tumor angiogenesis and lymphangiogenesis~\cite{Su:2006:flt4cancer}), yet the models generated here include structural motifs which are conserved and of known importance to other proteins of the same family.
 \color{black}
@@
     \caption{{\bf Src, Abl1, and Flt4 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
-    \color{blue}
+    \color{red}
     Two Src structures (PDB entries 1Y57~\cite{cowan-jacob:2005:1y57} and 2SRC~\cite{xu:1999:2src}) are projected onto the plots for reference, representing active and inactive states respectively.

### 2015-10-16 manuscript/ensembler.tex
Updated Ensembler release version and installation instructions

@@
 {\bf Ensembler} is written in Python, and can be used via a command-line tool ({\tt ensembler}) or via a flexible Python API to allow integration of its components into other applications.
-All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/tree/v1.0.2}{version 1.0.2 release of Ensembler}.
+All command-line and API information in this article refers to the \href{https://github.com/choderalab/ensembler/releases/tag/v1.0.6}{version 1.0.6 release of Ensembler}.
 Up-to-date documentation can be found at \href{http://ensembler.readthedocs.org/en/latest/}{ensembler.readthedocs.org}.
@@
 The code for {\bf Ensembler} is hosted on the collaborative open source software development platform GitHub (\href{http://github.com/choderalab/ensembler}{github.com/choderalab/ensembler}).
-The latest release can be installed via the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the two commands shown in Box~\ref{box:installation}.
-This will install all dependencies except for MODELLER and Rosetta, which are not available through the {\tt conda} package manager, and thus must be installed separately by the user.
+The latest release can be installed via the {\tt conda} package manager for Python (\href{http://conda.pydata.org}{conda.pydata.org}), using the commands shown in Box~\ref{box:installation}.
+This will install all required dependencies, though a license must first be obtained for MODELLER.
+The optional dependencies Rosetta and MolProbity are not available through the {\tt conda} package manager, and thus must be installed separately by the user according to the instructions for those packages..
 The latest source can be downloaded from the GitHub repository, which also contains up-to-date instructions for building and installing the code.
@@
 \begin{Verbatim}[frame=single]
-conda config -add channels https://conda.binstar.org/omnia
+export KEY_MODELLER="Modeller license key"
+conda config -add channels https://conda.anaconda.org/omnia
+conda config -add channels https://conda.anaconda.org/salilab
 conda install ensembler

### 2015-10-19 manuscript/ensembler.tex
Minor updates

@@
 \color{red}
-A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, guaranteeing homology and some degree of sequence identity between targets and templates.
+A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, establishing homology and some degree of sequence identity between targets and templates.
 \color{black}
@@
 
-\subsection*{{\color{blue} Model validation with MolProbity score}}
+\subsection*{{\color{red} Model validation with MolProbity score}}
 
@@
 \end{Verbatim}
-\caption{{\bf Ensembler installation using {\tt conda}.}
+\caption{{\bf Ensembler installation using {\tt conda} in a {\tt bash} shell.
+A license for MODELLER must first be obtained, and the license key should be used in place of the text {\tt "Modeller license key"}.}
 }

### 2015-12-09 manuscript/ensembler.tex
Appendix converted to Figure at request of PLoS Comp Biol.

@@
     \color{red}
-    Data is shown for each of the 93 TK kinase domains.
+    Data is shown for each of the 93 TK kinase domains, sorted in order of the number of available PDB structures for each domain. 
     The labels indicate the UniProt name for the target protein plus an index for the kinase domain (three of the selected proteins have two kinase domains). 
@@
 To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important order parameters for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Appendix \ref{si:residue-numbering-schemes}.
+We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Fig.~\ref{figure:residue-numbering-schemes}.
 \color{black}
@@
 \newpage
-\appendix
-\setcounter{figure}{0}
 
-\renewcommand{\thesection}{\arabic{section}}
-\section{Sequences and residue numbering schemes for Src and Abl1}
-\label{si:residue-numbering-schemes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Sequences and residue numbering schemes for Src and Abl1
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
+\begin{figure}[tb]
 
-\subsubsection*{Human Abl1 sequence}
+\begin{centering}
+\emph{Human Abl1 sequence}
+
+\end{centering}
 
@@
 
-\subsubsection*{Sequences for human and chicken Src, aligned using Clustal Omega}
+\begin{centering}
+\emph{Sequences for human and chicken Src, aligned using Clustal Omega}
+
+\end{centering}
 
@@
 
-% \section{Figures}
+    \caption{{\bf Sequences and residue numbering schemes for Src and Abl1.}
+Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
+  }
+  \label{figure:residue-numbering-schemes}
+\end{figure}
 

### 2016-01-07 manuscript-post-acceptance-submission/ensembler.tex
Added manuscript-post-acceptance-submission

--- 2015-12-09 manuscript/ensembler.tex
@@
 \usepackage{verbatim}
-\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[draft]{hyperref}
 \usepackage[group-separator={,}]{siunitx}
@@
 \author{Daniel L. Parton}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
+Corresponding author: john.chodera@choderalab.org
+}
   %\email{daniel.parton@choderalab.org}
 \author{Patrick B. Grinaway}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
+Corresponding author: john.chodera@choderalab.org
+}
 \author{Sonya M. Hanson}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
+Corresponding author: john.chodera@choderalab.org
+}
 \author{Kyle A. Beauchamp}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
+Corresponding author: john.chodera@choderalab.org
+}
   %\email{patrick.grinaway@choderalab.org}
 \author{John D. Chodera}
- \thanks{Corresponding author}
- \email{john.chodera@choderalab.org}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065}
-
-\date{\today}
+ %\thanks{Corresponding author}
+ %\email{john.chodera@choderalab.org}
+  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
+Corresponding author: john.chodera@choderalab.org
+}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
+% \begin{abstract}
+
+\maketitle
+
+\section*{abstract}
 
@@
 
-\end{abstract}
+% \end{abstract}
 
-\maketitle
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AUTHOR SUMMARY
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Author Summary}
+\label{section:author_summary}
+
+Proteins are the workhorses of the human body, and are involved in essentially every biological process.
+Many diseases are caused by proteins malfunctioning.
+To understand how a protein functions, it is necessary to know its physical properties.
+The field of structural biology provides many techniques for determining the three-dimensional structure of a protein.
+The dynamics of a protein, i.e. the way it moves, are of equal importance, but are more difficult to uncover with traditional experimental techniques.
+Computer simulations are an effective alternative method for understanding protein dynamics, but require experimental structural information as a starting point.
+While recent advances in genomics and experimental techniques have provided a wealth of such structural data, the appropriate software for using this data effectively has been lacking.
+To tackle this problem, we have developed a software package called {\bf Ensembler}, which allows a user to automatically select appropriate experimentally derived structures for a given protein or family of proteins, and to use them to prepare a series of simulations.
+The resultant simulation data can then used to investigate the dynamics of the protein(s) in question, and their involvement in disease.
 
@@
 MSM methods in particular have the advantage of being able to aggregate data from multiple independent MD trajectories, facilitating parallelization of production simulations and thus greatly alleviating overall computational cost.
-\color{red}
 There also exist a number of mature software packages for comparative modeling of protein structures, in which a target protein sequence is modeled using one or more structures as templates~\cite{moult:proteins:2014:casp10,baker:science:2001:structure-prediction-review,baker:structure:2013:rosettacm,yang:nature-methods:2015:i-tasser-suite}.
-\color{black}
 
-\color{red}
 However, it remains difficult for researchers to exploit the full variety of available protein sequence data (in simulating groups of related proteins) and structural data (exploiting multiple structures for each protein and its homologs/orthologs) in simulation studies in molecular simulations, largely due to limitations in software architecture.
@@
 It would also aid in studying protein families known to have multiple metastable conformations---such as kinases---for which the combined body of structural data for the family may cover a large range of these conformations, while the available structures for any individual member might encompass only one or two distinct conformations.
-\color{black}
 
@@
 
-\color{red}
 We anticipate that {\bf Ensembler} will prove to be useful in a number of other ways.
@@
 While the current version of {\bf Ensembler} only codifies some of these choices as default parameters, its modular nature allows additional stages to be easily added in the future.
-\color{black}
 
@@
 \begin{figure*}[tb]
-    \includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
+    %\includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
 
@@
   Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution via MPI reduces wall clock time nearly linearly.
+  \label{figure:pipeline}
   }
-  \label{figure:pipeline}
 \end{figure*}
 
-\subsection{Target selection and retrieval}
+\subsection{Target selection and retrieval (A)}
 
@@
 
-\subsection{Template selection and retrieval}
+\subsection{Template selection and retrieval (B)}
 
@@
 These templates can be specified manually, or using the {\tt ensembler gather\_templates} subcommand to automatically select templates based on a search of the Protein Data Bank (PDB) or UniProt.
-\color{red}
 A recommended approach is to select templates from UniProt which belong to the same protein family as the targets, establishing homology and some degree of sequence identity between targets and templates.
-\color{black}
 
 The {\tt ensembler gather\_templates} subcommand provides methods for selecting template structures from either UniProt or the PDB (\url{http://www.rcsb.org/pdb}), specified by the {\tt -{}-gather\_from} flag.
-\color{red}
 Both methods select templates at the level of PDB chains---a PDB structure containing multiple chains with identical sequence spans (e.g.~for crystals with non-crystallographic symmetry giving rise to independent conformations of the protein within the asymmetric unit) would thus give rise to multiple template structures.
-\color{black}
 
@@
 
-\subsection{Template refinement}
+\subsection{Template refinement (C)}
 
@@
 
-\subsection{Alignment and comparative modeling}
+\subsection{Alignment and comparative modeling (D)}
 
@@
 While MODELLER's automodel function can generate its own alignments automatically, a standalone function was preferable for reasons of programming convenience.
-\color{red}
 As such, we implemented pairwise alignment functionality using the BioPython {\tt pairwise2} module~\cite{cock:2009:biopython}---which uses a dynamic programming algorithm---with the PAM 250 scoring matrix of Gonnet \textit{et al.}~\cite{gonnet:science:1992:exhaustive-matching}, though other choices of scoring matrices available within the module can be selected.
-\color{black}
 The alignments are carried out with the {\tt align} subcommand, prior to the modeling step which is carried out with the {\tt build\_models} subcommand.
@@
 
-\color{red}
 At this time, the alignment and modeling functions cannot be used to model non-standard amino acids, though we plan to be able to provide this functionality in future versions.
 Note that the Ensembler functions for target and template selection only include standard amino acids which match the UniProt canonical isoform sequence, and thus any set of targets and templates selected this way should be compatible with the Ensembler alignment and modeling functions.
-\color{black}
 
@@
 
-\color{red}
-\subsection{Refinement of models and filtering of poor models by simulation}
+\subsection{Refinement of models and filtering of poor models by simulation (E)}
 
@@
 Exploration of conformational dynamics to derive MSMs, for example, will inevitably require orders of magnitude more simulation effort---very likely tens of microseconds to milliseconds of aggregate dynamics~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
-\color{black}
 
@@
 
-\color{red}
 For some studies, it may be useful to specify the protonation states of individual amino acids, rather than rely only on automatic protonation state assignment by OpenMM.
@@
 Any residues which do not have specific protonation states listed in the configuration file will have protonation states assigned automatically by OpenMM.
-\color{red}
 Note that {\bf Ensembler} currently only supports residue definitions provided by the forcefield definition files---it does not yet have the ability to derive new forcefield parameters for uncommon amino acids, cofactors, or ions not provided by the forcefield.
-\color{black}
 
-\subsection{Solvation and NPT equilibration}
+\subsection{Solvation and NPT equilibration (F)}
 
@@
 
-\subsection*{{\color{red} Model validation with MolProbity score}}
+\subsection*{{Model validation with MolProbity score}}
 
-\color{red}
 {\bf Ensembler} provides a function for validating model quality and filtering models using MolProbity~\cite{molprobity:2007,molprobity:2010}---a widely used tool for validation of protein models, which provides a numerical score derived from features such as steric clashes between atoms, bond geometry, Ramachandran angles, sidechain rotamer outliers, backbone deviations, and the presence of cis-peptides.
@@
 
-\color{black}
-
 \subsection*{Packaging}
 
-\color{red}
 {\bf Ensembler} provides a packaging module which can be used to prepare models for subsequent downstream use, such as the use of distributed or cluster computing resources for the generation of MSMs~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,noe:jcp:2011:msm-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
@@
 
-\color{red}
 An important use of the packaging stage is to filter models based on model quality.
@@
 
-\color{red}
 We stress that, despite evidence suggesting that there is a correspondence between solution-state dynamics and structural diversity of related template proteins~\cite{friedland:plos-comput-biol:2009:conformational-dynamics}, all models---especially those derived from low sequence identity templates---are not necessarily representative of conformations thermally accessible to the template proteins of interest.
 Care must be exercised in the use and analysis of these models.
-\color{black}
 
@@
 
-\color{red}
 Figure~\ref{figure:nstructures-per-tk-target} shows the number of PDB structures available for each of the 93 target TK domains.
@@
 {\bf Ensembler} thus helps to overcome this unequal distribution of structural information when building protein models for simulation by exploiting homologous structural data from a wider range of protein kinase domains and species.
-\color{black}
 
 \begin{figure*}[tbp]
-    \includegraphics[width=\textwidth]{nstructures_per_tk_target/nstructures_per_tk_target.pdf}
+    %\includegraphics[width=\textwidth]{nstructures_per_tk_target/nstructures_per_tk_target.pdf}
     \caption{{\bf Number of PDB structures available for each TK target.}
-    \color{red}
     Data is shown for each of the 93 TK kinase domains, sorted in order of the number of available PDB structures for each domain. 
@@
     Each PDB chain is counted individually, and only chains which contain the target domain are counted.
+    \label{figure:nstructures-per-tk-target}
 }
-    \label{figure:nstructures-per-tk-target}
 \end{figure*}
@@
 \begin{figure}[tbp]
-    \includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
+    %\includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
     \caption{{\bf Distributions for the number of missing residues in the TK templates.}
-    \color{red}
     \emph{Upper:} The number of missing residues per template, for all templates (blue) and for only those templates for which template remodeling with the {\tt loopmodel} subcommand failed (red).
@@
     \emph{Lower:} The number of residues in each missing loop, for all templates.    
+    \label{figure:loopmodel-nmissing-residues}
 }
-    \label{figure:loopmodel-nmissing-residues}
 \end{figure}
@@
 Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
-\color{red}
 There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful (when excluding templates with no missing residues).
-\color{black}
 
@@
 
+To evaluate the variety of template sequence similarities relative to each target sequence, we calculated sequence identity distributions, as shown in Fig.~\ref{figure:sequence-identity-distribution}.
+This suggests an intuitive division into three categories, with \num{355712} models in the 0--35\% sequence identity range, \num{51330} models in the 35--55\% range, and \num{5227} models in the 55--100\% range.
+We then computed the RMSD distributions for the models created for each target (relative to the model derived from the template with highest sequence identity) Fig.~\ref{figure:rmsd-distributions}, to assess the diversity of conformations captured by the modeling pipeline.
+Furthermore, to understand the influence of sequence identity on the conformational similarities of the resulting models, the RMSD distributions were stratified based on the three sequence identity categories described above.
+This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average, recapitulating the observation made years ago by Chothia and Lesk~\cite{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
+    %\includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
 
@@
     A kernel density estimate of the target-template sequence identity probability density function is shown as a solid line with shaded region, while the corresponding cumulative distribution function is shown as a dashed line.
+  \label{figure:sequence-identity-distribution}
     }
-  \label{figure:sequence-identity-distribution}
 \end{figure}
@@
     {\label{figure:rmsd-distribution-joint}%
-       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
+       %\includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
     }
@@
     {\label{figure:rmsd-distribution-by-sequence-identity}%
-       \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
+       %\includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
     }
@@
     The plotted distributions have been smoothed using kernel density estimation.
+  \label{figure:rmsd-distributions}
   }
-  \label{figure:rmsd-distributions}
 \end{figure}
 
-To evaluate the variety of template sequence similarities relative to each target sequence, we calculated sequence identity distributions, as shown in Fig.~\ref{figure:sequence-identity-distribution}.
-This suggests an intuitive division into three categories, with \num{355712} models in the 0--35\% sequence identity range, \num{51330} models in the 35--55\% range, and \num{5227} models in the 55--100\% range.
-We then computed the RMSD distributions for the models created for each target (relative to the model derived from the template with highest sequence identity) Fig.~\ref{figure:rmsd-distributions}, to assess the diversity of conformations captured by the modeling pipeline.
-Furthermore, to understand the influence of sequence identity on the conformational similarities of the resulting models, the RMSD distributions were stratified based on the three sequence identity categories described above.
-\color{red}
-This analysis indicates that higher sequence identity templates result in models with lower RMSDs, while templates with remote sequence identities result in larger RMSDs on average, recapitulating the observation made years ago by Chothia and Lesk~\cite{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}.
-
-\color{red}
 We also used the {\tt ensembler validate} subcommand to subject the refined models to analysis with MolProbity~\cite{molprobity:2007,molprobity:2010}.
@@
 
-\color{black}
 We also analyzed the potential energies of the models at the end of the implicit solvent MD refinement stage.
@@
 \begin{figure}[tbp]
-    \includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
+    %\includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
 
@@
     Refinement simulations were carried out at the default temperature of 300~K.
+  \label{figure:energies-implicit}
   }
-  \label{figure:energies-implicit}
 \end{figure}
@@
 
-\color{red}
 To provide a more detailed evaluation of the variety and utility of generated models, we have analyzed two specific TKs (\emph{Src} and \emph{Abl1}) in depth.
@@
 A recent large-scale MSM study has also studied the activation pathway of Src~\cite{shukla-pande:nature-commun:2014:src-activation-msm}, while a separate study employed biased sampling techniques to dissect the role of conformational changes in selectivity and affinity of imatinib recognition of Abl~\cite{roux:pnas:2013:abl-imatinib}.
-\color{black}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Superposition of Src and Abl1 sequence identity classes
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    \includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
-
-    \caption{{\bf Superposition of clustered models of Src and Abl1.}
-    Superposed renderings of nine models each for Src and Abl1, 
-    giving some indication the diversity of conformations generated by Ensembler.
-    The models for each target were divided into three sequence identity ranges (as in Fig.~\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed (using the msmbuilder clustering package~\cite{msmbuilder}) to select three clusters from each.
-    The models shown are the centroids of each cluster.
-    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
-  }
-  \label{figure:superposition}
-\end{figure}
-
-{\bf \color{red} Visualizing model structural diversity.}
+{\bf Visualizing model structural diversity.}
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
@@
 
-\color{red}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIGURE: Superposition of Src and Abl1 sequence identity classes
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[tb]
+    %\includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+
+    \caption{{\bf Superposition of clustered models of Src and Abl1.}
+    Superposed renderings of nine models each for Src and Abl1, 
+    giving some indication the diversity of conformations generated by Ensembler.
+    The models for each target were divided into three sequence identity ranges (as in Fig.~\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed (using the msmbuilder clustering package~\cite{msmbuilder}) to select three clusters from each.
+    The models shown are the centroids of each cluster.
+    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
+  \label{figure:superposition}
+  }
+\end{figure}
+
 {\bf Comparison with known biochemically relevant conformations.}
@@
 We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Fig.~\ref{figure:residue-numbering-schemes}.
-\color{black}
+
+Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC)~\cite{xu:1999:2src} and active (PDB code: 1Y57)~\cite{cowan-jacob:2005:1y57} states.
+One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
+These three residues are also well conserved~\cite{kannan:jmb:2005:kinase-regulation}, and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm, foda:nat-comm:2015:src, ozkirimli:prot-sci:2008:src}.
+As such, we have plotted the distance between these two residue pairs for the {\bf Ensembler} models for \emph{Src} and \emph{Abl1}, as well as \emph{Flt4} (Fig.~\ref{figure:pair-distances}).
+The models all show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions in-between (mainly models with low sequence identity).
+We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
+The \emph{Flt4} models (Fig.~\ref{figure:src-ref-structures}c) are of particular note, as there are no available crystal structures of the kinase domain of this TK protein (which is involved in tumor angiogenesis and lymphangiogenesis~\cite{Su:2006:flt4cancer}), yet the models generated here include structural motifs which are conserved and of known importance to other proteins of the same family.
 
@@
 \begin{figure}[tb]
-    \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
+    %\includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
 
@@
     To aid visualization of the active site in 1Y57, it has been included in the rendering by structurally aligning the surrounding homologous protein residues.
+  \label{figure:src-ref-structures}
   }
-  \label{figure:src-ref-structures}
 \end{figure}
-
-Fig.~\ref{figure:src-ref-structures} shows two structures of \emph{Src} believed to represent inactive (PDB code: 2SRC)~\cite{xu:1999:2src} and active (PDB code: 1Y57)~\cite{cowan-jacob:2005:1y57} states.
-One notable feature which distinguishes the two structures is the transfer of an electrostatic interaction of E310 from R409 (in the inactive state) to K295 (in the active state), brought about by a rotation of the $\alpha$C-helix.
-These three residues are also well conserved~\cite{kannan:jmb:2005:kinase-regulation}, and a number of experimental and simulation studies have suggested that this electrostatic switching process plays a role in a regulatory mechanism shared across the protein kinase family~\cite{shukla-pande:nature-commun:2014:src-activation-msm, foda:nat-comm:2015:src, ozkirimli:prot-sci:2008:src}.
-As such, we have plotted the distance between these two residue pairs for the {\bf Ensembler} models for \emph{Src} and \emph{Abl1}, as well as \emph{Flt4} (Fig.~\ref{figure:pair-distances}).
-The models all show strong coverage of regions in which either of the electrostatic interactions is fully formed (for models across all levels of target-template sequence identity), as well as a wide range of regions in-between (mainly models with low sequence identity).
-We thus expect that such a set of models, if used as starting configurations for highly parallel MD simulation, could greatly aid in sampling of functionally relevant conformational states.
-\color{red}
-The \emph{Flt4} models (Fig.~\ref{figure:src-ref-structures}c) are of particular note, as there are no available crystal structures of the kinase domain of this TK protein (which is involved in tumor angiogenesis and lymphangiogenesis~\cite{Su:2006:flt4cancer}), yet the models generated here include structural motifs which are conserved and of known importance to other proteins of the same family.
-\color{black}
 
@@
     \subfloat[Src]{\label{figure:src-distances}%
-       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_src.png}
+       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_src.png}
     }
     \subfloat[Abl]{\label{figure:abl-distances}%
-       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_abl.png}
+       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_abl.png}
     }
     \subfloat[Flt4]{\label{figure:flt4-distances}%
-       \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_flt4.png}
+       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_flt4.png}
     }
@@
     \caption{{\bf Src, Abl1, and Flt4 models projected onto the distances between two conserved residue pairs, colored by sequence identity.}
-    \color{red}
     Two Src structures (PDB entries 1Y57~\cite{cowan-jacob:2005:1y57} and 2SRC~\cite{xu:1999:2src}) are projected onto the plots for reference, representing active and inactive states respectively.
@@
     The atom names for these atoms, according to the PDB coordinate files for both reference structures, are---Lys: NZ, CD, CE (ethylamine); Glu: OE1, CD, OE2 (carboxylate); Arg: NH1, CZ, NH2 (part of guanidine).
+  \label{figure:pair-distances}
     }
-  \label{figure:pair-distances}
 \end{figure*}
@@
 A supplementary dataset can also be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
-This contains the TK models described in the \ref{section:results} section, general information on the targets and templates, plus a script and instructions for regenerating the same dataset.
+This contains the TK models described in the Results section, general information on the targets and templates, plus a script and instructions for regenerating the same dataset.
 
@@
 
-\color{red}
 We recognize that the current version of {\bf Ensembler} has a number of limitations that bound its domain of applicability: support for nonnatural amino acids is currently rudimentary and confined to those already appearing in the forcefield; cofactors cannot currently be automatically modeled in; ligands, cofactors, and nonnatural amino acids cannot yet be automatically parameterized; protonation state assignment is limited to selection of the most populated state based on the intrinsic p$K_a$ or user-specified overrides; the modeling of missing loops is rudimentary, relying on the subsequent dynamics for relaxation; there is not yet support for modeling of distinct domains from different templates, or the use of multiple templates to model a single domain.
@@
 
-\color{black}
-
 \subsection*{Conclusion}
@@
 The authors thank Nicholas M.~Levinson (University of Minnesota), Markus A.~Seeliger (Stony Brook), Diwakar Shukla (Stanford), and Avner Schlessinger (Mount Sinai) for helpful scientific feedback on modeling kinases.
-The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2, {\color{red} as well as the anonymous referees for constructive feedback on this manuscript.}
+The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2, as well as the anonymous referees for constructive feedback on this manuscript.
 All authors acknowledge support from the Sloan Kettering Institute.
@@
 
-\bibliographystyle{prsty} 
-\bibliography{ensembler.bib}
+\begin{thebibliography}{100}
+
+\bibitem{craik:science:2009:trapping-moving-targets}
+G.~M. Lee and C.~S. Craik, Science {\bf 324},  213  (2009).
+
+\bibitem{eastman:jctc:2012:openmm}
+P. Eastman, M.~S. Friedrichs, J.~D. Chodera, R.~J. Radmer, C.~M. Bruns, J.~P.
+  Ku, K.~A. Beauchamp, T.~J. Lane, L.-P. Wang, D. Shukla, and V.~S. Pande, J.
+  Chem. Theory Comput. {\bf 9},  461  (2012).
+
+\bibitem{walker:jctc:2013:amber-gpu}
+R. Salomon-Ferrer, A.~W. G\"{o}tz, D. Poole, S.~L. Grand, and R.~C. Walker, J.
+  Chem. Theor. Comput. {\bf 9},  3878  (2013).
+
+\bibitem{shirts-pande:science:2000:fah}
+M. Shirts and V.~S. Pande, Science {\bf 290},  1903  (2000).
+
+\bibitem{pronk:2011:copernicus}
+S. Pronk, P. Larsson, I. Pouya, G.~R. Bowman, I.~S. Haque, K. Beauchamp, B.
+  Hess, V.~S. Pande, P.~M. Kasson, and E. Lindahl,  in {\em Proceedings of 2011
+  {International} {Conference} for {High} {Performance} {Computing},
+  {Networking}, {Storage} and {Analysis}}, {\em {SC} '11} (ACM, New York, NY,
+  USA, 2011), pp.\ 60:1--60:10.
+
+\bibitem{pronk:2015:copernicus}
+S. Pronk, I. Pouya, M. Lundborg, G. Rotskoff, B. Wes\'{e}n, P.~M. Kasson, and
+  E. Lindahl, Journal of Chemical Theory and Computation  (2015).
+
+\bibitem{buch:2010:gpugrid}
+I. Buch, M.~J. Harvey, T. Giorgino, D.~P. Anderson, and G. De~Fabritiis,
+  Journal of Chemical Information and Modeling {\bf 50},  397  (2010).
+
+\bibitem{pande-beauchamp-bowman:2010:methods:markov-model-review}
+V.~S. Pande, K. Beauchamp, and G.~R. Bowman, Methods {\bf 52},  99  (2010).
+
+\bibitem{noe:jcp:2011:msm-review}
+J.-H. Prinz, H. Wu, M. Sarich, B. Keller, M. Fischbach, M. Held, J.~D. Chodera,
+  C. Sch\"{u}tte, and F. No\'{e}, J. Chem. Phys. {\bf 134},  174105  (2011).
+
+\bibitem{chodera-noe:curr-opin-struct-biol:2014:msm-review}
+J.~D. Chodera and F. No\'{e}, Curr. Opin. Struct. Biol. {\bf 25},  135  (2014).
+
+\bibitem{moult:proteins:2014:casp10}
+J. Moult, K. Fidelis, A. Kryshtafovych, T. Schwede, and A. Tramontano,
+  Proteins: Structure, Function, and Bioinformatics {\bf 82},  1  (2014).
+
+\bibitem{baker:science:2001:structure-prediction-review}
+D. Baker and A. {\v{S}ali}, Science {\bf 294},  93  (2001).
+
+\bibitem{baker:structure:2013:rosettacm}
+Y. Song, F. {DiMaio}, R.~Y.-R. Wang, D. Kim, C. Miles, T. Brunette, J.
+  Thompson, and D. Baker, Structure {\bf 21},  1735  (2013).
+
+\bibitem{yang:nature-methods:2015:i-tasser-suite}
+J. Yang, R. Yan, A. Roy, D. Xu, J. Poisson, and Y. Zhang, Nature Methods {\bf
+  12},  7  (2015).
+
+\bibitem{daggett:structure:2010:dynameomics}
+M.~W. {van der Kamp}, R.~D. Schaeffer, A.~L. Jonsson, A.~D. Scouras, A.~M.
+  Simms, R.~D. Toofanny, N.~C. Benson, P.~C. Anderson, E.~D. Merkley, S.
+  Rysavy, D. Bromley, D.~A.~C. Beck, and V. Daggett, Structure {\bf 18},  423
+  (2010).
+
+\bibitem{friedland:plos-comput-biol:2009:conformational-dynamics}
+G.~D. Friedland, N.-A. Lakomek, C. Griesinger, J. Meiler, and T. Kortemme, PLoS
+  Comput. Biol. {\bf 5},  e1000393  (2009).
+
+\bibitem{sali:2012:allostery-modeling}
+P. Weinkam, J. Pons, and A. Sali, Proceedings of the National Academy of
+  Sciences of the United States of America {\bf 109},  4875  (2012).
+
+\bibitem{msmbuilder}
+K.~A. Beauchamp, G.~R. Bowman, T.~J. Lane, L. Maibaum, I.~S. Haque, and V.~S.
+  Pande, Journal of Chemical Theory and Computation {\bf 7},  3412  (2011).
+
+\bibitem{caflisch:jpcb:2011:msm-subgraph}
+R. Scalco and A. Caflisch, The Journal of Physical Chemistry. B {\bf 115},
+  6358  (2011).
+
+\bibitem{uniprot:2015}
+T.~U. Consortium, Nucleic Acids Research {\bf 43},  D204  (2015).
+
+\bibitem{velankar:2013:sifts}
+S. Velankar, J.~M. Dana, J. Jacobsen, G. van Ginkel, P.~J. Gane, J. Luo, T.~J.
+  Oldfield, C. O'Donovan, M.-J. Martin, and G.~J. Kleywegt, Nucleic Acids
+  Research {\bf 41},  D483  (2013).
+
+\bibitem{qian:nature:2007:modeller}
+B. Qian, S. Raman, R. Das, P. Bradley, A.~J. McCoy, R.~J. Read, and D. Baker,
+  Nature {\bf 450},  259  (2007).
+
+\bibitem{wang:jmb:2007:modeller}
+C. Wang, P. Bradley, and D. Baker, Journal of Molecular Biology {\bf 373},  503
+   (2007).
+
+\bibitem{fiser:prot-sci:2000:modeller}
+A. Fiser, R.~K.~G. Do, and A. {\v{S}ali}, Protein Science {\bf 9},  1753
+  (2000).
+
+\bibitem{sali:jmb:1993:modeller}
+A. {\v{S}ali} and T.~L. Blundell, Journal of Molecular Biology {\bf 234},  779
+  (1993).
+
+\bibitem{cock:2009:biopython}
+P.~J.~A. Cock, T. Antao, J.~T. Chang, B.~A. Chapman, C.~J. Cox, A. Dalke, I.
+  Friedberg, T. Hamelryck, F. Kauff, B. Wilczynski, and M.~J.~L. de~Hoon,
+  Bioinformatics (Oxford, England) {\bf 25},  1422  (2009).
+
+\bibitem{gonnet:science:1992:exhaustive-matching}
+G.~H. Gonnet, M.~A. Cohen, and S.~A. Brenner, Science {\bf 256},  1443  (1992).
+
+\bibitem{thompson:2011:msa-review}
+J.~D. Thompson, B. Linard, O. Lecompte, and O. Poch, PLoS ONE {\bf 6},  e18093
+  (2011).
+
+\bibitem{pei:2008:promals3d}
+J. Pei, B.-H. Kim, and N.~V. Grishin, Nucleic Acids Research {\bf 36},  2295
+  (2008).
+
+\bibitem{expresso}
+F. Armougom, S. Moretti, O. Poirot, S. Audic, P. Dumas, B. Schaeli, V. Keduas,
+  and C. Notredame, Nucleic Acids Research {\bf 34},  W604  (2006).
+
+\bibitem{3dcoffee}
+O. Poirot, K. Suhre, C. Abergel, E. O'Toole, and C. Notredame, Nucleic Acids
+  Research {\bf 32},  W37  (2004).
+
+\bibitem{mdtraj}
+R.~T. McGibbon, K.~A. Beauchamp, C.~R. Schwantes, L.-P. Wang, C.~X.
+  Hern{\'a}ndez, M.~P. Harrigan, T.~J. Lane, J.~M. Swails, and V.~S. Pande,
+  bioRxiv  (2014).
+
+\bibitem{theobald:acta-cryst-a:2005:qcp}
+D.~L. Theobald, Acta Cryst. A {\bf 61},  478  (2005).
+
+\bibitem{theobald:j-comput-chem:2010:qcp}
+P. Liu, D.~K. Agrafiotis, and D.~L. Theobald, J. Comput. Chem. {\bf 31},  1561
+  (2010).
+
+\bibitem{theobald:j-comput-chem:2011:qcp}
+P. Liu, D.~K. Agrafiotis, and D.~L. Theobald, J. Comput. Chem. {\bf 32},  185
+  (2011).
+
+\bibitem{maccallum:prot:2011:casp-refinement}
+J.~L. MacCallum, A. P{\'e}rez, M.~J. Schnieders, L. Hua, M.~P. Jacobson, and
+  K.~A. Dill, Proteins: Structure, Function, and Bioinformatics {\bf 79},  74
+  (2011).
+
+\bibitem{zhang:curr-opin-struct-biol:2009:structure-prediction}
+Y. Zhang, Current Opinion in Structural Biology {\bf 19},  145  (2009).
+
+\bibitem{raval:prot:2012:long-timescale-md-refinement}
+A. Raval, S. Piana, M.~P. Eastwood, R.~O. Dror, and D.~E. Shaw, Proteins:
+  Structure, Function, and Bioinformatics {\bf 80},  2071  (2012).
+
+\bibitem{lbfgs}
+D.~C. Liu and J. Nocedal, Mathematical Programming {\bf 45},  503  (1989).
+
+\bibitem{amber99sb-ildn}
+K. Lindorff-Larsen, S.~P. anad Kim~Palmo, P. Maragakis, J.~L. Klepeis, R.~O.
+  Dror, and D.~E. Shaw, Proteins {\bf 78},  1950  (2010).
+
+\bibitem{gbsa-obc1}
+A. Onufriev, D. Bashford, and D.~A. Case, Proteins {\bf 55},  383  (2004).
+
+\bibitem{basconi:jctc:2013:collision-rates}
+J.~E. Basconi and M.~R. Shirts, Journal of Chemical Theory and Computation {\bf
+  9},  2887  (2013).
+
+\bibitem{tip3p}
+W.~L. Jorgensen, J. Chandrasekhar, J.~D. Madura, R.~W. Impey, and M.~L. Klein,
+  Journal of Chemical Physics {\bf 79},  926  (1983).
+
+\bibitem{tip4p-ew}
+H.~W. Horn, W.~C. Swope, J.~W. Pitera, J.~D. Madura, T.~J. Dick, G.~L. Hura,
+  and T. Head-Gordon, The Journal of Chemical Physics {\bf 120},  9665  (2004).
+
+\bibitem{molprobity:2007}
+I.~W. Davis, A. Leaver-Fay, V.~B. Chen, J.~N. Block, G.~J. Kapral, X. Wang,
+  L.~W. Murray, W.~B. Arendall, J. Snoeyink, J.~S. Richardson, and D.~C.
+  Richardson, Nucleic Acids Research {\bf 35},  W375  (2007).
+
+\bibitem{molprobity:2010}
+V.~B. Chen, W.~B. Arendall, J.~J. Headd, D.~A. Keedy, R.~M. Immormino, G.~J.
+  Kapral, L.~W. Murray, J.~S. Richardson, and D.~C. Richardson, Acta
+  Crystallographica. Section D, Biological Crystallography {\bf 66},  12
+  (2010).
+
+\bibitem{berendsen:1995:gromacs}
+H.~J.~C. Berendsen, D. van~der Spoel, and R. van Drunen, Comp. Phys. Comm. {\bf
+  91},  43  (1995).
+
+\bibitem{lindahl:2001:gromacs}
+E. Lindahl, B. Hess, and D. van~der Spoel, J. Mol. Model. {\bf 7},  306
+  (2001).
+
+\bibitem{krause:nejm:2005:tks-cancer}
+D.~S. Krause and R.~A. Van~Etten, New England Journal of Medicine {\bf 353},
+  172  (2005).
+
+\bibitem{greuber:2013:abl-cancer-review}
+E.~K. Greuber, P. Smith-Pearson, J. Wang, and A.~M. Pendergast, Nature Reviews
+  Cancer {\bf 13},  559  (2013).
+
+\bibitem{kim:2009:src-cancer-review}
+L.~C. Kim, L. Song, and E.~B. Haura, Nature Reviews Clinical Oncology {\bf 6},
+  587  (2009).
+
+\bibitem{liu:ncb:2006:kinase-inhibitors-inactive-conformations}
+Y. Liu and N.~S. Gray, Nature Chemical Biology {\bf 2},  358  (2006).
+
+\bibitem{shukla-pande:nature-commun:2014:src-activation-msm}
+D. Shukla, Y. Meng, B. Roux, and V.~S. Pande, Nature Commun. {\bf 5},  3397
+  (2014).
+
+\bibitem{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}
+C. Chothia and A.~M. Lesk, EMBO J. {\bf 5},  823  (1986).
+
+\bibitem{roux:pnas:2013:abl-imatinib}
+Y.-L. Lin, Y. Meng, W. Jiang, and B. Roux, Proc. Natl. Acad. Sci. USA {\bf
+  110},  1664  (2013).
+
+\bibitem{xu:1999:2src}
+W. Xu, A. Doshi, M. Lei, M.~J. Eck, and S.~C. Harrison, Molecular Cell {\bf 3},
+   629  (1999).
+
+\bibitem{cowan-jacob:2005:1y57}
+S.~W. Cowan-Jacob, G. Fendrich, P.~W. Manley, W. Jahnke, D. Fabbro, J.
+  Liebetanz, and T. Meyer, Structure {\bf 13},  861  (2005).
+
+\bibitem{young:2006:2f4j}
+M.~A. Young, N.~P. Shah, L.~H. Chao, M. Seeliger, Z.~V. Milanov, W.~H. Biggs,
+  D.~K. Treiber, H.~K. Patel, P.~P. Zarrinkar, D.~J. Lockhart, C.~L. Sawyers,
+  and J. Kuriyan, Cancer Research {\bf 66},  1007  (2006).
+
+\bibitem{cowan-jacob:2006:2hyy}
+S.~W. Cowan-Jacob, G. Fendrich, A. Floersheimer, P. Furet, J. Liebetanz, G.
+  Rummel, P. Rheinberger, M. Centeleghe, D. Fabbro, and P.~W. Manley, Acta
+  Crystallographica Section D: Biological Crystallography {\bf 63},  80
+  (2006).
+
+\bibitem{levinson:2006:2g1t}
+N.~M. Levinson, O. Kuchment, K. Shen, M.~A. Young, M. Koldobskiy, M. Karplus,
+  P.~A. Cole, and J. Kuriyan, PLoS Biol {\bf 4},  e144  (2006).
+
+\bibitem{kannan:jmb:2005:kinase-regulation}
+N. Kannan and A.~F. Neuwald, Journal of Molecular Biology {\bf 351},  956
+  (2005).
+
+\bibitem{foda:nat-comm:2015:src}
+Z.~H. Foda, Y. Shan, E.~T. Kim, D.~E. Shaw, and M.~A. Seeliger, Nature
+  Communications {\bf 6},  5939  (2015).
+
+\bibitem{ozkirimli:prot-sci:2008:src}
+E. Ozkirimli, S.~S. Yadav, W.~T. Miller, and C.~B. Post, Protein Science : A
+  Publication of the Protein Society {\bf 17},  1871  (2008).
+
+\bibitem{Su:2006:flt4cancer}
+J.-L. Su, P.-C. Yang, J.-Y. Shih, C.-Y. Yang, L.-H. Wei, C.-Y. Hsieh, C.-H.
+  Chou, Y.-M. Jeng, M.-Y. Wang, K.-J. Chang, M.-C. Hung, and M.-L. Kuo, Cancer
+  Cell {\bf 9},  209  (2006).
+
+\bibitem{schnieders:biophys-j:2010:x-ray-refinement}
+T.~D. Fenn, M.~J. Schnieders, A.~T. Brunger, and V.~S. Pande, Biophysical
+  Journal {\bf 98},  2984  (2010).
+
+\bibitem{schnieders:jctc:2011:space-group-pme}
+M.~J. Schnieders, T.~D. Fenn, and V.~S. Pande, Journal of Chemical Theory and
+  Computation {\bf 7},  1141  (2011).
+
+\bibitem{nagar:cell:2003:abl-1opl}
+B. Nagar, O. Hantschel, M.~A. Young, K. Scheffzek, D. Veach, W. Bornmann, B.
+  Clarkson, G. Superti-Furga, and J. Kuriyan, Cell {\bf 112},  859  (2003).
+
+\bibitem{shan:pnas:2009:abl}
+Y. Shan, M.~A. Seeliger, M.~P. Eastwood, F. Frank, H. Xu, M.~Ã. Jensen, R.~O.
+  Dror, J. Kuriyan, and D.~E. Shaw, Proceedings of the National Academy of
+  Sciences {\bf 106},  139  (2009).
+
+\bibitem{alexov-gunner:biophys-j:1997:mcce2}
+E.~G. Alexov and M.~R. Gunner, Biophys. J. {\bf 72},  2075  (1997).
+
+\bibitem{georgescu-alexov-gunner:biophys-j:2002:mcce2}
+R.~E. Georgescu, E.~G. Alexov, and M.~R. Gunner, Biophys. J. {\bf 83},  1731
+  (2002).
+
+\bibitem{song-mao-gunner:j-comput-chem:2009:mcce2}
+Y. Song, J. Mao, and M.~R. Gunner, J. Comput. Chem. {\bf 30},  2231  (2009).
+
+\bibitem{adams:prot-sci:1993:kinase-metal-ions}
+J.~A. Adams and S.~S. Taylor, Protein Science {\bf 2},  2177  (1993).
+
+\bibitem{sousa:book:2010:metal-ion-parameters-difficulties}
+S.~F. Sousa, R.~A. Fernandes, and M.~J. Ramos,  in {\em Kinetics and
+  {Dynamics}: {From} {Nano}- to {Bio}-{Scale}}, Vol.~12 of {\em Challenges and
+  {Advances} in {Computational} {Chemistry} and {Physics}}, edited by P.~a.
+  D.-D.~A. Paneth (Springer Science \& Business Media, Berlin, 2010), p.\ 530.
+
+\bibitem{gaff}
+J. Wang, R.~M. Wolf, J.~W. Caldwell, P.~A. Kollman, and D.~A. Case, J. Comput.
+  Chem. {\bf 25},  1157  (2004).
+
+\bibitem{gaff2}
+J. Wang, W. Wang, P.~A. Kollman, and D.~A. Case, J. Mol. Graph Model. {\bf 25},
+   247260  (2006).
+
+\end{thebibliography}
 

### 2016-01-11 manuscript-post-acceptance-submission/ensembler.tex
Updated post-acceptance version of manuscript to use PLoS template

@@
-% PRL look and style (easy on the eyes)
-%\documentclass[aps,pre,twocolumn,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
-% Two-column style (for submission/review/editing)
-\documentclass[aps,prl,preprint,nofootinbib,superscriptaddress,linenumbers]{revtex4-1}
-
-\pdfoutput=1
-\usepackage[pdftex]{graphicx}
-
-\usepackage{alltt}
+% Template for PLoS
+% Version 3.1 February 2015
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig." instead of "Figure".
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout or alignment
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% fixltx2e package for \textsubscript
+\usepackage{fixltx2e}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% rotating package for sideways tables
+\usepackage{rotating}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Additional packages used by authors
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% To allow code "Boxes" as well as figures
 \usepackage{fancyvrb}
-
-%\usepackage{palatino}
-
-%\usepackage{palatino}
-% Change to a sans serif font.
-\usepackage{sourcesanspro}
-\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
-\usepackage[T1]{fontenc}
-%\usepackage[font=sf,justification=justified]{caption}
-\usepackage[font=sf]{floatrow}
-
-% To allow code "Boxes" as well as figures
 \usepackage{newfloat}
 \DeclareFloatingEnvironment[name={Box}]{codebox}
-
-% Rework captions to use sans serif font.
-\makeatletter
-\renewcommand\@make@capt@title[2]{%
- \@ifx@empty\float@link{\@firstofone}{\expandafter\href\expandafter{\float@link}}%
-  {\sf\textbf{#1}}\sf\@caption@fignum@sep#2\quad
-}%
-\makeatother
-
 \usepackage{listings} % For code examples
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-%\usepackage[mathbf,mathcal]{euler}
-%\usepackage{citesort}
-\usepackage[caption=false]{subfig}
-\usepackage{dcolumn}
-\usepackage{boxedminipage}
 \usepackage{verbatim}
-%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
-\usepackage[draft]{hyperref}
-\usepackage[group-separator={,}]{siunitx}
-
-% Justification
-\captionsetup{singlelinecheck=off}
-
-% Pretty-printing of shell commands
-\newcommand{\shellcmd}[1]{\\\ \texttt{\scriptsize #1}}
-
-% The figures are in a figures/ subdirectory.
-\graphicspath{{../figures/}}
 
@@
 
-%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% For the \num command
+\usepackage[group-separator={,}]{siunitx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% End additional packages
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
+
 \begin{document}
-
-%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Ensembler: Enabling high-throughput molecular simulations at the superfamily scale}
-
-\author{Daniel L. Parton}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
-Corresponding author: john.chodera@choderalab.org
+\vspace*{0.35in}
+
+% Title must be 250 characters or less.
+% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
+\begin{flushleft}
+{\Large
+\textbf\newline{Ensembler: Enabling High-Throughput Molecular Simulations at the Superfamily Scale}
 }
-  %\email{daniel.parton@choderalab.org}
-\author{Patrick B. Grinaway}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
-Corresponding author: john.chodera@choderalab.org
-}
-\author{Sonya M. Hanson}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
-Corresponding author: john.chodera@choderalab.org
-}
-\author{Kyle A. Beauchamp}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
-Corresponding author: john.chodera@choderalab.org
-}
-  %\email{patrick.grinaway@choderalab.org}
-\author{John D. Chodera}
- %\thanks{Corresponding author}
- %\email{john.chodera@choderalab.org}
-  \affiliation{Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065 \\
-Corresponding author: john.chodera@choderalab.org
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABSTRACT/pacs
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{abstract}
-
-\maketitle
-
-\section*{abstract}
-
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Daniel L. Parton\textsuperscript{1,},
+Patrick B. Grinaway\textsuperscript{1},
+Sonya M. Hansom\textsuperscript{1},
+Kyle A. Beauchamp\textsuperscript{1},
+John D. Chodera\textsuperscript{1,*},
+\\
+\bigskip
+\bf{1} Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* john.chodera@choderalab.org
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 The rapidly expanding body of available genomic and protein structural data provides a rich resource for understanding protein dynamics with biomolecular simulation. 
@@
 
-\emph{Keywords: molecular dynamics simulation; comparative modeling; distributed simulation}
-
-% \end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AUTHOR SUMMARY
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
 \section*{Author Summary}
-\label{section:author_summary}
-
+% TODO improve
 Proteins are the workhorses of the human body, and are involved in essentially every biological process.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% INTRODUCTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{section:introduction}
-
+\linenumbers
+
+\section*{Introduction}
 Recent advances in genomics and structural biology have helped generate an enormous wealth of protein data at the level of amino-acid sequence and three-dimensional structure.
@@
 For example, the generated models could represent valuable data sets even without subsequent production simulation, allowing exploration of the conformational diversity present within the available structural data for a given protein family.
-Furthermore, automation of simulation preparation provides an excellent opportunity to make concrete certain "best practices", such as the choice of simulation parameters, approach to the treatment of protonation states, treatment of cofactors and structural ions, and pre-simulation refinement and equilibration procedures.
+Furthermore, automation of simulation preparation provides an excellent opportunity to make concrete certain ``best practices'', such as the choice of simulation parameters, approach to the treatment of protonation states, treatment of cofactors and structural ions, and pre-simulation refinement and equilibration procedures.
 While the current version of {\bf Ensembler} only codifies some of these choices as default parameters, its modular nature allows additional stages to be easily added in the future.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DESIGN AND IMPLEMENTATION
+% Design and Implementation
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design and Implementation}
+
+\section*{Design and Implementation}
 
@@
 
-\begin{figure*}[tb]
-    %\includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
+\begin{figure}[h]
+    % \includegraphics[width=1.0\textwidth]{pipeline/pipeline2}
 
@@
   Note that \emph{CPU-h} denotes the number of hours consumed by the equivalent of a single CPU hyperthread and \emph{GPU-h} on a single GPU---parallel execution via MPI reduces wall clock time nearly linearly.
+  }
   \label{figure:pipeline}
-  }
-\end{figure*}
-
-\subsection{Target selection and retrieval (A)}
+\end{figure}
+
+\subsection*{Target selection and retrieval (A)}
 
@@
 
+\begin{codebox}[h]
+\begin{adjustwidth}{-2.25in}{0in}
+\footnotesize
+\begin{Verbatim}[frame=single]
+ensembler gather_targets --query `family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes'
+                         --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
+ensembler gather_templates --gather_from uniprot --query `domain:"Protein kinase" AND reviewed:yes'
+                           --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
+\end{Verbatim}
+\caption{{\bf Ensembler command-line functions used to select targets and templates.}
+The commands retrieve target and template data by querying UniProt.
+The query string provided to the {\tt gather\_targets} command selects all human tyrosine protein kinases which have been reviewed by a curator, while the query string provided to the {\tt gather\_templates} command selects all reviewed protein kinases of any species.
+The {\tt-{}-uniprot\_domain\_regex} flag is used to select a subset of the domains belonging to the returned UniProt protein entries, by matching the domain annotations against a given regular expression.
+In this example, domains of type `Protein kinase', `Protein kinase 1', and `Protein kinase 2' were selected, while excluding many other domain types such as `Protein kinase; truncated', `Protein kinase; inactive', `SH2', `SH3', etc.
+Target selection simply entails the selection of sequences corresponding to each matching UniProt domain.
+Template selection entails the selection of the sequences and structures of any PDB entries corresponding to the matching UniProt domains.
+}
+\label{box:tk-gather-targets-templates}
+\end{adjustwidth}
+\end{codebox}
+
 In many cases, it will be desirable to build models of an isolated protein domain, rather than the full-length protein.
 The {\tt gather\_targets} subcommand allows protein domains to be selected from UniProt data by passing a regular expression string to the {\tt -{}-uniprot\_domain\_regex} flag. 
-For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including "Protein kinase", "Protein kinase 1", "Protein kinase 2", "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
+For example, the above {\tt -{}-query} flag for selecting all human protein kinases returns UniProt entries with domain annotations including `Protein kinase', `Protein kinase 1', `Protein kinase 2', `Protein kinase; truncated', `Protein kinase; inactive', `SH2', `SH3', etc.
 The regular expression shown in Box~\ref{box:tk-gather-targets-templates} selects only domains of the first three types.
@@
 
-\subsection{Template selection and retrieval (B)}
+\subsection*{Template selection and retrieval (B)}
 
@@
 
-\subsection{Template refinement (C)}
+\subsection*{Template refinement (C)}
 
@@
 
-% Because fewer loops need to be built during the subsequent target model-building stage, we find that prebuilding template loops tends to provide higher-quality models after completion of the {\bf Ensembler} pipeline.
-% [JDC: Should we cite our evidence for this with the TKs, or maybe tone back the claim a bit to say that it is possible this could make things easier?]
-% [DLP: Might be worth investigating an algorithm called \emph{pokefind} (or \emph{knotfind}, which I think is an earlier version) which aims to find knots in proteins, of the type which encouraged us to use Rosetta to reconstruct template loops.
-% DOI: 10.1093/bioinformatics/btp198
-% It sounds like these algorithms have actually been implemented in Rosetta, so this could explain why Rosetta seems to do better at avoiding making these knotted structures.
-% Would be useful to check this out further first, and then decide whether or not to discuss the knotted structures in the manuscript.]
-
-\subsection{Alignment and comparative modeling (D)}
+\subsection*{Alignment and comparative modeling (D)}
 
@@
 
-\subsection{Refinement of models and filtering of poor models by simulation (E)}
-
-A number of refinement methods have been developed to help guide comparative modeling techniques toward more "native-like" and physically consistent conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}.
+\subsection*{Refinement of models and filtering of poor models by simulation (E)}
+
+A number of refinement methods have been developed to help guide comparative modeling techniques toward more ``native-like'' and physically consistent conformations~\cite{maccallum:prot:2011:casp-refinement,zhang:curr-opin-struct-biol:2009:structure-prediction}.
 Both short~\cite{zhang:curr-opin-struct-biol:2009:structure-prediction} and long~\cite{raval:prot:2012:long-timescale-md-refinement} molecular dynamics simulations have been employed for this purpose.
@@
 
-The simulation protocol and default parameter values have been chosen to represent current "best practices" for the refinement simulations carried out here.
+The simulation protocol and default parameter values have been chosen to represent current ``best practices'' for the refinement simulations carried out here.
 As such, the simulation is performed using Langevin dynamics, with a default force field choice of Amber99SB-ILDN~\cite{amber99sb-ildn}, along with a modified generalized Born solvent model~\cite{gbsa-obc1} as implemented in the OpenMM package~\cite{eastman:jctc:2012:openmm}.
@@
 
-\subsection{Solvation and NPT equilibration (F)}
+\subsection*{Solvation and NPT equilibration (F)}
 
@@
 
-\subsection*{{Model validation with MolProbity score}}
+\subsection*{Model validation with MolProbity score}
 
@@
 
-\subsubsection*{Tracking provenance information}
-
+\paragraph{Tracking provenance information.}
 To aid the user in tracking the provenance of each model, each pipeline function also outputs a metadata file, which helps to link data to the software version used to generate it (both {\bf Ensembler} and its dependencies), and also provides timing and performance information, and other data such as hostname.
 
-\subsubsection*{Rapidly modeling a single template}
-
+\paragraph{Rapidly modeling a single template.}
 For users interested in simply using {\bf Ensembler} to rapidly generate a set of models for a single template sequence, {\bf Ensembler} provides a command-line tool {\tt quickmodel}, which performs the entire pipeline for a single target with a small number of templates.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section*{Results}
 \label{section:results}
@@
 
-\begin{figure*}[tbp]
-    %\includegraphics[width=\textwidth]{nstructures_per_tk_target/nstructures_per_tk_target.pdf}
+\begin{figure}[h]
+    % \includegraphics[width=\textwidth]{nstructures_per_tk_target/nstructures_per_tk_target.pdf}
     \caption{{\bf Number of PDB structures available for each TK target.}
@@
     Each PDB chain is counted individually, and only chains which contain the target domain are counted.
+}
     \label{figure:nstructures-per-tk-target}
-}
-\end{figure*}
-
-\begin{codebox*}[tbp]
-\scriptsize
-\begin{Verbatim}[frame=single]
-ensembler gather_targets --query `family:"tyr protein kinase family" AND organism:"homo sapiens" AND reviewed:yes'
-                         --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
-ensembler gather_templates --gather_from uniprot --query `domain:"Protein kinase" AND reviewed:yes'
-                           --uniprot_domain_regex `^Protein kinase(?!; truncated)(?!; inactive)'
-\end{Verbatim}
-\caption{{\bf Ensembler command-line functions used to select targets and templates.}
-The commands retrieve target and template data by querying UniProt.
-The query string provided to the {\tt gather\_targets} command selects all human tyrosine protein kinases which have been reviewed by a curator, while the query string provided to the {\tt gather\_templates} command selects all reviewed protein kinases of any species.
-The {\tt-{}-uniprot\_domain\_regex} flag is used to select a subset of the domains belonging to the returned UniProt protein entries, by matching the domain annotations against a given regular expression.
-In this example, domains of type "Protein kinase", "Protein kinase 1", and "Protein kinase 2" were selected, while excluding many other domain types such as "Protein kinase; truncated", "Protein kinase; inactive", "SH2", "SH3", etc.
-Target selection simply entails the selection of sequences corresponding to each matching UniProt domain.
-Template selection entails the selection of the sequences and structures of any PDB entries corresponding to the matching UniProt domains.
-}
-\label{box:tk-gather-targets-templates}
-\end{codebox*}
+\end{figure}
 
@@
 
-\begin{figure}[tbp]
-    %\includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
+Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, mean 14, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, mean 7, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
+To reduce the reliance on the MODELLER rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
+Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
+There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful (when excluding templates with no missing residues).
+
+\begin{figure}[h]
+    % \includegraphics[width=\textwidth]{loopmodel_analysis/nmissing_resis_distributions.pdf}
     \caption{{\bf Distributions for the number of missing residues in the TK templates.}
@@
     \emph{Lower:} The number of residues in each missing loop, for all templates.    
+}
     \label{figure:loopmodel-nmissing-residues}
-}
 \end{figure}
-
-Crystallographic structures of kinase catalytic domains generally contain a significant number of missing residues (median 11, mean 14, standard deviation 13, max 102) due to the high mobility of several loops (Fig.~\ref{figure:loopmodel-nmissing-residues}, top), with a number of these missing spans being significant in length (median 5, mean 7, standard deviation 6, max 82; Fig.~\ref{figure:loopmodel-nmissing-residues}, bottom).
-To reduce the reliance on the MODELLER rapid model construction stage to reconstruct very long unresolved loops, unresolved template residues were first remodeled using the {\tt loopmodel} subcommand.
-Out of \num{3666} templates with one or more missing residues, \num{3134} were successfully remodeled by the Rosetta loop modeling stage (with success defined simply as program termination without error); most remodeling failures were attributable to unsatisfiable spatial constraints imposed by the original template structure.
-There was some correlation between remodeling failures and the number of missing residues (Fig.~\ref{figure:loopmodel-nmissing-residues}, top); templates for which remodeling failed had a median of 20 missing residues, compared to a median of 14 missing residues for templates for which remodeling was successful (when excluding templates with no missing residues).
 
@@
 
-\subsubsection*{All tyrosine kinases}
+\paragraph{All tyrosine kinases.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Sequence identity distribution
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    %\includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
+We also used the {\tt ensembler validate} subcommand to subject the refined models to analysis with MolProbity~\cite{molprobity:2007,molprobity:2010}.
+\begin{figure}[h]
+    % \includegraphics[width=1.0\columnwidth]{seqid_dist/seqid_dist.pdf}
 
@@
     A kernel density estimate of the target-template sequence identity probability density function is shown as a solid line with shaded region, while the corresponding cumulative distribution function is shown as a dashed line.
+    }
   \label{figure:sequence-identity-distribution}
-    }
 \end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: RMSD distribution by sequence identity
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tbp]
-    \centering
-    {\label{figure:rmsd-distribution-joint}%
-       %\includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
-    }
-
-    {\label{figure:rmsd-distribution-by-sequence-identity}%
-       %\includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
-    }
+\begin{figure}[h]
+    % \centering
+    % {\label{figure:rmsd-distribution-joint}%
+    %    \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2-joint.pdf}
+    % }
+
+    % {\label{figure:rmsd-distribution-by-sequence-identity}%
+    %    \includegraphics[height=0.6\textwidth]{rmsddist/rmsddist2.pdf}
+    % }
 
@@
     The plotted distributions have been smoothed using kernel density estimation.
+  }
   \label{figure:rmsd-distributions}
-  }
 \end{figure}
 
-We also used the {\tt ensembler validate} subcommand to subject the refined models to analysis with MolProbity~\cite{molprobity:2007,molprobity:2010}.
 The MolProbity scores varied from 0.92 to 4.80, with a median of 3.84, a mean of 3.22, and a standard deviation of 1.07.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Final energies following implicit refinement
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tbp]
-    %\includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
+\begin{figure}[h]
+    % \includegraphics[width=1.0\columnwidth]{energies/energies.pdf}
 
@@
     Refinement simulations were carried out at the default temperature of 300~K.
+  }
   \label{figure:energies-implicit}
-  }
 \end{figure}
 
-\subsubsection*{Src and Abl1}
+\paragraph{Src and Abl1.}
 
@@
 
-{\bf Visualizing model structural diversity.}
+\paragraph{Visualizing model structural diversity.}
 Fig.~\ref{figure:superposition} shows a superposition of a set of representative models of \emph{Src} and \emph{Abl1}.
@@
 High sequence identity models (in opaque blue) tend to be quite structurally similar, with some variation in loops or changes in domain orientation.
+
+\begin{figure}[h]
+    % \includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
+
+    \caption{{\bf Superposition of clustered models of Src and Abl1.}
+    Superposed renderings of nine models each for Src and Abl1, 
+    giving some indication the diversity of conformations generated by Ensembler.
+    The models for each target were divided into three sequence identity ranges (as in Fig.~\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed (using the msmbuilder clustering package~\cite{msmbuilder}) to select three clusters from each.
+    The models shown are the centroids of each cluster.
+    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
+  }
+  \label{figure:superposition}
+\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Superposition of Src and Abl1 sequence identity classes
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    %\includegraphics[width=1.0\columnwidth]{superposition-src_abl/superposed-seqid_classes-clustered-one_fig}
-
-    \caption{{\bf Superposition of clustered models of Src and Abl1.}
-    Superposed renderings of nine models each for Src and Abl1, 
-    giving some indication the diversity of conformations generated by Ensembler.
-    The models for each target were divided into three sequence identity ranges (as in Fig.~\ref{figure:rmsd-distributions}), and RMSD-based $k$-medoids clustering was performed (using the msmbuilder clustering package~\cite{msmbuilder}) to select three clusters from each.
-    The models shown are the centroids of each cluster.
-    Models are colored and given transparency based on their sequence identity, so that high sequence identity models are blue and opaque, while lower sequence identity models are transparent and red.
-  \label{figure:superposition}
-  }
-\end{figure}
-
-{\bf Comparison with known biochemically relevant conformations.}
+\paragraph{Comparison with known biochemically relevant conformations.}
 To evaluate the models of \emph{Src} and \emph{Abl1} in the context of the published structural biology literature on functionally relevant conformations, we have focused on two residue pair distances thought to be important order parameters for the regulation of protein kinase domain activity.
-We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; the exact numbering schemes are provided in Fig.~\ref{figure:residue-numbering-schemes}.
+We use the residue numbering schemes for chicken Src (commonly employed in the literature even in reference to human Src)~\cite{xu:1999:2src, cowan-jacob:2005:1y57} and human Abl1 isoform A~\cite{young:2006:2f4j, cowan-jacob:2006:2hyy, levinson:2006:2g1t} respectively; see \nameref{S1_Text} for the exact sequences and numbering schemes used.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Reference structures for Src
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-    %\includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
+\begin{figure}[h]
+    % \includegraphics[width=1.0\textwidth]{residue_pair_distances/src/src_ref_structures}
 
@@
     To aid visualization of the active site in 1Y57, it has been included in the rendering by structurally aligning the surrounding homologous protein residues.
+  }
   \label{figure:src-ref-structures}
-  }
 \end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Residue pair distance plots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-
-    % \includegraphics[width=1.0\columnwidth]{residue_pair_distances/src_abl_combined/distances.pdf}
-    \subfloat[Src]{\label{figure:src-distances}%
-       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_src.png}
-    }
-    \subfloat[Abl]{\label{figure:abl-distances}%
-       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_abl.png}
-    }
-    \subfloat[Flt4]{\label{figure:flt4-distances}%
-       %\includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_flt4.png}
-    }
+\begin{figure}[h]
+
+    % \subfloat[Src]{\label{figure:src-distances}%
+    %    \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_src.png}
+    % }
+    % \subfloat[Abl]{\label{figure:abl-distances}%
+    %    \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_abl.png}
+    % }
+    % \subfloat[Flt4]{\label{figure:flt4-distances}%
+    %    \includegraphics[height=0.31\textwidth]{residue_pair_distances/src_abl_flt4_combined/distances_flt4.png}
+    % }
 
@@
     The atom names for these atoms, according to the PDB coordinate files for both reference structures, are---Lys: NZ, CD, CE (ethylamine); Glu: OE1, CD, OE2 (carboxylate); Arg: NH1, CZ, NH2 (part of guanidine).
+    }
   \label{figure:pair-distances}
-    }
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Availability and Future Directions}
+\end{figure}
+
+\section*{Availability and Future Directions}
 \label{section:availability}
@@
 
-A supplementary dataset can also be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
-This contains the TK models described in the Results section, general information on the targets and templates, plus a script and instructions for regenerating the same dataset.
-
-\begin{codebox}[tbp]
-\scriptsize
+\begin{codebox}[h]
+\begin{adjustwidth}{-2.25in}{0in}
+\footnotesize
 \begin{Verbatim}[frame=single]
@@
 \label{box:installation}
+\end{adjustwidth}
 \end{codebox}
+
+A supplementary dataset can also be downloaded from the Dryad Digital Repository (DOI: \href{https://dx.doi.org/10.5061/dryad.7fg32}{10.5061/dryad.7fg32}).
+This contains the TK models described in the \ref{section:results} section, general information on the targets and templates, plus a script and instructions for regenerating the same dataset.
 
@@
 
-{\bf Template remodeling.}
+\paragraph{Template remodeling.}
 The lack of crystallographically-resolved regions of template structures presents a challenge to deriving structures from these templates by comparative modeling, especially in kinases, where loops are frequently unresolved.
@@
 
-{\bf Comparative modeling.}
+\paragraph{Comparative modeling.}
 Comparative protein modeling can be approached in a number of different ways, with varying degrees of complexity.
@@
 
-{\bf Protonation states.}
+\paragraph{Protonation states.}
 Some amino acids can exist in different protonation states, depending on pH and on their local environment.
@@
 
-{\bf Cofactors, structural ions, and ligands.}
+\paragraph{Cofactors, structural ions, and ligands.}
 Many proteins require the presence of various types of non-protein atoms and molecules for proper function, such as metal ions (e.g.~Mg$^{+2}$), cofactors (e.g.~ATP) or post-translational modifications (e.g.~phosphorylation, methylation, glycosylation, etc.), and we thus plan for {\bf Ensembler} to eventually have the capability to include such entities in the generated models.
@@
 
-{\bf Post-translationally modified amino acids and other molecules without forcefield parameters.}
+\paragraph{Post-translationally modified amino acids and other molecules without forcefield parameters.}
 A major challenge in the preparation of simulations of proteins of interest is the wide variety of post-translational modifications possible that are often functionally or structurally relevant.
@@
 
-{\bf Long insertions and deletions.}
+\paragraph{Long insertions and deletions.}
 Another limitation with the present version of {\bf Ensembler} involves the treatment of members of a protein family with especially long residue insertions or deletions.
@@
 
-{\bf Markov state model (MSM) construction and model utility.}
+\paragraph{Markov state model (MSM) construction and model utility.}
 We are actively utilizing {\bf Ensembler}-generated models to seed the construction of Markov state models (MSMs)~\cite{pande-beauchamp-bowman:2010:methods:markov-model-review,chodera-noe:curr-opin-struct-biol:2014:msm-review}.
@@
 
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
+Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+Nam condimentum sem eget mollis euismod.
+Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
+Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
+Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
+For more information, see \nameref{S1_Text}.
+
+\section*{Supporting Information}
+
+\subsection*{S1 Text}
+\label{S1_Text}
+{\bf Sequences and residue numbering schemes for Src and Abl1.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
+\section*{Acknowledgments}
 \label{section:acknowledgments}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% BIBLIOGRAPHY
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\nolinenumbers
+
+%\section*{References}
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% OR
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here.
+%
+
+% DONE - when ready, comment line below and uncomment subsequent lines. Then paste in formatted bibliography from ensembler.bbl
+% \bibliography{ensembler.bib}
 
@@
 \bibitem{craik:science:2009:trapping-moving-targets}
-G.~M. Lee and C.~S. Craik, Science {\bf 324},  213  (2009).
+Lee GM, Craik CS.
+\newblock Trapping moving targets with small molecules.
+\newblock Science. 2009;324:213.
 
 \bibitem{eastman:jctc:2012:openmm}
-P. Eastman, M.~S. Friedrichs, J.~D. Chodera, R.~J. Radmer, C.~M. Bruns, J.~P.
-  Ku, K.~A. Beauchamp, T.~J. Lane, L.-P. Wang, D. Shukla, and V.~S. Pande, J.
-  Chem. Theory Comput. {\bf 9},  461  (2012).
+Eastman P, Friedrichs MS, Chodera JD, Radmer RJ, Bruns CM, Ku JP, et~al.
+\newblock {OpenMM} 4: A Reusable, Extensible, Hardware Independent Library for
+  High Performance Molecular Simulation.
+\newblock J Chem Theory Comput. 2012;9(1):461--469.
 
 \bibitem{walker:jctc:2013:amber-gpu}
-R. Salomon-Ferrer, A.~W. G\"{o}tz, D. Poole, S.~L. Grand, and R.~C. Walker, J.
-  Chem. Theor. Comput. {\bf 9},  3878  (2013).
+Salomon-Ferrer R, G\"{o}tz AW, Poole D, Grand SL, Walker RC.
+\newblock Routine microsecond molecular dynamics simulations with {AMBER} on
+  {GPUs}. 2. Explicit solvent particle mesh {Ewald}.
+\newblock J Chem Theor Comput. 2013;9(9):3878--3888.
 
 \bibitem{shirts-pande:science:2000:fah}
-M. Shirts and V.~S. Pande, Science {\bf 290},  1903  (2000).
+Shirts M, Pande VS.
+\newblock Screen {Savers} of the {World} {Unite}!
+\newblock Science. 2000 Dec;290(5498):1903--1904.
 
 \bibitem{pronk:2011:copernicus}
-S. Pronk, P. Larsson, I. Pouya, G.~R. Bowman, I.~S. Haque, K. Beauchamp, B.
-  Hess, V.~S. Pande, P.~M. Kasson, and E. Lindahl,  in {\em Proceedings of 2011
-  {International} {Conference} for {High} {Performance} {Computing},
-  {Networking}, {Storage} and {Analysis}}, {\em {SC} '11} (ACM, New York, NY,
-  USA, 2011), pp.\ 60:1--60:10.
+Pronk S, Larsson P, Pouya I, Bowman GR, Haque IS, Beauchamp K, et~al.
+\newblock Copernicus: {A} {New} {Paradigm} for {Parallel} {Adaptive}
+  {Molecular} {Dynamics}.
+\newblock In: Proceedings of 2011 {International} {Conference} for {High}
+  {Performance} {Computing}, {Networking}, {Storage} and {Analysis}. {SC} '11.
+  New York, NY, USA: ACM; 2011. p. 60:1--60:10.
 
 \bibitem{pronk:2015:copernicus}
-S. Pronk, I. Pouya, M. Lundborg, G. Rotskoff, B. Wes\'{e}n, P.~M. Kasson, and
-  E. Lindahl, Journal of Chemical Theory and Computation  (2015).
+Pronk S, Pouya I, Lundborg M, Rotskoff G, Wes\'{e}n B, Kasson PM, et~al.
+\newblock Molecular {Simulation} {Workflows} as {Parallel} {Algorithms}: {The}
+  {Execution} {Engine} of {Copernicus}, a {Distributed} {High}-{Performance}
+  {Computing} {Platform}.
+\newblock Journal of Chemical Theory and Computation. 2015 Apr;.
 
 \bibitem{buch:2010:gpugrid}
-I. Buch, M.~J. Harvey, T. Giorgino, D.~P. Anderson, and G. De~Fabritiis,
-  Journal of Chemical Information and Modeling {\bf 50},  397  (2010).
+Buch I, Harvey MJ, Giorgino T, Anderson DP, De~Fabritiis G.
+\newblock High-{Throughput} {All}-{Atom} {Molecular} {Dynamics} {Simulations}
+  {Using} {Distributed} {Computing}.
+\newblock Journal of Chemical Information and Modeling. 2010
+  Mar;50(3):397--403.
 
 \bibitem{pande-beauchamp-bowman:2010:methods:markov-model-review}
-V.~S. Pande, K. Beauchamp, and G.~R. Bowman, Methods {\bf 52},  99  (2010).
+Pande VS, Beauchamp K, Bowman GR.
+\newblock Everything you wanted to know about {Markov State Models} but were
+  afraid to ask.
+\newblock Methods. 2010;52:99--105.
 
 \bibitem{noe:jcp:2011:msm-review}
-J.-H. Prinz, H. Wu, M. Sarich, B. Keller, M. Fischbach, M. Held, J.~D. Chodera,
-  C. Sch\"{u}tte, and F. No\'{e}, J. Chem. Phys. {\bf 134},  174105  (2011).
+Prinz JH, Wu H, Sarich M, Keller B, Fischbach M, Held M, et~al.
+\newblock Markov models of molecular kinetics: Generation and validation.
+\newblock J Chem Phys. 2011;134:174105.
 
 \bibitem{chodera-noe:curr-opin-struct-biol:2014:msm-review}
-J.~D. Chodera and F. No\'{e}, Curr. Opin. Struct. Biol. {\bf 25},  135  (2014).
+Chodera JD, No\'{e} F.
+\newblock Markov state models of biomolecular conformational dynamics.
+\newblock Curr Opin Struct Biol. 2014;25:135--144.
 
 \bibitem{moult:proteins:2014:casp10}
-J. Moult, K. Fidelis, A. Kryshtafovych, T. Schwede, and A. Tramontano,
-  Proteins: Structure, Function, and Bioinformatics {\bf 82},  1  (2014).
+Moult J, Fidelis K, Kryshtafovych A, Schwede T, Tramontano A.
+\newblock Critical assessment of methods of protein structure prediction
+  ({CASP}) — round x.
+\newblock Proteins: Structure, Function, and Bioinformatics. 2014 Feb;82:1--6.
 
 \bibitem{baker:science:2001:structure-prediction-review}
-D. Baker and A. {\v{S}ali}, Science {\bf 294},  93  (2001).
+Baker D, {\v{S}ali} A.
+\newblock Protein {Structure} {Prediction} and {Structural} {Genomics}.
+\newblock Science. 2001 Oct;294(5540):93--96.
 
 \bibitem{baker:structure:2013:rosettacm}
-Y. Song, F. {DiMaio}, R.~Y.-R. Wang, D. Kim, C. Miles, T. Brunette, J.
-  Thompson, and D. Baker, Structure {\bf 21},  1735  (2013).
+Song Y, {DiMaio} F, Wang RYR, Kim D, Miles C, Brunette T, et~al.
+\newblock High-resolution comparative modeling with {RosettaCM}.
+\newblock Structure. 2013;21:1735--1742.
 
 \bibitem{yang:nature-methods:2015:i-tasser-suite}
-J. Yang, R. Yan, A. Roy, D. Xu, J. Poisson, and Y. Zhang, Nature Methods {\bf
-  12},  7  (2015).
+Yang J, Yan R, Roy A, Xu D, Poisson J, Zhang Y.
+\newblock The {I-TASSER Suite}: Protein structure and function prediction.
+\newblock Nature Methods. 2015;12:7.
 
 \bibitem{daggett:structure:2010:dynameomics}
-M.~W. {van der Kamp}, R.~D. Schaeffer, A.~L. Jonsson, A.~D. Scouras, A.~M.
-  Simms, R.~D. Toofanny, N.~C. Benson, P.~C. Anderson, E.~D. Merkley, S.
-  Rysavy, D. Bromley, D.~A.~C. Beck, and V. Daggett, Structure {\bf 18},  423
-  (2010).
+{van der Kamp} MW, Schaeffer RD, Jonsson AL, Scouras AD, Simms AM, Toofanny RD,
+  et~al.
+\newblock Dynameomics: A comprehensive database of protein dynamics.
+\newblock Structure. 2010;18:423--435.
 
 \bibitem{friedland:plos-comput-biol:2009:conformational-dynamics}
-G.~D. Friedland, N.-A. Lakomek, C. Griesinger, J. Meiler, and T. Kortemme, PLoS
-  Comput. Biol. {\bf 5},  e1000393  (2009).
+Friedland GD, Lakomek NA, Griesinger C, Meiler J, Kortemme T.
+\newblock A correspondence between solution-state dynamics of an individual
+  protein and the sequence and conformational diversity of its family.
+\newblock PLoS Comput Biol. 2009;5:e1000393.
 
 \bibitem{sali:2012:allostery-modeling}
-P. Weinkam, J. Pons, and A. Sali, Proceedings of the National Academy of
-  Sciences of the United States of America {\bf 109},  4875  (2012).
+Weinkam P, Pons J, Sali A.
+\newblock Structure-based model of allostery predicts coupling between distant
+  sites.
+\newblock Proceedings of the National Academy of Sciences of the United States
+  of America. 2012 Mar;109(13):4875--4880.
 
 \bibitem{msmbuilder}
-K.~A. Beauchamp, G.~R. Bowman, T.~J. Lane, L. Maibaum, I.~S. Haque, and V.~S.
-  Pande, Journal of Chemical Theory and Computation {\bf 7},  3412  (2011).
+Beauchamp KA, Bowman GR, Lane TJ, Maibaum L, Haque IS, Pande VS.
+\newblock {MSMBuilder}2: {Modeling} {Conformational} {Dynamics} on the
+  {Picosecond} to {Millisecond} {Scale}.
+\newblock Journal of Chemical Theory and Computation. 2011
+  Oct;7(10):3412--3419.
 
 \bibitem{caflisch:jpcb:2011:msm-subgraph}
-R. Scalco and A. Caflisch, The Journal of Physical Chemistry. B {\bf 115},
-  6358  (2011).
+Scalco R, Caflisch A.
+\newblock Equilibrium distribution from distributed computing (simulations of
+  protein folding).
+\newblock The Journal of Physical Chemistry B. 2011 May;115(19):6358--6365.
 
 \bibitem{uniprot:2015}
-T.~U. Consortium, Nucleic Acids Research {\bf 43},  D204  (2015).
+Consortium TU.
+\newblock {UniProt}: a hub for protein information.
+\newblock Nucleic Acids Research. 2015 Jan;43(D1):D204--D212.
 
 \bibitem{velankar:2013:sifts}
-S. Velankar, J.~M. Dana, J. Jacobsen, G. van Ginkel, P.~J. Gane, J. Luo, T.~J.
-  Oldfield, C. O'Donovan, M.-J. Martin, and G.~J. Kleywegt, Nucleic Acids
-  Research {\bf 41},  D483  (2013).
+Velankar S, Dana JM, Jacobsen J, van Ginkel G, Gane PJ, Luo J, et~al.
+\newblock {SIFTS}: {Structure} {Integration} with {Function}, {Taxonomy} and
+  {Sequences} resource.
+\newblock Nucleic Acids Research. 2013 Jan;41(D1):D483--D489.
 
 \bibitem{qian:nature:2007:modeller}
-B. Qian, S. Raman, R. Das, P. Bradley, A.~J. McCoy, R.~J. Read, and D. Baker,
-  Nature {\bf 450},  259  (2007).
+Qian B, Raman S, Das R, Bradley P, McCoy AJ, Read RJ, et~al.
+\newblock High-resolution structure prediction and the crystallographic phase
+  problem.
+\newblock Nature. 2007 Nov;450(7167):259--264.
 
 \bibitem{wang:jmb:2007:modeller}
-C. Wang, P. Bradley, and D. Baker, Journal of Molecular Biology {\bf 373},  503
-   (2007).
+Wang C, Bradley P, Baker D.
+\newblock Protein–{Protein} {Docking} with {Backbone} {Flexibility}.
+\newblock Journal of Molecular Biology. 2007 Oct;373(2):503--519.
 
 \bibitem{fiser:prot-sci:2000:modeller}
-A. Fiser, R.~K.~G. Do, and A. {\v{S}ali}, Protein Science {\bf 9},  1753
-  (2000).
+Fiser A, Do RKG, {\v{S}ali} A.
+\newblock Modeling of loops in protein structures.
+\newblock Protein Science. 2000 Jan;9(9):1753--1773.
 
 \bibitem{sali:jmb:1993:modeller}
-A. {\v{S}ali} and T.~L. Blundell, Journal of Molecular Biology {\bf 234},  779
-  (1993).
+{\v{S}ali} A, Blundell TL.
+\newblock Comparative {Protein} {Modelling} by {Satisfaction} of {Spatial}
+  {Restraints}.
+\newblock Journal of Molecular Biology. 1993 Dec;234(3):779--815.
 
 \bibitem{cock:2009:biopython}
-P.~J.~A. Cock, T. Antao, J.~T. Chang, B.~A. Chapman, C.~J. Cox, A. Dalke, I.
-  Friedberg, T. Hamelryck, F. Kauff, B. Wilczynski, and M.~J.~L. de~Hoon,
-  Bioinformatics (Oxford, England) {\bf 25},  1422  (2009).
+Cock PJA, Antao T, Chang JT, Chapman BA, Cox CJ, Dalke A, et~al.
+\newblock Biopython: freely available {Python} tools for computational
+  molecular biology and bioinformatics.
+\newblock Bioinformatics (Oxford, England). 2009 Jun;25(11):1422--1423.
 
 \bibitem{gonnet:science:1992:exhaustive-matching}
-G.~H. Gonnet, M.~A. Cohen, and S.~A. Brenner, Science {\bf 256},  1443  (1992).
+Gonnet GH, Cohen MA, Brenner SA.
+\newblock Exhaustive matching of the entire protein sequence database.
+\newblock Science. 1992;256:1443--1445.
 
 \bibitem{thompson:2011:msa-review}
-J.~D. Thompson, B. Linard, O. Lecompte, and O. Poch, PLoS ONE {\bf 6},  e18093
-  (2011).
+Thompson JD, Linard B, Lecompte O, Poch O.
+\newblock A {Comprehensive} {Benchmark} {Study} of {Multiple} {Sequence}
+  {Alignment} {Methods}: {Current} {Challenges} and {Future} {Perspectives}.
+\newblock PLoS ONE. 2011 Mar;6(3):e18093.
 
 \bibitem{pei:2008:promals3d}
-J. Pei, B.-H. Kim, and N.~V. Grishin, Nucleic Acids Research {\bf 36},  2295
-  (2008).
+Pei J, Kim BH, Grishin NV.
+\newblock {PROMALS}3D: a tool for multiple protein sequence and structure
+  alignments.
+\newblock Nucleic Acids Research. 2008 Apr;36(7):2295--2300.
 
 \bibitem{expresso}
-F. Armougom, S. Moretti, O. Poirot, S. Audic, P. Dumas, B. Schaeli, V. Keduas,
-  and C. Notredame, Nucleic Acids Research {\bf 34},  W604  (2006).
+Armougom F, Moretti S, Poirot O, Audic S, Dumas P, Schaeli B, et~al.
+\newblock Expresso: automatic incorporation of structural information in
+  multiple sequence alignments using 3D-{Coffee}.
+\newblock Nucleic Acids Research. 2006 Jul;34(Web Server issue):W604--608.
 
 \bibitem{3dcoffee}
-O. Poirot, K. Suhre, C. Abergel, E. O'Toole, and C. Notredame, Nucleic Acids
-  Research {\bf 32},  W37  (2004).
+Poirot O, Suhre K, Abergel C, O'Toole E, Notredame C.
+\newblock 3DCoffee@igs: a web server for combining sequences and structures
+  into a multiple sequence alignment.
+\newblock Nucleic Acids Research. 2004 Jul;32(Web Server issue):W37--40.
 
 \bibitem{mdtraj}
-R.~T. McGibbon, K.~A. Beauchamp, C.~R. Schwantes, L.-P. Wang, C.~X.
-  Hern{\'a}ndez, M.~P. Harrigan, T.~J. Lane, J.~M. Swails, and V.~S. Pande,
-  bioRxiv  (2014).
+McGibbon RT, Beauchamp KA, Schwantes CR, Wang LP, Hern{\'a}ndez CX, Harrigan
+  MP, et~al.
+\newblock MDTraj: a modern, open library for the analysis of molecular dynamics
+  trajectories.
+\newblock bioRxiv. 2014;.
 
 \bibitem{theobald:acta-cryst-a:2005:qcp}
-D.~L. Theobald, Acta Cryst. A {\bf 61},  478  (2005).
+Theobald DL.
+\newblock Rapid calculation of {RMSDs} using a quaternion-based characteristic
+  polynomial.
+\newblock Acta Cryst A. 2005;61:478--480.
 
 \bibitem{theobald:j-comput-chem:2010:qcp}
-P. Liu, D.~K. Agrafiotis, and D.~L. Theobald, J. Comput. Chem. {\bf 31},  1561
-  (2010).
+Liu P, Agrafiotis DK, Theobald DL.
+\newblock Fast determination of the optimal rotation matrix for macromolecular
+  superpositions.
+\newblock J Comput Chem. 2010;31:1561--1563.
 
 \bibitem{theobald:j-comput-chem:2011:qcp}
-P. Liu, D.~K. Agrafiotis, and D.~L. Theobald, J. Comput. Chem. {\bf 32},  185
-  (2011).
+Liu P, Agrafiotis DK, Theobald DL.
+\newblock Reply to comment on ``Fast determination of the optimal rotation
+  matrix for macromolecular superpositions''.
+\newblock J Comput Chem. 2011;32:185--186.
 
 \bibitem{maccallum:prot:2011:casp-refinement}
-J.~L. MacCallum, A. P{\'e}rez, M.~J. Schnieders, L. Hua, M.~P. Jacobson, and
-  K.~A. Dill, Proteins: Structure, Function, and Bioinformatics {\bf 79},  74
-  (2011).
+MacCallum JL, P{\'e}rez A, Schnieders MJ, Hua L, Jacobson MP, Dill KA.
+\newblock Assessment of protein structure refinement in {CASP}9.
+\newblock Proteins: Structure, Function, and Bioinformatics. 2011
+  Jan;79(S10):74--90.
 
 \bibitem{zhang:curr-opin-struct-biol:2009:structure-prediction}
-Y. Zhang, Current Opinion in Structural Biology {\bf 19},  145  (2009).
+Zhang Y.
+\newblock Protein structure prediction: when is it useful?
+\newblock Current Opinion in Structural Biology. 2009 Apr;19(2):145--155.
 
 \bibitem{raval:prot:2012:long-timescale-md-refinement}
-A. Raval, S. Piana, M.~P. Eastwood, R.~O. Dror, and D.~E. Shaw, Proteins:
-  Structure, Function, and Bioinformatics {\bf 80},  2071  (2012).
+Raval A, Piana S, Eastwood MP, Dror RO, Shaw DE.
+\newblock Refinement of protein structure homology models via long, all-atom
+  molecular dynamics simulations.
+\newblock Proteins: Structure, Function, and Bioinformatics. 2012
+  Aug;80(8):2071--2079.
 
 \bibitem{lbfgs}
-D.~C. Liu and J. Nocedal, Mathematical Programming {\bf 45},  503  (1989).
+Liu DC, Nocedal J.
+\newblock On the Limited Memory {BFGS} Method For Large Scale Optimization.
+\newblock Mathematical Programming. 1989;45:503--528.
 
 \bibitem{amber99sb-ildn}
-K. Lindorff-Larsen, S.~P. anad Kim~Palmo, P. Maragakis, J.~L. Klepeis, R.~O.
-  Dror, and D.~E. Shaw, Proteins {\bf 78},  1950  (2010).
+Lindorff-Larsen K, anad Kim~Palmo SP, Maragakis P, Klepeis JL, Dror RO, Shaw
+  DE.
+\newblock Improved side-chain torsion potentials for the {Amber} {ff99SB}
+  protein force field.
+\newblock Proteins. 2010;78:1950--1958.
 
 \bibitem{gbsa-obc1}
-A. Onufriev, D. Bashford, and D.~A. Case, Proteins {\bf 55},  383  (2004).
+Onufriev A, Bashford D, Case DA.
+\newblock Exploring protein native states and large-scale conformational
+  changes with a modified generalized {Born} model.
+\newblock Proteins. 2004;55(22):383--394.
 
 \bibitem{basconi:jctc:2013:collision-rates}
-J.~E. Basconi and M.~R. Shirts, Journal of Chemical Theory and Computation {\bf
-  9},  2887  (2013).
+Basconi JE, Shirts MR.
+\newblock Effects of {Temperature} {Control} {Algorithms} on {Transport}
+  {Properties} and {Kinetics} in {Molecular} {Dynamics} {Simulations}.
+\newblock Journal of Chemical Theory and Computation. 2013 Jul;9(7):2887--2899.
 
 \bibitem{tip3p}
-W.~L. Jorgensen, J. Chandrasekhar, J.~D. Madura, R.~W. Impey, and M.~L. Klein,
-  Journal of Chemical Physics {\bf 79},  926  (1983).
+Jorgensen WL, Chandrasekhar J, Madura JD, Impey RW, Klein ML.
+\newblock Comparison of simple potential functions for simulating liquid water.
+\newblock Journal of Chemical Physics. 1983;79:926--935.
 
 \bibitem{tip4p-ew}
-H.~W. Horn, W.~C. Swope, J.~W. Pitera, J.~D. Madura, T.~J. Dick, G.~L. Hura,
-  and T. Head-Gordon, The Journal of Chemical Physics {\bf 120},  9665  (2004).
+Horn HW, Swope WC, Pitera JW, Madura JD, Dick TJ, Hura GL, et~al.
+\newblock Development of an improved four-site water model for biomolecular
+  simulations: {TIP}4P-{Ew}.
+\newblock The Journal of Chemical Physics. 2004 May;120(20):9665--9678.
 
 \bibitem{molprobity:2007}
-I.~W. Davis, A. Leaver-Fay, V.~B. Chen, J.~N. Block, G.~J. Kapral, X. Wang,
-  L.~W. Murray, W.~B. Arendall, J. Snoeyink, J.~S. Richardson, and D.~C.
-  Richardson, Nucleic Acids Research {\bf 35},  W375  (2007).
+Davis IW, Leaver-Fay A, Chen VB, Block JN, Kapral GJ, Wang X, et~al.
+\newblock {MolProbity}: all-atom contacts and structure validation for proteins
+  and nucleic acids.
+\newblock Nucleic Acids Research. 2007 Jul;35(suppl 2):W375--W383.
 
 \bibitem{molprobity:2010}
-V.~B. Chen, W.~B. Arendall, J.~J. Headd, D.~A. Keedy, R.~M. Immormino, G.~J.
-  Kapral, L.~W. Murray, J.~S. Richardson, and D.~C. Richardson, Acta
-  Crystallographica. Section D, Biological Crystallography {\bf 66},  12
-  (2010).
+Chen VB, Arendall WB, Headd JJ, Keedy DA, Immormino RM, Kapral GJ, et~al.
+\newblock {MolProbity}: all-atom structure validation for macromolecular
+  crystallography.
+\newblock Acta Crystallographica Section D, Biological Crystallography. 2010
+  Jan;66(Pt 1):12--21.
 
 \bibitem{berendsen:1995:gromacs}
-H.~J.~C. Berendsen, D. van~der Spoel, and R. van Drunen, Comp. Phys. Comm. {\bf
-  91},  43  (1995).
+Berendsen HJC, van~der Spoel D, van Drunen R.
+\newblock {GROMACS}: {A} message-passing parallel molecular dynamics
+  implementation.
+\newblock Comp Phys Comm. 1995 Sep;91(1-3):43--56.
 
 \bibitem{lindahl:2001:gromacs}
-E. Lindahl, B. Hess, and D. van~der Spoel, J. Mol. Model. {\bf 7},  306
-  (2001).
+Lindahl E, Hess B, van~der Spoel D.
+\newblock {GROMACS} 3.0: a package for molecular simulation and trajectory
+  analysis.
+\newblock J Mol Model. 2001;7:306--317.
 
 \bibitem{krause:nejm:2005:tks-cancer}
-D.~S. Krause and R.~A. Van~Etten, New England Journal of Medicine {\bf 353},
-  172  (2005).
+Krause DS, Van~Etten RA.
+\newblock Tyrosine {Kinases} as {Targets} for {Cancer} {Therapy}.
+\newblock New England Journal of Medicine. 2005 Jul;353(2):172--187.
 
 \bibitem{greuber:2013:abl-cancer-review}
-E.~K. Greuber, P. Smith-Pearson, J. Wang, and A.~M. Pendergast, Nature Reviews
-  Cancer {\bf 13},  559  (2013).
+Greuber EK, Smith-Pearson P, Wang J, Pendergast AM.
+\newblock Role of {ABL} family kinases in cancer: from leukaemia to solid
+  tumours.
+\newblock Nature Reviews Cancer. 2013 Aug;13(8):559--571.
 
 \bibitem{kim:2009:src-cancer-review}
-L.~C. Kim, L. Song, and E.~B. Haura, Nature Reviews Clinical Oncology {\bf 6},
-  587  (2009).
+Kim LC, Song L, Haura EB.
+\newblock Src kinases as therapeutic targets for cancer.
+\newblock Nature Reviews Clinical Oncology. 2009 Oct;6(10):587--595.
 
 \bibitem{liu:ncb:2006:kinase-inhibitors-inactive-conformations}
-Y. Liu and N.~S. Gray, Nature Chemical Biology {\bf 2},  358  (2006).
+Liu Y, Gray NS.
+\newblock Rational design of inhibitors that bind to inactive kinase
+  conformations.
+\newblock Nature Chemical Biology. 2006 Jul;2(7):358--364.
 
 \bibitem{shukla-pande:nature-commun:2014:src-activation-msm}
-D. Shukla, Y. Meng, B. Roux, and V.~S. Pande, Nature Commun. {\bf 5},  3397
-  (2014).
+Shukla D, Meng Y, Roux B, Pande VS.
+\newblock Activation pathway of {Src} kinase reveals intermediate states as
+  targets for drug design.
+\newblock Nature Commun. 2014;5:3397.
 
 \bibitem{chothia-lesk:embo-j:1986:rmsd-sequence-identity-relation}
-C. Chothia and A.~M. Lesk, EMBO J. {\bf 5},  823  (1986).
+Chothia C, Lesk AM.
+\newblock The relation between the divergence of sequence and structure in
+  proteins.
+\newblock EMBO J. 1986;5:823--826.
 
 \bibitem{roux:pnas:2013:abl-imatinib}
-Y.-L. Lin, Y. Meng, W. Jiang, and B. Roux, Proc. Natl. Acad. Sci. USA {\bf
-  110},  1664  (2013).
+Lin YL, Meng Y, Jiang W, Roux B.
+\newblock Explaining why {Gleevec} is a specific and potent inhibitor of {Abl}
+  kinase.
+\newblock Proc Natl Acad Sci USA. 2013;110:1664.
 
 \bibitem{xu:1999:2src}
-W. Xu, A. Doshi, M. Lei, M.~J. Eck, and S.~C. Harrison, Molecular Cell {\bf 3},
-   629  (1999).
+Xu W, Doshi A, Lei M, Eck MJ, Harrison SC.
+\newblock Crystal {Structures} of c-{Src} {Reveal} {Features} of {Its}
+  {Autoinhibitory} {Mechanism}.
+\newblock Molecular Cell. 1999 May;3(5):629--638.
 
 \bibitem{cowan-jacob:2005:1y57}
-S.~W. Cowan-Jacob, G. Fendrich, P.~W. Manley, W. Jahnke, D. Fabbro, J.
-  Liebetanz, and T. Meyer, Structure {\bf 13},  861  (2005).
+Cowan-Jacob SW, Fendrich G, Manley PW, Jahnke W, Fabbro D, Liebetanz J, et~al.
+\newblock The {Crystal} {Structure} of a c-{Src} {Complex} in an {Active}
+  {Conformation} {Suggests} {Possible} {Steps} in c-{Src} {Activation}.
+\newblock Structure. 2005 Jun;13(6):861--871.
 
 \bibitem{young:2006:2f4j}
-M.~A. Young, N.~P. Shah, L.~H. Chao, M. Seeliger, Z.~V. Milanov, W.~H. Biggs,
-  D.~K. Treiber, H.~K. Patel, P.~P. Zarrinkar, D.~J. Lockhart, C.~L. Sawyers,
-  and J. Kuriyan, Cancer Research {\bf 66},  1007  (2006).
+Young MA, Shah NP, Chao LH, Seeliger M, Milanov ZV, Biggs WH, et~al.
+\newblock Structure of the {Kinase} {Domain} of an {Imatinib}-{Resistant} {Abl}
+  {Mutant} in {Complex} with the {Aurora} {Kinase} {Inhibitor} {VX}-680.
+\newblock Cancer Research. 2006 Jan;66(2):1007--1014.
 
 \bibitem{cowan-jacob:2006:2hyy}
-S.~W. Cowan-Jacob, G. Fendrich, A. Floersheimer, P. Furet, J. Liebetanz, G.
-  Rummel, P. Rheinberger, M. Centeleghe, D. Fabbro, and P.~W. Manley, Acta
-  Crystallographica Section D: Biological Crystallography {\bf 63},  80
-  (2006).
+Cowan-Jacob SW, Fendrich G, Floersheimer A, Furet P, Liebetanz J, Rummel G,
+  et~al.
+\newblock Structural biology contributions to the discovery of drugs to treat
+  chronic myelogenous leukaemia.
+\newblock Acta Crystallographica Section D: Biological Crystallography. 2006
+  Dec;63(Pt 1):80--93.
 
 \bibitem{levinson:2006:2g1t}
-N.~M. Levinson, O. Kuchment, K. Shen, M.~A. Young, M. Koldobskiy, M. Karplus,
-  P.~A. Cole, and J. Kuriyan, PLoS Biol {\bf 4},  e144  (2006).
+Levinson NM, Kuchment O, Shen K, Young MA, Koldobskiy M, Karplus M, et~al.
+\newblock A {Src}-{Like} {Inactive} {Conformation} in the {Abl} {Tyrosine}
+  {Kinase} {Domain}.
+\newblock PLoS Biol. 2006 May;4(5):e144.
 
 \bibitem{kannan:jmb:2005:kinase-regulation}
-N. Kannan and A.~F. Neuwald, Journal of Molecular Biology {\bf 351},  956
-  (2005).
+Kannan N, Neuwald AF.
+\newblock Did {Protein} {Kinase} {Regulatory} {Mechanisms} {Evolve} {Through}
+  {Elaboration} of a {Simple} {Structural} {Component}?
+\newblock Journal of Molecular Biology. 2005 Sep;351(5):956--972.
 
 \bibitem{foda:nat-comm:2015:src}
-Z.~H. Foda, Y. Shan, E.~T. Kim, D.~E. Shaw, and M.~A. Seeliger, Nature
-  Communications {\bf 6},  5939  (2015).
+Foda ZH, Shan Y, Kim ET, Shaw DE, Seeliger MA.
+\newblock A dynamically coupled allosteric network underlies binding
+  cooperativity in {Src} kinase.
+\newblock Nature Communications. 2015 Jan;6:5939.
 
 \bibitem{ozkirimli:prot-sci:2008:src}
-E. Ozkirimli, S.~S. Yadav, W.~T. Miller, and C.~B. Post, Protein Science : A
-  Publication of the Protein Society {\bf 17},  1871  (2008).
+Ozkirimli E, Yadav SS, Miller WT, Post CB.
+\newblock An electrostatic network and long-range regulation of {Src} kinases.
+\newblock Protein Science : A Publication of the Protein Society. 2008
+  Nov;17(11):1871--1880.
 
 \bibitem{Su:2006:flt4cancer}
-J.-L. Su, P.-C. Yang, J.-Y. Shih, C.-Y. Yang, L.-H. Wei, C.-Y. Hsieh, C.-H.
-  Chou, Y.-M. Jeng, M.-Y. Wang, K.-J. Chang, M.-C. Hung, and M.-L. Kuo, Cancer
-  Cell {\bf 9},  209  (2006).
+Su JL, Yang PC, Shih JY, Yang CY, Wei LH, Hsieh CY, et~al.
+\newblock The {VEGF}-{C}/{Flt}-4 axis promotes invasion and metastasis of
+  cancer cells.
+\newblock Cancer Cell. 2006 Mar;9(3):209--223.
 
 \bibitem{schnieders:biophys-j:2010:x-ray-refinement}
-T.~D. Fenn, M.~J. Schnieders, A.~T. Brunger, and V.~S. Pande, Biophysical
-  Journal {\bf 98},  2984  (2010).
+Fenn TD, Schnieders MJ, Brunger AT, Pande VS.
+\newblock Polarizable {Atomic} {Multipole} {X}-{Ray} {Refinement}: {Hydration}
+  {Geometry} and {Application} to {Macromolecules}.
+\newblock Biophysical Journal. 2010 Jun;98(12):2984--2992.
 
 \bibitem{schnieders:jctc:2011:space-group-pme}
-M.~J. Schnieders, T.~D. Fenn, and V.~S. Pande, Journal of Chemical Theory and
-  Computation {\bf 7},  1141  (2011).
+Schnieders MJ, Fenn TD, Pande VS.
+\newblock Polarizable {Atomic} {Multipole} {X}-{Ray} {Refinement}: {Particle}
+  {Mesh} {Ewald} {Electrostatics} for {Macromolecular} {Crystals}.
+\newblock Journal of Chemical Theory and Computation. 2011 Apr;7(4):1141--1156.
 
 \bibitem{nagar:cell:2003:abl-1opl}
-B. Nagar, O. Hantschel, M.~A. Young, K. Scheffzek, D. Veach, W. Bornmann, B.
-  Clarkson, G. Superti-Furga, and J. Kuriyan, Cell {\bf 112},  859  (2003).
+Nagar B, Hantschel O, Young MA, Scheffzek K, Veach D, Bornmann W, et~al.
+\newblock Structural {Basis} for the {Autoinhibition} of c-{Abl} {Tyrosine}
+  {Kinase}.
+\newblock Cell. 2003 Mar;112(6):859--871.
 
 \bibitem{shan:pnas:2009:abl}
-Y. Shan, M.~A. Seeliger, M.~P. Eastwood, F. Frank, H. Xu, M.~Ã. Jensen, R.~O.
-  Dror, J. Kuriyan, and D.~E. Shaw, Proceedings of the National Academy of
-  Sciences {\bf 106},  139  (2009).
+Shan Y, Seeliger MA, Eastwood MP, Frank F, Xu H, Jensen M{\O}, et~al.
+\newblock A conserved protonation-dependent switch controls drug binding in the
+  {Abl} kinase.
+\newblock Proceedings of the National Academy of Sciences. 2009
+  Jan;106(1):139--144.
 
 \bibitem{alexov-gunner:biophys-j:1997:mcce2}
-E.~G. Alexov and M.~R. Gunner, Biophys. J. {\bf 72},  2075  (1997).
+Alexov EG, Gunner MR.
+\newblock Incorporating protein conformational flexibility into the calculation
+  of {pH}-dependent protein properties.
+\newblock Biophys J. 1997;72:2075--2093.
 
 \bibitem{georgescu-alexov-gunner:biophys-j:2002:mcce2}
-R.~E. Georgescu, E.~G. Alexov, and M.~R. Gunner, Biophys. J. {\bf 83},  1731
-  (2002).
+Georgescu RE, Alexov EG, Gunner MR.
+\newblock Combining conformational flexibility and continuum electrostatics for
+  calculating p$K_a$s in proteins.
+\newblock Biophys J. 2002;83:1731--1748.
 
 \bibitem{song-mao-gunner:j-comput-chem:2009:mcce2}
-Y. Song, J. Mao, and M.~R. Gunner, J. Comput. Chem. {\bf 30},  2231  (2009).
+Song Y, Mao J, Gunner MR.
+\newblock {MCCE2}: Improving protein p$K_a$ calculations with extensive
+  sidechain rotamer sampling.
+\newblock J Comput Chem. 2009;30:2231--2247.
 
 \bibitem{adams:prot-sci:1993:kinase-metal-ions}
-J.~A. Adams and S.~S. Taylor, Protein Science {\bf 2},  2177  (1993).
+Adams JA, Taylor SS.
+\newblock Divalent metal ions influence catalysis and active-site accessibility
+  in the {cAMP}-dependent protein kinase.
+\newblock Protein Science. 1993 Dec;2(12):2177--2186.
 
 \bibitem{sousa:book:2010:metal-ion-parameters-difficulties}
-S.~F. Sousa, R.~A. Fernandes, and M.~J. Ramos,  in {\em Kinetics and
-  {Dynamics}: {From} {Nano}- to {Bio}-{Scale}}, Vol.~12 of {\em Challenges and
-  {Advances} in {Computational} {Chemistry} and {Physics}}, edited by P.~a.
-  D.-D.~A. Paneth (Springer Science \& Business Media, Berlin, 2010), p.\ 530.
+Sousa SF, Fernandes RA, Ramos MJ.
+\newblock Molecular {Dynamics} {Simulations}: {Difficulties}, {Solutions} and
+  {Strategies} for {Treating} {Metalloenzymes}.
+\newblock In: Paneth PaDDA, editor. Kinetics and {Dynamics}: {From} {Nano}- to
+  {Bio}-{Scale}. vol.~12 of Challenges and {Advances} in {Computational}
+  {Chemistry} and {Physics}. Berlin: Springer Science \& Business Media; 2010.
+  p. 530.
 
 \bibitem{gaff}
-J. Wang, R.~M. Wolf, J.~W. Caldwell, P.~A. Kollman, and D.~A. Case, J. Comput.
-  Chem. {\bf 25},  1157  (2004).
+Wang J, Wolf RM, Caldwell JW, Kollman PA, Case DA.
+\newblock Development and testing of a general {AMBER} force field.
+\newblock J Comput Chem. 2004;25:1157--1174.
 
 \bibitem{gaff2}
-J. Wang, W. Wang, P.~A. Kollman, and D.~A. Case, J. Mol. Graph Model. {\bf 25},
-   247260  (2006).
+Wang J, Wang W, Kollman PA, Case DA.
+\newblock Automatic atom type and bond type perception in molecular mechanics
+  calculations.
+\newblock J Mol Graph Model. 2006;25:247260.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% APPENDICES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\onecolumngrid
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE: Sequences and residue numbering schemes for Src and Abl1
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[tb]
-
-\begin{centering}
-\emph{Human Abl1 sequence}
-
-\end{centering}
-
-\begin{alltt}
-1     MLEICLKLVG CKSKKGLSSS SSCYLEEALQ RPVASDFEPQ GLSEAARWNS KENLLAGPSE    60
-61    NDPNLFVALY DFVASGDNTL SITKGEKLRV LGYNHNGEWC EAQTKNGQGW VPSNYITPVN   120
-121   SLEKHSWYHG PVSRNAAEYL LSSGINGSFL VRESESSPGQ RSISLRYEGR VYHYRINTAS   180
-181   DGKLYVSSES RFNTLAELVH HHSTVADGLI TTLHYPAPKR NKPTVYGVSP NYDKWEMERT   240
-241   D{\color{red}ITMKHKLGG GQYGEVYEGV WKKYSLTVAV {\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TLKEDTMEV EEFLK{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AAVM KEIKHPNLVQ}   300
-301   {\color{red}LLGVCTREPP FYIITEFMTY GNLLDYLREC NRQEVNAVVL LYMATQISSA MEYLEKKNFI}   360
-361   {\color{red}HRDLAARNCL VGENHLVKVA DFGLS{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LMTG DTYTAHAGAK FPIKWTAPES LAYNKFSIKS}   420
-421   {\color{red}DVWAFGVLLW EIATYGMSPY PGIDLSQVYE LLEKDYRMER PEGCPEKVYE LMRACWQWNP}   480
-481   {\color{red}SDRPSFAEIH QAF}ETMFQES SISDEVEKEL GKQGVRGAVS TLLQAPELPT KTRTSRRAAE   540
-541   HRDTTDVPEM PHSKGQGESD PLDHEPAVSP LLPRKERGPP EGGLNEDERL LPKDKKTNLF   600
-601   SALIKKKKKT APTPPKRSSS FREMDGQPER RGAGEEEGRD ISNGALAFTP LDTADPAKSP   660
-661   KPSNGAGVPN GALRESGGSG FRSPHLWKKS STLTSSRLAT GEEEGGGSSS KRFLRSCSAS   720
-721   CVPHGAKDTE WRSVTLPRDL QSTGRQFDSS TFGGHKSEKP ALPRKRAGEN RSDQVTRGTV   780
-781   TPPPRLVKKN EEAADEVFKD IMESSPGSSP PNLTPKPLRR QVTVAPASGL PHKEEAGKGS   840
-841   ALGTPAAAEP VTPTSKAGSG APGGTSKGPA EESRVRRHKH SSESPGRDKG KLSRLKPAPP   900
-901   PPPAASAGKA GGKPSQSPSQ EAAGEAVLGA KTKATSLVDA VNSDAAKPSQ PGEGLKKPVL   960
-961   PATPKPQSAK PSGTPISPAP VPSTLPSASS ALAGDQPSST AFIPLISTRV SLRKTRQPPE  1020
-1021  RIASGAITKG VVLDSTEALC LAISRNSEQM ASHSAVLEAG KNLYTFCVSY VDSIQQMRNK  1080
-1081  FAFREAINKL ENNLRELQIC PATAGSGPAA TQDFSKLLSS VKEISDIVQR             1130
-\end{alltt}
-
-\begin{centering}
-\emph{Sequences for human and chicken Src, aligned using Clustal Omega}
-
-\end{centering}
-
-\begin{alltt}
-SRC_HUMAN   1     MGSNKSKPKD ASQRRRSLEP AENVHGAGGG AFPASQTPSK PASADGHRGP SAAFAPAAAE    60
-SRC_CHICK   1     MGSSKSKPKD PSQRRRSLEP PDSTH---HG GFPASQTPNK TAAPDTHRTP SRSFGTVATE    57
-                  ***.******  *********  :..*    * .*******.*  *: * ** * * :*. .*:*  
-SRC_HUMAN   61    PKLFGGFNSS DTVTSPQRAG PLAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   120
-SRC_CHICK   58    PKLFGGFNTS DTVTSPQRAG ALAGGVTTFV ALYDYESRTE TDLSFKKGER LQIVNNTEGD   117
-                  ********:* **********  ********* ********** ********** **********  
-SRC_HUMAN   121   WWLAHSLSTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNAEN PRGTFLVRES   180
-SRC_CHICK   118   WWLAHSLTTG QTGYIPSNYV APSDSIQAEE WYFGKITRRE SERLLLNPEN PRGTFLVRES   177
-                  *******:** ********** ********** ********** ******* ** **********  
-SRC_HUMAN   181   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFNSLQQLV AYYSKHADGL   240
-SRC_CHICK   178   ETTKGAYCLS VSDFDNAKGL NVKHYKIRKL DSGGFYITSR TQFSSLQQLV AYYSKHADGL   237
-                  ********** ********** ********** ********** ***.****** **********  
-SRC_HUMAN   241   CHRLTTVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   300
-SRC_CHICK   238   CHRLTNVCPT SKPQTQGLAK DAWEIPRES{\color{red}L RLEVKLGQGC FGEVWMGTWN GTTRVAI{\setlength{\fboxsep}{0pt}\colorbox{yellow}{K}}TL}   297
-                  *****.**** ********** ********** ********** ********** **********  
-SRC_HUMAN   301   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGETGKY}   360
-SRC_CHICK   298   {\color{red}KPGTMSPEAF LQ{\setlength{\fboxsep}{0pt}\colorbox{yellow}{E}}AQVMKKL RHEKLVQLYA VVSEEPIYIV TEYMSKGSLL DFLKGEMGKY}   357
-                  ********** ********** ********** ********** ********** ****** ***  
-SRC_HUMAN   361   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   420
-SRC_CHICK   358   {\color{red}LRLPQLVDMA AQIASGMAYV ERMNYVHRDL RAANILVGEN LVCKVADFGL A{\setlength{\fboxsep}{0pt}\colorbox{yellow}{R}}LIEDNEYT}   417
-                  ********** ********** ********** ********** ********** **********  
-SRC_HUMAN   421   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   480
-SRC_CHICK   418   {\color{red}ARQGAKFPIK WTAPEAALYG RFTIKSDVWS FGILLTELTT KGRVPYPGMV NREVLDQVER}   477
-                  ********** ********** ********** ********** ********** **********  
-SRC_HUMAN   481   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKEPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       536
-SRC_CHICK   478   {\color{red}GYRMPCPPEC PESLHDLMCQ CWRKDPEERP TFEYLQAFLE DYF}TSTEPQY QPGENL       533
-                  ********** ********** ****:***** ********** ********** ******
-\end{alltt}
-
-    \caption{{\bf Sequences and residue numbering schemes for Src and Abl1.}
-Kinase catalytic domains are highlighted in red, and the conserved residues analyzed in the main text (Figs.~\ref{figure:src-ref-structures} and~\ref{figure:pair-distances}) are highlighted with yellow background.
-  }
-  \label{figure:residue-numbering-schemes}
-\end{figure}
-
 \end{document}

### 2016-01-11 manuscript-post-acceptance-submission/compile.sh

@@ -1,0 +1,5 @@
+latex ensembler
+bibtex ensembler
+latex ensembler
+latex ensembler
+dvipdf ensembler

### 2016-01-15 manuscript-post-acceptance-submission/ensembler.tex
Minor edits to fix up manuscript before resubmission for production

@@
 \caption{{\bf Ensembler installation using {\tt conda} in a {\tt bash} shell.
-A license for MODELLER must first be obtained, and the license key should be used in place of the text {\tt "Modeller license key"}.}
+A license for MODELLER (free for academic use) must first be obtained, and the license key should be used in place of the text {\tt "Modeller license key"}.}
 }
@@
 
-Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
-Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
-Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-Nam condimentum sem eget mollis euismod.
-Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
-Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
-Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
-For more information, see \nameref{S1_Text}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% SUPPORTING INFORMATION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-01-17 manuscript-post-acceptance-submission/ensembler.tex
Remove support statements from Acknowledgments (PLoS has a separate section for this).

@@
 The authors are grateful to Benjamin Webb and Andrej \v{S}ali (UCSF) for help with the MODELLER package, Peter Eastman and Vijay Pande (Stanford) for assistance with OpenMM, and Marilyn Gunner (CCNY) for assistance with MCCE2, as well as the anonymous referees for constructive feedback on this manuscript.
-All authors acknowledge support from the Sloan Kettering Institute.
-JDC, KAB, and DLP acknowledge partial support from NIH grant P30 CA008748.
-JDC and DLP also acknowledge the generous support of a Louis V.~Gerstner Young Investigator Award.
-KAB was also supported in part by Starr Foundation grant I8-A8-058.
-PBG acknowledges partial funding support from the Weill Cornell Graduate School of Medical Sciences.
 

### 2016-01-17 manuscript-post-acceptance-submission/ensembler.tex
Fix typo in Sonya's name.

@@
 \\
-Daniel L. Parton\textsuperscript{1,},
-Patrick B. Grinaway\textsuperscript{1},
-Sonya M. Hansom\textsuperscript{1},
-Kyle A. Beauchamp\textsuperscript{1},
-John D. Chodera\textsuperscript{1,*},
+Daniel L.~Parton\textsuperscript{1,},
+Patrick B.~Grinaway\textsuperscript{2},
+Sonya M.~Hanson\textsuperscript{1},
+Kyle A.~Beauchamp\textsuperscript{1},
+John D.~Chodera\textsuperscript{1,*},
 \\
@@
 \bf{1} Computational Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York, NY 10065
+\\
+\bf{2} Graduate Program in Physiology, Biophysics, and Systems Biology, Weill Cornell Medical College, New York, NY 10065
 \\
